---
title: "02_Model_Selection"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)


# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  ),
  Deployment = case_when(
      Year_Collected == "2022" ~ "21/22",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2026" ~ "25/26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")

Pocilloporidae <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Pocilloporidae, Top_Pocil, 
         Under_Pocil, Side_Pocil, total)

Pocilloporidae_orientation <- Pocilloporidae %>% 
  rename(Top = Top_Pocil,
                   Bottom = Under_Pocil,
                   Side = Side_Pocil,
         Recruits_per_tile = total)

Pocilloporidae_orientation <- Pocilloporidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Acroporidae <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Acroporidae, Top_Acr, 
         Under_Acr, Side_Acr, total)

Acroporidae_orientation <- Acroporidae %>% 
  rename(Top = Top_Acr,
                   Bottom = Under_Acr,
                   Side = Side_Acr,
         Recruits_per_tile = total)

Acroporidae_orientation <- Acroporidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Poritid <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Poritid, Top_Por, Under_Por, 
         Side_Por, total)

Poritidae_orientation <- Poritid %>% 
  rename(Top = Top_Por,
                   Bottom = Under_Por,
                   Side = Side_Por,
         Recruits_per_tile = total) 

Poritidae_orientation <- Poritidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Other <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Other, Top_other, 
         Under_oth, Side_other, total)

# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(prop_acro),
    sd_aco = sd(prop_acro),
    mean_poci = mean(prop_poci),
    sd_poci = sd(prop_poci),
    mean_por = mean(prop_pori),
    sd_por = sd(prop_pori),
    mean_macroalgae = mean(Prop_macroalgae),
    sd_macroalgae = sd(Prop_macroalgae),
    PIT = n()
  )

Combined <- left_join(Coral_settlement_nGBR, join)


## Total recuits
total_recruits <- Coral_settlement_nGBR %>%
  select(Year_Collected, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 5:8, names_to = "Family", values_to = "Total")

#Order so that Other is last in Family
total_recruits <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, Reef, Site, Obs, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 6:9, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))

total_recruits$Year_Collected <- as.factor(total_recruits$Year_Collected)

## Family recruitment data
family_total_recruits <- Coral_settlement_nGBR %>%
  select(2:13) %>% 
  pivot_longer(cols = 9:12, names_to = "Family", values_to = "Total") |> 
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))
```

# Total Recruitment
```{r warning=FALSE, message=FALSE}
pois_total_yr <- glmmTMB(total ~ Shelf + Year + (1 | Reef/Site), 
      family = "poisson", data = Coral_settlement_nGBR)

nbin_total_yr <- glmmTMB(total ~ Shelf +  Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

zbin_total_yr <- glmmTMB(total ~ Shelf + Year + (1 | Reef/Site),
                      ziformula = ~1,
        family = "nbinom2", data = Coral_settlement_nGBR)

pois_total_yr2 <- glmmTMB(total ~ Shelf * Year + (1 | Reef/Site), 
      family = "poisson", data = Coral_settlement_nGBR)

nbin_total_yr2 <- glmmTMB(total ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

zbin_total_yr2 <- glmmTMB(total ~ Shelf * Year + (1 | Reef/Site),
                      ziformula = ~1,
        family = "nbinom2", data = Coral_settlement_nGBR)
```

```{r eval=FALSE}
testResiduals(sim1.1) #not good fit
plot(sim1.1)
plotResiduals(sim1.1, quantreg = FALSE, plot = FALSE) # Returns ggplot object

testResiduals(sim1.2) #not good fit
plot(sim1.2)
plotResiduals(sim1.2, quantreg = FALSE, plot = FALSE) # Returns ggplot object

testResiduals(sim1.5) #good
plot(sim1.5)
plotResiduals(sim1.5, quantreg = FALSE, plot = FALSE) # Returns ggplot object

testResiduals(sim1.6) #good
plot(sim1.6)
 plotResiduals(sim1.6, quantreg = FALSE, plot = FALSE) # Returns ggplot object
```

```{r}
sim1.3 <- simulateResiduals(nbin_total_yr)
testResiduals(sim1.3) #good
plot(sim1.3)
sim1.3_plot <- plotResiduals(sim1.3, quantreg = FALSE, plot = FALSE) # Returns ggplot object

sim1.4 <- simulateResiduals(nbin_total_yr2)
testResiduals(sim1.4) #good
plot(sim1.4)
plotResiduals(sim1.4, quantreg = FALSE, plot = FALSE) # Returns ggplot object


```

model.sel(nbin_total_yr, zbin_total_yr, nbin_total_yr2, zbin_total_yr2)
```{r}
model.sel(nbin_total_yr, zbin_total_yr, nbin_total_yr2, zbin_total_yr2)
```

# Family Recruitment
```{r poisson of total Recruits, message=FALSE}
pois_Poritid_fixed <- glmmTMB(Poritid ~ Shelf + Year+(1 | Reef/Site) , 
      family = "poisson", data = Poritid) 

pois_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
      family = "poisson", data = Poritid)

pois_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + Year+ (1 | Reef/Site), 
      family = "poisson", data = Acroporidae)

pois_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
      family = "poisson", data = Acroporidae)

pois_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf + Year  + (1 | Reef/Site), 
      family = "poisson", data = Pocilloporidae)

pois_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf*Year + (1 | Reef/Site), 
      family = "poisson", data = Pocilloporidae)

pois_other_interaction <- glmmTMB(Other ~ Shelf*Year + (1 | Reef/Site), 
      family = "poisson", data = Other)
```

Negative Binomial of Total Recruits (for a family)

```{r}
## Fixed
nbin_Poritid_fixed <- glmmTMB(Poritid ~ Shelf + (1 | Reef/Site) + Year, 
        family = "nbinom2", data = Poritid)

nbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf + (1 | Reef/Site) + Year, 
        family = "nbinom2", data = Pocilloporidae)

nbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + (1 | Reef/Site) + Year, 
        family = "nbinom2", data = Acroporidae)

## Interaction
nbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Poritid)

nbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf * Year + (1 | Reef/Site), 
        family = "nbinom2", data = Pocilloporidae)

nbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Acroporidae)

nbin_other_interaction <- glmmTMB(Other ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Other)
```


Negative Binomial Zero Inflated of Total recruits (per family)

```{r}
#Fixed
# Fit model with zero-inflated negative binomial distribution 
zbin_Poritid_fixed <- glmmTMB(Poritid ~ Shelf+ Year + (1 | Reef/Site) ,
                        ziformula = ~1,
                        family = "nbinom2",
                        data = Poritid)
zbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf * Year + (1 | Reef/Site),
                        ziformula = ~1,
                        family = "nbinom2",
                        data = Poritid)

zbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf+ Year + (1 | Reef/Site),
                               ziformula = ~1,
                        family = "nbinom2",
                        data = Pocilloporidae)

zbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf*Year + (1 | Reef/Site),
                               ziformula = ~1,
                        family = "nbinom2",
                        data = Pocilloporidae)

zbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + Year+ (1 | Reef/Site),
                            ziformula = ~1,
                        family = "nbinom2",
                        data = Acroporidae)
zbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site),
                            ziformula = ~1,
                        family = "nbinom2",
                        data = Acroporidae)
```

## Acroporidae

Simulate residuals for models

```{r}
sim2.1 <- simulateResiduals(nbin_Acroporidae_fixed) 
testResiduals(sim2.1) #fine
plot(sim2.1) #bad
sim2.1_plot <- plotResiduals(sim2.1, quantreg = FALSE, plot = FALSE) # Returns ggplot object

sim2.2 <- simulateResiduals(nbin_Acroporidae_interaction)
testResiduals(sim2.2) #fine
plot(sim2.2) #fine
sim2.2_plot <- plotResiduals(sim2.2, quantreg = FALSE, plot = FALSE) # Returns ggplot object

sim2.3 <- simulateResiduals(zbin_Acroporidae_fixed)
testResiduals(sim2.3) #fine
plot(sim2.3) #bad
sim2.3_plot <- plotResiduals(sim2.3, quantreg = FALSE, plot = FALSE) # Returns ggplot object

sim2.4 <- simulateResiduals(zbin_Acroporidae_interaction)
testResiduals(sim2.4) #fine
plot(sim2.4) #fine
sim2.4_plot <- plotResiduals(sim2.4, quantreg = FALSE, plot = FALSE) # Returns ggplot object

sim2.5 <- simulateResiduals(pois_Acroporidae_fixed) #not a good fit
testResiduals(sim2.5) #bad
plot(sim2.5) #bad
sim2.5_plot <- plotResiduals(sim2.5, quantreg = FALSE, plot = FALSE) # Returns ggplot object

sim2.6 <- simulateResiduals(pois_Acroporidae_interaction) #not a good fit
testResiduals(sim2.6) #bad
plot(sim2.6) #bad
sim2.6_plot <- plotResiduals(sim2.6, quantreg = FALSE, plot = FALSE) # Returns ggplot object
```

Compare AIC of those models

```{r}
AIC( nbin_Acroporidae_interaction, zbin_Acroporidae_interaction)

model.sel(nbin_Acroporidae_interaction, zbin_Acroporidae_interaction)
```

## Pocilloporidae
```{r}
sim3.1 <- simulateResiduals(pois_Pocilloporidae_fixed)
testResiduals(sim3.1)
sim3.1_plot <- plotResiduals(sim3.1, quantreg = FALSE, plot = FALSE) # Returns ggplot object

sim3.2 <- simulateResiduals(pois_Pocilloporidae_interaction)
testResiduals(sim3.2)
sim3.2_plot <- plotResiduals(sim3.2, quantreg = FALSE, plot = FALSE) # Returns ggplot object

sim3.3 <- simulateResiduals(nbin_Pocilloporidae_interaction)
testResiduals(sim3.3)
sim3.3_plot <- plotResiduals(sim3.3, quantreg = FALSE, plot = FALSE) # Returns ggplot object

sim3.4 <- simulateResiduals(nbin_Pocilloporidae_fixed)
testResiduals(sim3.4)
sim3.4_plot <- plotResiduals(sim3.4, quantreg = FALSE, plot = FALSE) # Returns ggplot object

sim3.5 <- simulateResiduals(zbin_Pocilloporidae_fixed)
testResiduals(sim3.5)
sim3.5_plot <- plotResiduals(sim3.5, quantreg = FALSE, plot = FALSE) # Returns ggplot object

sim3.6 <- simulateResiduals(zbin_Pocilloporidae_interaction)
testResiduals(sim3.6)
sim3.6_plot <- plotResiduals(sim3.6, quantreg = FALSE, plot = FALSE) # Returns ggplot object
```

```{r}
AIC(pois_Pocilloporidae_fixed, nbin_Pocilloporidae_fixed, zbin_Pocilloporidae_fixed, nbin_Pocilloporidae_interaction)

model.sel(pois_Pocilloporidae_fixed, nbin_Pocilloporidae_fixed, zbin_Pocilloporidae_fixed, nbin_Pocilloporidae_interaction)
```

## Poritidae
```{r}
sim4.1 <- simulateResiduals(pois_Poritid_fixed)
testResiduals(sim4.1) # bad fit

sim4.2 <- simulateResiduals(pois_Poritid_interaction)
testResiduals(sim4.2) # bad fit

sim4.3 <- simulateResiduals(nbin_Poritid_interaction)
testResiduals(sim4.3)

sim4.4 <- simulateResiduals(nbin_Poritid_fixed)
testResiduals(sim4.4)

sim4.5 <- simulateResiduals(zbin_Poritid_fixed)
testResiduals(sim4.5)

sim4.6 <- simulateResiduals(zbin_Poritid_interaction)
testResiduals(sim4.6)
```


Both nbin_interaction, and zbin_interaction for Portidae have significant dispersion.
Don't see much of a difference in Poritid across shelf position. Although slightly less in mid shore reefs (p<0.1)

```{r echo=FALSE}
AIC(nbin_Poritid_fixed, zbin_Poritid_fixed, nbin_Poritid_interaction, zbin_Poritid_interaction)

model.sel(nbin_Poritid_fixed, zbin_Poritid_fixed, nbin_Poritid_interaction, zbin_Poritid_interaction)
```


## All family option
Models of family total recruits

```{r}
pois_total_family_fixed <- glmmTMB(Total ~ Family +Shelf + Year_Collected+ (1 | Reef/Site), 
      family = "poisson",
      data = family_total_recruits)

pois_total_family_interaction <-  glmmTMB(Total ~ Family * Shelf * Year_Collected + (1 | Reef/Site), 
      family = "poisson",
      data = family_total_recruits)

nb_total_family_fixed <- glmmTMB(Total ~ Family + Shelf + Year_Collected + (1 | Reef/Site) ,
                        family = "nbinom2",
                        data = family_total_recruits)

nb_total_family_interaction <- glmmTMB(Total ~ Family * Shelf * Year_Collected + (1 | Reef/Site) ,
                        family = "nbinom2",
                        data = family_total_recruits)

zi_total_family_fixed <- glmmTMB(Total ~ Family + Shelf + Year_Collected + (1 | Reef/Site) ,
                        ziformula = ~1,
                        family = "nbinom2",
                        data = family_total_recruits)

zi_total_family_interaction <-  glmmTMB(Total ~ Family * Shelf * Year_Collected + (1 | Reef/Site) ,
                        ziformula = ~1,
                        family = "nbinom2",
                        data = family_total_recruits)
```

simulate and test residuals

```{r}
sim4.1 <- simulateResiduals(pois_total_family_fixed)
testResiduals(sim4.1) 

sim4.2 <- simulateResiduals(pois_total_family_interaction) 
testResiduals(sim4.2)

sim4.3 <- simulateResiduals(nb_total_family_fixed) #middle plot doesn't look great, but okay
testResiduals(sim4.3)

sim4.4 <- simulateResiduals(nb_total_family_interaction) #good
testResiduals(sim4.4)

sim4.5 <- simulateResiduals(zi_total_family_fixed) # middle plot doesn't look great but okay
testResiduals(sim4.5)

sim4.6 <- simulateResiduals(zi_total_family_interaction)  #good
testResiduals(sim4.6)
```

```{r}
AIC(nb_total_family_fixed, nb_total_family_interaction, zi_total_family_fixed, zi_total_family_interaction) #nb family interaction is best

model.sel(nb_total_family_fixed, nb_total_family_interaction, zi_total_family_fixed, zi_total_family_interaction)
```



# Brood Stock and Recruitment

Does Adult brood stock relate to the number of recruits, accounting for shelf position and year collected
```{r}
## Fixed
nbin_Poritid_fixed <- glmmTMB(Poritid ~ mean_por + (1 | Shelf/Site) + (1 | Year_Collected), 
        family = "nbinom2", data = Combined)

nbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ mean_poci + (1 | Shelf/Site) + (1 | Year_Collected), 
        family = "nbinom2", data = Combined)

nbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ mean_aco +(1 | Shelf/Site) + (1 | Year_Collected), 
        family = "nbinom2", data = Combined)

## Interaction
nbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*mean_por + (1|Year_Collected) + (1 | Reef/Site), 
        family = "nbinom2", data = Combined)

nbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf *mean_poci + (1|Year_Collected) + (1 | Reef/Site), 
        family = "nbinom2", data = Combined)

nbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*mean_aco + (1|Year_Collected) + (1 | Reef/Site), 
        family = "nbinom2", data = Combined)
```

```{r}
# Fixed
zbin_Poritid_fixed <- glmmTMB(Poritid ~ mean_por + (1 | Shelf/Site) + (1 | Year_Collected), 
                              ziformula = ~1,
        family = "nbinom2", data = Combined)

zbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ mean_poci + (1 | Shelf/Site) + (1 | Year_Collected),
                                     ziformula = ~1,
        family = "nbinom2", data = Combined)

zbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ mean_aco +(1 | Shelf/Site) + (1 | Year_Collected), 
                                  ziformula = ~1,
        family = "nbinom2", data = Combined)

# Fixed
pois_Poritid_fixed <- glmmTMB(Poritid ~ mean_por + (1 | Shelf/Site) + (1 | Year_Collected),
        family = "poisson", data = Combined)

pois_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ mean_poci + (1 | Shelf/Site) + (1 | Year_Collected),
        family = "poisson", data = Combined)

pois_Acroporidae_fixed <- glmmTMB(Acroporidae ~ mean_aco +(1 | Shelf/Site) + (1 | Year_Collected), 
        family = "poisson", data = Combined)
```


```{r}
# Random Slope models
nbin_Poritid_rs <- glmmTMB(Poritid ~ mean_por + (mean_por | Shelf/Site) + (mean_por | Year_Collected), 
        family = "nbinom2", data = Combined)

nbin_Pocilloporidae_rs <- glmmTMB(Pocilloporidae ~ mean_poci + (1 | Shelf/Site) + (1 | Year_Collected), 
        family = "nbinom2", data = Combined)

nbin_Acroporidae_rs <- glmmTMB(Acroporidae ~ mean_aco +(mean_aco | Shelf/Site) + (mean_aco | Year_Collected), 
        family = "nbinom2", data = Combined)
```

### Acroporidae


```{r}
sim10.1 <- simulateResiduals(nbin_Acroporidae_interaction)
testResiduals(sim10.1) # passes

sim10.2 <- simulateResiduals(nbin_Acroporidae_fixed) 
testResiduals(sim10.2) # deviation significant

sim10.3 <- simulateResiduals(zbin_Acroporidae_fixed)
testResiduals(sim10.3) # deviation significant

sim10.4 <- simulateResiduals(pois_Acroporidae_fixed) # deviation significant
testResiduals(sim10.4)# deviation significant

sim10.5 <- simulateResiduals(nbin_Acroporidae_rs)
testResiduals(sim10.5) # deviation significant

model.sel(nbin_Acroporidae_interaction)
```
```{r echo=FALSE}
Combined |> 
  ggplot()+
  geom_point(aes(x=mean_aco, y= Acroporidae))
```

The only way the Acroporidae model passes DHARMA is when there is Shelf as a interaction term, not as a random effect.

### Pocilloporidae


```{r}
sim11.1 <- simulateResiduals(nbin_Pocilloporidae_interaction)
testResiduals(sim11.1) #fine

sim11.2 <- simulateResiduals(nbin_Pocilloporidae_fixed)
testResiduals(sim11.2) #fine

sim11.3 <- simulateResiduals(nbin_Pocilloporidae_rs)
testResiduals(sim11.3) #passes

model.sel(nbin_Pocilloporidae_fixed, nbin_Pocilloporidae_interaction, nbin_Pocilloporidae_rs)
```

Interaction model performs best for Pocilloporidae

### Poritidae

```{r}
sim12.1 <- simulateResiduals(nbin_Poritid_interaction) #outlier test significant and the test on the bottom of qqplot
testResiduals(sim12.1)

sim12.2 <- simulateResiduals(nbin_Poritid_fixed)  # passes
testResiduals(sim12.2)

sim12.3 <- simulateResiduals(nbin_Poritid_rs)
testResiduals(sim12.3) # passes

model.sel(nbin_Poritid_fixed, nbin_Poritid_rs )
```
Random slope performs better for Poritidae.


# Recruitment and Juveniles
These models will need to be lagged. 
One question I should potentially ask Andy and Victor, and then if not sure on answer ask Murray is whether I should 

