---
title: "02_Model_Selection"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)


# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected),
         Shelf = factor(Shelf))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))


Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  ),
  Deployment = case_when(
      Year_Collected == "2022" ~ "21/22",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2026" ~ "25/26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")

Pocilloporidae <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Pocilloporidae, Top_Pocil, 
         Under_Pocil, Side_Pocil, total)

Acroporidae <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Acroporidae, Top_Acr, 
         Under_Acr, Side_Acr, total)


Poritid <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Poritid, Top_Por, Under_Por, 
         Side_Por, total)


Other <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Other, Top_other, 
         Under_oth, Side_other, total)

# Mean or mean no. tiles per site / year
Join_Coral <- Coral_settlement_nGBR |> 
  group_by(Shelf, Year, Site, Reef) |> 
  summarise(
    st_mean_acro = mean(Acroporidae),
    st_sd_acro = sd(Acroporidae),
    st_median_acro = median(Acroporidae),
    st_mean_poci = mean(Pocilloporidae),
    st_sd_poci = sd(Pocilloporidae),
    st_median_poci = median(Pocilloporidae),
    st_mean_por = mean(Poritid),
    st_sd_por = sd(Poritid),
    st_median_por = median(Poritid),
    st_mean_total = mean(total),
    st_sd_total = sd(total),
    Tiles = n()
  )

# Benthic Data
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    )  )


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))


PIT <- PIT %>%
  mutate(
    Acroporidae = rowSums(dplyr::select(., `Acropora - branching`:`Montipora`), na.rm = TRUE), #grouping Acroporidae
    Pocilloporidae = rowSums(dplyr::select(., Pocillopora:Stylophora), na.rm = TRUE), #grouping Pocilloporidae genera
    Poritidae = rowSums(dplyr::select(., `Porites - branching`:Goniopora), na.rm = TRUE), #grouping Poritidae genera
    `Other Hard coral` = rowSums(dplyr::select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE), # Grouping other hard corals
    `Soft coral` = rowSums(dplyr::select(., Lobophytum:`Soft-other`), na.rm = TRUE), #soft coral categories
    `Brown MA` = rowSums(dplyr::select(., Sargassum:Lobophora), na.rm = TRUE), #brown macroalgae categories
    `Red MA` = rowSums(dplyr::select(., Asparagopsis:Jania), na.rm = TRUE), #red macroalgae
    `Green MA`= rowSums(dplyr::select(., Halimeda:Valonia), na.rm = TRUE), #green macroalgae
    `Sand and Rubble` = rowSums(dplyr::select(., Sand:Rubble), na.rm = TRUE), #sand, rubble, and combo
    Turf = rowSums(dplyr::select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE), #Turfing substrata (dead corals + EAM)
    Other = rowSums(dplyr::select(., Sponge:Clam), na.rm = TRUE), #other invertes
    Zoantharian = rowSums(dplyr::select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  dplyr::select(Year, Shelf, Reef, Site, Transect, Acroporidae:Zoantharian,CCA, Cyanobacteria, Halophila)

PIT <- PIT |> 
  mutate(
    sum = rowSums(across(Acroporidae:Halophila), na.rm = TRUE) 
  )

#summary(PIT$sum) #check that the number of points = 50 per transect

PIT <- PIT |> 
  mutate(
    across(Acroporidae:Halophila, ~ .x * 2) #multiply by two to get out of 100%
  )


join_pit <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    PIT = n()
  )

Combined <- left_join(Join_Coral, join_pit)

# Juvenile data
Belt_10m <- read_csv("Belt_10m_final.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    ),
    Site = factor(Site, levels = levels(PIT$Site))
  )

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))
Belt_10m <- Belt_10m |>
  mutate(
    Other_HC = Favia + Favites + Goniastrea + Astrea +
                 Echinopora + Galaxea + Turbinaria +
                 Psammacora + Pavona
  )

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(Total_juveniles),
    sd_total = sd(Total_juveniles),
    Belt = n()
  )

## Total recuits
total_recruits <- Coral_settlement_nGBR %>%
  dplyr::select(Year_Collected, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 5:8, names_to = "Family", values_to = "Total")

#Order so that Other is last in Family
total_recruits <- Coral_settlement_nGBR %>% 
  dplyr::select(Year_Collected, Shelf, Reef, Site, Obs, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 6:9, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))

total_recruits$Year_Collected <- as.factor(total_recruits$Year_Collected)

## Family recruitment data
family_total_recruits <- Coral_settlement_nGBR %>%
  dplyr::select(2:13) %>% 
  pivot_longer(cols = 9:12, names_to = "Family", values_to = "Total") |> 
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))
```

# 1. Recruitment

## 1.1. Spatio-temporal varitaion

### 1.1.1. Total Recruitment

```{r warning=FALSE, message=FALSE}
pois_total_yr <- glmmTMB(total ~ Shelf + Year + (1 | Reef/Site), 
      family = "poisson", data = Coral_settlement_nGBR)
# sim <- simulateResiduals(pois_total_yr)
# testResiduals(sim) #not good fit Ks deviation and outlier
# testZeroInflation(sim) # p =0

zi_pois_total_yr <- glmmTMB(total ~ Shelf + Year + (1 | Reef/Site), 
                            ziformula = ~1,
      family = "poisson", data = Coral_settlement_nGBR) #exclude from model.sel
# sim <- simulateResiduals(zi_pois_total_yr)
# testResiduals(sim) #not good fit Ks deviation and outlier

pois_total_yr2 <- glmmTMB(total ~ Shelf * Year + (1 | Reef/Site), 
      family = "poisson", data = Coral_settlement_nGBR)
# sim <- simulateResiduals(pois_total_yr2)
# testResiduals(sim) #not good fit, KS deviation significant and outliers significant
# testZeroInflation(sim) # p =0

zi_pois_total_yr2 <- glmmTMB(total ~ Shelf * Year + (1 | Reef/Site), 
                             ziformula = ~1,
      family = "poisson", data = Coral_settlement_nGBR) #excluded from model.sel
# sim <- simulateResiduals(zi_pois_total_yr)
# testResiduals(sim) #outlier and KS significant

nbin_total_yr <- glmmTMB(total ~ Shelf +  Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)
# sim <- simulateResiduals(nbin_total_yr)
# testResiduals(sim) #good
# testZeroInflation(sim) #p 0.752


nbin_total_yr2 <- glmmTMB(total ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)
sim1.4 <- simulateResiduals(nbin_total_yr2)
testResiduals(sim1.4) #good
testZeroInflation(sim1.4) # p = 0.52
```


model.sel(nbin_total_yr, nbin_total_yr2)
```{r}
model.sel(nbin_total_yr, nbin_total_yr2)
```

 Family Recruitment
```{r poisson of total Recruits, message=FALSE}
# Poritidae
pois_Poritid_fixed <- glmmTMB(Poritid ~ Shelf + Year +(1 | Reef/Site) , 
      family = "poisson", data = Poritid) 

pois_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
      family = "poisson", data = Poritid)

# Acroporidae 
pois_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + Year+ (1 | Reef/Site), 
      family = "poisson", data = Acroporidae)

pois_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
      family = "poisson", data = Acroporidae)

#Pocillioporidae

pois_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf + Year  + (1 | Reef/Site), 
      family = "poisson", data = Pocilloporidae)

pois_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf*Year + (1 | Reef/Site), 
      family = "poisson", data = Pocilloporidae)

# pois_other_interaction <- glmmTMB(Other ~ Shelf*Year + (1 | Reef/Site), 
#       family = "poisson", data = Other)
```

Negative Binomial of Total Recruits (for a family)

```{r}
## Fixed
nbin_Poritid_fixed <- glmmTMB(Poritid ~ Shelf + (1 | Reef/Site) + Year, 
        family = "nbinom2", data = Poritid)

nbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf + (1 | Reef/Site) + Year, 
        family = "nbinom2", data = Pocilloporidae)

nbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + (1 | Reef/Site) + Year, 
        family = "nbinom2", data = Acroporidae)

## Interaction
nbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Poritid)

nbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf * Year + (1 | Reef/Site), 
        family = "nbinom2", data = Pocilloporidae)

nbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Acroporidae)

```


### 1.1.2. Acroporidae

Simulate residuals for models

```{r}
sim2.1 <- simulateResiduals(nbin_Acroporidae_fixed) 
testResiduals(sim2.1) #fine
testZeroInflation(sim2.1) #P<0.06
# plot(sim2.1) #bad

sim2.2 <- simulateResiduals(nbin_Acroporidae_interaction)
testResiduals(sim2.2) #fine
testZeroInflation(sim2.2) # P = 0.768
#plot(sim2.2) #fine

sim2.5 <- simulateResiduals(pois_Acroporidae_fixed) #not a good fit
testResiduals(sim2.5) #bad
testZeroInflation(sim2.5) #p <0.05, therefore i should fit a zi pois
# plot(sim2.5) #bad

sim2.6 <- simulateResiduals(pois_Acroporidae_interaction) #not a good fit
testResiduals(sim2.6) #significant deviation
testZeroInflation(sim2.6) # p = 0.064 which is pretty close
#plot(sim2.6) #bad
```


```{r}
zi_pois_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + Year+ (1 | Reef/Site),
                                     ziformula = ~1,
      family = "poisson", data = Acroporidae)

sim2.7 <- simulateResiduals(zi_pois_Acroporidae_fixed)
testResiduals(sim2.7) #significant deviation
```

Compare AIC of those models

```{r}
model.sel(nbin_Acroporidae_interaction, nbin_Acroporidae_fixed)
```

### 1.1.3 Pocilloporidae
```{r}
sim3.1 <- simulateResiduals(pois_Pocilloporidae_fixed)
testResiduals(sim3.1) #significant deviation
testZeroInflation(sim3.1) # p<0.05 so need to run a zi pois

sim3.2 <- simulateResiduals(pois_Pocilloporidae_interaction)
testResiduals(sim3.2) #significant deviation
testZeroInflation(sim3.2) #p<0.05

sim3.3 <- simulateResiduals(nbin_Pocilloporidae_interaction)
testResiduals(sim3.3) #passes
testZeroInflation(sim3.3) #passes

sim3.4 <- simulateResiduals(nbin_Pocilloporidae_fixed)
testResiduals(sim3.4) #passes
testZeroInflation(sim3.4) #p =0.72

```

```{r}
zi_pois_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf + Year  + (1 | Reef/Site), 
                                        ziformula = ~1,
      family = "poisson", data = Pocilloporidae)
sim3.7 <- simulateResiduals(zi_pois_Pocilloporidae_fixed)
testResiduals(sim3.7) #significant deviation

zi_pois_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf * Year + (1 | Reef/Site), 
                                              ziformula = ~1,
        family = "poisson", data = Pocilloporidae)

sim3.8 <- simulateResiduals(zi_pois_Pocilloporidae_interaction)
testResiduals(sim3.8)  #fails
```


```{r}
model.sel(nbin_Pocilloporidae_fixed, nbin_Pocilloporidae_interaction)
```

### 1.1.4. Poritidae
```{r}
sim4.1 <- simulateResiduals(pois_Poritid_fixed)
testResiduals(sim4.1) # bad fit
testZeroInflation(sim4.1) # P<0.05, therefore need to run zi

sim4.2 <- simulateResiduals(pois_Poritid_interaction)
testResiduals(sim4.2) # bad fit
testZeroInflation(sim4.2) # p<0.05

sim4.3 <- simulateResiduals(nbin_Poritid_interaction)
testResiduals(sim4.3) #passed
testZeroInflation(sim4.3) # p = 0.848

sim4.4 <- simulateResiduals(nbin_Poritid_fixed)
testResiduals(sim4.4) #passed
testZeroInflation(sim4.4) #p=0.88

```

Zero inflated models where required
```{r}
zi_pois_Poritid_fixed <- glmmTMB(Poritid ~ Shelf + Year+(1 | Reef/Site) ,
                                 ziformula = ~1,
      family = "poisson", data = Poritid) 

zi_pois_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
                                       ziformula = ~1,
      family = "poisson", data = Poritid)

sim4.7 <- simulateResiduals(zi_pois_Poritid_fixed)
testResiduals(sim4.7) #outliers detected

sim4.8 <- simulateResiduals(zi_pois_Poritid_interaction)
testResiduals(sim4.8) #outlier significant

```


Both nbin_interaction, and zbin_interaction for Portidae have significant dispersion.
Don't see much of a difference in Poritid across shelf position. Although slightly less in mid shore reefs (p<0.1)

```{r echo=FALSE}
model.sel(nbin_Poritid_fixed, nbin_Poritid_interaction)
```

```{r}
summary(nbin_Poritid_interaction)
```

##1.2. Composition

```{r}
# Vegetated tiles 
Recruits <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  select(Acroporidae, Pocilloporidae, Poritid, Other)

environment_data <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  select(Year_Collected, Site, Reef, Shelf)
```

```{r}
# Filtered recruit data
Recruits_sqrt <- sqrt(Recruits) #Squareroot transformed because Acroporidae is more abundant than the other taxa

# To stratify by year (this asks "do shelf differences exist after accounting for year structure?") 
h <- how(blocks = environment_data$Year_Collected)                                                   # restrict permutations within years

#Bray curtis PERMANOVA
adonis2(Recruits_sqrt ~ Shelf, 
        data = environment_data, 
        permutations = h,
        method = "bray")   # PERMANOVA

```

```{r}
#PERMDISP
# Distance matrix
bray_sqrt_recruits<- vegdist(Recruits_sqrt, method = "bray")

bd <- betadisper(bray_sqrt_recruits, environment_data$Shelf)
anova(bd)
permutest(bd) 
```

# 2. Juveniles
## 2.1. Spatio-temporal variation

### 2.1.1. Total

```{r}
pois <- glmmTMB(Total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Belt_10m)

nbin <- glmmTMB(Total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_10m)

pois2 <- glmmTMB(Total_juveniles ~ Year + Shelf+(1 | Reef/Site), 
      family = "poisson", data = Belt_10m)

nbin2 <- glmmTMB(Total_juveniles ~ Year + Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_10m)
```

```{r eval = FALSE}
sim14.1 <- simulateResiduals(pois)
testResiduals(sim14.1) #passes
testZeroInflation(sim14.1) # Is the number of observed zeros significantly different from what the model predicts? no

sim14.2 <- simulateResiduals(nbin)
testResiduals(sim14.2) #passes
testZeroInflation(sim14.2) #nope

sim <- simulateResiduals(pois2)
testResiduals(sim) #passes
testZeroInflation(sim) # Is the number of observed zeros significantly different from what the model predicts? p = 0.208

sim <- simulateResiduals(nbin2)
testResiduals(sim) #passes
testZeroInflation(sim) #p=0.584

```

```{r}
model.sel(pois, nbin, pois2, nbin2)
summary(nbin)
```

### 2.1.2. Acroporidae

```{r}
Acro <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Acroporidae)

pois <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Acro)

nbin <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Acro)

pois2 <- glmmTMB(Acroporidae ~ Year + Shelf+(1 | Reef/Site), 
      family = "poisson", data = Acro)

zi_pois2 <-  glmmTMB(Acroporidae ~ Year + Shelf+(1 | Reef/Site), 
                     ziformula =  ~ 1,
      family = "poisson", data = Acro)

nbin2 <- glmmTMB(Acroporidae ~ Year + Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Acro)
```

```{r}
sim15.1 <- simulateResiduals(pois)
testResiduals(sim15.1) # passes tests
testZeroInflation(sim15.1) # p= 0.12

sim15.2 <- simulateResiduals(nbin)
testResiduals(sim15.2) #passes
testZeroInflation(sim15.2) #p=0.416

sim <- simulateResiduals(pois2)
testResiduals(sim) #passes
testZeroInflation(sim) # Is the number of observed zeros significantly different from what the model predicts? p = 0.008, therefore will require zero inflation

sim <- simulateResiduals(zi_pois2)
testResiduals(sim) #passes
testZeroInflation(sim) #passes

sim <- simulateResiduals(nbin2)
testResiduals(sim) #passes
testZeroInflation(sim) #p=0.32
```

Both meet assumptions
```{r}
model.sel(pois, nbin, zi_pois2, nbin2) #nbin behaves better
```

### 2.1.3. Pocilloporidae

```{r}
Poc <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Pocilloporidae)

pois <- glmmTMB(Pocilloporidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Poc)

nbin <- glmmTMB(Pocilloporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Poc)

pois2 <- glmmTMB(Pocilloporidae ~ Year + Shelf+(1 | Reef/Site), 
      family = "poisson", data = Poc)

nbin2 <- glmmTMB(Pocilloporidae ~ Year + Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Poc)
```

```{r}
sim16.1 <- simulateResiduals(pois)
testResiduals(sim16.1) #passed
testZeroInflation(sim16.1) #p=0.968

sim16.2 <- simulateResiduals(nbin)
testResiduals(sim16.2) #passed
testZeroInflation(sim16.2) #p =1

sim <- simulateResiduals(pois2)
testResiduals(sim) #outlier test significant
testZeroInflation(sim) # Is the number of observed zeros significantly different from what the model predicts? p = 0.784

sim <- simulateResiduals(nbin2)
testResiduals(sim) #passes
testZeroInflation(sim) #p=0.928
```
```{r}
model.sel(pois, nbin, pois2, nbin2) #nbin2 behaves better
```

### 2.1.4. Poritidae

```{r}
Por <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Poritidae)

pois <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Por)

nbin <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Por)

pois2 <- glmmTMB(Poritidae ~ Year + Shelf+(1 | Reef/Site), 
      family = "poisson", data = Por)

nbin2 <- glmmTMB(Poritidae ~ Year + Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Por)
```

```{r}
sim <- simulateResiduals(pois)
testResiduals(sim) # passes
testZeroInflation(sim) #p=0.248

sim <- simulateResiduals(nbin)
testResiduals(sim) #passes
testZeroInflation(sim) # p=0.784

sim <- simulateResiduals(pois2)
testResiduals(sim) #passes
testZeroInflation(sim) # Is the number of observed zeros significantly different from what the model predicts? p = 0.416

sim <- simulateResiduals(nbin2)
testResiduals(sim) #passes
testZeroInflation(sim) #p=0.904
```

Both models have outliers as significant.
```{r}
model.sel(pois, nbin, pois2, nbin2) #nbin behaves better
```

## 2.2. Composition
```{r}
Belt_10m <- Belt_10m |> 
  rename("Other Hard Coral" = Other_HC
  )
# Step 1: Create site-by-species matrix with Site_ID
site_by_belt <- Belt_10m %>%
  filter(Total_juveniles > 0) |> #filter to make sure that dissimilarities are non zero
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, Acroporidae, Pocilloporidae, Poritidae, `Other Hard Coral`
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))


# Step 2: Metadata (KEEP Site_ID)
metadata_belt <- site_by_belt %>%
  dplyr::select(Shelf, Year, Reef)

# Step 3: Species matrix
belt_matrix_filtered <- site_by_belt %>%
  dplyr::select(
    6:9
  )

# Check whether any family has much higher abundances than other families
summary(belt_matrix_filtered$Acroporidae)
summary(belt_matrix_filtered$Pocilloporidae)
summary(belt_matrix_filtered$Poritidae)
summary(belt_matrix_filtered$`Other Hard Coral`)

#Think Ill just do raw abundances then

```

```{r}
# create a distance matrix
belt_dist <- vegdist(belt_matrix_filtered, method = "bray")

# With respect to Shelf
bd <- betadisper(belt_dist, metadata_belt$Shelf)
anova(bd) # Significant was <0.001
permutest(bd) #0.049 on Permutest, then 0.06
```

```{r}
# To stratify by year (this asks "do shelf differences exist after accounting for year structure?") 
h <- how(blocks = metadata_belt$Year)                                                   # restrict permutations within years

#Bray curtis PERMANOVA
permanova_result <- adonis2(belt_matrix_filtered ~ Shelf, 
        data = metadata_belt, 
        permutations = h,
        method = "bray")   # PERMANOVA



print(permanova_result) #p =0.001
```

### 2.1.2. NMDS

```{r}
nmds <- metaMDS(belt_matrix_filtered, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress #0.16
```

```{r}
# Creating Vectors for driving the patterns
belt_envfit <- envfit(
  nmds,
  belt_matrix_filtered,
  permutations = 999
) 

# Extract p-values
pvals <- belt_envfit$vectors$pvals

# Extract r-squared values
r2 <- belt_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(belt_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- belt_envfit$vectors$r
vec_scores$pval <- belt_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

belt_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")
```

# 3. Benthic Cover

## 3.1. Composition

### 3.1.1. PERMANOVA
```{r}
# Step 1: Create site-by-species matrix with Site_ID
site_by_PIT <- PIT %>%
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, 6:20
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  dplyr::select(Shelf, Year, Reef)

# Step 3: Species matrix
PIT_matrix_filtered <- site_by_PIT %>%
  dplyr::select(
    Acroporidae:Halophila
  )
```

```{r}
h <- how(blocks = metadata_PIT$Year)   
# create a distance matrix
PIT_dist <- vegdist(PIT_matrix_filtered, method = "bray")

# With respect to Shelf
betadisper(PIT_dist, metadata_PIT$Shelf)
anova(betadisper(PIT_dist, metadata_PIT$Shelf)) # Significant was 0.01143 * ***
permutest(betadisper(PIT_dist, metadata_PIT$Shelf), permutations = h) #0.049 on Permutest, then 0.06
```

```{r}

# To stratify by year (this asks "do shelf differences exist after accounting for year structure?") 
h <- how(blocks = metadata_PIT$Year)                                                   # restrict permutations within years
permanova_result <- adonis2(PIT_matrix_filtered ~ Shelf,
  data = site_by_PIT,
  method = "bray",
  permutations = h)

print(permanova_result)
```

### 3.1.2. NMDS

```{r}
nmds <- metaMDS(PIT_matrix_filtered, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress
```

```{r}
# Creating Vectors for driving the patterns
Benthic_envfit <- envfit(
  nmds,
  PIT_matrix_filtered,
  permutations = h
) 

# Extract p-values
pvals <- Benthic_envfit$vectors$pvals

# Extract r-squared values
r2 <- Benthic_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(Benthic_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- Benthic_envfit$vectors$r
vec_scores$pval <- Benthic_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

Benthic_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")
```

## 3.2. Family benthic cover
```{r}
PIT_adult <- PIT |> 
  dplyr::select(Year, Shelf, Reef, Site, Transect,Acroporidae, Pocilloporidae, Poritidae, `Other Hard coral`)
```

### 3.2.1. Acroporidae

```{r}
# Acroporidae adult Model
pois_acro_adult <- glmmTMB(Acroporidae ~ Shelf + Year + (1 | Reef/Site), 
      family = "poisson", data = PIT_adult)
# sim <- simulateResiduals(pois_acro_adult)
# testResiduals(sim) #passes
# testZeroInflation(sim) #p = 0.112

nbin_acro_adult <- glmmTMB(Acroporidae ~ Shelf +  Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT_adult)
# sim <- simulateResiduals(nbin_acro_adult)
# testResiduals(sim) #passes
# testZeroInflation(sim) #p = 0.224

pois_acro_adult_int <- glmmTMB(Acroporidae ~ Shelf * Year + (1 | Reef/Site), 
      family = "poisson", data = PIT_adult)
# sim <- simulateResiduals(pois_acro_adult_int)
# testResiduals(sim)  #passes
# testZeroInflation(sim) #p = 0.16

nbin_acro_adult_int <- glmmTMB(Acroporidae ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT_adult)
# sim <- simulateResiduals(nbin_acro_adult_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #p=0.16
```

```{r}
model.sel(pois_acro_adult, pois_acro_adult_int, nbin_acro_adult, nbin_acro_adult_int)
```

```{r}
summary(nbin_acro_adult_int)
```
High variability among sites (56.7% of variance), whereas reef only accounts for 0.5% on top of what site within reef.

### 3.2.2. Pocilloporidae

```{r}
pois_poci_adult <- glmmTMB(Pocilloporidae ~ Shelf + Year + (1 | Reef/Site), 
      family = "poisson", data = PIT_adult)
# sim <- simulateResiduals(pois_poci_adult)
# testResiduals(sim)#passes
# testZeroInflation(sim) # p =0.6

nbin_poci_adult <- glmmTMB(Pocilloporidae ~ Shelf +  Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT_adult)
# sim <- simulateResiduals(nbin_poci_adult)
# testResiduals(sim) #passes
# testZeroInflation(sim) # p = 0.672

pois_poci_adult_int <- glmmTMB(Pocilloporidae ~ Shelf * Year + (1 | Reef/Site), 
      family = "poisson", data = PIT_adult)
# sim <- simulateResiduals(pois_poci_adult_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #p=0.376

nbin_poci_adult_int <- glmmTMB(Pocilloporidae ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT_adult)
# sim <- simulateResiduals(nbin_poci_adult_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) # p = 0.456
```

```{r}
model.sel(pois_poci_adult, pois_poci_adult_int, nbin_poci_adult, nbin_poci_adult_int)
```

```{r}
summary(nbin_poci_adult_int)
```

Sites and reef account  for very little of the variation seen in shelf and year.

### 3.2.3. Poritidae

```{r}
pois_por_adult <- glmmTMB(Poritidae ~ Shelf + Year + (1 | Reef/Site), 
      family = "poisson", data = PIT_adult)
#  sim <- simulateResiduals(pois_por_adult)
#  testResiduals(sim)#passes
# testZeroInflation(sim) # p =1

nbin_por_adult <- glmmTMB(Poritidae ~ Shelf +  Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT_adult)
# sim <- simulateResiduals(nbin_por_adult)
# testResiduals(sim) #passes
# testZeroInflation(sim) # p = 0.888

pois_por_adult_int <- glmmTMB(Poritidae ~ Shelf * Year + (1 | Reef/Site), 
      family = "poisson", data = PIT_adult)
# sim <- simulateResiduals(pois_por_adult_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #p=0.816

nbin_por_adult_int <- glmmTMB(Poritidae ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT_adult)
# sim <- simulateResiduals(nbin_por_adult_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) # p = 0.816
```


```{r}
model.sel(nbin_por_adult_int, nbin_por_adult, pois_por_adult, pois_por_adult_int)
```
Interaction term doesn't significantly improve model for Poritidae

```{r}
summary(nbin_por_adult)
```


### 3.2.4. Total Hard coral
```{r}
PIT <- PIT |> 
  mutate(
    Total_HC_Cover = Acroporidae+ Pocilloporidae + Poritidae + `Other Hard coral`
  )

PIT$Total_HC_Cover |> summary() #range of Total HC

PIT |> summarise(mean_HC = mean(Total_HC_Cover),
                 se_HC = sd(Total_HC_Cover)/n()^0.5)
```
```{r}
PIT$Total_HC_Cover |> summary()
PIT$Acroporidae |> summary()

PIT$Pocilloporidae |> summary()

PIT$Poritidae |> summary()
```


```{r}
pois_HC_adult <- glmmTMB(Total_HC_Cover ~ Shelf + Year + (1 | Reef/Site), 
      family = "poisson", data = PIT)
#  sim <- simulateResiduals(pois_HC_adult)
#  testResiduals(sim)#passes
# testZeroInflation(sim) # p = 0.88, consider zero inflation

nbin_HC_adult <- glmmTMB(Total_HC_Cover ~ Shelf +  Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)
# sim <- simulateResiduals(nbin_HC_adult)
# testResiduals(sim) #passes
# testZeroInflation(sim) # p = 0.072

pois_HC_adult_int <- glmmTMB(Total_HC_Cover ~ Shelf * Year + (1 | Reef/Site), 
      family = "poisson", data = PIT)
# sim <- simulateResiduals(pois_HC_adult_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #p=0.104

nbin_HC_adult_int <- glmmTMB(Total_HC_Cover ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)
# sim <- simulateResiduals(nbin_HC_adult_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) # p = 0.136

```

```{r}
model.sel(pois_HC_adult_int, pois_HC_adult, nbin_HC_adult, nbin_HC_adult_int)
```
```{r}
summary(nbin_HC_adult_int)
```


# 4. Brood stock (PIT) and Recruitment (Tiles)

```{r}
st_join <- Coral_settlement_nGBR |> 
  group_by(Site, Year) |> 
  summarise(
    acro_mean_st = mean(Acroporidae),
    poci_mean_st = mean(Pocilloporidae),
    por_mean_st = mean(Poritid),
    acro_sd_st = sd(Acroporidae),
    poci_sd_st = sd(Pocilloporidae),
    por_sd_st = sd(Poritid)
    # acro_median_st = median(Acroporidae),
    # poci_median_st = median(Pocilloporidae),
    # por_median_st = median(Poritid),
  )

combined_averages <- left_join(join_pit, st_join)
```

Now these are both averages, it can't follow an integar and would need to follow a different family of error distribution.

## 4.1. Acroporidae

Despite not being interested in across the shelf differences, as shelf has only three levels there are two few levels for a stable random effect.
We are now interested in shelf effect, so add as a fixed and interaction term and then compare the AICC.
```{r}
Acro <- combined_averages |> 
  filter(acro_mean_st > 0)
```

```{r}
gamma_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco+(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Acro)
# sim <- simulateResiduals(gamma_avg_Acroporidae_fixed)
# testResiduals(sim) #passes

ln_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Reef/Site), 
        family = "lognormal", data = Acro)
# sim <- simulateResiduals(ln_avg_Acroporidae_fixed)
# testResiduals(sim) #passes


gau_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco + (1 | Reef/Site), 
        family = "gaussian", data = Acro)
# sim <- simulateResiduals(gau_avg_Acroporidae_fixed)
# testResiduals(sim) #fails

#interaction with Shelf
gamma_avg_Acroporidae_int <- glmmTMB(acro_mean_st ~ mean_aco*Shelf +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Acro)
# sim <- simulateResiduals(gamma_avg_Acroporidae_int)
# testResiduals(sim) #passes


ln_avg_Acroporidae_int <- glmmTMB(acro_mean_st ~ mean_aco*Shelf +(1 | Reef/Site), 
        family = "lognormal", data = Acro)
# sim <- simulateResiduals(ln_avg_Acroporidae_int)
# testResiduals(sim) #passes


gaus_avg_Acroporidae_int <- glmmTMB(acro_mean_st ~ mean_aco*Shelf +(1 | Reef/Site), 
        family = "gaussian", data = Acro)
# sim <- simulateResiduals(gaus_avg_Acroporidae_int)
# testResiduals(sim) #terrible, truly.
# testZeroInflation(sim)

gamma_avg_Acroporidae_fixed2 <- glmmTMB(acro_mean_st ~ mean_aco + Shelf+ (1 | Reef/Site), 
         family = Gamma(link = "log"), data = Acro)
# sim <- simulateResiduals(gamma_avg_Acroporidae_fixed2)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

ln_avg_Acroporidae_fixed2 <- glmmTMB(acro_mean_st ~ mean_aco +Shelf +(1 | Reef/Site), 
        family = "lognormal", data = Acro)
# sim <- simulateResiduals(ln_avg_Acroporidae_fixed2)
# testResiduals(sim) #passes
# testZeroInflation(sim)

gau_avg_Acroporidae_fixed2 <- glmmTMB(acro_mean_st ~ mean_aco +Shelf + (1 | Reef/Site), 
        family = "gaussian", data = Acro)
# sim <- simulateResiduals(gau_avg_Acroporidae_fixed2)
# testResiduals(sim) #fails
# testZeroInflation(sim) #passes
```

Log normal models worked best for all of these
```{r}
model.sel(gamma_avg_Acroporidae_fixed, ln_avg_Acroporidae_fixed, 
          gamma_avg_Acroporidae_fixed2, ln_avg_Acroporidae_fixed2,
          gamma_avg_Acroporidae_int, ln_avg_Acroporidae_int)

summary(gamma_avg_Acroporidae_int)
```


## 4.2. Pocilloporidae

```{r}
## Pocilloporidae
Poci <- combined_averages |> 
  filter(poci_mean_st > 0)

ln_avg_poci_fixed <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Reef/Site), 
        family = "lognormal", data = Poci)
# sim <- simulateResiduals(ln_avg_poci_fixed)
# testResiduals(sim) #passes

ln_avg_poci_fixed2 <- glmmTMB(poci_mean_st ~ mean_poci +Shelf +(1 | Reef/Site),
        family = "lognormal", data = Poci)
# # sim <- simulateResiduals(ln_avg_poci_fixed2)
# # testResiduals(sim) #passes

gaus_avg_poci_fixed <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Reef/Site), 
        family = "gaussian", data = Poci)
# sim <- simulateResiduals(gaus_avg_poci_fixed)
# testResiduals(sim) #outlier test significant

gaus_avg_poci_fixed2 <- glmmTMB(poci_mean_st ~ mean_poci + Shelf + (1 | Reef/Site), 
        family = "gaussian", data = Poci)
# sim <- simulateResiduals(gaus_avg_poci_fixed)
# testResiduals(sim) #passes

gamma_avg_Pocilloporidae_fixed <- glmmTMB(poci_mean_st ~ mean_poci + (1 | Reef/Site),
        family = Gamma(link = "log"), data = Poci)
# sim <- simulateResiduals(gamma_avg_Pocilloporidae_fixed)
# testResiduals(sim) #passes

gamma_avg_Pocilloporidae_fixed2 <- glmmTMB(poci_mean_st ~ mean_poci + Shelf+ (1 | Reef/Site),
        family = Gamma(link = "log"), data = Poci)
# # sim <- simulateResiduals(gamma_avg_Pocilloporidae_fixed2)
# # testResiduals(sim) #passes

ln_avg_Pocilloporidae_int <- glmmTMB(poci_mean_st ~ mean_poci * Shelf +(1 | Reef/Site),
         family = "lognormal", data = Poci)
# sim <- simulateResiduals(ln_avg_Pocilloporidae_int)
# # testResiduals(sim) #passes
# 
gau_avg_Pocilloporidae_int <- glmmTMB(poci_mean_st ~ mean_poci * Shelf +(1 | Reef/Site),
         family = "gaussian", data = Poci)
# # sim <- simulateResiduals(gau_avg_Pocilloporidae_int)
# # testResiduals(sim) #passes
# 
gamma_avg_Pocilloporidae_int <- glmmTMB(poci_mean_st ~ mean_poci * Shelf +(1 | Reef/Site),
         family = Gamma(link = "log"), data = Poci)
# # sim <- simulateResiduals(gamma_avg_Pocilloporidae_int)
# # testResiduals(sim) #passes
```

```{r}
model.sel(gamma_avg_Pocilloporidae_fixed, ln_avg_poci_fixed,
          ln_avg_poci_fixed2, gamma_avg_Pocilloporidae_fixed2, gaus_avg_poci_fixed2,
          gamma_avg_Pocilloporidae_int, gau_avg_Pocilloporidae_int, ln_avg_Pocilloporidae_int)
```
Without shelf performs slightly better

```{r}
summary(ln_avg_poci_fixed)
summary(ln_avg_poci_fixed2)
```

## 4.3. Poritidae

```{r}
## Poritidae
Por <- combined_averages |> 
  filter(por_mean_st > 0)

gamma_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Reef/Site),
        family = Gamma(link = "log"), data = Por)
# sim <- simulateResiduals(gamma_avg_Poritid_fixed) #passed
# testResiduals(sim)


ln_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Reef/Site), 
        family = "lognormal", data = Por)
# sim <- simulateResiduals(ln_avg_Poritid_fixed) # very good
# testResiduals(sim)


gaus_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Reef/Site), 
        family = "gaussian", data = Por)
# sim <- simulateResiduals(gaus_avg_Poritid_fixed)
# testResiduals(sim) # definitely not this one

gamma_avg_Poritid_fixed2 <- glmmTMB(por_mean_st ~ mean_por + Shelf + (1 | Reef/Site),
        family = Gamma(link = "log"), data = Por)
# sim <- simulateResiduals(gamma_avg_Poritid_fixed2)
# testResiduals(sim) #passes

ln_avg_Poritid_fixed2 <- glmmTMB(por_mean_st ~ mean_por + Shelf +(1 | Reef/Site), 
        family = "lognormal", data = Por)
# sim <- simulateResiduals(ln_avg_Poritid_fixed2)
# testResiduals(sim) #passes

gaus_avg_Poritid_fixed2 <- glmmTMB(por_mean_st ~ mean_por + Shelf+ (1 | Reef/Site), 
        family = "gaussian", data = Por)
# sim <- simulateResiduals(gaus_avg_Poritid_fixed2)
# testResiduals(sim) # definitely not this one


gamma_avg_Poritid_int <- glmmTMB(por_mean_st ~ mean_por*Shelf + (1 | Reef/Site),
        family = Gamma(link = "log"), data = Por)
# sim <- simulateResiduals(gamma_avg_Poritid_int)
# testResiduals(sim) #passes

ln_avg_Poritid_int <- glmmTMB(por_mean_st ~ mean_por * Shelf + (1 | Reef/Site), 
        family = "lognormal", data = Por)
# sim <- simulateResiduals(ln_avg_Poritid_int)
# testResiduals(sim) #passes

gaus_avg_Poritid_int <- glmmTMB(por_mean_st ~ mean_por * Shelf + (1 | Reef/Site), 
        family = "gaussian", data = Por)
# sim <- simulateResiduals(gaus_avg_Poritid_int)
# testResiduals(sim) # definitely not this one

```


```{r}
model.sel(gamma_avg_Poritid_fixed, ln_avg_Poritid_fixed, ln_avg_Poritid_fixed2, gamma_avg_Poritid_fixed2, ln_avg_Poritid_int, gamma_avg_Poritid_int)
```


```{r}
summary(ln_avg_Poritid_fixed)
```

All of the models had significant intercepts, but the estimate for mean_por was not significant in any of the models.

*Note the very small variances explained by the conditional model. Perhaps should not include them, but then the model structure would be different to other families.*


# 5. Recruitment and Juveniles
These models will need to be lagged. 

```{r}
combine_belt_st <- left_join(join_belt, Join_Coral)


# Only including years in which both data is included (once lagging has happened)
combine_belt_st <- combine_belt_st |> 
  mutate(Year_start = as.numeric(substr(Year, 1, 4)))

gropd_df2 <- combine_belt_st %>%
  arrange(Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_mean_acro_lag1 = lag(st_mean_acro, 1),
    st_mean_poci_lag1 = lag(st_mean_poci, 1),
    st_mean_por_lag1 = lag(st_mean_por, 1),
  ) %>%
  ungroup()
```


## 5.1. Acroporidae

Remove first year row
```{r}
model_df_acro <- gropd_df2 %>%
  filter(!is.na(mean_acro))
```

Fit models
```{r}
ln_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)

# sim <- simulateResiduals(ln_acro_juv_st_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim)

gau_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 +Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)

# sim <- simulateResiduals(gau_acro_juv_st_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim)

gamma_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 +Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_acro)

# sim <- simulateResiduals(gau_acro_juv_st_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passed


ln_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)

# sim <- simulateResiduals(ln_acro_juv_st_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

gau_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)

# sim <- simulateResiduals(gau_acro_juv_st_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

gamma_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_acro)

# sim <- simulateResiduals(gamma_acro_juv_st_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

# Not including shelf, accounting for reef and site

ln_acro_juv_st <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)

# sim <- simulateResiduals(ln_acro_juv_st)
# testResiduals(sim) #passes
# testZeroInflation(sim)

gau_acro_juv_st <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)

# sim <- simulateResiduals(gau_acro_juv_st)
# testResiduals(sim) #passes
# testZeroInflation(sim)

gamma_acro_juv_st <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_acro)
# sim <- simulateResiduals(gamma_acro_juv_st)
# testResiduals(sim) #passes
# testZeroInflation(sim)
```

```{r}
model.sel(gau_acro_juv_st_int, 
          gau_acro_juv_st_fixed, gamma_acro_juv_st_fixed,
          gamma_acro_juv_st_int,  ln_acro_juv_st_int,  ln_acro_juv_st_fixed,
          gau_acro_juv_st, gamma_acro_juv_st, ln_acro_juv_st)
```

```{r}
ln_acro_juv_st_int

summary(ln_acro_juv_st_int)
```

## 5.2. Pocilloporidae

Remove first years rows
```{r}
model_df_poci <- gropd_df2 %>%
  filter(!is.na(mean_poci))
```


```{r}
ln_poci_juv_st_fixed <- glmmTMB(mean_poci ~ st_mean_poci_lag1 + Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)

# sim <- simulateResiduals(ln_poci_juv_st_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim)


gau_poci_juv_st_fixed <- glmmTMB(mean_poci ~ st_mean_poci_lag1+Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)

# sim <- simulateResiduals(gau_poci_juv_st_fixed)
# testResiduals(sim) #passes but doesn't look great
# testZeroInflation(sim)

gamma_poci_juv_st_fixed <- glmmTMB(mean_poci ~ st_mean_poci_lag1+ Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_poci)

# sim <- simulateResiduals(gau_poci_juv_st_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim)

#Interaction with Shelf
ln_poci_juv_st_int <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)
# sim <- simulateResiduals(ln_poci_juv_st_int)
# testResiduals(sim) # passed
# testZeroInflation(sim)

gau_poci_juv_st_int<- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)

# sim <- simulateResiduals(gau_poci_juv_st_int)
# testResiduals(sim) # passed but doesn't look great
# testZeroInflation(sim)

gamma_poci_juv_st_int<- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_poci)

# sim <- simulateResiduals(gamma_poci_juv_st_int)
# testResiduals(sim) # passed
# testZeroInflation(sim)


ln_poci_juv_st <- glmmTMB(mean_poci ~ st_mean_poci_lag1  + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)
# sim <- simulateResiduals(ln_poci_juv_st)
# testResiduals(sim) # passed
# testZeroInflation(sim)

gau_poci_juv_st <- glmmTMB(mean_poci ~ st_mean_poci_lag1 + (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)

# sim <- simulateResiduals(gau_poci_juv_st)
# testResiduals(sim) # passed but doesn't look great
# testZeroInflation(sim)

gamma_poci_juv_st <- glmmTMB(mean_poci ~ st_mean_poci_lag1 + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_poci)

# sim <- simulateResiduals(gamma_poci_juv_st)
# testResiduals(sim) # significant deviation
# testZeroInflation(sim)
```


```{r}
model.sel(gau_poci_juv_st_fixed, gau_poci_juv_st_int, 
          #gau_poci_juv_st_nested, gamma_poci_juv_st_nested,ln_poci_juv_st_nested,
gamma_poci_juv_st_fixed, gamma_poci_juv_st_int,  ln_poci_juv_st_fixed, ln_poci_juv_st_int,
ln_poci_juv_st, gau_poci_juv_st)
```

```{r}
ln_poci_juv_st_fixed

summary(ln_poci_juv_st_fixed)
```

## 5.3. Poritidae

Remove first year rows
```{r}
model_df_por <- gropd_df2 %>%
  filter(!is.na(mean_por))
```

```{r}
model_df_por <- model_df_por |> 
  filter(mean_por > 0)

ln_por_juv_st_fixed <- glmmTMB(mean_por ~ st_mean_por_lag1 + Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_por)

sim <- simulateResiduals(ln_por_juv_st_fixed)
testResiduals(sim) # passsed
testZeroInflation(sim)

gau_por_juv_st_fixed <- glmmTMB(mean_por ~ st_mean_por_lag1 + Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_por)

sim <- simulateResiduals(gau_por_juv_st_fixed)
testResiduals(sim) # passed but not great
testZeroInflation(sim)

gamma_por_juv_st_fixed <- glmmTMB(mean_por ~ st_mean_por_lag1 + Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_por)

sim <- simulateResiduals(gamma_por_juv_st_fixed)
testResiduals(sim) # passsed
testZeroInflation(sim)

#Interaction
ln_por_juv_st_int <- glmmTMB(mean_por ~ st_mean_por_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_por)
sim <- simulateResiduals(ln_por_juv_st_int)
testResiduals(sim) # passed
testZeroInflation(sim)

gau_por_juv_st_int<- glmmTMB(mean_por ~ st_mean_por_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_por)

sim <- simulateResiduals(gau_por_juv_st_int)
testResiduals(sim) # passed but outlier test very close to being significant
testZeroInflation(sim)

gamma_por_juv_st_int <- glmmTMB(mean_por ~ st_mean_por_lag1 * Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_por)
sim <- simulateResiduals(gamma_por_juv_st_int)
testResiduals(sim) # passed
testZeroInflation(sim)

# Shelf/Reef/Site
ln_por_juv_st_nested <- glmmTMB(mean_por ~ st_mean_por_lag1+ (1|Shelf/Reef/Site),
              family = "lognormal",
           data = model_df_por)

sim <- simulateResiduals(ln_por_juv_st_nested)
testResiduals(sim) # passsed
testZeroInflation(sim)

gau_por_juv_st_nested <- glmmTMB(mean_por ~ st_mean_por_lag1 + (1|Shelf/Reef/Site),
              family = "gaussian",
           data = model_df_por)

sim <- simulateResiduals(gau_por_juv_st_nested)
testResiduals(sim) # passed but not great
testZeroInflation(sim)

gamma_por_juv_st_nested <- glmmTMB(mean_por ~ st_mean_por_lag1+ (1|Shelf/Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_por)

sim <- simulateResiduals(gamma_por_juv_st_nested)
testResiduals(sim) # passsed
testZeroInflation(sim)

```

```{r}
model.sel(gau_por_juv_st_fixed, gau_por_juv_st_int, gau_por_juv_st_nested, ln_por_juv_st_fixed, ln_por_juv_st_int, ln_por_juv_st_nested, gamma_por_juv_st_fixed, gamma_por_juv_st_int, gamma_por_juv_st_nested)

```

```{r}
gamma_por_juv_st_fixed

summary(gamma_por_juv_st_fixed)
```

```{r}
library(broom.mixed)
library(openxlsx)
library(dplyr)
tidy_acro <- tidy(ln_acro_juv_st_int, effects = "fixed", conf.int = TRUE)
tidy_poci <- tidy(ln_poci_juv_st_fixed, effects = "fixed", conf.int = TRUE)
tidy_por <- tidy(gamma_por_juv_st_fixed, effects = "fixed", conf.int = TRUE)
glance_acro <- glance(ln_acro_juv_st_int)
glance_poci <- glance(ln_poci_juv_st_fixed)
glance_por  <- glance(gamma_por_juv_st_fixed)
wb <- createWorkbook()

addWorksheet(wb, "Acropora")
writeData(wb, "Acropora", tidy_acro)
writeData(wb, "Acropora", glance_acro, startRow = nrow(tidy_acro) + 3)

addWorksheet(wb, "Pocillopora")
writeData(wb, "Pocillopora", tidy_poci)
writeData(wb, "Pocillopora", glance_poci, startRow = nrow(tidy_poci) + 3)

addWorksheet(wb, "Porites")
writeData(wb, "Porites", tidy_por)
writeData(wb, "Porites", glance_por, startRow = nrow(tidy_por) + 3)

saveWorkbook(wb, "output/coral_model_summaries.xlsx", overwrite = TRUE)
```


# 6. Brood stock and juveniles

```{r}
# Benthic Data
PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    )  )


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

PIT_detailed <- PIT %>%
  mutate(
    Acropora = rowSums(pick(`Acropora - branching`:`Acropora - tables`), na.rm = TRUE),
    Porites  = rowSums(pick(`Porites - branching`:`Porites - massive`), na.rm = TRUE)
  ) %>%
  dplyr::select(Year, Shelf, Reef, Site, Transect, Acropora, Isopora, Montipora, Pocillopora, 
                Seriatopora, Stylophora, Porites, Goniopora) %>%
  mutate(across(Acropora:Goniopora, ~ .x * 2))

join_PIT_detailed <- PIT_detailed |> pivot_longer(cols = Acropora:Goniopora, names_to = "genus", values_to = "Percent_cover")
join_PIT_detailed <- join_PIT_detailed |> 
  group_by(Year, Shelf, Reef, Site, genus) |> 
  summarise(
    mean_cover = mean(Percent_cover),
    sd_cover = sd(Percent_cover)
  )

```

```{r}
Belt_detailed <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Transect, Acropora, Isopora, Montipora, Pocillopora, 
                Seriatopora, Stylophora, Porites)

belt_detailed_long <- Belt_detailed |>  pivot_longer(cols = Acropora:Porites, names_to = "genus", values_to = "juvenile_density")

belt_detailed_long <- belt_detailed_long |> 
  group_by(Year, Shelf, Reef, Site, genus) |> 
  summarise(
    mean_juv = mean(juvenile_density),
    sd_juv = sd(juvenile_density)
  )
```

```{r}
combined_juv_cover <- left_join(join_PIT_detailed, belt_detailed_long)
```

Despite not being interested in across the shelf differences, as shelf has only three levels there are two few levels for a stable random effect
```{r}
combined_juv_cover |> group_by(genus) |> 
  summarise(
    min_cover = min(mean_cover),
    max_cover = max(mean_cover),
    median_cover = median(mean_cover),
    min_juv = min(mean_juv),
    max_juv = max(mean_juv),
    median_juv = median(mean_juv)
  )
```


From this I would say we can only really do Acropora, Porites, and Pocillopora analysis because they have medians not equal to 0

## 6.1. Acropora

```{r}
Acro_j_bs <- combined_juv_cover |>  filter(genus == "Acropora")

gamma_avg_Acropora_fixed <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site) + (1 | Year), 
         family = Gamma(link = "log"), data = Acro_j_bs)
# sim <- simulateResiduals(gamma_avg_Acropora_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

ln_avg_Acropora_fixed <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site) + (1 | Year), 
        family = "lognormal", data = Acro_j_bs)
# sim <- simulateResiduals(ln_avg_Acropora_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim)

gau_avg_Acropora_fixed <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site) + (1 | Year), 
        family = "gaussian", data = Acro_j_bs)
# sim <- simulateResiduals(gau_avg_Acropora_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

#interaction with Shelf
gamma_avg_Acropora_int <- glmmTMB(mean_juv ~ mean_cover*Shelf +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Acro_j_bs)
# sim <- simulateResiduals(gamma_avg_Acropora_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

ln_avg_Acropora_int <- glmmTMB(mean_juv ~ mean_cover*Shelf +(1 | Reef/Site), 
        family = "lognormal", data = Acro_j_bs)
# sim <- simulateResiduals(ln_avg_Acropora_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

gaus_avg_Acropora_int <- glmmTMB(mean_juv ~ mean_cover*Shelf +(1 | Reef/Site), 
        family = "gaussian", data = Acro_j_bs)
# sim <- simulateResiduals(gaus_avg_Acropora_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

gamma_avg_Acropora_fixed2 <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Acro_j_bs)
# sim <- simulateResiduals(gamma_avg_Acropora_fixed2)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

ln_avg_Acropora_fixed2 <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site), 
        family = "lognormal", data = Acro_j_bs)
# sim <- simulateResiduals(ln_avg_Acropora_fixed2)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

gau_avg_Acropora_fixed2 <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site), 
        family = "gaussian", data = Acro_j_bs)
sim <- simulateResiduals(gau_avg_Acropora_fixed2)
testResiduals(sim) #passes
testZeroInflation(sim) #passes
```

```{r}
model.sel(gamma_avg_Acropora_fixed, gamma_avg_Acropora_fixed2, gamma_avg_Acropora_int,
          ln_avg_Acropora_fixed, ln_avg_Acropora_fixed2, ln_avg_Acropora_int,
          gau_avg_Acropora_fixed, gau_avg_Acropora_fixed2, gaus_avg_Acropora_int)
```

```{r}
summary(ln_avg_Acropora_fixed2)
```


## 6.2. Pocillopora

Issue is that we have zeros for the averages here in both cover and juveniles. cover has one site/year where there was no Pocillopora, whereas there are two site/year that have 0 pocillopora juveniles.
Will likely need to filter the 0 ones out to be able to run models but will just try
```{r}
Poci_j_bs <- combined_juv_cover |>  filter(genus == "Pocillopora")
Poci_j_bs <- combined_juv_cover |> filter(mean_juv > 0)

gamma_avg_Pocillopora_fixed <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site) + (1 | Year), 
         family = Gamma(link = "log"), data = Poci_j_bs)
# sim <- simulateResiduals(gamma_avg_Pocillopora_fixed)
# testResiduals(sim) #passes

ln_avg_Pocillopora_fixed <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site) + (1 | Year), 
        family = "lognormal", data = Poci_j_bs)
# sim <- simulateResiduals(ln_avg_Pocillopora_fixed)
# testResiduals(sim) #passes

gau_avg_Pocillopora_fixed <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site) + (1 | Year), 
        family = "gaussian", data = Poci_j_bs)
# sim <- simulateResiduals(gau_avg_Pocillopora_fixed)
# testResiduals(sim) #significant deviation
# testZeroInflation(sim) #signficant zero inflation

zi_gau_avg_Pocillopora_fixed <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site) + (1 | Year),
                                        ziformula = ~1,
        family = "gaussian", data = Poci_j_bs)
# sim <- simulateResiduals(zi_gau_avg_Pocillopora_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim) #still significant..

#interaction with Shelf
gamma_avg_Pocillopora_int <- glmmTMB(mean_juv ~ mean_cover*Shelf +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Poci_j_bs)
# sim <- simulateResiduals(gamma_avg_Pocillopora_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

ln_avg_Pocillopora_int <- glmmTMB(mean_juv ~ mean_cover*Shelf +(1 | Reef/Site), 
        family = "lognormal", data = Poci_j_bs)
# sim <- simulateResiduals(ln_avg_Pocillopora_int)
# testResiduals(sim) #passes

gaus_avg_Pocillopora_int <- glmmTMB(mean_juv ~ mean_cover*Shelf +(1 | Reef/Site), 
        family = "gaussian", data = Poci_j_bs)
# sim <- simulateResiduals(gaus_avg_Pocillopora_int)
# testResiduals(sim) #outlier significant, significant KS deviation


gamma_avg_Pocillopora_fixed2 <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Poci_j_bs)
# sim <- simulateResiduals(gamma_avg_Pocillopora_fixed2)
# testResiduals(sim) #passes

ln_avg_Pocillopora_fixed2 <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site), 
        family = "lognormal", data = Poci_j_bs)
# sim <- simulateResiduals(ln_avg_Pocillopora_fixed2)
# testResiduals(sim) #passes


gau_avg_Pocillopora_fixed2 <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site), 
        family = "gaussian", data = Poci_j_bs)
# sim <- simulateResiduals(gau_avg_Pocillopora_fixed2)
# testResiduals(sim) #KS and outlier significant

```

```{r}
model.sel(gamma_avg_Pocillopora_fixed, gamma_avg_Pocillopora_fixed2, gamma_avg_Pocillopora_int,
          ln_avg_Pocillopora_fixed, ln_avg_Pocillopora_fixed2, ln_avg_Pocillopora_int #,
          #gau_avg_Porites_fixed, gau_avg_Porites_fixed2, gaus_avg_Porites_int
          )
```

```{r}
summary(gamma_avg_Pocillopora_fixed2)
```

I don't love the amount of varaince in the conditional model


## 6.3. Porites
Lots of site/year without Porites cover, but only 2 without porites juveniles. One of the ones without coral cover also did not have juveniles (21/22 TS crest). The other did have some coral cover...
```{r}
Por_j_bs <- combined_juv_cover |>  filter(genus == "Porites")
Por_j_bs <- Por_j_bs |> filter(mean_juv > 0)
```

```{r}
gamma_avg_Porites_fixed <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site) + (1 | Year), 
         family = Gamma(link = "log"), data = Por_j_bs)
sim <- simulateResiduals(gamma_avg_Porites_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

ln_avg_Porites_fixed <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site) + (1 | Year), 
        family = "lognormal", data = Por_j_bs)
sim <- simulateResiduals(ln_avg_Porites_fixed)
testResiduals(sim) #passes
# testZeroInflation(sim)

gau_avg_Porites_fixed <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site) + (1 | Year), 
        family = "gaussian", data = Por_j_bs)
# sim <- simulateResiduals(gau_avg_Porites_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim) #significant zero inflation

#interaction with Shelf
gamma_avg_Porites_int <- glmmTMB(mean_juv ~ mean_cover*Shelf +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Por_j_bs)
# sim <- simulateResiduals(gamma_avg_Porites_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

ln_avg_Porites_int <- glmmTMB(mean_juv ~ mean_cover*Shelf +(1 | Reef/Site), 
        family = "lognormal", data = Por_j_bs)
# sim <- simulateResiduals(ln_avg_Porites_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

gaus_avg_Porites_int <- glmmTMB(mean_juv ~ mean_cover*Shelf +(1 | Reef/Site), 
        family = "gaussian", data = Por_j_bs)
# sim <- simulateResiduals(gaus_avg_Porites_int)
# testResiduals(sim) #significant outlier
# # testZeroInflation(sim) #passes

gamma_avg_Porites_fixed2 <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Por_j_bs)
# sim <- simulateResiduals(gamma_avg_Porites_fixed2)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

ln_avg_Porites_fixed2 <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site), 
        family = "lognormal", data = Por_j_bs)
# sim <- simulateResiduals(ln_avg_Porites_fixed2)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

gau_avg_Porites_fixed2 <- glmmTMB(mean_juv ~ mean_cover +(1 | Reef/Site), 
        family = "gaussian", data = Por_j_bs)
# sim <- simulateResiduals(gau_avg_Porites_fixed2)
# testResiduals(sim) #outliers
# testZeroInflation(sim) #passes
```

```{r}
model.sel(gamma_avg_Porites_fixed, gamma_avg_Porites_fixed2, gamma_avg_Porites_int,
          ln_avg_Porites_fixed, ln_avg_Porites_fixed2, ln_avg_Porites_int #,
          #gau_avg_Porites_fixed, gau_avg_Porites_fixed2, gaus_avg_Porites_int
          )
```

```{r}
summary(gamma_avg_Porites_fixed)
```
```{r}
summary(gamma_avg_Porites_int)
```

