---
title: "02_Model_Selection"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)


# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected),
         Shelf = factor(Shelf))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))


Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  ),
  Deployment = case_when(
      Year_Collected == "2022" ~ "21/22",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2026" ~ "25/26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")

Pocilloporidae <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Pocilloporidae, Top_Pocil, 
         Under_Pocil, Side_Pocil, total)

# Pocilloporidae_orientation <- Pocilloporidae %>% 
#   rename(Top = Top_Pocil,
#                    Bottom = Under_Pocil,
#                    Side = Side_Pocil,
#          Recruits_per_tile = total)

# Pocilloporidae_orientation <- Pocilloporidae_orientation %>% 
#   pivot_longer(cols = 7:9,
#                names_to = "Tile_Surface",
#                values_to = "Recruits")

Acroporidae <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Acroporidae, Top_Acr, 
         Under_Acr, Side_Acr, total)

# Acroporidae_orientation <- Acroporidae %>% 
#   rename(Top = Top_Acr,
#                    Bottom = Under_Acr,
#                    Side = Side_Acr,
#          Recruits_per_tile = total)
# 
# Acroporidae_orientation <- Acroporidae_orientation %>% 
#   pivot_longer(cols = 7:9,
#                names_to = "Tile_Surface",
#                values_to = "Recruits")

Poritid <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Poritid, Top_Por, Under_Por, 
         Side_Por, total)

# Poritidae_orientation <- Poritid %>% 
#   rename(Top = Top_Por,
#                    Bottom = Under_Por,
#                    Side = Side_Por,
#          Recruits_per_tile = total) 

# Poritidae_orientation <- Poritidae_orientation %>% 
#   pivot_longer(cols = 7:9,
#                names_to = "Tile_Surface",
#                values_to = "Recruits")

Other <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Other, Top_other, 
         Under_oth, Side_other, total)

# Mean or mean no. tiles per site / year
Join_Coral <- Coral_settlement_nGBR |> 
  group_by(Shelf, Year, Site, Reef) |> 
  summarise(
    st_mean_acro = mean(Acroporidae),
    st_sd_acro = sd(Acroporidae),
    st_median_acro = median(Acroporidae),
    st_mean_poci = mean(Pocilloporidae),
    st_sd_poci = sd(Pocilloporidae),
    st_median_poci = median(Pocilloporidae),
    st_mean_por = mean(Poritid),
    st_sd_por = sd(Poritid),
    st_median_por = median(Poritid),
    st_mean_total = mean(total),
    st_sd_total = sd(total),
    Tiles = n()
  )

# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:109, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

join_pit <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    PIT = n()
  )

Combined <- left_join(Join_Coral, join_pit)

# Juvenile data
Belt_10m <- read_csv("Belt_10m.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))
Belt_10m <- Belt_10m |>
  mutate(
    Other_HC = Favia + Favites + Goniastrea + Astrea +
                 Echinopora + Galaxea + Turbinaria +
                 Psammacora + Pavona
  )

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(Total_juveniles),
    sd_total = sd(Total_juveniles),
    Belt = n()
  )

## Total recuits
total_recruits <- Coral_settlement_nGBR %>%
  select(Year_Collected, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 5:8, names_to = "Family", values_to = "Total")

#Order so that Other is last in Family
total_recruits <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, Reef, Site, Obs, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 6:9, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))

total_recruits$Year_Collected <- as.factor(total_recruits$Year_Collected)

## Family recruitment data
family_total_recruits <- Coral_settlement_nGBR %>%
  select(2:13) %>% 
  pivot_longer(cols = 9:12, names_to = "Family", values_to = "Total") |> 
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))
```

# 1. Recruitment

## 1.1. Spatio-temporal varitaion

### 1.1.1. Total Recruitment

```{r warning=FALSE, message=FALSE}
pois_total_yr <- glmmTMB(total ~ Shelf + Year + (1 | Reef/Site), 
      family = "poisson", data = Coral_settlement_nGBR)

nbin_total_yr <- glmmTMB(total ~ Shelf +  Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

zbin_total_yr <- glmmTMB(total ~ Shelf + Year + (1 | Reef/Site),
                      ziformula = ~1,
        family = "nbinom2", data = Coral_settlement_nGBR)

pois_total_yr2 <- glmmTMB(total ~ Shelf * Year + (1 | Reef/Site), 
      family = "poisson", data = Coral_settlement_nGBR)

nbin_total_yr2 <- glmmTMB(total ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

zbin_total_yr2 <- glmmTMB(total ~ Shelf * Year + (1 | Reef/Site),
                      ziformula = ~1,
        family = "nbinom2", data = Coral_settlement_nGBR)
```

```{r eval=FALSE}
sim1.1 <- simulateResiduals(pois_total_yr)
testResiduals(sim1.1) #not good fit
testZeroInflation(sim1.1)

sim1.2 <- simulateResiduals(pois_total_yr2)
testResiduals(sim1.2) #not good fit
testZeroInflation(sim1.2)
```

```{r}
sim1.3 <- simulateResiduals(nbin_total_yr)
testResiduals(sim1.3) #good
testZeroInflation(sim1.3)
plot(sim1.3)

sim1.4 <- simulateResiduals(nbin_total_yr2)
testResiduals(sim1.4) #good
testZeroInflation(sim1.4)
plot(sim1.4)
```

model.sel(nbin_total_yr, zbin_total_yr, nbin_total_yr2, zbin_total_yr2)
```{r}
model.sel(nbin_total_yr, zbin_total_yr, nbin_total_yr2, zbin_total_yr2)
```

 Family Recruitment
```{r poisson of total Recruits, message=FALSE}
pois_Poritid_fixed <- glmmTMB(Poritid ~ Shelf + Year+(1 | Reef/Site) , 
      family = "poisson", data = Poritid) 

pois_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
      family = "poisson", data = Poritid)

pois_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + Year+ (1 | Reef/Site), 
      family = "poisson", data = Acroporidae)

pois_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
      family = "poisson", data = Acroporidae)

pois_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf + Year  + (1 | Reef/Site), 
      family = "poisson", data = Pocilloporidae)

pois_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf*Year + (1 | Reef/Site), 
      family = "poisson", data = Pocilloporidae)

pois_other_interaction <- glmmTMB(Other ~ Shelf*Year + (1 | Reef/Site), 
      family = "poisson", data = Other)
```

Negative Binomial of Total Recruits (for a family)

```{r}
## Fixed
nbin_Poritid_fixed <- glmmTMB(Poritid ~ Shelf + (1 | Reef/Site) + Year, 
        family = "nbinom2", data = Poritid)

nbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf + (1 | Reef/Site) + Year, 
        family = "nbinom2", data = Pocilloporidae)

nbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + (1 | Reef/Site) + Year, 
        family = "nbinom2", data = Acroporidae)

## Interaction
nbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Poritid)

nbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf * Year + (1 | Reef/Site), 
        family = "nbinom2", data = Pocilloporidae)

nbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Acroporidae)

nbin_other_interaction <- glmmTMB(Other ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Other)
```


Negative Binomial Zero Inflated of Total recruits (per family)

```{r}
#Fixed
# Fit model with zero-inflated negative binomial distribution 
zbin_Poritid_fixed <- glmmTMB(Poritid ~ Shelf+ Year + (1 | Reef/Site) ,
                        ziformula = ~1,
                        family = "nbinom2",
                        data = Poritid)
zbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf * Year + (1 | Reef/Site),
                        ziformula = ~1,
                        family = "nbinom2",
                        data = Poritid)

zbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf+ Year + (1 | Reef/Site),
                               ziformula = ~1,
                        family = "nbinom2",
                        data = Pocilloporidae)

zbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf*Year + (1 | Reef/Site),
                               ziformula = ~1,
                        family = "nbinom2",
                        data = Pocilloporidae)

zbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + Year+ (1 | Reef/Site),
                            ziformula = ~1,
                        family = "nbinom2",
                        data = Acroporidae)
zbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site),
                            ziformula = ~1,
                        family = "nbinom2",
                        data = Acroporidae)
```

### 1.1.2. Acroporidae

Simulate residuals for models

```{r}
sim2.1 <- simulateResiduals(nbin_Acroporidae_fixed) 
testResiduals(sim2.1) #fine
testZeroInflation(sim2.1) #P<0.05
plot(sim2.1) #bad

sim2.2 <- simulateResiduals(nbin_Acroporidae_interaction)
testResiduals(sim2.2) #fine
testZeroInflation(sim2.2) # P = 0.768
plot(sim2.2) #fine

# sim2.3 <- simulateResiduals(zbin_Acroporidae_fixed)
# testResiduals(sim2.3) #fine
# plot(sim2.3) #bad


# sim2.4 <- simulateResiduals(zbin_Acroporidae_interaction)
# testResiduals(sim2.4) #fine
# plot(sim2.4) #fine

sim2.5 <- simulateResiduals(pois_Acroporidae_fixed) #not a good fit
testResiduals(sim2.5) #bad
testZeroInflation(sim2.5) #p <0.05, therefore i should fit a zi pois
plot(sim2.5) #bad

sim2.6 <- simulateResiduals(pois_Acroporidae_interaction) #not a good fit
testResiduals(sim2.6) #significant deviation
testZeroInflation(sim2.6) # p = 0.64 which is pretty close
plot(sim2.6) #bad
```


```{r}
zi_pois_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + Year+ (1 | Reef/Site),
                                     ziformula = ~1,
      family = "poisson", data = Acroporidae)

sim2.7 <- simulateResiduals(zi_pois_Acroporidae_fixed)
testResiduals(sim2.7) #significant deviation
```

Compare AIC of those models

```{r}
AIC( nbin_Acroporidae_interaction, zbin_Acroporidae_interaction)

model.sel(nbin_Acroporidae_interaction, nbin_Acroporidae_fixed)
```

### 1.1.3 Pocilloporidae
```{r}
sim3.1 <- simulateResiduals(pois_Pocilloporidae_fixed)
testResiduals(sim3.1) #significant deviation
testZeroInflation(sim3.1) # p<0.05 so need to run a zi pois

sim3.2 <- simulateResiduals(pois_Pocilloporidae_interaction)
testResiduals(sim3.2) #significant deviation
testZeroInflation(sim3.2) #p<0.05

sim3.3 <- simulateResiduals(nbin_Pocilloporidae_interaction)
testResiduals(sim3.3) #passes
testZeroInflation(sim3.3) #passes

sim3.4 <- simulateResiduals(nbin_Pocilloporidae_fixed)
testResiduals(sim3.4) #passes
testZeroInflation(sim3.4) #p =0.72

# sim3.5 <- simulateResiduals(zbin_Pocilloporidae_fixed)
# testResiduals(sim3.5)
# 
# sim3.6 <- simulateResiduals(zbin_Pocilloporidae_interaction)
# testResiduals(sim3.6)

```

```{r}
zi_pois_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf + Year  + (1 | Reef/Site), 
                                        ziformula = ~1,
      family = "poisson", data = Pocilloporidae)
sim3.7 <- simulateResiduals(zi_pois_Pocilloporidae_fixed)
testResiduals(sim3.7) #significant deviation

zi_pois_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf * Year + (1 | Reef/Site), 
                                              ziformula = ~1,
        family = "poisson", data = Pocilloporidae)

sim3.8 <- simulateResiduals(zi_pois_Pocilloporidae_interaction)
testResiduals(sim3.8)  #fails
```


```{r}
model.sel(nbin_Pocilloporidae_fixed, nbin_Pocilloporidae_interaction)
```

### 1.1.4. Poritidae
```{r}
sim4.1 <- simulateResiduals(pois_Poritid_fixed)
testResiduals(sim4.1) # bad fit
testZeroInflation(sim4.1) # P<0.05, therefore need to run zi

sim4.2 <- simulateResiduals(pois_Poritid_interaction)
testResiduals(sim4.2) # bad fit
testZeroInflation(sim4.2) # p<0.05

sim4.3 <- simulateResiduals(nbin_Poritid_interaction)
testResiduals(sim4.3) #passed
testZeroInflation(sim4.3) # p = 0.848

sim4.4 <- simulateResiduals(nbin_Poritid_fixed)
testResiduals(sim4.4) #passed
testZeroInflation(sim4.4) #p=0.88

# sim4.5 <- simulateResiduals(zbin_Poritid_fixed)
# testResiduals(sim4.5)
# 
# sim4.6 <- simulateResiduals(zbin_Poritid_interaction)
# testResiduals(sim4.6)
```

Zero inflated models where required
```{r}
zi_pois_Poritid_fixed <- glmmTMB(Poritid ~ Shelf + Year+(1 | Reef/Site) ,
                                 ziformula = ~1,
      family = "poisson", data = Poritid) 

zi_pois_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
                                       ziformula = ~1,
      family = "poisson", data = Poritid)

sim4.7 <- simulateResiduals(zi_pois_Poritid_fixed)
testResiduals(sim4.7) #outliers detected

sim4.8 <- simulateResiduals(zi_pois_Poritid_interaction)
testResiduals(sim4.8) #outlier significant

```


Both nbin_interaction, and zbin_interaction for Portidae have significant dispersion.
Don't see much of a difference in Poritid across shelf position. Although slightly less in mid shore reefs (p<0.1)

```{r echo=FALSE}
AIC(nbin_Poritid_fixed, zbin_Poritid_fixed, nbin_Poritid_interaction, zbin_Poritid_interaction)

model.sel(nbin_Poritid_fixed, nbin_Poritid_interaction)
```


All family option
Models of family total recruits

```{r}
pois_total_family_fixed <- glmmTMB(Total ~ Family +Shelf + Year_Collected+ (1 | Reef/Site), 
      family = "poisson",
      data = family_total_recruits)

pois_total_family_interaction <-  glmmTMB(Total ~ Family * Shelf * Year_Collected + (1 | Reef/Site), 
      family = "poisson",
      data = family_total_recruits)

nb_total_family_fixed <- glmmTMB(Total ~ Family + Shelf + Year_Collected + (1 | Reef/Site) ,
                        family = "nbinom2",
                        data = family_total_recruits)

nb_total_family_interaction <- glmmTMB(Total ~ Family * Shelf * Year_Collected + (1 | Reef/Site) ,
                        family = "nbinom2",
                        data = family_total_recruits)

zi_total_family_fixed <- glmmTMB(Total ~ Family + Shelf + Year_Collected + (1 | Reef/Site) ,
                        ziformula = ~1,
                        family = "nbinom2",
                        data = family_total_recruits)

zi_total_family_interaction <-  glmmTMB(Total ~ Family * Shelf * Year_Collected + (1 | Reef/Site) ,
                        ziformula = ~1,
                        family = "nbinom2",
                        data = family_total_recruits)
```

simulate and test residuals

```{r}
sim4.1 <- simulateResiduals(pois_total_family_fixed)
testResiduals(sim4.1) 

sim4.2 <- simulateResiduals(pois_total_family_interaction) 
testResiduals(sim4.2)

sim4.3 <- simulateResiduals(nb_total_family_fixed) #middle plot doesn't look great, but okay
testResiduals(sim4.3)

sim4.4 <- simulateResiduals(nb_total_family_interaction) #good
testResiduals(sim4.4)

sim4.5 <- simulateResiduals(zi_total_family_fixed) # middle plot doesn't look great but okay
testResiduals(sim4.5)

sim4.6 <- simulateResiduals(zi_total_family_interaction)  #good
testResiduals(sim4.6)
```

```{r}
AIC(nb_total_family_fixed, nb_total_family_interaction, zi_total_family_fixed, zi_total_family_interaction) #nb family interaction is best

model.sel(nb_total_family_fixed, nb_total_family_interaction, zi_total_family_fixed, zi_total_family_interaction)
```

##1.2. Composition

```{r}
Recruits <- Coral_settlement_nGBR %>%
  filter(total > 0) %>%                                                          # ensure only vegetated tiles included 
  group_by(Reef, Year_Collected, Shelf) %>%               # this was missing in your code
  summarise(across(Acroporidae:Other, mean), .groups = "drop")

recruit_hellinger <- decostand(Recruits %>% select(Acroporidae, Pocilloporidae, Poritid, Other),
                                                         method = "hellinger")

# reef year?
reef_year <- Recruits |> 
  select(Reef, Year_Collected, Shelf)

permanova_result <- adonis2(recruit_hellinger ~ Shelf,
  data = Recruits,
  method = "euclidean",    # the Hellinger transformation is designed for Euclidean distnaces. Bray-Curtis is not “wrong” but it’s redundant if you’ve already done a Hellinger transformation.
  by = "margin",
  permutations = 999)

print(permanova_result)

```

```{r}
Recruits <- Coral_settlement_nGBR %>%
  filter(total > 0) %>%                                                          # ensure only vegetated tiles included 
  group_by(Reef, Year_Collected, Shelf) %>%               # this was missing in your code
  summarise(across(Acroporidae:Other, mean), .groups = "drop")

recruit_hellinger <- decostand(Recruits %>% select(Acroporidae, Pocilloporidae, Poritid, Other),
                                                         method = "hellinger")

# reef year?
reef_year <- Recruits |> 
  select(Reef, Year_Collected, Shelf)

permanova_result <- adonis2(recruit_hellinger ~ Year_Collected * Shelf,
  data = Recruits,
  method = "euclidean",    # the Hellinger transformation is designed for Euclidean distnaces. Bray-Curtis is not “wrong” but it’s redundant if you’ve already done a Hellinger transformation.
  by = "margin",
  permutations = 999)

print(permanova_result)
```


# 2. Juveniles
## 2.1. Spatio-temporal variation

### 2.1.1. Total

```{r}
pois <- glmmTMB(Total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Belt_10m)

nbin <- glmmTMB(Total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_10m)
```

```{r}
sim14.1 <- simulateResiduals(pois)
testResiduals(sim14.1)
testZeroInflation(sim14.1) # Is the number of observed zeros significantly different from what the model predicts? no

sim14.2 <- simulateResiduals(nbin)
testResiduals(sim14.2) #passes
testZeroInflation(sim14.2) #nope
```

```{r}
model.sel(pois, nbin)
summary(nbin)
```

### 2.1.2. Acroporidae

```{r}
Acro <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Acroporidae)

pois <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Acro)

nbin <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Acro)
```

```{r}
sim15.1 <- simulateResiduals(pois)
testResiduals(sim15.1) # passes tests
testZeroInflation(sim15.1) # p= 0.12

sim15.2 <- simulateResiduals(nbin)
testResiduals(sim15.2) #passes
testZeroInflation(sim15.2) #p=0.416
```

Both meet assumptions
```{r}
model.sel(pois, nbin) #nbin behaves better
```

### 2.1.3. Pocilloporidae

```{r}
Poc <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Pocilloporidae)

pois <- glmmTMB(Pocilloporidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Poc)

nbin <- glmmTMB(Pocilloporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Poc)
```

```{r}
sim16.1 <- simulateResiduals(pois)
testResiduals(sim16.1) #passed
testZeroInflation(sim16.1) #p=0.968

sim16.2 <- simulateResiduals(nbin)
testResiduals(sim16.2) #passed
testZeroInflation(sim16.2) #p =1
```
```{r}
model.sel(pois, nbin) #nbin behaves better
```

### 2.1.4. Poritidae

```{r}
Por <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Poritidae)

pois <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Por)

nbin <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Por)
```

```{r}
sim17.1 <- simulateResiduals(pois)
testResiduals(sim17.1) # passes
testZeroInflation(sim17.1) #p=0.248

sim17.2 <- simulateResiduals(nbin)
testResiduals(sim17.2) #passes
testZeroInflation(sim17.2) # p=0.784
```

Both models have outliers as significant.
```{r}
model.sel(pois, nbin) #nbin behaves better
```

### 2.1.5. Other

```{r}
Other_juv <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Other_HC)

pois <- glmmTMB(Other_HC ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Other_juv)
z_pois <- glmmTMB(Other_HC ~ Year*Shelf+(1 | Reef/Site),ziformula = ~1,
      family = "poisson", data = Other_juv)

nbin <- glmmTMB(Other_HC ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Other_juv)
```

```{r}
sim <- simulateResiduals(pois)
testResiduals(sim) # KS deviation significant
testZeroInflation(sim)

sim <- simulateResiduals(z_pois)
testResiduals(sim) # passes

sim <- simulateResiduals(nbin)
testResiduals(sim) #passes
testZeroInflation(sim)
```

```{r}
model.sel(z_pois, nbin)
```



## 2.2. Composition
```{r}
Belt_10m <- Belt_10m |> 
  rename("Other Hard Coral" = Other_HC
  )
# Step 1: Create site-by-species matrix with Site_ID
site_by_belt <- Belt_10m %>%
  select(
    Year, Reef, Shelf, Site, Transect, Acroporidae, Pocilloporidae, Poritidae, `Other Hard Coral`
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

site_by_belt <- site_by_belt |>                   
  group_by(Reef, Year, Shelf) %>%               # this was missing in your code
  summarise(across(Acroporidae:`Other Hard Coral`, mean), .groups = "drop")



# Step 2: Metadata (KEEP Site_ID)
metadata_belt <- site_by_belt %>%
  select(Shelf, Year, Reef)

# Step 3: Species matrix
belt_matrix_filtered <- site_by_belt %>%
  select(
    4:7
  )
```

```{r}
# create a distance matrix
belt_dist <- vegdist(belt_matrix_filtered, method = "bray")

# With respect to Shelf
betadisper(belt_dist, metadata_belt$Shelf)
anova(betadisper(belt_dist, metadata_belt$Shelf)) # Significant was 0.03139 *
```

```{r}
permanova_result <- adonis2(belt_matrix_filtered ~ Shelf,
  data = site_by_belt,
  method = "bray",
  by = "margin",
  permutations = 999)

print(permanova_result)
```

### 2.1.2. NMDS

```{r}
nmds <- metaMDS(belt_matrix_filtered, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress
```

```{r}
# Creating Vectors for driving the patterns
belt_envfit <- envfit(
  nmds,
  belt_matrix_filtered,
  permutations = 999
) 

# Extract p-values
pvals <- belt_envfit$vectors$pvals

# Extract r-squared values
r2 <- belt_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(belt_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- belt_envfit$vectors$r
vec_scores$pval <- belt_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

belt_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")
```

# 3. Benthic Cover

## 3.1. Composition

### 3.1.1. PERMANOVA
```{r}
# Step 1: Create site-by-species matrix with Site_ID
site_by_PIT <- PIT %>%
  select(
    Year, Reef, Shelf, Site, Transect, 95:108
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

site_by_PIT <- site_by_PIT |>                                                         # ensure only vegetated tiles included 
  group_by(Reef, Year, Shelf) %>%               # this was missing in your code
  summarise(across(`Soft Coral`:CCA, mean), .groups = "drop")

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  select(Shelf, Year, Reef)

# Step 3: Species matrix
PIT_matrix_filtered <- site_by_PIT %>%
  select(
    4:17
  )
```

```{r}
# create a distance matrix
PIT_dist <- vegdist(PIT_matrix_filtered, method = "bray")

# With respect to Shelf
betadisper(PIT_dist, metadata_PIT$Shelf)
anova(betadisper(PIT_dist, metadata_PIT$Shelf)) # Significant was 0.01143 * ***
```

```{r}
permanova_result <- adonis2(PIT_matrix_filtered ~ Shelf,
  data = site_by_PIT,
  method = "bray",
  by = "margin",
  permutations = 999)

print(permanova_result)
```

### 3.1.2. NMDS

```{r}
nmds <- metaMDS(PIT_matrix_filtered, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress
```

```{r}
# Creating Vectors for driving the patterns
Benthic_envfit <- envfit(
  nmds,
  PIT_matrix_filtered,
  permutations = 999
) 

# Extract p-values
pvals <- Benthic_envfit$vectors$pvals

# Extract r-squared values
r2 <- Benthic_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(Benthic_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- Benthic_envfit$vectors$r
vec_scores$pval <- Benthic_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

Benthic_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")
```


# 4. Brood stock (PIT) and Recruitment (Tiles)

```{r}
st_join <- Coral_settlement_nGBR |> 
  group_by(Site, Year) |> 
  summarise(
    acro_mean_st = mean(Acroporidae),
    poci_mean_st = mean(Pocilloporidae),
    por_mean_st = mean(Poritid),
    acro_sd_st = sd(Acroporidae),
    poci_sd_st = sd(Pocilloporidae),
    por_sd_st = sd(Poritid),
    acro_median_st = median(Acroporidae),
    poci_median_st = median(Pocilloporidae),
    por_median_st = median(Poritid),
  )

combined_averages <- left_join(join_pit, st_join)
```

Now these are both averages, it can't follow an integar and would need to follow a different family of error distribution.

## 4.1. Acroporidae

Despite not being insteresting in across the shelf differences, as shelf has only three levels there are two few levels for a stable randm effect
Despite not being interested in across the shelf differences, as shelf has only three levels there are two few levels for a stable randm effect
```{r}
Acro <- combined_averages |> 
  filter(acro_mean_st > 0)
```

```{r}
gamma_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Reef/Site) + (1 | Year), 
         family = Gamma(link = "log"), data = Acro)
sim <- simulateResiduals(gamma_avg_Acroporidae_fixed)
testResiduals(sim) #passes
testZeroInflation(sim) #passes

ln_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Reef/Site) + (1 | Year), 
        family = "lognormal", data = Acro)
sim <- simulateResiduals(ln_avg_Acroporidae_fixed)
testResiduals(sim) #passes
testZeroInflation(sim)

gau_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Reef/Site) + (1 | Year), 
        family = "gaussian", data = Acro)
sim <- simulateResiduals(gau_avg_Acroporidae_fixed)
testResiduals(sim) #fails
testZeroInflation(sim) #passes

gamma_avg_Acroporidae_nested <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Shelf/Reef/Site), 
         family = Gamma(link = "log"), data = Acro)
sim <- simulateResiduals(gamma_avg_Acroporidae_nested)
testResiduals(sim) #passes
testZeroInflation(sim) #passes

ln_avg_Acroporidae_nested <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Shelf/Reef/Site), 
        family = "lognormal", data = Acro)
sim <- simulateResiduals(ln_avg_Acroporidae_nested)
testResiduals(sim) #passes
testZeroInflation(sim) #passes

gaus_avg_Acroporidae_nested <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Shelf/Reef/Site), 
        family = "gaussian", data = Acro)
sim <- simulateResiduals(gaus_avg_Acroporidae_nested)
testResiduals(sim) #fails miserably
testZeroInflation(sim)

#interaction with Shelf
gamma_avg_Acroporidae_int <- glmmTMB(acro_mean_st ~ mean_aco*Shelf +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Acro)
sim <- simulateResiduals(gamma_avg_Acroporidae_int)
testResiduals(sim) #passes
testZeroInflation(sim) #passes

ln_avg_Acroporidae_int <- glmmTMB(acro_mean_st ~ mean_aco*Shelf +(1 | Reef/Site), 
        family = "lognormal", data = Acro)
sim <- simulateResiduals(ln_avg_Acroporidae_int)
testResiduals(sim) #passes
testZeroInflation(sim) #passes

gaus_avg_Acroporidae_int <- glmmTMB(acro_mean_st ~ mean_aco*Shelf +(1 | Reef/Site), 
        family = "gaussian", data = Acro)
sim <- simulateResiduals(gaus_avg_Acroporidae_int)
testResiduals(sim) #terrible, truly.
testZeroInflation(sim)
```

Log normal models worked best for all of these
```{r}
model.sel(gamma_avg_Acroporidae_fixed, ln_avg_Acroporidae_fixed, gamma_avg_Acroporidae_nested, ln_avg_Acroporidae_nested, gamma_avg_Acroporidae_int, ln_avg_Acroporidae_int)

summary(gamma_avg_Acroporidae_fixed)
```


## 4.2. Pocilloporidae

```{r}
## Pocilloporidae
Poci <- combined_averages |> 
  filter(poci_mean_st > 0)

ln_avg_poci_fixed <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Reef/Site) + (1 | Year), 
        family = "lognormal", data = Poci)
ln_avg_poci_fixed2 <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Reef/Site), 
        family = "lognormal", data = Poci)

gaus_avg_poci_fixed <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Reef/Site) + (1 | Year), 
        family = "gaussian", data = Poci)

gamma_avg_Pocilloporidae_fixed <- glmmTMB(poci_mean_st ~ mean_poci + (1 | Reef/Site) + (1 | Year),
        family = Gamma(link = "log"), data = Poci)
gamma_avg_Pocilloporidae_fixed2 <- glmmTMB(poci_mean_st ~ mean_poci + (1 | Reef/Site),
        family = Gamma(link = "log"), data = Poci)

ln_avg_poci_nested <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Shelf/Reef/Site), 
        family = "lognormal", data = Poci)

gaus_avg_poci_nested <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Shelf/Reef/Site), 
        family = "gaussian", data = Poci)

gamma_avg_Pocilloporidae_nested <- glmmTMB(poci_mean_st ~ mean_poci + (1 | Shelf/Reef/Site),
        family = Gamma(link = "log"), data = Poci)

ln_avg_Pocilloporidae_int <- glmmTMB(poci_mean_st ~ mean_poci * Shelf +(1 | Reef/Site), 
         family = "lognormal", data = Poci)

gau_avg_Pocilloporidae_int <- glmmTMB(poci_mean_st ~ mean_poci * Shelf +(1 | Reef/Site), 
         family = "gaussian", data = Poci)

gamma_avg_Pocilloporidae_int <- glmmTMB(poci_mean_st ~ mean_poci * Shelf +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Poci)
```

```{r}
sim14.1 <- simulateResiduals(gamma_avg_Pocilloporidae_fixed) #passed
testResiduals(sim14.1)
testZeroInflation(sim14.1)
sim14.1 <- simulateResiduals(gamma_avg_Pocilloporidae_fixed2) #passed
testResiduals(sim14.1)
testZeroInflation(sim14.1)
sim14.2 <- simulateResiduals(ln_avg_poci_fixed) # very good
testResiduals(sim14.2)
testZeroInflation(sim14.2)

sim14.3 <- simulateResiduals(gaus_avg_poci_fixed)
testResiduals(sim14.3) # definitelty not this one
testZeroInflation(sim14.3)

sim14.4 <- simulateResiduals(gamma_avg_Pocilloporidae_nested) #passed
testResiduals(sim14.4)
testZeroInflation(sim14.4)

sim14.5 <- simulateResiduals(ln_avg_poci_nested) # very good
testResiduals(sim14.5)
testZeroInflation(sim14.5)

sim14.6 <- simulateResiduals(gaus_avg_poci_fixed)
testResiduals(sim14.6) # Outliers
testZeroInflation(sim14.6)

sim14.7 <- simulateResiduals(ln_avg_Pocilloporidae_int)
testResiduals(sim14.7) # passes
testZeroInflation(sim14.7) #passes

sim14.8 <- simulateResiduals(gamma_avg_Pocilloporidae_int)
testResiduals(sim14.8) # passes
testZeroInflation(sim14.8) #passes

sim14.9 <- simulateResiduals(gau_avg_Pocilloporidae_int)
testResiduals(sim14.9) # passes
testZeroInflation(sim14.9) #passes
```

```{r}
model.sel(gamma_avg_Pocilloporidae_fixed, gamma_avg_Pocilloporidae_nested, ln_avg_poci_fixed, ln_avg_poci_nested, gamma_avg_Pocilloporidae_int, gau_avg_Pocilloporidae_int, ln_avg_Pocilloporidae_int, gamma_avg_Pocilloporidae_fixed2, ln_avg_poci_fixed2)
```
Without year performs slightly better
```{r}
summary(gamma_avg_Pocilloporidae_fixed)
```

## 4.3. Poritidae

```{r}
## Poritidae
Por <- combined_averages |> 
  filter(por_mean_st > 0)

gamma_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Reef/Site) + (1 | Year),
        family = Gamma(link = "log"), data = Por)
ln_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Reef/Site) + (1 | Year), 
        family = "lognormal", data = Por)
gaus_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Reef/Site) + (1 | Year), 
        family = "gaussian", data = Por)

```

```{r}
sim15.1 <- simulateResiduals(gamma_avg_Poritid_fixed) #passed
testResiduals(sim15.1)
testZeroInflation(sim15.1)
sim15.2 <- simulateResiduals(ln_avg_Poritid_fixed) # very good
testResiduals(sim15.2)
testZeroInflation(sim15.2)

sim15.3 <- simulateResiduals(gaus_avg_Poritid_fixed)
testResiduals(sim15.3) # definitely not this one
testZeroInflation(sim15.3)
```
```{r}
model.sel(gamma_avg_Poritid_fixed, ln_avg_Poritid_fixed)
```


```{r}
summary(ln_avg_Poritid_fixed)
```

All of the models had significant intercepts, but the estimate for mean_por was not significant in any of the models.


# 5. Recruitment and Juveniles
These models will need to be lagged. 

```{r}
combine_belt_st <- left_join(join_belt, Join_Coral)

# Only including years in which both data is included (once lagging has happened)
combine_belt_st <- combine_belt_st |> 
  mutate(Year_start = as.numeric(substr(Year, 1, 4)))

gropd_df2 <- combine_belt_st %>%
  arrange(Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_mean_acro_lag1 = lag(st_mean_acro, 1),
    st_mean_poci_lag1 = lag(st_mean_poci, 1),
    st_mean_por_lag1 = lag(st_mean_por, 1),
  ) %>%
  ungroup()
```


## 5.1. Acroporidae

Remove first year row
```{r}
model_df_acro <- gropd_df2 %>%
  filter(!is.na(mean_acro))
```

Fit models
```{r}
ln_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)

sim <- simulateResiduals(ln_acro_juv_st_fixed)
testResiduals(sim) #passes
testZeroInflation(sim)

gau_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 +Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)

sim <- simulateResiduals(gau_acro_juv_st_fixed)
testResiduals(sim) #passes
testZeroInflation(sim)

gamma_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 +Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_acro)

sim <- simulateResiduals(gau_acro_juv_st_fixed)
testResiduals(sim) #passes
testZeroInflation(sim) #passed

#Nested
ln_acro_juv_st_nested <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + (1|Shelf/Reef/Site),
              family = "lognormal",
           data = model_df_acro)
sim <- simulateResiduals(ln_acro_juv_st_nested)
testResiduals(sim) # passed but not amazing
testZeroInflation(sim) #passed

gau_acro_juv_st_nested<- glmmTMB(mean_acro ~ st_mean_acro_lag1 + (1| Shelf/Reef/Site),
              family = "gaussian",
           data = model_df_acro)

sim <- simulateResiduals(gau_acro_juv_st_nested)
testResiduals(sim) # passed but doesn't look great
testZeroInflation(sim) #passed

gamma_acro_juv_st_nested <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + (1|Shelf/Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_acro)

sim <- simulateResiduals(gau_acro_juv_st_nested)
testResiduals(sim) #passes but not amazing
testZeroInflation(sim) #passes

# interaction between shelf and st

ln_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)

sim <- simulateResiduals(ln_acro_juv_st_int)
testResiduals(sim) #passes
testZeroInflation(sim) #passes

gau_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)

sim <- simulateResiduals(gau_acro_juv_st_int)
testResiduals(sim) #passes
testZeroInflation(sim) #passes

gamma_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_acro)

sim <- simulateResiduals(gamma_acro_juv_st_int)
testResiduals(sim) #passes
testZeroInflation(sim) #passes
```

```{r}
model.sel(gau_acro_juv_st_int, gau_acro_juv_st_nested, gau_acro_juv_st_fixed, gamma_acro_juv_st_fixed, gamma_acro_juv_st_int, gamma_acro_juv_st_nested, ln_acro_juv_st_int, ln_acro_juv_st_nested, ln_acro_juv_st_fixed)
```

```{r}
ln_acro_juv_st_int

summary(ln_acro_juv_st_int)
```

## 5.2. Pocilloporidae

Remove first years rows
```{r}
model_df_poci <- gropd_df2 %>%
  filter(!is.na(mean_poci))
```


```{r}
ln_poci_juv_st_fixed <- glmmTMB(mean_poci ~ st_mean_poci_lag1 + Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)

sim <- simulateResiduals(ln_poci_juv_st_fixed)
testResiduals(sim) #passes
testZeroInflation(sim)


gau_poci_juv_st_fixed <- glmmTMB(mean_poci ~ st_mean_poci_lag1+Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)

sim <- simulateResiduals(gau_poci_juv_st_fixed)
testResiduals(sim) #passes but doesn't look great
testZeroInflation(sim)

gamma_poci_juv_st_fixed <- glmmTMB(mean_poci ~ st_mean_poci_lag1+ Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_poci)

sim <- simulateResiduals(gau_poci_juv_st_fixed)
testResiduals(sim) #passes
testZeroInflation(sim)

#Interaction with Shelf
ln_poci_juv_st_int <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)
sim <- simulateResiduals(ln_poci_juv_st_int)
testResiduals(sim) # passed
testZeroInflation(sim)

gau_poci_juv_st_int<- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)

sim <- simulateResiduals(gau_poci_juv_st_int)
testResiduals(sim) # passed but doesn't look great
testZeroInflation(sim)

gamma_poci_juv_st_int<- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_poci)

sim <- simulateResiduals(gamma_poci_juv_st_int)
testResiduals(sim) # passed
testZeroInflation(sim)

#Nested
ln_poci_juv_st_nested <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Shelf/Reef/Site),
              family = "lognormal",
           data = model_df_poci)
sim <- simulateResiduals(ln_poci_juv_st_nested)
testResiduals(sim) # passed
testZeroInflation(sim)

gau_poci_juv_st_nested <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Shelf/Reef/Site),
              family = "gaussian",
           data = model_df_poci)
sim <- simulateResiduals(ln_poci_juv_st_nested)
testResiduals(sim) # passed
testZeroInflation(sim)

gamma_poci_juv_st_nested <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Shelf/Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_poci)
sim <- simulateResiduals(gamma_poci_juv_st_nested)
testResiduals(sim) # passed
testZeroInflation(sim)
```


```{r}
model.sel(gau_poci_juv_st_fixed, gau_poci_juv_st_int, gau_poci_juv_st_nested, gamma_poci_juv_st_fixed, gamma_poci_juv_st_int, gamma_poci_juv_st_nested, ln_poci_juv_st_fixed, ln_poci_juv_st_int, ln_poci_juv_st_nested)
```


```{r}
ln_poci_juv_st_fixed

summary(ln_poci_juv_st_fixed)
```

## 5.3. Poritidae

Remove first year rows
```{r}
model_df_por <- gropd_df2 %>%
  filter(!is.na(mean_por))
```

```{r}
model_df_por <- model_df_por |> 
  filter(mean_por > 0)

ln_por_juv_st_fixed <- glmmTMB(mean_por ~ st_mean_por_lag1 + Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_por)

sim <- simulateResiduals(ln_por_juv_st_fixed)
testResiduals(sim) # passsed
testZeroInflation(sim)

gau_por_juv_st_fixed <- glmmTMB(mean_por ~ st_mean_por_lag1 + Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_por)

sim <- simulateResiduals(gau_por_juv_st_fixed)
testResiduals(sim) # passed but not great
testZeroInflation(sim)

gamma_por_juv_st_fixed <- glmmTMB(mean_por ~ st_mean_por_lag1 + Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_por)

sim <- simulateResiduals(gamma_por_juv_st_fixed)
testResiduals(sim) # passsed
testZeroInflation(sim)

#Interaction
ln_por_juv_st_int <- glmmTMB(mean_por ~ st_mean_por_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_por)
sim <- simulateResiduals(ln_por_juv_st_int)
testResiduals(sim) # passed
testZeroInflation(sim)

gau_por_juv_st_int<- glmmTMB(mean_por ~ st_mean_por_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_por)

sim <- simulateResiduals(gau_por_juv_st_int)
testResiduals(sim) # passed but outlier test very close to being significant
testZeroInflation(sim)

gamma_por_juv_st_int <- glmmTMB(mean_por ~ st_mean_por_lag1 * Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_por)
sim <- simulateResiduals(gamma_por_juv_st_int)
testResiduals(sim) # passed
testZeroInflation(sim)

# Shelf/Reef/Site
ln_por_juv_st_nested <- glmmTMB(mean_por ~ st_mean_por_lag1+ (1|Shelf/Reef/Site),
              family = "lognormal",
           data = model_df_por)

sim <- simulateResiduals(ln_por_juv_st_nested)
testResiduals(sim) # passsed
testZeroInflation(sim)

gau_por_juv_st_nested <- glmmTMB(mean_por ~ st_mean_por_lag1 + (1|Shelf/Reef/Site),
              family = "gaussian",
           data = model_df_por)

sim <- simulateResiduals(gau_por_juv_st_nested)
testResiduals(sim) # passed but not great
testZeroInflation(sim)

gamma_por_juv_st_nested <- glmmTMB(mean_por ~ st_mean_por_lag1+ (1|Shelf/Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_por)

sim <- simulateResiduals(gamma_por_juv_st_nested)
testResiduals(sim) # passsed
testZeroInflation(sim)

```

```{r}
model.sel(gau_por_juv_st_fixed, gau_por_juv_st_int, gau_por_juv_st_nested, ln_por_juv_st_fixed, ln_por_juv_st_int, ln_por_juv_st_nested, gamma_por_juv_st_fixed, gamma_por_juv_st_int, gamma_por_juv_st_nested)

```

```{r}
gamma_por_juv_st_fixed

summary(gamma_por_juv_st_fixed)
```

```{r eval=FALSE, echo=FALSE}
write.csv(Poci_juv_st_df, 
          file = "Poci_juv_st_model_sel.csv")

write.csv(Acro_juv_st_df, 
          file = "Acro_juv_st_model_sel.csv")

write.csv(Poritidae_juv_st_df, 
          file = "Por_juv_st_model_sel.csv")
```

