---
title: "Figures"
author: "Monique White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)

# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  ),
  Deployment = case_when(
      Year_Collected == "2022" ~ "21/22",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2026" ~ "25/26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")

Pocilloporidae <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Pocilloporidae, Top_Pocil, 
         Under_Pocil, Side_Pocil, total)

Pocilloporidae_orientation <- Pocilloporidae %>% 
  rename(Top = Top_Pocil,
                   Bottom = Under_Pocil,
                   Side = Side_Pocil,
         Recruits_per_tile = total)

Pocilloporidae_orientation <- Pocilloporidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Acroporidae <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Acroporidae, Top_Acr, 
         Under_Acr, Side_Acr, total)

Acroporidae_orientation <- Acroporidae %>% 
  rename(Top = Top_Acr,
                   Bottom = Under_Acr,
                   Side = Side_Acr,
         Recruits_per_tile = total)

Acroporidae_orientation <- Acroporidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Poritid <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Poritid, Top_Por, Under_Por, 
         Side_Por, total)

Poritidae_orientation <- Poritid %>% 
  rename(Top = Top_Por,
                   Bottom = Under_Por,
                   Side = Side_Por,
         Recruits_per_tile = total) 

Poritidae_orientation <- Poritidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Other <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Other, Top_other, 
         Under_oth, Side_other, total)

# Mean or mean no. tiles per site / year
Join_Coral <- Coral_settlement_nGBR |> 
  group_by(Shelf, Year, Site, Reef) |> 
  summarise(
    st_mean_acro = mean(Acroporidae),
    st_sd_acro = sd(Acroporidae),
    st_median_acro = median(Acroporidae),
    st_mean_poci = mean(Pocilloporidae),
    st_sd_poci = sd(Pocilloporidae),
    st_median_poci = median(Pocilloporidae),
    st_mean_por = mean(Poritid),
    st_sd_por = sd(Poritid),
    st_median_por = median(Poritid),
    st_mean_total = mean(total),
    st_sd_total = sd(total),
    Tiles = n()
  )

# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae)
  )

Combined <- left_join(Coral_settlement_nGBR, join)
# Juvenile data
Belt_10m <- read_csv("Belt_10m.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(Total_juveniles),
    sd_total = sd(Total_juveniles),
    Belt = n()
  )

Combined1 <- left_join(Coral_settlement_nGBR, join_belt)

## Total recuits
total_recruits <- Coral_settlement_nGBR %>%
  select(Year_Collected, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 5:8, names_to = "Family", values_to = "Total")

#Order so that Other is last in Family
total_recruits <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, Reef, Site, Obs, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 6:9, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))

total_recruits$Year_Collected <- as.factor(total_recruits$Year_Collected)

## Position on tile

position_total_recruits <- Coral_settlement_nGBR %>%
  select(2:9, 15:29, 32) %>% 
  pivot_longer(cols = 21:23, names_to = "Pos", values_to = "Total")

position_total_recruits <- position_total_recruits |> 
  mutate(Pos = case_when(
      Pos == "Top.Total" ~ "Top",
      Pos == "Side.Total" ~ "Side",
      Pos == "Under.Total" ~ "Bottom"
  ))
 
position_total_recruits <- position_total_recruits |>                           
mutate(Pos = factor(Pos, levels = c("Bottom", "Side", "Top")))

position_total_recruits <- left_join(position_total_recruits, join)

## Family recruitment data
family_total_recruits <- Coral_settlement_nGBR %>%
  select(2:13) %>% 
  pivot_longer(cols = 9:12, names_to = "Family", values_to = "Total") |> 
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))
```

# 1. Recruitment

## 1.1 Spatio-temporal
Figure 1

```{r}
nbin_total_yr2 <- glmmTMB(total ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

total_model_emm <- nbin_total_yr2|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) 

total_mean_site <- Coral_settlement_nGBR |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(total),
    se_site = sd(total)/(n()^0.5)
  )
```

```{r}
A <- ggplot()+
    geom_jitter(data = total_mean_site, aes(x=Year, y = mean_site, group = Shelf), 
                position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.85),
    size = 0.9,
    alpha = 0.5,
    colour = "#A8A8A8")+
  geom_pointrange(data = total_model_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin = asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
  labs(y = expression("Total recruits (" * tile^{-1} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1),
        text = element_text(size =9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
A
```

### 1.1.1 Acroporidae

Create data
```{r}
nbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Acroporidae)

nbin_acroa_data <- nbin_Acroporidae_interaction|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  )

acro_mean_site <- Acroporidae |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Acroporidae),
    se_site = sd(Acroporidae)/(n()^0.5)
  )
```

```{r}
B <- ggplot()+
     geom_jitter(data =acro_mean_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 0.9,
    alpha = 0.5,
    colour = "grey"
  )+
  geom_pointrange(data = nbin_acroa_data,aes(ymin = asymp.LCL, ymax = asymp.UCL,x=Year,
             y=response, 
             color=Shelf),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(y=expression("Acroporidae recruits ( " * tile^{-1} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(
    #legend.position = c(1,1), legend.justification = c(1,1),
        legend = element_blank(),
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
B
```

### 1.1.2 Pocilloporidae

Model
```{r}
nbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf * Year + (1 | Reef/Site), 
        family = "nbinom2", data = Pocilloporidae)
```

Create data
```{r}
nbin_poci_data <- nbin_Pocilloporidae_interaction|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  )

poci_mean_site <- Pocilloporidae |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Pocilloporidae),
    se_site = sd(Pocilloporidae)/(n()^0.5)
  )
```


Generate figure
```{r}
C <- ggplot()+
     geom_jitter(data =poci_mean_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 0.9,
    alpha = 0.5,
    colour = "#A8A8A8"
  )+
  geom_pointrange(data = nbin_poci_data,aes(ymin = asymp.LCL, ymax = asymp.UCL,x=Year,
             y=response, 
             color=Shelf),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(y=expression("Pocilloporidae recruits ( " * tile^{-1} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(
    #legend.position = c(1,1), legend.justification = c(1,1),
        legend= element_blank(),
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
C
```

### 1.1.3. Poritidae
Model
```{r}
nbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Poritid)
```

Create data
```{r}
nbin_pori_data <- nbin_Poritid_interaction|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  )

pori_mean_site <- Poritid |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Poritid),
    se_site = sd(Poritid)/(n()^0.5)
  )
```

Generate figure
```{r}
D <- ggplot()+
     geom_jitter(data =pori_mean_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 0.9,
    alpha = 0.5,
    colour = "#A8A8A8"
  )+
  geom_pointrange(data = nbin_pori_data,aes(ymin = asymp.LCL, ymax = asymp.UCL,x=Year,
             y=response, 
             color=Shelf),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(y=expression("Poritidae recruits ( " * tile^{-1} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(
    #legend.position = c(1,1), legend.justification = c(1,1),
        legend = element_blank(),
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
D
```

### 1.1.4. Other 
Model
```{r}
nbin_other_interaction <- glmmTMB(Other ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Other)
```
Create data
```{r}
nbin_other_data <- nbin_other_interaction|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  )

other_mean_site <- Other |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Other),
    se_site = sd(Other)/(n()^0.5)
  )
```

```{r}
E <- ggplot()+
     geom_jitter(data =other_mean_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 0.9,
    alpha = 0.5,
    colour = "#A8A8A8"
  )+
  geom_pointrange(data = nbin_other_data,aes(ymin = asymp.LCL, ymax = asymp.UCL,x=Year,
             y=response, 
             color=Shelf),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(y=expression("Other recruits ( " * tile^{-1} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(
    #legend.position = c(1,1), legend.justification = c(1,1),
        legend = element_blank(),
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
E
```

Combined figure

```{r}
bottom <- ggarrange(B, C, D, E, ncol =2, nrow = 2, labels = c("B", "C", "D", "E"), common.legend = TRUE, legend = "none")
bottom

ggarrange(A, bottom, ncol = 1, labels = c("A", ""), heights = c(1, 1.5)) |> 
  ggsave(filename = "Figure2.PNG", 
         #width = 189,
         height = 240, units = "mm"
        )
```

## 1.2. Composition of Vegetated tiles
```{r}
Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  rename("Other Hard\nCorals" = Other)
```


```{r}
Recruits <- Coral_settlement_nGBR %>%
  filter(total > 0) %>%                                                          # ensure only vegetated tiles included 
  group_by(Reef, Year_Collected, Shelf) %>%               # this was missing in your code
  summarise(across(Acroporidae:`Other Hard
Corals`, mean), .groups = "drop")

recruit_hellinger <- decostand(Recruits %>% select(Acroporidae, Pocilloporidae, Poritid, `Other Hard
Corals`),
                                                         method = "hellinger")

# reef year?
reef_year <- Recruits |> 
  select(Reef, Year_Collected, Shelf)

permanova_result <- adonis2(recruit_hellinger ~ Year_Collected * Shelf,
  data = Recruits,
  method = "euclidean",    # the Hellinger transformation is designed for Euclidean distnaces. Bray-Curtis is not “wrong” but it’s redundant if you’ve already done a Hellinger transformation.
  by = "margin",
  permutations = 999)

print(permanova_result)


euc_hellinger <- vegdist(recruit_hellinger, method = "euclidean")
# With respect to Shelf
betadisper(euc_hellinger, Recruits$Shelf)
anova(betadisper(euc_hellinger, Recruits$Shelf))
```

```{r}
set.seed(999)     # for reproducibility

nmds <- metaMDS(recruit_hellinger, distance = "euclidean", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, reef_year[, c("Year_Collected", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2022" ~ "21/22")
  )
# Creating Vectors for driving the patterns
Recruits_envfit <- envfit(
  nmds,
  recruit_hellinger,
  permutations = 999
) 

#Saving the vectors as a data frame
Recruits_envfit <- as.data.frame(scores(Recruits_envfit, display = "vectors"))

Recruits_envfit <- Recruits_envfit |> 
  rownames_to_column(var = "label_family")
```

```{r}
p1 <- ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 1.5, alpha = 0.8) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.5, 0.05),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
)  +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 0.95,
         y = -0.95,
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1,
         vjust = 0,
         size = 3)+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))

p2 <- ggplot()+
  geom_segment(data = Recruits_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Recruits_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_family),
                  inherit.aes = FALSE,
                  size = 2.5,
                  box.padding = 0.5, 
                  min.segment.length = Inf,
                  nudge_x = -0.1,
                  nudge_y = 0.1)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year Collected"
    )+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))

ST_NMDS <- ggarrange(p1, p2, ncol = 1, heights = c(1, 1))
```

```{r}
ggsave(filename = "Figure3.PNG",
       plot = ST_NMDS,
       height = 220,
       unit = "mm")
```


# 2. Juveniles

## 2.1. Spatio-temporal

### 2.1.1. Total

```{r}
nbin <- glmmTMB(Total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_10m)

total_juv_site <- Belt_10m |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Total_juveniles),
    se_site = sd(Total_juveniles)/(n()^0.5))
  
total_juv_emm <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


juv_total_plot <-  ggplot()+
   geom_point(data =total_juv_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 0.9,
    alpha = 0.5,
    colour = "#A8A8A8")+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = total_juv_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(
   y = expression(atop("Predicted Scleractinian juveniles",
                    "(per 10"~m^{2}*")")),
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "none",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
juv_total_plot
```

### 2.1.2. Acroporidae

```{r}
Acro <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Acroporidae)

nbin <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Acro)

acro_juv_site <- Belt_10m |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Acroporidae),
    se_site = sd(Acroporidae)/(n()^0.5))
  
acro_juv_emm <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


juv_acro_plot <-  ggplot()+
   geom_point(data =acro_juv_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 0.9,
    alpha = 0.5,
    colour = "#A8A8A8")+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = acro_juv_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(
    y = expression(atop("Predicted Acroporidae juveniles",
                    "(per 10"~m^{2}*")")),
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
juv_acro_plot
```

### 2.1.3. Pocilloporidae

```{r}
Poc <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Pocilloporidae)

nbin <- glmmTMB(Pocilloporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Poc)

poci_juv_site <- Belt_10m |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Pocilloporidae),
    se_site = sd(Pocilloporidae)/(n()^0.5))
  
poci_juv_emm <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


juv_poci_plot <-  ggplot()+
   geom_point(data =poci_juv_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 0.9,
    alpha = 0.5,
    colour = "#A8A8A8")+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = poci_juv_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(
    y = expression(atop("Predicted Pocilloporidae juveniles",
                    "(per 10"~m^{2}*")")),
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
juv_poci_plot
```


### 2.1.4. Poritidae

```{r}
Por <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Poritidae)

nbin <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Por)

por_juv_site <- Belt_10m |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Poritidae),
    se_site = sd(Poritidae)/(n()^0.5))
  
por_juv_emm <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


juv_por_plot <-  ggplot()+
   geom_point(data =por_juv_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 0.9,
    alpha = 0.5,
    colour = "#A8A8A8")+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = por_juv_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(
    y = expression(atop("Predicted Poritidae juveniles",
                    "(per 10"~m^{2}*")")),
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
juv_por_plot
```


### 2.1.5. Other Hard Corals

```{r}
Other_juv <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Other_HC)


nbin <- glmmTMB(Other_HC ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Other_juv)

oth_juv_site <- Belt_10m |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Other_HC),
    se_site = sd(Other_HC)/(n()^0.5))
  
oth_juv_emm <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


juv_oth_plot <-  ggplot()+
   geom_point(data =oth_juv_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 0.9,
    alpha = 0.5,
    colour = "#A8A8A8")+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = oth_juv_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(
    y = expression(atop("Predicted other hard coral juveniles",
                    "(per 10"~m^{2}*")")),
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
juv_oth_plot
```

 combined figure

```{r}
juv_bottom <- ggarrange( juv_acro_plot, juv_poci_plot, juv_por_plot, juv_oth_plot,
                         labels = c("B", "C", "D", "E"),
          ncol = 2, 
          nrow = 2,common.legend = TRUE,
          legend = "bottom")

juv_complete <- ggarrange(juv_total_plot, juv_bottom, 
                          labels = c("A", ""),
                          ncol = 1,
          heights = c(0.9, 1.5))
```

```{r}
ggsave(juv_complete,
       file = "Juvenile_emm.PNG",
       height = 250,
       width = 189,
  units = "mm"
       )
```

## 2.2. Composition
```{r}
Belt_10m <- Belt_10m |> 
  rename("Other Hard\nCoral" =Other_HC
  )
# Step 1: Create site-by-species matrix with Site_ID
site_by_belt <- Belt_10m %>%
  select(
    Year, Reef, Shelf, Site, Transect, 26:29
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

site_by_belt <- site_by_belt |>                                                         # ensure only vegetated tiles included 
  group_by(Reef, Year, Shelf) %>%               # this was missing in your code
  summarise(across(Acroporidae:`Other Hard
Coral`, mean), .groups = "drop")

# Step 2: Metadata (KEEP Site_ID)
metadata_belt <- site_by_belt %>%
  select(Shelf, Year, Reef)

# Step 3: Species matrix
belt_matrix_filtered <- site_by_belt %>%
  select(
    4:7
  )
```

```{r}
nmds <- metaMDS(belt_matrix_filtered, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, metadata_belt[, c("Year", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year == "2024-25" ~ "24/25",
      Year == "2023-24" ~ "23/24",
      Year == "2022-23" ~ "22/23",
      Year == "2021-22" ~ "21/22")
  )

belt_A <- ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 1.5, alpha = 0.8) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.5, 0.05),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 0.95,     # right limit
         y = -0.75,       # bottom limit
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1.05, # push slightly inward
         vjust = -0.5, # lift slightly above bottom
         size = 3)+
  scale_x_continuous(limits = c(-0.9, 1.))+
  scale_y_continuous(limits = c(-0.8, 0.5))

belt_A
```

```{r}
# Creating Vectors for driving the patterns
belt_envfit <- envfit(
  nmds,
  belt_matrix_filtered,
  permutations = 999
) 

# Extract p-values
pvals <- belt_envfit$vectors$pvals

# Extract r-squared values
r2 <- belt_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(belt_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- belt_envfit$vectors$r
vec_scores$pval <- belt_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

belt_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")
```

```{r}
belt_B <- ggplot()+
  geom_segment(data = belt_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = belt_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_category),
                  inherit.aes = FALSE,
                  size = 2.5,
                  box.padding = 0.1, 
                  min.segment.length = Inf,
                  nudge_x = -0.05,
                  nudge_y = -0.1)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year"
    )+
  scale_x_continuous(limits = c(-0.9, 1))+
  scale_y_continuous(limits = c(-0.8, 0.5))
belt_B
```
```{r}
belt_Figure <- ggarrange(
  belt_A, belt_B, labels = c("A", "B"), heights = c(1, 1), ncol = 1
)

belt_Figure
```

```{r}
ggsave(belt_Figure,
  filename = "belt_NMDS.png",
  height = 200, 
  units = "mm"
)
```

# 3. Benthic Cover

## 3.1. Composition across the shelf

```{r}
PIT <- PIT |> 
  rename("Other Hard\nCoral" =`Other Hard Coral`,
         "Red\nMacroalgae" = `Red Macroalgae`,
         "Sand and\nRubble" = `Sand and Rubble`)
# Step 1: Create site-by-species matrix with Site_ID
site_by_PIT <- PIT %>%
  select(
    Year, Reef, Shelf, Site, Transect, 95:108
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

site_by_PIT <- site_by_PIT |>                                                         # ensure only vegetated tiles included 
  group_by(Reef, Year, Shelf) %>%               # this was missing in your code
  summarise(across(`Soft Coral`:CCA, mean), .groups = "drop")

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  select(Shelf, Year, Reef)

# Step 3: Species matrix
PIT_matrix_filtered <- site_by_PIT %>%
  select(
    4:17
  )
```

```{r}
# create a distance matrix
PIT_dist <- vegdist(PIT_matrix_filtered, method = "bray")

# With respect to Shelf
betadisper(PIT_dist, metadata_PIT$Shelf)
anova(betadisper(PIT_dist, metadata_PIT$Shelf)) # Significant was 0.01143 * ***
```

```{r}
nmds <- metaMDS(PIT_matrix_filtered, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, metadata_PIT[, c("Year", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year == "2024-25" ~ "24/25",
      Year == "2023-24" ~ "23/24",
      Year == "2022-23" ~ "22/23",
      Year == "2021-22" ~ "21/22")
  )

Benthic_A <- ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 1.5, alpha = 0.8) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.5, 0.05),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 8),
  title = element_text(size = 8)
)+
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 1.05,     # right limit
         y = -1,       # bottom limit
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1.05, # push slightly inward
         vjust = -0.5, # lift slightly above bottom
         size = 3)+
  scale_x_continuous(limits = c(-0.9, 1.05))+
  scale_y_continuous(limits = c(-1, 1))
```


```{r}
# Creating Vectors for driving the patterns
Benthic_envfit <- envfit(
  nmds,
  PIT_matrix_filtered,
  permutations = 999
) 

# Extract p-values
pvals <- Benthic_envfit$vectors$pvals

# Extract r-squared values
r2 <- Benthic_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(Benthic_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- Benthic_envfit$vectors$r
vec_scores$pval <- Benthic_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

vec_scores <- vec_scores |> 
  filter(pval < 0.05) # filter to only show vectors that are significant drivers.

Benthic_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")

#Other, Zoantharian, Poritidae...
```


```{r}
Benthic_envfit$NMDS1_lab <- Benthic_envfit$NMDS1 * 1.1

Benthic_envfit$NMDS2_lab <- Benthic_envfit$NMDS2 * 1.1

Benthic_B <- ggplot()+
  geom_segment(data = Benthic_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Benthic_envfit,
                  aes(x = NMDS1_lab, y= NMDS2_lab, label = label_category),
                  inherit.aes = FALSE,
                   size = 2.5,
    max.overlaps = Inf,
    force = 2,
    max.iter = 20000,
    box.padding = 0.3,
    point.padding = 0.2)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year"
    )+
  scale_x_continuous(limits = c(-0.9, 1.05))+
  scale_y_continuous(limits = c(-1, 1))
```

```{r}
Benthic_Figure <- ggarrange(
  Benthic_A, Benthic_B, labels = c("A", "B"), heights = c(0.9, 1), ncol = 1
  
)

Benthic_Figure
```

Need to move the stress to somewhere less weird, and change CCA name because it looks dumb currently.

```{r}
ggsave(
  filename = "Benthic_NMDS.png",
  height = 240, 
  units = "mm"
)
```

## 3.2. Brood Stock and Recruits

# 4. Larval supply and Juveniles

## 4.1. Spatial Patterns

Load data for these models
```{r}
combine_belt_st <- left_join(join_belt, Join_Coral)
```

```{r}
combine_belt_st <- combine_belt_st |> 
  mutate(Year_start = as.numeric(substr(Year, 1, 4)))
```

```{r}
gropd_df2 <- combine_belt_st %>%
  arrange(Shelf, Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_mean_acro_lag1 = lag(st_mean_acro, 1),
    st_mean_poci_lag1 = lag(st_mean_poci, 1),
    st_mean_por_lag1 = lag(st_mean_por, 1),
  ) %>%
  ungroup()
gropd_df2 <- gropd_df2 |> 
  mutate(Shelf, factor(Shelf, levels = c("Inner", "Mid", "Outer"), ordered = TRUE))
```


### 4.1.1. Acroporidae

```{r}
#model
model_df_acro <- gropd_df2 %>%
  filter(!is.na(mean_acro))

ln_acro_juv_st_int_wo_yr <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)
summary(model_df_acro$st_mean_acro_lag1)
```

```{r}

library(ggplot2)
model_df_acro$st_mean_acro_lag1_z <-
  scale(model_df_acro$st_mean_acro_lag1)

ln_acro_juv_st_int_wo_yr <- glmmTMB(
  mean_acro ~ st_mean_acro_lag1_z * Shelf + (1|Reef/Site),
  family = "lognormal",
  data = model_df_acro
)

pred <- ggpredict(
  ln_acro_juv_st_int_wo_yr,
  terms = c("st_mean_acro_lag1_z [all]", "Shelf")
)

acro_juv_st <- ggplot(pred, aes(x = x, y = predicted, colour = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, colour = NA) +
  labs(x= expression(atop("Lagged Acroporidae larval supply",
                    "(t)")),
       y = expression(atop("Predicted Acroporidae", "juveniles (t + 1)")),
       colour = "Shelf",
       fill = "Shelf") +
  facet_wrap(~group, scale = "free_y")+
  theme_classic()+scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+scale_fill_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

acro_juv_st
```

### 4.1.2. Pocilloporidae

```{r}
model_df_poci <- gropd_df2 %>%
  filter(!is.na(mean_poci))
ln_poci_juv_st_int_wo_yr <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)

pred <- ggpredict(
  ln_poci_juv_st_int_wo_yr,
  terms = c("st_mean_poci_lag1 [all]", "Shelf")
)

poci_juv_st <- ggplot(pred, aes(x = x, y = predicted, colour = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, colour = NA) +
  labs(x= expression(atop("Lagged Pocilloporidae larval supply",
                    "(t)")),
       y = expression(atop("Predicted Pocilloporidae", "juveniles (t + 1)")),
       colour = "Shelf",
       fill = "Shelf") +
  facet_wrap(~group, scale = "free_y")+
  theme_classic()+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_fill_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

poci_juv_st
```

### 4.1.3. Poritidae

```{r}
model_df_por <- gropd_df2 %>%
  filter(!is.na(mean_por))

model_df_por <- model_df_por |> 
  filter(mean_por > 0)

ln_por_juv_st_int_wo_yr <- glmmTMB(mean_por ~ st_mean_por_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_por)

pred <- ggpredict(
  ln_por_juv_st_int_wo_yr,
  terms = c("st_mean_por_lag1 [all]", "Shelf")
)

por_juv_st <- ggplot(pred, aes(x = x, y = predicted, colour = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, colour = NA) +
  labs(x= expression(atop("Lagged Poritidae larval supply",
                    "(t)")),
       y = expression(atop("Predicted Poritidae", "juveniles (t + 1)")),
       colour = "Shelf",
       fill = "Shelf") +
  facet_wrap(~group, scale = "free_y")+
  theme_classic()+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_fill_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

por_juv_st
```

```{r}
Juv_st_relationship <- ggarrange(acro_juv_st, poci_juv_st, por_juv_st, ncol =1, nrow =3, common.legend = TRUE,
          labels = c("A", "B", "C"))

ggsave(Juv_st_relationship,
       file = "juv_st_relationship_fig.PNG",
       height = 220,
       width = 189,
       unit = "mm")
```



# 5. Composition Plot Combined

```{r}
# Make all nmds one column

NMDS_combined <- ggarrange(p1, belt_A, Benthic_A, ncol =3, labels = c("A", "C", "E"), common.legend = TRUE, legend = "bottom", widths = c(0.8, 0.8, 1))
NMDS_combined

Envfit_combined <- ggarrange(p2, belt_B, Benthic_B, ncol =3, labels = c("B", "D", "F"), widths = c(0.8, 0.8, 1))
Envfit_combined

Complete_Composition <- ggarrange(NMDS_combined, Envfit_combined, ncol = 1, heights = c(1.1,0.9))
Complete_Composition
```


```{r}
ST_NMDS <- ggarrange(p1, p2, ncol = 1, labels = c("A", "B"))

belt_Figure <- ggarrange(belt_A, belt_B, labels = c("C", "D"), ncol = 1)

Benthic_Figure <- ggarrange(
  Benthic_A, Benthic_B, labels = c("E", "F"), ncol = 1
)

complete_composition <- ggarrange(ST_NMDS, belt_Figure, Benthic_Figure, widths = c(0.8, 0.8, 1), ncol = 3)
complete_composition

ggsave(filename = "Complete_composition.PDF",
       plot = complete_composition,
       width = 250,
       height = 200,
       units = "mm")

```

