---
title: "Figures"
author: "Monique White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)
library(ggordiplots) #nmds and ordination plots

# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  ),
  Deployment = case_when(
      Year_Collected == "2022" ~ "21/22",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2026" ~ "25/26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")

Pocilloporidae <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Pocilloporidae, Top_Pocil, 
         Under_Pocil, Side_Pocil, total)


Acroporidae <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Acroporidae, Top_Acr, 
         Under_Acr, Side_Acr, total)




Poritid <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Poritid, Top_Por, Under_Por, 
         Side_Por, total)


Other <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Other, Top_other, 
         Under_oth, Side_other, total)

# Mean or mean no. tiles per site / year
Join_Coral <- Coral_settlement_nGBR |> 
  group_by(Shelf, Year, Site, Reef) |> 
  summarise(
    st_mean_acro = mean(Acroporidae),
    st_sd_acro = sd(Acroporidae),
    st_median_acro = median(Acroporidae),
    st_mean_poci = mean(Pocilloporidae),
    st_sd_poci = sd(Pocilloporidae),
    st_median_poci = median(Pocilloporidae),
    st_mean_por = mean(Poritid),
    st_sd_por = sd(Poritid),
    st_median_por = median(Poritid),
    st_mean_total = mean(total),
    st_sd_total = sd(total),
    Tiles = n()
  )

# Benthic Data

# Benthic Data
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    )
  )


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

PIT_detailed <- PIT |> 
  dplyr::select(1:93)

PIT <- PIT %>%
  mutate(
    Acroporidae = rowSums(dplyr::select(., `Acropora - branching`:`Montipora`), na.rm = TRUE),
    
    Pocilloporidae = rowSums(dplyr::select(., Pocillopora:Stylophora), na.rm = TRUE),
    
    Poritidae = rowSums(dplyr::select(., `Porites - branching`:Goniopora), na.rm = TRUE),
    
    `Other Hard coral` = rowSums(dplyr::select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE),
    
    `Soft coral` = rowSums(dplyr::select(., Lobophytum:`Soft-other`), na.rm = TRUE),
    `Brown MA` = rowSums(dplyr::select(., Sargassum:Lobophora), na.rm = TRUE),
    `Red MA` = rowSums(dplyr::select(., Asparagopsis:Jania), na.rm = TRUE),
    `Green MA`= rowSums(dplyr::select(., Halimeda:Valonia), na.rm = TRUE),
    `Sand and Rubble` = rowSums(dplyr::select(., Sand:Rubble), na.rm = TRUE),
    Turf = rowSums(dplyr::select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE),
    Other = rowSums(dplyr::select(., Sponge:Clam), na.rm = TRUE),
    Zoantharian = rowSums(dplyr::select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  dplyr::select(Year, Shelf, Reef, Site, Transect, Acroporidae:Other,CCA, Cyanobacteria, Halophila)

PIT <- PIT |> 
  mutate(
    sum = rowSums(across(Acroporidae:Halophila), na.rm = TRUE)
  )

summary(PIT$sum)

PIT <- PIT |> 
  mutate(
    across(Acroporidae:Halophila, ~ .x * 2)
  )

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae)
  )

Combined <- left_join(Coral_settlement_nGBR, join)
# Juvenile data
Belt_10m <- read_csv("Belt_10m_final.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
 mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))
Belt_10m <- Belt_10m |>
  mutate(
    Other_HC = Favia + Favites + Goniastrea + Astrea +
                 Echinopora + Galaxea + Turbinaria +
                 Psammacora + Pavona
  )

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(Total_juveniles),
    sd_total = sd(Total_juveniles),
    Belt = n()
  )

Combined1 <- left_join(Coral_settlement_nGBR, join_belt)

## Total recuits
total_recruits <- Coral_settlement_nGBR %>%
  dplyr::select(Year_Collected, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 5:8, names_to = "Family", values_to = "Total")

#Order so that Other is last in Family
total_recruits <- Coral_settlement_nGBR %>% 
  dplyr::select(Year_Collected, Shelf, Reef, Site, Obs, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 6:9, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))

total_recruits$Year_Collected <- as.factor(total_recruits$Year_Collected)

## Position on tile

position_total_recruits <- Coral_settlement_nGBR %>%
  dplyr::select(2:9, 15:29, 32) %>% 
  pivot_longer(cols = 21:23, names_to = "Pos", values_to = "Total")

position_total_recruits <- position_total_recruits |> 
  mutate(Pos = case_when(
      Pos == "Top.Total" ~ "Top",
      Pos == "Side.Total" ~ "Side",
      Pos == "Under.Total" ~ "Bottom"
  ))
 
position_total_recruits <- position_total_recruits |>                           
mutate(Pos = factor(Pos, levels = c("Bottom", "Side", "Top")))

position_total_recruits <- left_join(position_total_recruits, join)

## Family recruitment data
family_total_recruits <- Coral_settlement_nGBR %>%
  dplyr::select(2:13) %>% 
  pivot_longer(cols = 9:12, names_to = "Family", values_to = "Total") |> 
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))
```

# 1. Recruitment

## 1.1 Spatio-temporal
Figure 1

```{r}
nbin_total_yr2 <- glmmTMB(total ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

total_model_emm <- nbin_total_yr2|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) 

total_mean_site <- Coral_settlement_nGBR |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(total),
    se_site = sd(total)/(n()^0.5)
  )
```

```{r}
library(scales)
A <- ggplot()+
    geom_jitter(data = total_mean_site, aes(x=Year, y = mean_site, group = Shelf), 
                position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1)+
  geom_pointrange(data = total_model_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin = asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  #scale_y_log10()+ #TBD whether to include
  #scale_y_continuous(trans = "log1p")+
  scale_y_continuous(
    trans = log1p_trans(),
    breaks = c(0, 10, 50, 100, 200, 300)
  )+
  theme_classic()+
  labs(y = expression("Predicted recruits (" * tile^{-1} * ")"),
       title = "Total")+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1),
        text = element_text(size =12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
A

A.2 <- ggplot()+
    geom_jitter(data = total_mean_site, aes(x=Year, y = mean_site, group = Shelf), 
                position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1)+
  geom_pointrange(data = total_model_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin = asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  scale_y_log10()+ #TBD whether to include
  theme_classic()+
  labs(y = expression("Predicted recruits (" * tile^{-1} * ")"),
       title = "Total")+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1),
        text = element_text(size =12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
A.2
```

### 1.1.1 Acroporidae

Create data
```{r}
nbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Acroporidae)

nbin_acroa_data <- nbin_Acroporidae_interaction|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  )

acro_mean_site <- Acroporidae |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Acroporidae),
    se_site = sd(Acroporidae)/(n()^0.5)
  )
```

```{r}
B <- ggplot()+
     geom_jitter(data =acro_mean_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060",
    shape = 1
  )+
  geom_pointrange(data = nbin_acroa_data,aes(ymin = asymp.LCL, ymax = asymp.UCL,x=Year,
             y=response, 
             color=Shelf),
                  position = position_dodge(width = 0.85))+
  scale_y_continuous(
    trans = log1p_trans(),
    breaks = c(0, 10, 50, 100, 200, 300)
  )+
  theme_classic()+
  labs(y=expression("Predicted recruits (" * tile^{-1} * ")"),
       title = "Acroporidae")+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(
    #legend.position = c(1,1), legend.justification = c(1,1),
        legend = element_blank(),
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
B
```

### 1.1.2 Pocilloporidae

Model
```{r}
nbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf * Year + (1 | Reef/Site), 
        family = "nbinom2", data = Pocilloporidae)
```

Create data
```{r}
nbin_poci_data <- nbin_Pocilloporidae_interaction|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  )

poci_mean_site <- Pocilloporidae |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Pocilloporidae),
    se_site = sd(Pocilloporidae)/(n()^0.5)
  )
```


Generate figure
```{r}
C <- ggplot()+
     geom_jitter(data =poci_mean_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1
  )+
  geom_pointrange(data = nbin_poci_data,aes(ymin = asymp.LCL, ymax = asymp.UCL,x=Year,
             y=response, 
             color=Shelf),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
  labs(y=expression("Predicted recruits (" * tile^{-1} * ")"),
       title = "Pocilloporidae")+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(
    #legend.position = c(1,1), legend.justification = c(1,1),
        legend= element_blank(),
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
C
```

### 1.1.3. Poritidae
Model
```{r}
nbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Poritid)
```

Create data
```{r}
nbin_pori_data <- nbin_Poritid_interaction|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  )

pori_mean_site <- Poritid |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Poritid),
    se_site = sd(Poritid)/(n()^0.5)
  )
```

Generate figure
```{r}
D <- ggplot()+
     geom_jitter(data =pori_mean_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1
  )+
  geom_pointrange(data = nbin_pori_data,aes(ymin = asymp.LCL, ymax = asymp.UCL,x=Year,
             y=response, 
             color=Shelf),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
  labs(y=expression("Predicted recruits (" * tile^{-1} * ")"),
       title = "Poritidae")+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(
    #legend.position = c(1,1), legend.justification = c(1,1),
        legend = element_blank(),
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
D
```

### 1.1.4. Other 
Model
```{r}
nbin_other_interaction <- glmmTMB(Other ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Other)
```
Create data
```{r}
nbin_other_data <- nbin_other_interaction|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  )

other_mean_site <- Other |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Other),
    se_site = sd(Other)/(n()^0.5)
  )
```

```{r}
E <- ggplot()+
     geom_jitter(data =other_mean_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1
  )+
  geom_pointrange(data = nbin_other_data,aes(ymin = asymp.LCL, ymax = asymp.UCL,x=Year,
             y=response, 
             color=Shelf),
                  position = position_dodge(width = 0.85), )+
  theme_classic()+
  labs(y=expression("Other recruits ( " * tile^{-1} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(
    #legend.position = c(1,1), legend.justification = c(1,1),
        legend = element_blank(),
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
E
```

Combined figure

```{r}
# recruits_wo_other <- ggarrange(B, C, D, ncol =1, nrow = 3, labels = c("B", "C", "D"), common.legend = TRUE, legend = "none")
# 
# recruits_wo_other <- ggarrange(A, recruits_wo_other, ncol =1, labels = c("A", ""), heights = c(1, 3))
# 
# recruits_wo_other


# recruits_wo_other <- ggarrange(A, B, C, D, ncol =1, nrow = 4, labels = c("A","B", "C", "D"), common.legend = TRUE, legend = "bottom")
# 
#   ggsave(recruits_wo_other, filename = "Figure2.PDF", 
#          width = 189,
#          height = 325, units = "mm"
#         )
  
recruits_wo_other <- ggarrange(A, B, C, D, ncol =1, nrow = 4, labels = c("A","B", "C", "D"), common.legend = TRUE, legend = "bottom")
   # ggsave(recruits_wo_other, filename = "Figure2.PNG",
   #       width = 320,
   #       height = 220, units = "mm"
   #      )

recruits_wo_other
```

```{r eval = FALSE}
   ggsave(recruits_wo_other, filename = "figures/Figure2.PDF",
         width = 180,
         height = 300,
         units = "mm"
        )
   
```

```{r}
bottom <- ggarrange(B, C, D, E, ncol =2, nrow = 2, labels = c("B", "C", "D", "E"), common.legend = TRUE, legend = "none")
bottom

ggarrange(A, bottom, ncol = 1, labels = c("A", ""), heights = c(1, 1.5)) |> 
  ggsave(filename = "Figure2.PNG", 
         #width = 189,
         height = 240, units = "mm"
        )
```

### Presentation 

```{r}
ggsave(A, filename = "Presentation_fig/Total_recruits.PNG",
         width = 180,
         height = 120,
         units = "mm"
        )

ggsave(B, filename = "Presentation_fig/Acro_recruits.PNG",
         width = 250,
         height = 100,
         units = "mm"
        )

Families_recruits <- ggarrange(B, C, D, ncol =3, nrow = 1, common.legend = TRUE, legend = "bottom")


ggsave(Families_recruits, filename = "Presentation_fig/Families_recruits.PNG",
         width = 325,
         height = 120,
         units = "mm"
        )

Families_recruits <- ggarrange(B, C, D, ncol =1, nrow = 3, common.legend = TRUE, legend = "none")


ggsave(Families_recruits, filename = "Presentation_fig/Families_recruits_portrait.PNG",
         width = 180,
         height = 200,
         units = "mm"
        )
```



## 1.2. Composition of Vegetated tiles
```{r}
Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  rename("Other Hard Corals" = Other)
```

```{r}
# Vegetated tiles 
Recruits <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  dplyr::select(Acroporidae, Pocilloporidae, Poritid, Other)

environment_data <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  dplyr::select(Year_Collected, Site, Reef, Shelf)

environment_data |> count(Shelf, Year_Collected)
```

```{r}
# Filtered recruit data
Recruits_sqrt <- sqrt(Recruits) #Squareroot transformed because Acroporidae is more abundant than the other taxa

recruits_sqrt_enviro <- bind_cols(environment_data, Recruits_sqrt)
#write.csv(Recruits, file = "Matrices_for_PRIMER/Recruits_raw.csv")
# write.csv(Recruits_sqrt, file = "Matrices_for_PRIMER/Recruits_sqrt.csv")
# write.csv(environment_data, file = "Matrices_for_PRIMER/recruit_enviro.csv")
#write.csv(recruits_sqrt_enviro, file = "Matrices_for_PRIMER/recruit_sqrt_enviro.csv")
# To stratify by year (this asks "do shelf differences exist after accounting for year structure?") 
h <- how(blocks = environment_data$Reef)                                                   # restrict permutations within years
```

```{r}
set.seed(999)     # for reproducibility

nmds <- metaMDS(Recruits_sqrt, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress #0.15

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, environment_data[, c("Year_Collected", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Year =
    case_when(
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2022" ~ "21/22")
  )
# Creating Vectors for driving the patterns
Recruits_envfit <- envfit(
  nmds,
  Recruits_sqrt,
  permutations = 999 
) 

#Saving the vectors as a data frame
Recruits_envfit <- as.data.frame(scores(Recruits_envfit, display = "vectors"))

Recruits_envfit <- Recruits_envfit |> 
  rownames_to_column(var = "label_family")
```


```{r}
library(MASS)  # for mvrnorm
library(vegan)

grp <- interaction(scores$Shelf, scores$Year, drop = TRUE)
# Function to generate ellipse points for ggplot
ellipse_points <- function(ell_obj, npoints = 100) {
  do.call(rbind,
    lapply(names(ell_obj), function(g) {
      center <- ell_obj[[g]]$center      # centroid
      cov_mat <- ell_obj[[g]]$cov        # covariance matrix
      angles <- seq(0, 2*pi, length.out = npoints)
      circle <- cbind(cos(angles), sin(angles))
      # scale unit circle by sqrt of covariance (1 SD)
      ellipse <- t(center + t(circle %*% chol(cov_mat)))
      data.frame(NMDS1 = ellipse[,1],
                 NMDS2 = ellipse[,2],
                 Shelf = g)
    })
  )
}

ell <- ordiellipse(nmds, groups = grp, kind = "sd", draw = "none")
ell_df <- ellipse_points(ell, npoints = 100)

ell_df <- ell_df |>
  tidyr::separate(Shelf, into = c("Shelf", "Year"), sep = "\\.")
```


```{r}
p1 <- ggplot(data = scores, aes(x = NMDS1, y = NMDS2,colour = Shelf)) +
     geom_point(size = 1, alpha = 0.8) +
  geom_path(data = ell_df,
            aes(x = NMDS1, y = NMDS2, colour = Shelf, group = Shelf),
            linewidth = 0.6) +
    # stat_ellipse(aes(group = Shelf, colour = Shelf),
    #              level = 0.75, linewidth = 0.5) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.5, 0.05),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
)  +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 0.95,
         y = -0.95,
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1,
         vjust = 0,
         size = 3)+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))+
  facet_wrap(~Year)



p2 <- ggplot()+
  geom_segment(data = Recruits_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Recruits_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_family),
                  inherit.aes = FALSE,
                  size = 2.5,
                  box.padding = 0.5, 
                  min.segment.length = Inf,
                  nudge_x = -0.1,
                  nudge_y = 0.1)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year Collected"
    )+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))

ST_NMDS <- ggarrange(p1, p2, ncol = 1, heights = c(1, 1), legend = "none")
ST_NMDS
```

```{r eval = FALSE}
# ggsave(filename = "Figure3.PNG",
#        plot = ST_NMDS,
#        height = 220,
#        unit = "mm")
# 
ggsave(filename = "figures/Figure3.PDF",
       plot = ST_NMDS,
       height = 220,
       unit = "mm")
# 
# ggsave(filename = "figures/recruit_vec_nmds.PDF",
#        plot = p1.5,
#        height = 125, width = 150,
#        unit = "mm")
```

### 1.2.2. Sqrt
```{r}
Recruits <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  dplyr::select(Acroporidae, Pocilloporidae, Poritid, Other)

environment_data <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  dplyr::select(Year_Collected, Site, Reef, Shelf)

recruits_sqrt <- sqrt(Recruits)
set.seed(999)     # for reproducibility

nmds <- metaMDS(recruits_sqrt, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress #0.15

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, environment_data[, c("Year_Collected", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Year =
    case_when(
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2022" ~ "21/22")
  )
# Creating Vectors for driving the patterns
Recruits_envfit <- envfit(
  nmds,
  recruits_sqrt,
  permutations = 999#,
  # permutations = h #restrict permutations to year, change if using PERMANOVA from PRIMER
) 

# #Saving the vectors as a data frame
# Recruits_envfit <- as.data.frame(scores(Recruits_envfit, display = "vectors"))
# 
# Recruits_envfit <- Recruits_envfit |> 
#   rownames_to_column(var = "label_family")

# Extract p-values
pvals <- Recruits_envfit$vectors$pvals

# Extract r-squared values
r2 <- Recruits_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(Recruits_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- Recruits_envfit$vectors$r
vec_scores$pval <- Recruits_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

vec_scores <- vec_scores |> 
  filter(pval < 0.05, r2 > 0.25) # filter to only show vectors that are significant drivers.

Recruits_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")
```
```{r}
library(MASS)  # for mvrnorm
library(vegan)

grp <- interaction(scores$Shelf, scores$Year, drop = TRUE)
# Function to generate ellipse points for ggplot
ellipse_points <- function(ell_obj, npoints = 100) {
  do.call(rbind,
    lapply(names(ell_obj), function(g) {
      center <- ell_obj[[g]]$center      # centroid
      cov_mat <- ell_obj[[g]]$cov        # covariance matrix
      angles <- seq(0, 2*pi, length.out = npoints)
      circle <- cbind(cos(angles), sin(angles))
      # scale unit circle by sqrt of covariance (1 SD)
      ellipse <- t(center + t(circle %*% chol(cov_mat)))
      data.frame(NMDS1 = ellipse[,1],
                 NMDS2 = ellipse[,2],
                 Shelf = g)
    })
  )
}

ell <- ordiellipse(nmds, groups = grp, kind = "sd", draw = "none")
ell_df <- ellipse_points(ell, npoints = 100)

ell_df <- ell_df |>
  tidyr::separate(Shelf, into = c("Shelf", "Year"), sep = "\\.")
```


```{r}
p1 <- ggplot(data = scores, aes(x = NMDS1, y = NMDS2,colour = Shelf)) +
     geom_point(size = 1, alpha = 0.8) +
  geom_path(data = ell_df,
            aes(x = NMDS1, y = NMDS2, colour = Shelf, group = Shelf),
            linewidth = 0.6) +
    # stat_ellipse(aes(group = Shelf, colour = Shelf),
    #              level = 0.75, linewidth = 0.5) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.5, 0.05),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
)  +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 0.95,
         y = -0.95,
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1,
         vjust = 0,
         size = 3)+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))+
  facet_wrap(~Year)

p2 <- ggplot()+
  geom_segment(data = Recruits_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Recruits_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_family),
                  inherit.aes = FALSE,
                  size = 2.5,
                  box.padding = 0.5, 
                  min.segment.length = Inf,
                  nudge_x = -0.1,
                  nudge_y = 0.1)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year Collected"
    )+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))

# ST_NMDS <- ggarrange(p1, p2, ncol = 1, heights = c(1, 1), legend = "none")
# ST_NMDS
```

```{r}
library(patchwork)

# Create spacers
spacer <- plot_spacer()

# Bottom row: spacer | plot | spacer
bottom_centered <- spacer + p2 + spacer +
  plot_layout(ncol = 3, widths = c(0.9, 2.1, 1))  # adjust middle width

# Combine with top plot
final_plot_st <- p1 / bottom_centered +
  plot_layout(heights = c(2, 0.9))  # adjust vertical proportions

final_plot_st

ggsave(final_plot_st,
       file = "figures/Recruits_NMDS.PDF",
       height = 220,
       width = 189,
       unit = "mm")

```


# 2. Juveniles

## 2.1. Spatio-temporal

### 2.1.1. Total

```{r}
nbin <- glmmTMB(Total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_10m)

total_juv_site <- Belt_10m |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Total_juveniles),
    se_site = sd(Total_juveniles)/(n()^0.5))
  
total_juv_emm <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


juv_total_plot <-  ggplot()+
   geom_point(data =total_juv_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1)+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = total_juv_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
  labs(
   y = expression(atop("Predicted juvenile",  "colonies (10"~m^{-2}*")")),
   title = "Total",
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "none",
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
juv_total_plot
```

### 2.1.2. Acroporidae

```{r}
Acro <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Acroporidae)

nbin <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Acro)

acro_juv_site <- Belt_10m |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Acroporidae),
    se_site = sd(Acroporidae)/(n()^0.5))
  
acro_juv_emm <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


juv_acro_plot <-  ggplot()+
   geom_point(data =acro_juv_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1)+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = acro_juv_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
 labs(
   y = expression(atop("Predicted juvenile",  "colonies (10"~m^{-2}*")")),
   title = "Acroporidae",
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
juv_acro_plot
```

*text box of the error family is an option for when models are from different distributions*

### 2.1.3. Pocilloporidae

```{r}
Poc <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Pocilloporidae)

nbin <- glmmTMB(Pocilloporidae ~ Year+Shelf+(1 | Reef/Site), #doing fixed model because that performs the best
        family = "nbinom2", data = Poc)


poci_juv_site <- Belt_10m |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Pocilloporidae),
    se_site = sd(Pocilloporidae)/(n()^0.5))
  
poci_juv_emm <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


juv_poci_plot <-  ggplot()+
   geom_point(data =poci_juv_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1)+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = poci_juv_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
  labs(
   y = expression(atop("Predicted juvenile",  "colonies (10"~m^{-2}*")")),
   title = "Pocilloporidae",
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
juv_poci_plot
```


### 2.1.4. Poritidae

```{r}
Por <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Poritidae)

nbin <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Por)

por_juv_site <- Belt_10m |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Poritidae),
    se_site = sd(Poritidae)/(n()^0.5))
  
por_juv_emm <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


juv_por_plot <-  ggplot()+
   geom_point(data =por_juv_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1)+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = por_juv_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
 labs(
   y = expression(atop("Predicted juvenile",  "colonies (10"~m^{-2}*")")),
   title = "Poritidae",
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
juv_por_plot
```

```{r}
# ggarrange(A, juv_total_plot, B, juv_acro_plot, C,juv_poci_plot, D, juv_por_plot, ncol = 2, nrow = 4, common.legend = TRUE, legend = "bottom", labels = c("A", "B", "C", "D", "E", "F", "G", "H") )

Total_st_juv <- ggarrange(A, juv_total_plot, ncol = 1, labels = c("A", "B"), common.legend = TRUE)
# Total_st_juv
# 
ggsave(Total_st_juv,
       filename = "Total_ST_juv.PDF",
       height = 200,
       width = 189,
       unit = "mm")
# 
# ggsave(Total_st_juv,
#        filename = "Total_ST_juv.PDF",
#        height = 200,
#        width = 189,
#        unit = "mm")
```


```{r}
# Juv_4 <- ggarrange(juv_total_plot, juv_acro_plot, juv_poci_plot, juv_por_plot, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom", labels = c("A", "B", "C", "D") )
# 
# ggsave(Juv_4,
#        filename = "Figure4.PNG",
#        width = 320,
#       height = 220,
#        unit = "mm")
# 
# ggsave(Juv_4,
#        filename = "figures/Figure4.PDF",
#        width = 320,
#       height = 220,
#        unit = "mm")

Juv_4 <- ggarrange(juv_total_plot, juv_acro_plot, juv_poci_plot, juv_por_plot, ncol = 1, nrow = 4, common.legend = TRUE, legend = "bottom", labels = c("A", "B", "C", "D") )

Juv_4
```

```{r eval = FALSE}
# ggsave(Juv_4,
#        filename = "Figure4.PNG",
#        height = 200,
#        width = 275,
#        unit = "mm")

# ggsave(Juv_4,
#        filename = "Figure4.PNG",
#        width = 320,
#       height = 220,
#        unit = "mm")

ggsave(Juv_4,
       filename = "figures/Figure4.PDF",
       width = 180,
      height = 275,
       unit = "mm")
```

### 2.1.5. Other Hard Corals

```{r}
Other_juv <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Other_HC)


nbin <- glmmTMB(Other_HC ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Other_juv)

oth_juv_site <- Belt_10m |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Other_HC),
    se_site = sd(Other_HC)/(n()^0.5))
  
oth_juv_emm <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


juv_oth_plot <-  ggplot()+
   geom_point(data =oth_juv_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1)+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = oth_juv_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
  labs(
    y = expression(atop("Predicted other hard coral juveniles",
                    "(per 10"~m^{2}*")")),
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
juv_oth_plot
```

 combined figure

```{r}
juv_bottom <- ggarrange( juv_acro_plot, juv_poci_plot, juv_por_plot, juv_oth_plot,
                         labels = c("B", "C", "D", "E"),
          ncol = 2, 
          nrow = 2,common.legend = TRUE,
          legend = "bottom")

juv_complete <- ggarrange(juv_total_plot, juv_bottom, 
                          labels = c("A", ""),
                          ncol = 1,
          heights = c(0.9, 1.5))
```

```{r eval = FALSE}
ggsave(juv_complete,
       file = "Juvenile_emm.PNG",
       height = 250,
       width = 189,
  units = "mm"
       )
```

### Presentation 

```{r}
ggsave(juv_total_plot, filename = "Presentation_fig/Total_juv.PNG",
         width = 180,
         height = 120,
         units = "mm"
        )

Families_juv <- ggarrange(juv_acro_plot, juv_poci_plot, juv_por_plot, ncol =3, nrow = 1, common.legend = TRUE, legend = "bottom")


ggsave(Families_juv, filename = "Presentation_fig/Families_juv.PNG",
         width = 325,
         height = 120,
         units = "mm"
        )

Families_juv <- ggarrange(juv_acro_plot, juv_poci_plot, juv_por_plot, ncol = 1, nrow = 3, common.legend = TRUE, legend = "bottom")


ggsave(Families_juv, filename = "Presentation_fig/Families_juv_portrait.PNG",
         width = 180,
         height = 200,
         units = "mm"
        )
```

## 2.2. Composition
 sqrt

```{r}
Belt_10m <- Belt_10m |> 
  rename("Other Hard Coral" = Other_HC
  )
 
site_by_belt <- Belt_10m %>%
  filter(Total_juveniles > 0) |> #filter to make sure that dissimilarities are non zero
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, Acroporidae, Pocilloporidae, Poritidae, `Other Hard Coral`
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"),
         Shelf_Year = paste(Year, Shelf, sep = "_"))

# Step 2: Metadata (KEEP Site_ID)
metadata_belt <- site_by_belt %>%
  dplyr::select(Shelf, Year, Reef, Site, Shelf_Year)

# metadata_belt|> group_by(Shelf, Reef, Site, Year) |> 
# summarise(reps = n()) |>
#   pivot_wider(id_cols = c("Shelf", "Site"), names_from = Year, values_from = reps) |> write.csv("output/belt_transects_with_juveniles.csv")

# Step 3: Species matrix
belt_matrix_filtered <- site_by_belt %>%
  dplyr::select(
    Acroporidae:`Other Hard Coral`
  )
belt_matrix_filtered_sqrt <- sqrt(belt_matrix_filtered)
belt_detailed_sqrt_dist <- vegdist(belt_matrix_filtered_sqrt, method = "bray")
```

```{r}
nmds <- metaMDS(belt_matrix_filtered_sqrt, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress #0.195

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, metadata_belt[, c("Year", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year == "2024-25" ~ "24/25",
      Year == "2023-24" ~ "23/24",
      Year == "2022-23" ~ "22/23",
      Year == "2021-22" ~ "21/22")
  )

summary(scores$NMDS1) #ranfe NMDS1 -1.15, 1.65
summary(scores$NMDS2) #range NMDS2 -1.25, 0.75
```

```{r}
library(MASS)  # for mvrnorm
library(vegan)
grp <- interaction(scores$Shelf, scores$Year, drop = TRUE)
# Function to generate ellipse points for ggplot
ellipse_points <- function(ell_obj, npoints = 100) {
  do.call(rbind,
    lapply(names(ell_obj), function(g) {
      center <- ell_obj[[g]]$center      # centroid
      cov_mat <- ell_obj[[g]]$cov        # covariance matrix
      angles <- seq(0, 2*pi, length.out = npoints)
      circle <- cbind(cos(angles), sin(angles))
      # scale unit circle by sqrt of covariance (1 SD)
      ellipse <- t(center + t(circle %*% chol(cov_mat)))
      data.frame(NMDS1 = ellipse[,1],
                 NMDS2 = ellipse[,2],
                 Shelf = g)
    })
  )
}

ell <- ordiellipse(nmds, groups = grp, kind = "sd", draw = "none")
ell_df <- ellipse_points(ell, npoints = 100)

ell_df <- ell_df |>
  tidyr::separate(Shelf, into = c("Shelf", "Year"), sep = "\\.")
# # Function to generate ellipse points for ggplot
# ellipse_points <- function(ell_obj, npoints = 100) {
#   do.call(rbind,
#     lapply(names(ell_obj), function(g) {
#       center <- ell_obj[[g]]$center      # centroid
#       cov_mat <- ell_obj[[g]]$cov        # covariance matrix
#       angles <- seq(0, 2*pi, length.out = npoints)
#       circle <- cbind(cos(angles), sin(angles))
#       # scale unit circle by sqrt of covariance (1 SD)
#       ellipse <- t(center + t(circle %*% chol(cov_mat)))
#       data.frame(NMDS1 = ellipse[,1],
#                  NMDS2 = ellipse[,2],
#                  Shelf = g)
#     })
#   )
# }
# 
# ell <- ordiellipse(nmds, groups = scores$Shelf, kind = "sd", draw = "none")
# ell_df <- ellipse_points(ell, npoints = 100)
```

```{r}
belt_A_sqrt <- ggplot(scores, aes(x = NMDS1, y = NMDS2,
                                           colour = Shelf)) +
     geom_point(size = 1, alpha = 0.8) +
  geom_path(data = ell_df,
            aes(x = NMDS1, y = NMDS2, colour = Shelf, group = Shelf),
            linewidth = 0.6) +
  
    # stat_ellipse(aes(group = Shelf, colour = Shelf),
    #              level = 0.75, linewidth = 0.5) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = "none", # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 0.95,     # right limit
         y = -0.75,       # bottom limit
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1.05, # push slightly inward
         vjust = -0.75, # lift slightly above bottom
         size = 3)+
  facet_wrap(~Year)+
  scale_x_continuous(limits = c(-1.15, 1.6))+
  scale_y_continuous(limits = c(-1.1, 0.85))

belt_A_sqrt
```

```{r}
# To stratify by year (this asks "do shelf differences exist after accounting for year structure?") 
#h <- how(blocks = metadata_belt$Year)                                                   # restrict permutations within years

# Creating Vectors for driving the patterns
belt_envfit <- envfit(
  nmds,
  belt_matrix_filtered_sqrt #,
  #permutations = h #restricts permuatations to a year
) 

# Extract p-values
pvals <- belt_envfit$vectors$pvals

# Extract r-squared values
r2 <- belt_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(belt_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- belt_envfit$vectors$r
vec_scores$pval <- belt_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

belt_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")
```
```{r}
belt_vectors_sqrt <- ggplot() +
  geom_segment(data = belt_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = belt_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_category),
                  inherit.aes = FALSE,
                  size = 2.5,
                  box.padding = 0.1, 
                  min.segment.length = Inf,
                  nudge_x = -0.05,
                  nudge_y = -0.1)+
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.75, 0.85),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_x_continuous(limits = c(-1.15, 1.6))+
  scale_y_continuous(limits = c(-1.1, 0.85))


belt_vectors_sqrt

```

```{r}
library(patchwork)

# Create spacers
spacer <- plot_spacer()

# Bottom row: spacer | plot | spacer
bottom_centered <- spacer + belt_vectors_sqrt + spacer +
  plot_layout(ncol = 3, widths = c(0.9, 2.1, 1))  # adjust middle width

# Combine with top plot
final_plot <- belt_A_sqrt / bottom_centered +
  plot_layout(heights = c(2, 0.9))  # adjust vertical proportions

final_plot

ggsave(final_plot,
       file = "figures/Juvenile_NMDS.PDF",
       height = 220,
       width = 189,
       unit = "mm")

```


# 3. Benthic Cover

## 3.1. Composition across the shelf

###3.1.1 Updated to have detailed coral information, but grouped other categories

```{r}
# Benthic Data
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

PIT <- PIT %>%
  mutate(
    # Acroporidae = rowSums(select(., `Acropora - branching`:`Montipora`), na.rm = TRUE),
    # 
    # Pocilloporidae = rowSums(select(., Pocillopora:Stylophora), na.rm = TRUE),
    # 
    # Poritidae = rowSums(select(., `Porites - branching`:Goniopora), na.rm = TRUE),
    
    `Other Hard coral` = rowSums(dplyr::select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE),
    
    `Soft coral` = rowSums(dplyr::select(., Lobophytum:`Soft-other`), na.rm = TRUE),
    `Phaeophyceae` = rowSums(dplyr::select(., Sargassum:Lobophora), na.rm = TRUE),
    `Rhodophyta` = rowSums(dplyr::select(., Asparagopsis:Jania), na.rm = TRUE),
    `Chlorophyta`= rowSums(dplyr::select(., Halimeda:Valonia), na.rm = TRUE),
    `Sand and Rubble` = rowSums(dplyr::select(., Sand:Rubble), na.rm = TRUE),
    Turf = rowSums(dplyr::select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE),
    Other = rowSums(dplyr::select(., Sponge:Clam), na.rm = TRUE),
    Zoantharian = rowSums(dplyr::select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  dplyr::select(Year, Shelf, Reef, Site, Transect, `Acropora - branching`:Goniopora, `Other Hard coral`:Other, CCA, Cyanobacteria, Halophila)

PIT <- PIT |> 
  mutate(
    sum = rowSums(across(`Acropora - branching`:Halophila), na.rm = TRUE)
  )

PIT <- PIT |> 
  mutate(
    across(`Acropora - branching`:Halophila, ~ .x * 2)
  )
# Step 1: Create site-by-species matrix with Site_ID
site_by_PIT <- PIT %>%
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, 6:29
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  dplyr::select(Shelf, Year, Reef, Site)

# Step 3: Species matrix
PIT_matrix_filtered <- site_by_PIT %>%
  dplyr::select(
    `Acropora - branching`:Halophila
  )

PIT_matrix_filtered_sqrt <- sqrt(PIT_matrix_filtered)
```

```{r}
set.seed(999)
nmds <- metaMDS(PIT_matrix_filtered_sqrt, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress #0.196

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, metadata_PIT[, c("Year", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year == "2024-25" ~ "24/25",
      Year == "2023-24" ~ "23/24",
      Year == "2022-23" ~ "22/23",
      Year == "2021-22" ~ "21/22")
  )
```

```{r}
library(MASS)  # for mvrnorm
library(vegan)
grp <- interaction(scores$Shelf, scores$Year, drop = TRUE)
# Function to generate ellipse points for ggplot
ellipse_points <- function(ell_obj, npoints = 100) {
  do.call(rbind,
    lapply(names(ell_obj), function(g) {
      center <- ell_obj[[g]]$center      # centroid
      cov_mat <- ell_obj[[g]]$cov        # covariance matrix
      angles <- seq(0, 2*pi, length.out = npoints)
      circle <- cbind(cos(angles), sin(angles))
      # scale unit circle by sqrt of covariance (1 SD)
      ellipse <- t(center + t(circle %*% chol(cov_mat)))
      data.frame(NMDS1 = ellipse[,1],
                 NMDS2 = ellipse[,2],
                 Shelf = g)
    })
  )
}

ell <- ordiellipse(nmds, groups = grp, kind = "sd", draw = "none")
ell_df <- ellipse_points(ell, npoints = 100)

ell_df <- ell_df |>
  tidyr::separate(Shelf, into = c("Shelf", "Year"), sep = "\\.")
```


```{r}
summary(scores$NMDS1)
summary(scores$NMDS2)
Benthic_A <- ggplot(scores, aes(x = NMDS1, y = NMDS2,                          
                                colour = Shelf)) +
     geom_point(size = 1, alpha = 0.8) +
  geom_path(data = ell_df,
            aes(x = NMDS1, y = NMDS2, colour = Shelf, group = Shelf),
            linewidth = 0.6) +
    # stat_ellipse(aes(group = Shelf, colour = Shelf, groups = Shelf,),
    #              level = 0.75, linewidth = 0.5) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
   legend.position = "none",  
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
)+
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 1.05,     # right limit
         y = -1,       # bottom limit
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1.05, # push slightly inward
         vjust = -0.5, # lift slightly above bottom
         size = 3)+
  scale_x_continuous(limits = c(-1.3, 1.05))+
  scale_y_continuous(limits = c(-1.1, 1.1))+
  facet_wrap(~Year)
Benthic_A
```


```{r}
Benthic_envfit <- envfit(
  nmds,
  PIT_matrix_filtered_sqrt,
  permutations = 999
) 

# Extract p-values
pvals <- Benthic_envfit$vectors$pvals

# Extract r-squared values
r2 <- Benthic_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(Benthic_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- Benthic_envfit$vectors$r
vec_scores$pval <- Benthic_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

vec_scores <- vec_scores |> 
  filter(pval < 0.05, r2 > 0.25) # filter to only show vectors that are significant drivers.

Benthic_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")

```

```{r}
# Benthic_envfit$NMDS1_lab <- Benthic_envfit$NMDS1 * 1.1
# 
# Benthic_envfit$NMDS2_lab <- Benthic_envfit$NMDS2 * 1.1

Benthic_B <- ggplot()+
  geom_segment(data = Benthic_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Benthic_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_category),
                  inherit.aes = FALSE,
                   size = 2.5,
    max.overlaps = Inf,
    force = 2,
    max.iter = 20000,
    box.padding = 0.3,
    point.padding = 0.2)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year"
     ) +
  scale_x_continuous(limits = c(-1.3, 1.05))+
  scale_y_continuous(limits = c(-1.1, 1.1))
Benthic_B
```
```{r}
library(patchwork)

# Create spacers
spacer <- plot_spacer()

# Bottom row: spacer | plot | spacer
bottom_centered <- spacer + Benthic_B + spacer +
  plot_layout(ncol = 3, widths = c(0.85, 2.2, 0.95))  # adjust middle width

# Combine with top plot
final_plot <- Benthic_A / bottom_centered +
  plot_layout(heights = c(2, 1.05))  # adjust vertical proportions

final_plot

ggsave(final_plot,
       file = "figures/PIT_NMDS.PDF",
       height = 220,
       width = 189,
       unit = "mm")

```

### 3.1.2. Family of corals
```{r}
# Benthic Data
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

PIT <- PIT %>%
  mutate(
    Acroporidae = rowSums(dplyr::select(., `Acropora - branching`:`Montipora`), na.rm = TRUE),
    Pocilloporidae = rowSums(dplyr::select(., Pocillopora:Stylophora), na.rm = TRUE),
     Poritidae = rowSums(dplyr::select(., `Porites - branching`:Goniopora), na.rm = TRUE),
    `Other Hard coral` = rowSums(dplyr::select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE),
    `Soft coral` = rowSums(dplyr::select(., Lobophytum:`Soft-other`), na.rm = TRUE),
    `Phaeophyceae` = rowSums(dplyr::select(., Sargassum:Lobophora), na.rm = TRUE),
    `Rhodophyta` = rowSums(dplyr::select(., Asparagopsis:Jania), na.rm = TRUE),
    `Chlorophyta`= rowSums(dplyr::select(., Halimeda:Valonia), na.rm = TRUE),
    `Sand and Rubble` = rowSums(dplyr::select(., Sand:Rubble), na.rm = TRUE),
    Turf = rowSums(dplyr::select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE),
    Other = rowSums(dplyr::select(., Sponge:Clam), na.rm = TRUE),
    Zoantharian = rowSums(dplyr::select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  dplyr::select(Year, Shelf, Reef, Site, Transect, Acroporidae:Other, CCA, Cyanobacteria, Halophila)

# PIT <- PIT |> 
#   mutate(
#     sum = rowSums(across(`Acropora - branching`:Halophila), na.rm = TRUE)
#   )

PIT <- PIT |> 
  mutate(
    across(Acroporidae:Halophila, ~ .x * 2)
  )
# Step 1: Create site-by-species matrix with Site_ID
site_by_PIT <- PIT %>%
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, Acroporidae:Halophila
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  dplyr::select(Shelf, Year, Reef, Site)

# Step 3: Species matrix
PIT_matrix_filtered <- site_by_PIT %>%
  dplyr::select(
    Acroporidae:Halophila
  )

#write.csv(PIT_matrix_filtered, file = "Matrices_for_PRIMER/Family_PIT.csv")

PIT_matrix_filtered_sqrt <- sqrt(PIT_matrix_filtered)
```

```{r}
set.seed(999)
nmds <- metaMDS(PIT_matrix_filtered_sqrt, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress #0.200

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, metadata_PIT[, c("Year", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year == "2024-25" ~ "24/25",
      Year == "2023-24" ~ "23/24",
      Year == "2022-23" ~ "22/23",
      Year == "2021-22" ~ "21/22")
  )
```

```{r}
library(MASS)  # for mvrnorm
library(vegan)
grp <- interaction(scores$Shelf, scores$Year, drop = TRUE)
# Function to generate ellipse points for ggplot
ellipse_points <- function(ell_obj, npoints = 100) {
  do.call(rbind,
    lapply(names(ell_obj), function(g) {
      center <- ell_obj[[g]]$center      # centroid
      cov_mat <- ell_obj[[g]]$cov        # covariance matrix
      angles <- seq(0, 2*pi, length.out = npoints)
      circle <- cbind(cos(angles), sin(angles))
      # scale unit circle by sqrt of covariance (1 SD)
      ellipse <- t(center + t(circle %*% chol(cov_mat)))
      data.frame(NMDS1 = ellipse[,1],
                 NMDS2 = ellipse[,2],
                 Shelf = g)
    })
  )
}

ell <- ordiellipse(nmds, groups = grp, kind = "sd", draw = "none")
ell_df <- ellipse_points(ell, npoints = 100)

ell_df <- ell_df |>
  tidyr::separate(Shelf, into = c("Shelf", "Year"), sep = "\\.")
```


```{r}
summary(scores$NMDS1)
summary(scores$NMDS2)
Benthic_A <- ggplot(scores, aes(x = NMDS1, y = NMDS2,                          
                                colour = Shelf)) +
     geom_point(size = 1, alpha = 0.8) +
  geom_path(data = ell_df,
            aes(x = NMDS1, y = NMDS2, colour = Shelf, group = Shelf),
            linewidth = 0.6) +
    # stat_ellipse(aes(group = Shelf, colour = Shelf, groups = Shelf,),
    #              level = 0.75, linewidth = 0.5) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
   legend.position = "bottom",  
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
)+
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 1.05,     # right limit
         y = -1,       # bottom limit
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1.05, # push slightly inward
         vjust = -0.5, # lift slightly above bottom
         size = 3)+
  scale_x_continuous(limits = c(-1.3, 0.95))+
  scale_y_continuous(limits = c(-1.1, 1.1))+
  facet_wrap(~Year)
Benthic_A
```


```{r}
Benthic_envfit <- envfit(
  nmds,
  PIT_matrix_filtered_sqrt,
  permutations = 999
) 

# Extract p-values
pvals <- Benthic_envfit$vectors$pvals

# Extract r-squared values
r2 <- Benthic_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(Benthic_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- Benthic_envfit$vectors$r
vec_scores$pval <- Benthic_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

vec_scores <- vec_scores |> 
  filter(pval < 0.05, r2 > 0.2) # filter to only show vectors that are significant drivers.

Benthic_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")

#Other, Zoantharian, Poritidae...
```

```{r}
Benthic_B <- ggplot()+
  geom_segment(data = Benthic_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Benthic_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_category),
                  inherit.aes = FALSE,
                   size = 2.5,
    max.overlaps = Inf,
    force = 2,
    max.iter = 20000,
    box.padding = 0.3,
    point.padding = 0.2)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year"
     ) +
  scale_x_continuous(limits = c(-1.3, 0.95))+
  scale_y_continuous(limits = c(-1.1, 1.1))
Benthic_B
```

```{r}
library(patchwork)

# Create spacers
spacer <- plot_spacer()

# Bottom row: spacer | plot | spacer
bottom_centered <- spacer + Benthic_B + spacer +
  plot_layout(ncol = 3, widths = c(0.9, 2.1, 1))  # adjust middle width

# Combine with top plot
final_plot <- Benthic_A / bottom_centered +
  plot_layout(heights = c(2, 0.9))  # adjust vertical proportions

final_plot

ggsave(final_plot,
       file = "figures/PIT_NMDS_family.PDF",
       height = 220,
       width = 189,
       unit = "mm")

```

## 3.2. Brood stock
```{r}
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

PIT <- PIT |> 
  mutate(
    across(`Acropora - branching`:Halophila, ~ .x * 2)
  )

PIT <- PIT %>%
  mutate(
    Acroporidae = rowSums(dplyr::select(., `Acropora - branching`:`Montipora`), na.rm = TRUE),
    Pocilloporidae = rowSums(dplyr::select(., Pocillopora:Stylophora), na.rm = TRUE),
    Poritidae = rowSums(dplyr::select(., `Porites - branching`:Goniopora), na.rm = TRUE),
    `Other Hard coral` = rowSums(dplyr::select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE) #,
    
    # `Soft coral` = rowSums(dplyr::select(., Lobophytum:`Soft-other`), na.rm = TRUE),
    # `Phaeophyceae` = rowSums(dplyr::select(., Sargassum:Lobophora), na.rm = TRUE),
    # `Rhodophyta` = rowSums(dplyr::select(., Asparagopsis:Jania), na.rm = TRUE),
    # `Chlorophyta`= rowSums(dplyr::select(., Halimeda:Valonia), na.rm = TRUE),
    # `Sand and Rubble` = rowSums(dplyr::select(., Sand:Rubble), na.rm = TRUE),
    # Turf = rowSums(dplyr::select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE),
    # Other = rowSums(dplyr::select(., Sponge:Clam), na.rm = TRUE),
    # Zoantharian = rowSums(dplyr::select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )
```


###3.2.1. Acroporidae
```{r}
nbin_acro_adult_int <- glmmTMB(Acroporidae ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)

acro_adult_site <- PIT |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Acroporidae),
    se_site = sd(Acroporidae)/(n()^0.5))
  
acro_adult_emm <-  nbin_acro_adult_int |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


acro_adult_plot <-  ggplot()+
   geom_point(data =acro_adult_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1)+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = acro_adult_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
  labs(
   y = expression("Predicted cover (%)"),
   title = "Acroporidae",
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "none",
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
acro_adult_plot
```

```{r eval = FALSE}
acro_st_juv_cc <- ggarrange(B, juv_acro_plot, acro_adult_plot, ncol = 1, labels = c("A", "B", "C"), common.legend = TRUE)
# Total_st_juv
# 
ggsave(acro_st_juv_cc,
       filename = "acrp_ST_juv.PDF",
       height = 200,
       width = 189,
       unit = "mm")
```

###3.2.2. Pocilloporidae

```{r}
nbin_poci_adult_int <- glmmTMB(Pocilloporidae ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)

poci_adult_site <- PIT |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Pocilloporidae),
    se_site = sd(Pocilloporidae)/(n()^0.5))
  
poci_adult_emm <-  nbin_poci_adult_int |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()


poci_adult_plot <-  ggplot()+
   geom_point(data =poci_adult_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1)+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = poci_adult_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
  labs(
   y = expression("Predicted cover (%)"),
   title = "Pocilloporidae",
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "none",
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
poci_adult_plot
```

```{r eval = FALSE}
Poci_st_juv_cc <- ggarrange(C, juv_poci_plot, poci_adult_plot, ncol = 1, labels = c("A", "B", "C"), common.legend = TRUE)
# Total_st_juv
# 
ggsave(Poci_st_juv_cc,
       filename = "Poci_ST_juv.PDF",
       height = 200,
       width = 189,
       unit = "mm")
```

###3.2.3. Poritidae
```{r}
# nbin_por_adult_int <- glmmTMB(Poritidae ~ Shelf * Year +(1 | Reef/Site), 
#         family = "nbinom2", data = PIT)
nbin_por_adult <- glmmTMB(Poritidae ~ Shelf +  Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)

por_adult_site <- PIT |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Poritidae),
    se_site = sd(Poritidae)/(n()^0.5))
  
# por_adult_emm <-  nbin_por_adult_int |> 
#   emmeans(~Shelf | Year, type = "response") |> 
#   as.data.frame()

por_adult_emm <-  nbin_por_adult |>  #using the fixed term model because AICc was lower
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()

por_adult_plot <-  ggplot()+
   geom_point(data =por_adult_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.85),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1)+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = por_adult_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
  labs(
   y = expression("Predicted cover (%)"),
   title = "Poritidae",
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "none",
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
por_adult_plot

```

```{r eval = FALSE}
Por_st_juv_cc <- ggarrange(D, juv_por_plot, por_adult_plot, ncol = 1, labels = c("A", "B", "C"), common.legend = TRUE)
# Total_st_juv
# 
ggsave(Por_st_juv_cc,
       filename = "Por_ST_juv.PDF",
       height = 200,
       width = 189,
       unit = "mm")
```



###3.2.4. Total hard coral cover
```{r}
PIT <- PIT |> 
  mutate(
    Total_HC_Cover = Acroporidae+ Pocilloporidae + Poritidae + `Other Hard coral`
  )

nbin_HC_adult_int <- glmmTMB(Total_HC_Cover ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)

HC_adult_site <- PIT |> 
  group_by(Shelf, Site, Year, Reef) |> 
  summarise(
    mean_site = mean(Total_HC_Cover),
    se_site = sd(Total_HC_Cover)/(n()^0.5))
  
HC_adult_emm <-  nbin_HC_adult_int |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()

HC_adult_plot <-  ggplot()+
   geom_point(data =HC_adult_site, aes(x=Year, y=mean_site, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
    size = 1.5,
    alpha = 0.75,
    colour = "#606060", shape = 1)+
 # geom_point(data = total_juv_emm, aes(x=Year,
 #             y=response, 
 #             color=Shelf))+
  geom_pointrange(data = HC_adult_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
  labs(
   y = expression("Predicted cover (%)"),
   title = "Total Coral Cover",
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "none",
        text = element_text(size = 12))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
HC_adult_plot
```

```{r eval = FALSE}
Total_st_juv_cc <- ggarrange(A, juv_total_plot, HC_adult_plot, ncol = 1, labels = c("A", "B", "C"), common.legend = TRUE)
# Total_st_juv
# 
ggsave(Total_st_juv_cc,
       filename = "Total_ST_juv.PDF",
       height = 200,
       width = 189,
       unit = "mm")
```

```{r}
ggarrange(HC_adult_plot, acro_adult_plot, poci_adult_plot, por_adult_plot, labels = c("A", "B", "C", "D"), ncol = 1, nrow = 4)
```

```{r eval=FALSE}
#ggarrange(HC_adult_plot, acro_adult_plot, poci_adult_plot, por_adult_plot, labels = c("A", "B", "C", "D"), ncol = 2, nrow = 2)
# 
# ggsave(filename = "Figure6.PNG",
#        height = 200,
#        width = 275,
#        unit = "mm")

ggarrange(HC_adult_plot, acro_adult_plot, poci_adult_plot, por_adult_plot, labels = c("A", "B", "C", "D"), ncol = 1, nrow = 4)

ggsave(filename = "figures/Figure6.PDF",
       width = 180,
      height = 275,
       unit = "mm")
```

### Presentation 

```{r}
ggsave(HC_adult_plot, filename = "Presentation_fig/Total_cover.PNG",
         width = 180,
         height = 120,
         units = "mm"
        )

Families_cover <- ggarrange(acro_adult_plot, poci_adult_plot, por_adult_plot, ncol =3, nrow = 1, common.legend = TRUE, legend = "bottom")


ggsave(Families_cover, filename = "Presentation_fig/Families_cover.PNG",
         width = 325,
         height = 120,
         units = "mm"
        )

Families_cover <- ggarrange(acro_adult_plot, poci_adult_plot, por_adult_plot, ncol = 1, nrow = 3, common.legend = TRUE, legend = "bottom")


ggsave(Families_cover, filename = "Presentation_fig/Families_cover_portrait.PNG",
         width = 180,
         height = 200,
         units = "mm"
        )
```

## 3.4. Brood Stock and Recruits
Brood stock is estimated from PITs, and Recruits estimated from settlement tiles.
*Not sure this is worth making figures for, given that all models did not find a significant effect of brood stock*

```{r}
# Benthic Data
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    )  )


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))


PIT <- PIT %>%
  mutate(
    Acroporidae = rowSums(dplyr::select(., `Acropora - branching`:`Montipora`), na.rm = TRUE), #grouping Acroporidae
    Pocilloporidae = rowSums(dplyr::select(., Pocillopora:Stylophora), na.rm = TRUE), #grouping Pocilloporidae genera
    Poritidae = rowSums(dplyr::select(., `Porites - branching`:Goniopora), na.rm = TRUE), #grouping Poritidae genera
    `Other Hard coral` = rowSums(dplyr::select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE), # Grouping other hard corals
    `Soft coral` = rowSums(dplyr::select(., Lobophytum:`Soft-other`), na.rm = TRUE), #soft coral categories
    `Brown MA` = rowSums(dplyr::select(., Sargassum:Lobophora), na.rm = TRUE), #brown macroalgae categories
    `Red MA` = rowSums(dplyr::select(., Asparagopsis:Jania), na.rm = TRUE), #red macroalgae
    `Green MA`= rowSums(dplyr::select(., Halimeda:Valonia), na.rm = TRUE), #green macroalgae
    `Sand and Rubble` = rowSums(dplyr::select(., Sand:Rubble), na.rm = TRUE), #sand, rubble, and combo
    Turf = rowSums(dplyr::select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE), #Turfing substrata (dead corals + EAM)
    Other = rowSums(dplyr::select(., Sponge:Clam), na.rm = TRUE), #other invertes
    Zoantharian = rowSums(dplyr::select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  dplyr::select(Year, Shelf, Reef, Site, Transect, Acroporidae:Zoantharian,CCA, Cyanobacteria, Halophila)

# PIT <- PIT |> 
#   mutate(
#     sum = rowSums(across(Acroporidae:Halophila), na.rm = TRUE) 
#   )

#summary(PIT$sum) #check that the number of points = 50 per transect

PIT <- PIT |> 
  mutate(
    across(Acroporidae:Halophila, ~ .x * 2) #multiply by two to get out of 100%
  )


join_pit <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    PIT = n()
  )

st_join <- Coral_settlement_nGBR |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    acro_mean_st = mean(Acroporidae),
    poci_mean_st = mean(Pocilloporidae),
    por_mean_st = mean(Poritid),
    acro_sd_st = sd(Acroporidae),
    poci_sd_st = sd(Pocilloporidae),
    por_sd_st = sd(Poritid)
  )


combined_averages <- left_join(join_pit, st_join)
```

### 3.2.1. Acroporidae

```{r}
Acro <- combined_averages |> 
  filter(acro_mean_st > 0)
gamma_avg_Acroporidae_int <- glmmTMB(acro_mean_st ~ mean_aco*Shelf +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Acro)
gamma_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Reef/Site) + (1 | Year), 
         family = Gamma(link = "log"), data = Acro)
# rng <- range(Acro$mean_aco, na.rm = TRUE)
# 
# pred <- ggpredict(
#   gamma_avg_Acroporidae_int,
#   terms = c(sprintf("mean_aco [%f:%f]", rng[1], rng[2]), "Shelf")
# )

pred <- ggpredict(
  gamma_avg_Acroporidae_int,
  terms = c("mean_aco", "Shelf")
)

#produces a pred with 9 values of acroporidae cover, and represents the population-level prediction (random effects set to zero) is usually what you want for interpretation and plotting the relationship.

acro_benthic_st <- ggplot(pred, aes(x = x, y = predicted, colour = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, colour = NA) +
 labs(x= expression("Observed cover (%)"),
       y = expression("Predicted recruits ( " * tile^{-1} * ")"),
       title = expression(atop("Acroporidae", "gamma (log-link) average % cover x shelf" )),
       colour = "Shelf",
       fill = "Shelf") +
  facet_wrap(~group, scale = "free_y")+
  theme_classic()+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_fill_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

  

acro_benthic_st
```

### 3.2.2 Pocilloporidae

```{r}
Poci <- combined_averages |> 
  filter(poci_mean_st > 0)

ln_avg_poci_fixed <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Reef/Site), 
        family = "lognormal", data = Poci)

pred <- ggpredict(
  ln_avg_poci_fixed,
  terms = c("mean_poci")
)

#produces a pred with 9 values of pocilloporidae cover, and represents the population-level prediction (random effects set to zero) is usually what you want for interpretation and plotting the relationship.

poci_benthic_st <- ggplot(pred, aes(x = x, y = predicted, 
                                    #colour = group
                                    )) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, 
                  #fill = group
                  ),
              alpha = 0.2, colour = NA) +
  labs(x= expression("Observed cover (%)"),
       y = expression("Predicted recruits ( " * tile^{-1} * ")"),
       title = expression(atop("Pocilloporidae", "Log-normal"))) +
  theme_classic()

poci_benthic_st
```

### 3.2.3. Poritidae

```{r}
Por <- combined_averages |> 
  filter(por_mean_st > 0)
ln_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Reef/Site), 
        family = "lognormal", data = Por)

pred <- ggpredict(
  ln_avg_Poritid_fixed,
  terms = c("mean_por")
)

#produces a pred with 9 values of acroporidae cover, and represents the population-level prediction (random effects set to zero) is usually what you want for interpretation and plotting the relationship.

por_benthic_st <- ggplot(pred, aes(x = x, y = predicted, 
                                   #colour = group
                                   )) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, 
                  #fill = group
                  ),
              alpha = 0.2, 
              colour = NA) +
  labs(x= expression("Observed cover (%)"),
       y = expression("Predicted recruits ( " * tile^{-1} * ")"),
       title = expression(atop("Poritidae", "lognormal"))) +
  theme_classic()

por_benthic_st
```

```{r}
ggarrange(acro_benthic_st, poci_benthic_st, por_benthic_st, 
          labels = c("A", "B", "C"),
          ncol = 1,
          nrow = 3
          )

ggsave(filename = "figures/benthic_and_recruitment.PDF",
       width = 189,
       height = 220,
       unit = "mm")
```


# 4. Larval supply and Juveniles

## 4.1. Spatial Patterns

Load data for these models
```{r}
combine_belt_st <- left_join(join_belt, Join_Coral)
```

```{r}
combine_belt_st <- combine_belt_st |> 
  mutate(Year_start = as.numeric(substr(Year, 1, 4)))
```

```{r}
gropd_df2 <- combine_belt_st %>%
  arrange(Shelf, Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_mean_acro_lag1 = lag(st_mean_acro, 1),
    st_mean_poci_lag1 = lag(st_mean_poci, 1),
    st_mean_por_lag1 = lag(st_mean_por, 1),
  ) %>%
  ungroup()
gropd_df2 <- gropd_df2 |> 
  mutate(Shelf, factor(Shelf, levels = c("Inner", "Mid", "Outer"), ordered = TRUE))
```


### 4.1.1. Acroporidae

*note set maximum value for mid and outer shelf*

```{r}
#model
model_df_acro <- gropd_df2 %>%
  filter(!is.na(mean_acro))


ln_acro_juv_st_int <- glmmTMB(mean_acro ~ scale(st_mean_acro_lag1, scale = FALSE) * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)
summary(ln_acro_juv_st_int)
summary(model_df_acro$st_mean_acro_lag1)
emmeans(ln_acro_juv_st_int, ~Shelf, type = "response") #note for victor, pls help
regrid


pred <- ggpredict(
  ln_acro_juv_st_int,
  terms = c("st_mean_acro_lag1 [all]", "Shelf")
)
# pred <- ggpredict(
#   ln_acro_juv_st_int,
#   terms = c("st_mean_acro_lag1 [all]", "Shelf"),
#   type = "link"
# )

acro_juv_st <- ggplot(pred, aes(x = x, y = predicted, colour = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, colour = NA) +
  labs(x= expression("Recruits (" * tile^{-1} * ") (t)"),
       y = expression(atop("Predicted juvenile colonies ",  "(10"~m^{-2}*") (t + 1)")),
       colour = "Shelf",
       title = expression(atop("Acroporidae", "Log-normal average Acroporidae recruits x Shelf")),
       fill = "Shelf") +
  facet_wrap(~group, scale = "free_y")+
  theme_classic()+scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+scale_fill_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
  

acro_juv_st
```

```{r}
pred <- ggpredict(
  ln_acro_juv_st_int,
  terms = c("st_mean_acro_lag1 [all]", "Shelf")
)

inner_df <- pred %>% filter(group == "Inner")
mid_df   <- pred %>% filter(group == "Mid")
outer_df <- pred %>% filter(group == "Outer")

plot_fun <- function(df, xlim_vals = NULL, title_text = NULL) {
  p <- ggplot(df, aes(x = x, y = predicted, colour = group)) +
    geom_line(size = 1.2) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
                alpha = 0.2, colour = NA) +
    theme_classic() +
    labs(title = title_text) +
    scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F")) +
    scale_fill_manual(values = c("#A8C85A", "#9e379f", "#C97B5F")) +
    theme(legend.position = "none")  # avoid repeated legends
  
  if (!is.null(xlim_vals)) {
    p <- p + coord_cartesian(xlim = xlim_vals)
  }
  
  return(p)
}

model_df_acro |> filter(Shelf == "Outer") |> summarise(min = min(st_mean_acro), max = max(st_mean_acro))
model_df_acro |> filter(Shelf == "Mid") |> summarise(min = min(st_mean_acro), max = max(st_mean_acro))

p_inner <- plot_fun(inner_df, title_text = "Inner shelf")

p_mid   <- plot_fun(mid_df, xlim_vals = c(0, 100), title_text = "Mid shelf")

p_outer <- plot_fun(outer_df, xlim_vals = c(0, 20), ylim_vals = c(0, 100), title_text = "Outer shelf")

final_plot <- (p_inner | p_mid | p_outer) +
  plot_annotation(
    title = "Acroporidae: Log-normal recruits × Shelf",
    theme = theme(plot.title = element_text(hjust = 0.5))
  )

final_plot & 
  labs(
    x = expression("Recruits (" * tile^{-1} * ") (t)"),
    y = expression(atop("Predicted juvenile colonies ", "(10"~m^{-2}*") (t + 1)"))
  )
```


*Try to work out what to do with those whack outer shelf values*
### 4.1.2. Pocilloporidae

```{r}
model_df_poci <- gropd_df2 %>%
  filter(!is.na(mean_poci))

ln_poci_juv_st_fixed <- glmmTMB(mean_poci ~ scale(st_mean_poci_lag1, scale = FALSE) + Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)

pred <- ggpredict(
  ln_poci_juv_st_fixed,
  terms = c("st_mean_poci_lag1 [all]", "Shelf")
)

poci_juv_st <- ggplot(pred, aes(x = x, y = predicted, colour = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, colour = NA) +
  labs(x= expression("Recruits (" * tile^{-1} * ") (t)"),
       y = expression(atop("Predicted juvenile colonies", "(10"~m^{-2}*") (t + 1)")),
       colour = "Shelf",
       title = expression(atop("Pocilloporidae", "Log-normal average Pocilloporidae recruits + Shelf")),
       fill = "Shelf") +
  facet_wrap(~group)+
  theme_classic()+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_fill_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

poci_juv_st
```

### 4.1.3. Poritidae

```{r}
model_df_por <- gropd_df2 %>%
  filter(!is.na(mean_por))

model_df_por <- model_df_por |> 
  filter(mean_por > 0)

gamma_por_juv_st_fixed <- glmmTMB(mean_por ~ scale(st_mean_por_lag1, scale = FALSE) + Shelf + (1|Reef/Site),
              family = Gamma(link = "log"),
           data = model_df_por)

pred <- ggpredict(
  gamma_por_juv_st_fixed,
  terms = c("st_mean_por_lag1 [all]", "Shelf")
)

por_juv_st <- ggplot(pred, aes(x = x, y = predicted, colour = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, colour = NA) +
  labs(x= expression("Recruits (" * tile^{-1} * ") (t)"),
       y = expression(atop("Predicted juvenile colonies", " (10"~m^{-2}*") (t + 1)")),
       colour = "Shelf",
       title = expression(atop("Poritidae", "Gamma (log-link) average Poritidae recruits + Shelf")),,
       fill = "Shelf") +
  facet_wrap(~group)+
  theme_classic()+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_fill_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

por_juv_st
```

```{r eval=FALSE}
Juv_st_relationship <- ggarrange(acro_juv_st, poci_juv_st, por_juv_st, ncol =1, nrow =3, common.legend = TRUE,
          labels = c("A", "B", "C"))

ggsave(Juv_st_relationship,
       file = "figures/juv_st_relationship_fig.PDF",
       height = 240,
       width = 189,
       unit = "mm")
```


# 5. Composition Plot Combined

```{r}
# Make all nmds one column

NMDS_combined <- ggarrange(p1, belt_A, Benthic_A, ncol =3, labels = c("A", "C", "E"), common.legend = TRUE, legend = "bottom", widths = c(0.8, 0.8, 1))
NMDS_combined

Envfit_combined <- ggarrange(p2, belt_B, Benthic_B, ncol =3, labels = c("B", "D", "F"), widths = c(0.8, 0.8, 1))
Envfit_combined

Complete_Composition <- ggarrange(NMDS_combined, Envfit_combined, ncol = 1, heights = c(1.1,0.9))
Complete_Composition
```

```{r eval=FALSE}
ggsave(filename = "Complete_composition.PDF",
       plot = Complete_Composition,
       width = 250,
       height = 200,
       units = "mm")
```


```{r}
ST_NMDS <- ggarrange(p1, p2, ncol = 1, labels = c("A", "B"))

belt_Figure <- ggarrange(belt_A, belt_B, labels = c("C", "D"), ncol = 1)

Benthic_Figure <- ggarrange(
  Benthic_A, Benthic_B, labels = c("E", "F"), ncol = 1
)

# complete_composition <- ggarrange(ST_NMDS, belt_Figure, Benthic_Figure, widths = c(0.8, 0.8, 1), ncol = 3)
# complete_composition
```

# 6. Brood stock and juveniles

```{r}
# Benthic Data
PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    )  )


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

PIT <- PIT |> 
  mutate(
    Acroporidae_PIT = rowSums(across(`Acropora - branching`:Montipora), na.rm = TRUE),
    Pocilloporidae_PIT = rowSums(across(Pocillopora:Stylophora), na.rm = TRUE),
    Poritidae_PIT = rowSums(across(`Porites - branching`:Goniopora), na.rm = TRUE),
    `Other Hard coral` = rowSums(across(`Favia (plocoid)`:`other submassive`), na.rm = TRUE),
    `Soft coral` = rowSums(across(Lobophytum:`Soft-other`), na.rm = TRUE),
    `Brown MA` = rowSums(across(Sargassum:Lobophora), na.rm = TRUE),
    `Red MA` = rowSums(across(Asparagopsis:Jania), na.rm = TRUE),
    `Green MA` = rowSums(across(Halimeda:Valonia), na.rm = TRUE),
    `Sand and Rubble` = rowSums(across(Sand:Rubble), na.rm = TRUE),
    Turf = rowSums(across(`Pavement/EAM`:`Dead massive`), na.rm = TRUE),
    Other = rowSums(across(Sponge:Clam), na.rm = TRUE),
    Zoantharian = rowSums(across(`Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  mutate(
    across(Acroporidae_PIT:Poritidae_PIT, ~ .x * 2) #multiply by two to get out of 100%
  )

PIT_adult <- PIT |> select(Year, Shelf, Reef, Site, Transect, Acroporidae_PIT, Pocilloporidae_PIT, Poritidae_PIT)

```

```{r}
Belt_family <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Transect, Acroporidae, Pocilloporidae, Poritidae)

# belt_family_long <- Belt_family |>  pivot_longer(cols = Acroporidae:Poritidae, names_to = "Family", values_to = "juvenile_density")
# 
# belt_detailed_long <- belt_detailed_long |> 
#   group_by(Year, Shelf, Reef, Site, genus) |> 
#   summarise(
#     mean_juv = mean(juvenile_density),
#     sd_juv = sd(juvenile_density)
#   )
```

```{r}
combined_juv_cover <- left_join(PIT_adult, Belt_family)

combined_juv_cover <- combined_juv_cover |> 
  mutate(
    Shelf = factor(Shelf, levels = c("Inner", "Mid", "Outer"))
  )
```

Despite not being interested in across the shelf differences, as shelf has only three levels there are two few levels for a stable random effect


From this I would say we can only really do Acropora, Porites, and Pocillopora analysis because they have medians not equal to 0

## 6.1. Acropora

```{r}
nbin_avg_Acro_int <- glmmTMB(Acroporidae ~ Acroporidae_PIT * Shelf + (1|Reef/Site/Transect), 
         family = "nbinom2", data = combined_juv_cover)

pred <- ggpredict(
  nbin_avg_Acro_int,
  terms = c("Acroporidae_PIT", "Shelf")
)

#produces a pred with 9 values of acroporidae cover, and represents the population-level prediction (random effects set to zero) is usually what you want for interpretation and plotting the relationship.

acro_benthic_juv <- ggplot(pred, aes(x = x, y = predicted, colour = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, colour = NA) +
 labs(x= expression("Observed cover (%)"),
       y = expression("Predicted juvenile colonies (10 "~m^{-2}*")"),
       title = expression(atop("Acroporidae", "nbin % cover x shelf" )),
       colour = "Shelf",
       fill = "Shelf") +
  facet_wrap(~group)+
  theme_classic()+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_fill_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

acro_benthic_juv
```


## 6.2. Pocillopora

```{r}
nbin_avg_Poci_fixed2 <- glmmTMB(Pocilloporidae ~ Pocilloporidae_PIT + Shelf + (1|Reef/Site/Transect), 
         family = "nbinom2", data = combined_juv_cover)

pred <- ggpredict(
  nbin_avg_Poci_fixed2,
  terms = c("Pocilloporidae_PIT", "Shelf")
)


#produces a pred with 9 values of acroporidae cover, and represents the population-level prediction (random effects set to zero) is usually what you want for interpretation and plotting the relationship.

poci_benthic_juv <- ggplot(pred, aes(x = x, y = predicted, colour = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, colour = NA) +
 labs(x= expression("Observed cover (%)"),
       y = expression("Predicted juvenile colonies (10 "~m^{-2}*")"),
       title = expression(atop("Pocilloporidae", "nbin % cover x shelf" )),
       colour = "Shelf",
       fill = "Shelf") +
  facet_wrap(~group)+
  theme_classic()+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_fill_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

poci_benthic_juv

```


## 6.3. Poritidae

```{r}
nbin_avg_Por_int <- glmmTMB(Poritidae ~ Poritidae_PIT * Shelf + (1|Site), 
         family = "nbinom2", data = combined_juv_cover)

pred <- ggpredict(
  nbin_avg_Por_int,
  terms = c("Poritidae_PIT", "Shelf")
)

#produces a pred with 9 values of acroporidae cover, and represents the population-level prediction (random effects set to zero) is usually what you want for interpretation and plotting the relationship.

por_benthic_juv <- ggplot(pred, aes(x = x, y = predicted, colour = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, colour = NA) +
 labs(x= expression("Observed cover (%)"),
       y = expression("Predicted juvenile colonies (10 "~m^{-2}*")"),
       title = expression(atop("Poritidae", "nbin % cover x shelf" )),
       colour = "Shelf",
       fill = "Shelf") +
  facet_wrap(~group)+
  theme_classic()+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_fill_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

por_benthic_juv
```


```{r}
benthic_juv<- ggarrange(acro_benthic_juv, poci_benthic_juv, por_benthic_juv, ncol = 1,
          common.legend = TRUE,
          labels = c("A", "B", "C"))

ggsave(filename = "benthic_juvenile.PDF",
  plot = benthic_juv,
       height = 240, width = 189, unit = "mm")
```

