---
title: "Composition of Vegetated Tiles"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
#library(lme4)
#library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(data.table) #not sure if required at this point
library(plotly) #interactive nmds

#import and clean data
Coral_settlement_nGBR_CS <- read_csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(`Shelf posn` != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR |> 
  filter(notes != "NA")

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= `Shelf posn`,
         "Year_Collected"= `Year (Jan)`,
         "Site_#"= `Site #`,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)
Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  )) |> 
  filter(notes != "Dislodged")

# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))

PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_macroalgae = mean(Macroalgae),
    sd_macroalgae = sd(Macroalgae),
    mean_sand =mean(Rubble_and_Sand),
    sd_sand = sd(Rubble_and_Sand),
    mean_turf = mean(Turf),
    sd_turf = sd(Turf),
    PIT = n()
  )

Combined <- left_join(Coral_settlement_nGBR, join)
```

```{r}
Recruits <- Coral_settlement_nGBR %>%
  filter(total >0) |>  #ensure only vegetated tiles included
  select(Acroporidae, Pocilloporidae, Poritid, Other)


recruit_hellinger <- decostand(Recruits, method = "hellinger")

environment_data <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  select(Year_Collected, Site, Reef, Shelf) #worried because environment data is nested here.
```

#PERMANOVA

As the permdisp was better for the sqrt transformation, I chose to do that transformation instead.

Here we are analysing the hellinger recruit matrix as a function of year and shelf, and their interaction
```{r}
# Bray–Curtis PERMANOVA
permanova_result <- adonis2(recruit_hellinger ~ Year_Collected * Shelf, data = environment_data, method = "bray", by = "term", permutations = 999) 

print(permanova_result)
```
What this output suggests to me is that Shelf has the largest F value, so is the main driver. Year collected has a significant effect on community composition, but it accounts for a relatively small (0.022 * 100)% of variation in recruitment.
- Shelf drives 37% of the variation in composition (statistically significant p<0.001)
- Year drives 2.2% of the observed variation in composition.
- Residual is 56.6% of the variation in composition - so perhaps need to explore how to improve this.
does have somewhat of an effect, but its much smaller.
Residual is 0.566

##Extract the Pseudo-F 

```{r}
psuedo_f <- permanova_result$F[1]
psuedo_f
```

# reorder because the DF is different, and that means you need to do ADONis twice

```{r}
permanova_result <- adonis2(recruit_hellinger ~ Shelf * Year_Collected, data = environment_data, method = "bray", by = "term", permutations = 999) 

print(permanova_result)
```


##Extract the Pseudo-F 

```{r}
psuedo_f <- permanova_result$F[1]
psuedo_f
```

### Tried to add benthic community to see if that would imrpove how much variation was explained by the model.

```{r}
# environment_data2 <- Combined %>%
#   filter(total >0) |> 
#   select(Year_Collected, Site, Reef, Shelf, mean_turf, mean_macroalgae, mean_sand) 
```

```{r}
# # Bray–Curtis PERMANOVA
# permanova_result2 <- adonis2(recruit_hellinger ~ Year_Collected * Shelf *mean_turf * mean_macroalgae, data = environment_data2, method = "bray", by = "term", permutations = 999) 
# 
# print(permanova_result2)
# When I ran this they were significant, but hard to ignore that the cover is related to the shelf/site, so not sure this is appropriate
```

Run with Site as well, to see if that is explaining some of the variation
```{r}
permanova_result3 <- adonis2(recruit_hellinger ~ Year_Collected * Shelf/Site, data = environment_data, method = "bray", by = "term", permutations = 999) 
print(permanova_result3)
```

Can see here that site is pretty important driver here in the composition. Not sure this is the correct way to nest though, or what to do about this.
May also mean I need to include site as the shape or something in the NMDS.

```{r}
 #Nested PERMANOVA with site as strata
n_p <- adonis2(recruit_hellinger ~ Shelf/Site *Year_Collected, data = environment_data, method = "bray",  strata = environment_data$Site, permutations = 999 ) 

print(n_p)
```

Issue with this is that I don't know how to get the individual predictors in the model, I can just see that the model is significant. Doesn't particularly help me writing my results.
```{r}
 #Nested PERMANOVA with shelf as strata
n_p <- adonis2(recruit_hellinger ~ Shelf/Site *Year_Collected, data = environment_data, method = "bray",  strata = environment_data$Shelf, permutations = 999 ) 

print(n_p)
```

The key assumption of PERMANOVA is that the dispersion among levels is homogenous.
I will test this using betadisp
```{r}
bray_hellinger <- vegdist(recruit_hellinger, method = "bray")

# With respect to Shelf
betadisper(bray_hellinger, environment_data$Shelf)
anova(betadisper(bray_hellinger, environment_data$Shelf))

# with respect to year
betadisper(bray_hellinger, environment_data$Year_Collected)
anova(betadisper(bray_hellinger, environment_data$Year_Collected))
```

Well that isn't ideal. The significant p value indicates that dispersion is significantly different among levels. Thus, we can't determine whether the significant effect on composition (i.e., the significant pvalue in the PERMANOVA ) is driven by differences in dispersion or differences among the groups.

As the permdisp was better for the sqrt transformation, I chose to do that transformation instead.

### SQRT

```{r}
# Filtered recruit data
Recruits_sqrt <- sqrt(Recruits)
```

Run PERMDISP
```{r}
bray <- vegdist(Recruits_sqrt, method = "bray")
# With respect to Shelf
betadisper(bray, environment_data$Shelf)
anova(betadisper(bray, environment_data$Shelf))

# with respect to year
betadisper(bray, environment_data$Year_Collected)
anova(betadisper(bray, environment_data$Year_Collected))
```


```{r}
# Bray–Curtis PERMANOVA
adonis2(Recruits_sqrt ~ Year_Collected * Shelf, data = environment_data, method = "bray")

#Nested PERMANOVA
adonis2(Recruits_sqrt ~ Year_Collected * Shelf, data = environment_data, method = "bray",  strata = environment_data$Reef) |> 
print()

adonis2(Recruits_sqrt~ Year_Collected * Shelf/Site, data = environment_data, method = "bray", by = "term", permutations = 999,
        strata = environment_data$Site) |> print()
```

# NMDs
```{r}
set.seed(999)     # for reproducibility

nmds <- metaMDS(Recruits_sqrt, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, environment_data[, c("Year_Collected", "Shelf", "Reef", "Site")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2022" ~ "21/22")
  )
```

```{r}
# Decide limits of x and y axis, so that both plots will be the same
scores |> 
  summarise(
    x_min = min(NMDS1),
    x_max = max(NMDS1),
    y_min = min(NMDS2),
    y_max = max(NMDS2)
  )
```


```{r}
# Creating Vectors for driving the patterns
Recruits_envfit <- envfit(
  nmds,
  Recruits_sqrt,
  permutations = 999
) 

#Saving the vectors as a data frame
Recruits_envfit <- as.data.frame(scores(Recruits_envfit, display = "vectors"))

Recruits_envfit <- Recruits_envfit |> 
  rownames_to_column(var = "label_family")
```


```{r}
set.seed(999)     # for reproducibility

nmds_hellinger <- metaMDS(recruit_hellinger, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds_hellinger$stress

# Extract NMDS for plot
site_scores_hellinger <- as.data.frame(vegan::scores(nmds_hellinger, display = "sites"))
scores_hellinger <- cbind(site_scores_hellinger, environment_data[, c("Year_Collected", "Shelf", "Reef", "Site")])
```

```{r}
Figure5A1 <-  ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 1.5, alpha = 0.8) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "top",
         text = element_text(size = 8),
        title = element_text(size = 8)) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
             x = min(scores$NMDS1),
             y = min(scores$NMDS2),              
             label = paste0("Stress = ", round(nmds$stress, 3)),
              hjust = -2, vjust = 0.5, size = 3)+
   facet_wrap(~ Deployment, ncol = 2)+
  scale_x_continuous(limits = c(-1, 1.45))+
  scale_y_continuous(limits = c(-1.3, 1.1))
Figure5A1
```


```{r}
Fig5.B <- ggplot()+
  geom_segment(data = Recruits_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Recruits_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_family),
                  inherit.aes = FALSE,
                  size = 3,
                  box.padding = 0.5, 
                  min.segment.length = Inf,
                  nudge_x = -0.1,
                  nudge_y = 0.1)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year Collected"
    ) +
  scale_x_continuous(limits = c(-1, 1.45))+
  scale_y_continuous(limits = c(-1.3, 1.1))
Fig5.B
```


```{r}
Figure5 <- ggpubr::ggarrange(Figure5A1, Fig5.B, ncol =1, labels = c("A", "B"), heights = c(1.2, 0.5))
Figure5


ggsave(filename = "Fig5.png",
       plot = Figure5,
       height = 250,
       width = 169,
       unit = "mm")
```

