---
title: "04_Tables"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)


# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  ),
  Deployment = case_when(
      Year_Collected == "2022" ~ "21/22",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2026" ~ "25/26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")

Pocilloporidae <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Pocilloporidae, Top_Pocil, 
         Under_Pocil, Side_Pocil, total)

Pocilloporidae_orientation <- Pocilloporidae %>% 
  rename(Top = Top_Pocil,
                   Bottom = Under_Pocil,
                   Side = Side_Pocil,
         Recruits_per_tile = total)

Pocilloporidae_orientation <- Pocilloporidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Acroporidae <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Acroporidae, Top_Acr, 
         Under_Acr, Side_Acr, total)

Acroporidae_orientation <- Acroporidae %>% 
  rename(Top = Top_Acr,
                   Bottom = Under_Acr,
                   Side = Side_Acr,
         Recruits_per_tile = total)

Acroporidae_orientation <- Acroporidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Poritid <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Poritid, Top_Por, Under_Por, 
         Side_Por, total)

Poritidae_orientation <- Poritid %>% 
  rename(Top = Top_Por,
                   Bottom = Under_Por,
                   Side = Side_Por,
         Recruits_per_tile = total) 

Poritidae_orientation <- Poritidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Other <- Coral_settlement_nGBR %>% 
  select(Year, Shelf, 
         Reef, Site, Tile, 
         Other, Top_other, 
         Under_oth, Side_other, total)

# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    PIT = n()
  )

Combined <- left_join(Coral_settlement_nGBR, join)

# Juvenile data
Belt_10m <- read_csv("Belt_10m.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))

## Total recuits
total_recruits <- Coral_settlement_nGBR %>%
  select(Year_Collected, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 5:8, names_to = "Family", values_to = "Total")

#Order so that Other is last in Family
total_recruits <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, Reef, Site, Obs, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 6:9, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))

total_recruits$Year_Collected <- as.factor(total_recruits$Year_Collected)

## Position on tile

position_total_recruits <- Coral_settlement_nGBR %>%
  select(2:9, 15:29, 32) %>% 
  pivot_longer(cols = 21:23, names_to = "Pos", values_to = "Total")

position_total_recruits <- position_total_recruits |> 
  mutate(Pos = case_when(
      Pos == "Top.Total" ~ "Top",
      Pos == "Side.Total" ~ "Side",
      Pos == "Under.Total" ~ "Bottom"
  ))
 
position_total_recruits <- position_total_recruits |>                           
mutate(Pos = factor(Pos, levels = c("Bottom", "Side", "Top")))

position_total_recruits <- left_join(position_total_recruits, join)

## Family recruitment data
family_total_recruits <- Coral_settlement_nGBR %>%
  select(2:13) %>% 
  pivot_longer(cols = 9:12, names_to = "Family", values_to = "Total") |> 
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))
```

# 1. Methods
## 1.1. Table 1, sample sizes
### 1.1.1 PIT

```{r}
PIT |> group_by(Shelf,Reef,Site, Year) |> 
  summarise(
    n = n()
  )
```

### 1.1.2. Settlement Tiles

```{r}
Coral_settlement_nGBR |> group_by(Shelf,Reef,Site, Year) |> 
  summarise(
    n = n()
  )
```

### 1.1.3. Juveniles

```{r}
Belt_10m |>  group_by(Shelf,Reef,Site, Year) |> 
  summarise(
   n = n()
  )
```

# 2. Recruitment

## 2.1. Spatio-temporal Recruitment
### 2.1.1 Total Recruits Tukey HSD contrasts


```{r}
nbin_total_yr2 <- glmmTMB(total ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)
```

Shelf
```{r}
emm_shelf <- emmeans(nbin_total_yr2, ~Shelf | Year, type = "response")
emm_shelf

pairs(emm_shelf)
```

Year

```{r}
emm_year_response <- emmeans(nbin_total_yr2, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

### 2.1.2. Acroporidae Recruits Tukey HSD

```{r}
nbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Acroporidae)
```

```{r}
emm_shelf <- emmeans(nbin_Acroporidae_interaction, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

Year

```{r}
emm_year_response <- emmeans(nbin_Acroporidae_interaction, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```


### 2.1.3. Pocilloporidae Recruits Tukey HSD

```{r}
nbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf * Year + (1 | Reef/Site), 
        family = "nbinom2", data = Pocilloporidae)
```

```{r}
emm_shelf <- emmeans(nbin_Pocilloporidae_interaction, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

Year

```{r}
emm_year_response <- emmeans(nbin_Pocilloporidae_interaction, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```


### 2.1.4. Poritidae Recruits Tukey HSD

```{r}
nbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Poritid)
```

```{r}
emm_shelf <- emmeans(nbin_Poritid_interaction, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

Year

```{r}
emm_year_response <- emmeans(nbin_Poritid_interaction, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```


## 2.2. Composition (PERMANOVA Table)


# 3. Juveniles

## 3.1. Spatio-temporal patterns

### 3.1.1. Total juveniles

```{r}
nbin_total_juv <- glmmTMB(total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_10m)
```

```{r}
emm_shelf <- emmeans(nbin_total_juv, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

```{r}
emm_year_response <- emmeans(nbin_total_juv, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

### 3.1.2. Acroporidae

```{r}
Acro <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Acroporidae)


nbin_acro_juv <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Acro)
```

```{r}
emm_shelf <- emmeans(nbin_acro_juv, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

```{r}
emm_year_response <- emmeans(nbin_acro_juv, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

### 3.1.3. Pocilloporidae

```{r}
Poc <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Pocilloporidae)

nbin_poci_juv <- glmmTMB(Pocilloporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Poc)
```

```{r}
emm_shelf <- emmeans(nbin_poci_juv, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

```{r}
emm_year_response <- emmeans(nbin_poci_juv, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

### 3.1.4. Poritidae
```{r}
Por <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Poritidae)

nbin_por_juv <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Por)
```

```{r}
emm_shelf <- emmeans(nbin_por_juv, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

```{r}
emm_year_response <- emmeans(nbin_por_juv, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

### 3.1.5. Other Hard  corals
```{r}
Other_juv <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Other_HC)

nbin_oth_juv <- glmmTMB(Other_HC ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Other_juv)
```

```{r}
emm_shelf <- emmeans(nbin_oth_juv, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

```{r}
emm_year_response <- emmeans(nbin_oth_juv, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

## 3.2. Composition

# 4. Brood stock and larval supply
## 4.1. Acroporidae
Model selection Table

## 4.2. Pocilloporidae
Model selection Table

## 4.3. Poritidae
Model selection Table

# 5. Larval Supply and Juveniles

```{r}
combine_belt_st <- left_join(join_belt, Join_Coral)

# Only including years in which both data is included (once lagging has happened)
combine_belt_st <- combine_belt_st |> 
  mutate(Year_start = as.numeric(substr(Year, 1, 4)))

gropd_df2 <- combine_belt_st %>%
  arrange(Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_mean_acro_lag1 = lag(st_mean_acro, 1),
    st_mean_poci_lag1 = lag(st_mean_poci, 1),
    st_mean_por_lag1 = lag(st_mean_por, 1),
  ) %>%
  ungroup()
```

## 5.1. Spatial patterns

### 5.1.1 Acroporidae
Model selection
Remove first year row
```{r}
model_df_acro <- gropd_df2 %>%
  filter(!is.na(mean_acro))

ln_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + Shelf + Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)


gau_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 +Shelf + Year + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)


#without year
ln_acro_juv_st_int_wo_yr <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)


gau_acro_juv_st_int_wo_yr<- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)


ln_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf * Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)

gau_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf * Year + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)
```

```{r}
model.sel(gau_acro_juv_st_int, gau_acro_juv_st_fixed, ln_acro_juv_st_fixed, gau_acro_juv_st_int_wo_yr, ln_acro_juv_st_int_wo_yr)

Acro_juv_st_df <- model.sel(gau_acro_juv_st_int, gau_acro_juv_st_fixed, ln_acro_juv_st_fixed, gau_acro_juv_st_int_wo_yr, ln_acro_juv_st_int_wo_yr) |> as.data.frame()
```

Tukey HSD among Shelf
```{r}
emm_shelf <- emmeans(ln_acro_juv_st_int_wo_yr, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```


### 5.1.2. Pocilloporidae

Model selection
```{r}
model_df_poci <- gropd_df2 %>%
  filter(!is.na(mean_poci))

ln_poci_juv_st_fixed <- glmmTMB(mean_poci ~ st_mean_poci_lag1 + Shelf + Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)

gau_poci_juv_st_fixed <- glmmTMB(mean_poci ~ st_mean_poci_lag1+Shelf + Year+ (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)

#without year
ln_poci_juv_st_int_wo_yr <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)

gau_poci_juv_st_int_wo_yr<- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)

ln_poci_juv_st_int <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf * Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)

gau_poci_juv_st_int <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf * Year + (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)
```


```{r}
model.sel(gau_poci_juv_st_fixed, gau_poci_juv_st_int, ln_poci_juv_st_fixed, ln_poci_juv_st_int, gau_poci_juv_st_int_wo_yr, ln_poci_juv_st_int_wo_yr)

Poci_juv_st_df <- model.sel(gau_poci_juv_st_fixed, gau_poci_juv_st_int, ln_poci_juv_st_fixed, ln_poci_juv_st_int, gau_poci_juv_st_int_wo_yr, ln_poci_juv_st_int_wo_yr) |> as.data.frame()
```

Tukey HSD among Shelf
```{r}
emm_shelf <- emmeans(ln_poci_juv_st_int_wo_yr, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

### 5.1.3. Poritidae
Model selection

Tukey HSD among Shelf