---
title: "04_Tables"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)


# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  ),
  Deployment = case_when(
      Year_Collected == "2022" ~ "21/22",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2026" ~ "25/26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")


Pocilloporidae <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Pocilloporidae, Top_Pocil, 
         Under_Pocil, Side_Pocil, total)



Acroporidae <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Acroporidae, Top_Acr, 
         Under_Acr, Side_Acr, total)

# Acroporidae_orientation <- Acroporidae_orientation %>% 
#   pivot_longer(cols = 7:9,
#                names_to = "Tile_Surface",
#                values_to = "Recruits")

Poritid <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Poritid, Top_Por, Under_Por, 
         Side_Por, total)


Other <- Coral_settlement_nGBR %>% 
  dplyr::select(Year, Shelf, 
         Reef, Site, Tile, 
         Other, Top_other, 
         Under_oth, Side_other, total)

# Mean or mean no. tiles per site / year
Join_Coral <- Coral_settlement_nGBR |> 
  group_by(Shelf, Year, Site, Reef) |> 
  summarise(
    st_mean_acro = mean(Acroporidae),
    st_sd_acro = sd(Acroporidae),
    st_median_acro = median(Acroporidae),
    st_mean_poci = mean(Pocilloporidae),
    st_sd_poci = sd(Pocilloporidae),
    st_median_poci = median(Pocilloporidae),
    st_mean_por = mean(Poritid),
    st_sd_por = sd(Poritid),
    st_median_por = median(Poritid),
    st_mean_total = mean(total),
    st_sd_total = sd(total),
    Tiles = n()
  )



# Benthic Data

# Benthic Data
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    ) )


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))


PIT <- PIT %>%
  mutate(
    Acroporidae = rowSums(dplyr::select(., `Acropora - branching`:`Montipora`), na.rm = TRUE),
    
    Pocilloporidae = rowSums(dplyr::select(., Pocillopora:Stylophora), na.rm = TRUE),
    
    Poritidae = rowSums(dplyr::select(., `Porites - branching`:Goniopora), na.rm = TRUE),
    
    `Other Hard coral` = rowSums(dplyr::select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE),
    
    `Soft coral` = rowSums(dplyr::select(., Lobophytum:`Soft-other`), na.rm = TRUE),
    `Brown MA` = rowSums(dplyr::select(., Sargassum:Lobophora), na.rm = TRUE),
    `Red MA` = rowSums(dplyr::select(., Asparagopsis:Jania), na.rm = TRUE),
    `Green MA`= rowSums(dplyr::select(., Halimeda:Valonia), na.rm = TRUE),
    `Sand and Rubble` = rowSums(dplyr::select(., Sand:Rubble), na.rm = TRUE),
    Turf = rowSums(dplyr::select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE),
    Other = rowSums(dplyr::select(., Sponge:Clam), na.rm = TRUE),
    Zoantharian = rowSums(dplyr::select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  dplyr::select(Year, Shelf, Reef, Site, Transect, Acroporidae:Zoantharian,CCA, Cyanobacteria, Halophila)

# PIT <- PIT |> 
#   mutate(
#     sum = rowSums(across(Acroporidae:Halophila), na.rm = TRUE)
#   )
# 
# summary(PIT$sum)

PIT <- PIT |> 
  mutate(
    across(Acroporidae:Halophila, ~ .x * 2)
  )

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    PIT = n()
  )

Combined <- left_join(Coral_settlement_nGBR, join)

# Juvenile data
Belt_10m <- read_csv("Belt_10m_final.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    )
  )

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))
Belt_10m <- Belt_10m |>
  mutate(
    Other_HC = Favia + Favites + Goniastrea + Astrea +
                 Echinopora + Galaxea + Turbinaria +
                 Psammacora + Pavona
  )

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(Total_juveniles),
    sd_total = sd(Total_juveniles),
    Belt = n()
  )

## Total recuits
total_recruits <- Coral_settlement_nGBR %>%
  dplyr::select(Year_Collected, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 5:8, names_to = "Family", values_to = "Total")

#Order so that Other is last in Family
total_recruits <- Coral_settlement_nGBR %>% 
  dplyr::select(Year_Collected, Shelf, Reef, Site, Obs, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 6:9, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))

total_recruits$Year_Collected <- as.factor(total_recruits$Year_Collected)

## Position on tile

position_total_recruits <- Coral_settlement_nGBR %>%
  dplyr::select(2:9, 15:29, 32) %>% 
  pivot_longer(cols = 21:23, names_to = "Pos", values_to = "Total")

position_total_recruits <- position_total_recruits |> 
  mutate(Pos = case_when(
      Pos == "Top.Total" ~ "Top",
      Pos == "Side.Total" ~ "Side",
      Pos == "Under.Total" ~ "Bottom"
  ))
 
position_total_recruits <- position_total_recruits |>                           
mutate(Pos = factor(Pos, levels = c("Bottom", "Side", "Top")))

position_total_recruits <- left_join(position_total_recruits, join)

## Family recruitment data
family_total_recruits <- Coral_settlement_nGBR %>%
  dplyr::select(2:13) %>% 
  pivot_longer(cols = 9:12, names_to = "Family", values_to = "Total") |> 
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))
```

# 1. Methods
## 1.1. Table 1, sample sizes
### 1.1.1 PIT

```{r}
PIT |> group_by(Shelf,Reef,Site, Year) |> 
  summarise(
    n = n()
  )
```

### 1.1.2. Settlement Tiles

```{r}
Coral_settlement_nGBR |> group_by(Shelf,Reef,Site, Year) |> 
  summarise(
    n = n()
  )
```

### 1.1.3. Juveniles

```{r}
Belt_10m |>  group_by(Shelf,Reef,Site, Year) |> 
  summarise(
   n = n()
  )
```

# 2. Recruitment

## 2.1. Spatio-temporal Recruitment
### 2.1.1 Total Recruits Tukey HSD contrasts


```{r}
nbin_total_yr2 <- glmmTMB(total ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

summary(nbin_total_yr2)
```

After discussion with Murray Logan, I think this is the correct way to go. Using Holm adjustment as I can't use Tukey.
Example with shelf year
```{r}
Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    shelfyear = paste(Shelf, Year)
  )
unique(Coral_settlement_nGBR$shelfyear)
nbin_total_yr2 <- glmmTMB(total ~ shelfyear +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

summary(nbin_total_yr2)

cmat <- cbind(
  "Inner v Mid 21/22" = c(1, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
  "Inner v Out 21/22" = c(1, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0),
  "Mid v Out 21/22" = c(0, 1, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0),
  "Inner v Mid 22/23" = c(0, 0, 0, 1, -1, 0, 0, 0, 0, 0, 0, 0),
  "Inner v Out 22/23" = c(0, 0, 0, 1, 0, -1, 0, 0, 0, 0, 0, 0),
  "Mid v Out 22/23" = c(0, 0, 0, 0, 1, -1, 0, 0, 0, 0, 0, 0),
  "Inner v Mid 23/24" = c(0, 0, 0, 0, 0, 0, 1, -1, 0, 0, 0, 0),
  "Inner v Out 23/24" = c(0, 0, 0, 0, 0, 0, 1, 0, -1, 0, 0, 0),
  "Mid v Out 23/24" = c(0, 0, 0, 0, 0, 0, 0, 1, -1, 0, 0, 0),
  "Inner v Mid 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, -1, 0),
  "Inner v Out 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, -1),
  "Mid v Out 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, -1),
  "Inner 21/22 v 22/23" = c(1, 0, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0),
  "Inner 21/22 v 23/24" = c(1, 0, 0, 0, 0, 0, -1, 0, 0, 0, 0, 0),
  "Inner 21/22 v 24/25" = c(1, 0, 0, 0, 0, 0, 0, 0, 0, -1, 0, 0),
  "Inner 22/23 v 23/24" = c(0, 0, 0, 1, 0, 0, -1, 0, 0, 0, 0, 0),
  "Inner 22/23 v 24/25" = c(0, 0, 0, 1, 0, 0, 0, 0, 0, -1, 0, 0),
  "Inner 23/24 v 24/25" = c(0, 0, 0, 0, 0, 0, 1, 0, 0, -1, 0, 0),
  "Mid 21/22 v 22/23" = c(0, 1, 0, 0, -1, 0, 0, 0, 0, 0, 0, 0),
  "Mid 21/22 v 23/24" = c(0, 1, 0, 0, 0, 0, 0, -1, 0, 0, 0, 0),
  "Mid 21/22 v 24/25" = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, -1, 0),
  "Mid 22/23 v 23/24" = c(0, 0, 0, 0, 1, 0, 0, -1, 0, 0, 0, 0),
  "Mid 22/23 v 24/25" = c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, -1, 0),
  "Mid 23/24 v 24/25" = c(0, 0, 0, 0, 0, 0, 0, 1, 0, 0, -1, 0),
  "Outer 21/22 v 22/23" = c(0, 0, 1, 0, 0, -1, 0, 0, 0, 0, 0, 0),
  "Outer 21/22 v 23/24" = c(0, 0, 1, 0, 0, 0, 0, 0, -1, 0, 0, 0),
  "Outer 21/22 v 24/25" = c(0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, -1),
  "Outer 22/23 v 23/24" = c(0, 0, 0, 0, 0, 1, 0, 0, -1, 0, 0, 0),
  "Outer 22/23 v 24/25" = c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, -1),
  "Outer 23/24 v 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, -1)
)
nbin_total_yr2 |> emmeans(~shelfyear, type="response") |> 
  contrast(method=list(shelfyear=cmat), adjust= "holm")
```

```{r}
emm <- emmeans(nbin_total_yr2, ~ shelfyear, type = "response")

levels(emm)
```


### 2.1.2. Acroporidae Recruits Tukey HSD

```{r}
nbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Acroporidae)
```






### 2.1.3. Pocilloporidae Recruits Tukey HSD

```{r}
nbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf * Year + (1 | Reef/Site), 
        family = "nbinom2", data = Pocilloporidae)
```

```{r}
emm_shelf <- emmeans(nbin_Pocilloporidae_interaction, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

Year

```{r}
emm_year_response <- emmeans(nbin_Pocilloporidae_interaction, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

```{r}
pairs(emm_year_response, reverse = TRUE)
```


### 2.1.4. Poritidae Recruits Tukey HSD

```{r}
nbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = Poritid)
```

```{r}
emm_shelf <- emmeans(nbin_Poritid_interaction, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

Year

```{r}
emm_year_response <- emmeans(nbin_Poritid_interaction, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

```{r}
pairs(emm_year_response, reverse = TRUE)
```


## 2.2. Composition (PERMANOVA Table)


# 3. Juveniles

## 3.1. Spatio-temporal patterns

### 3.1.1. Total juveniles

```{r}
nbin_total_juv <- glmmTMB(Total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_10m)
```

```{r}
emm_shelf <- emmeans(nbin_total_juv, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

```{r}
pairs(emm_shelf, reverse = TRUE)
```


```{r}
emm_year_response <- emmeans(nbin_total_juv, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

### 3.1.2. Acroporidae

```{r}
Acro <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Acroporidae)


nbin_acro_juv <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Acro)
```

```{r}
emm_shelf <- emmeans(nbin_acro_juv, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

```{r}
pairs(emm_shelf, reverse = TRUE)
```

```{r}
emm_year_response <- emmeans(nbin_acro_juv, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

```{r}
pairs(emm_year_response, reverse = TRUE)
```


### 3.1.3. Pocilloporidae

```{r}
Poc <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Pocilloporidae)

nbin_poci_juv <- glmmTMB(Pocilloporidae ~ Year+Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Poc)
```

```{r}
emm_shelf <- emmeans(nbin_poci_juv, ~Shelf, type = "response")

pairs(emm_shelf)
```

```{r}
emm_year_response <- emmeans(nbin_poci_juv, ~Year, type = "response")
pairs(emm_year_response)
```

### 3.1.4. Poritidae
```{r}
Por <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Poritidae)

nbin_por_juv <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Por)
```

```{r}
emm_shelf <- emmeans(nbin_por_juv, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

```{r}
emm_year_response <- emmeans(nbin_por_juv, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

### 3.1.5. Other Hard  corals
```{r}
Other_juv <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Other_HC)

nbin_oth_juv <- glmmTMB(Other_HC ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Other_juv)
```

```{r}
emm_shelf <- emmeans(nbin_oth_juv, ~Shelf | Year, type = "response")

pairs(emm_shelf)
```

```{r}
emm_year_response <- emmeans(nbin_oth_juv, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

## 3.2. Composition
```{r}
# Step 1: Create site-by-species matrix with Site_ID
site_by_belt <- Belt_10m %>%
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, Acroporidae, Pocilloporidae, Poritidae, Other_HC
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

site_by_belt <- site_by_belt |>                                                         # ensure only vegetated tiles included 
  group_by(Reef, Year, Shelf) %>%               # this was missing in your code
  summarise(across(Acroporidae:Other_HC, mean), .groups = "drop")

# Step 2: Metadata (KEEP Site_ID)
metadata_belt <- site_by_belt %>%
  dplyr::select(Shelf, Year, Reef)

# Step 3: Species matrix
belt_matrix_filtered <- site_by_belt %>%
  dplyr::select(
    4:7
  )
```

```{r}
# # create a distance matrix
# belt_dist <- vegdist(belt_matrix_filtered, method = "bray")
# 
# # With respect to Shelf
# betadisper(belt_dist, metadata_belt$Shelf)
# anova(betadisper(belt_dist, metadata_belt$Shelf)) # Significant was 0.03139 *
```

Test whether hellinger transformation has less dispersion
```{r}
# create a distance matrix
belt_dist <- vegdist(belt_matrix_filtered, method = "hellinger")

# With respect to Shelf
betadisper(belt_dist, metadata_belt$Shelf)
anova(betadisper(belt_dist, metadata_belt$Shelf)) # Significant was 0.1688 *
```

Because the hellinger transformed meets assumptions for dispersion not significantly different, I think maybe I should use that..

### 3.2.2 Hellinger

```{r}
belt_dist <- vegdist(belt_matrix_filtered, method = "hellinger")
permanova_result <- adonis2(belt_dist ~ Shelf,
  data = site_by_belt,
  method = "euclidean",
  by = "margin",
  permutations = 999)

print(permanova_result)
```

45.9 percent of the variation in hellinger transformed averages was driven by shelf position.
F = 8.06

Indicates strong separation among Shelf groups relative to within-group variation.

```{r}
# permanova_result <- adonis2(belt_matrix_filtered ~ Shelf,
#   data = site_by_belt,
#   method = "bray",
#   by = "margin",
#   permutations = 999)
# 
# print(permanova_result)
```


# 4. Benthic Cover

```{r}
# Step 1: Create site-by-species matrix with Site_ID
site_by_PIT <- PIT %>%
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, 6:20
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

site_by_PIT <- site_by_PIT |>                                                         
  group_by(Reef, Year, Shelf) %>%               
  summarise(across(Acroporidae:Halophila, mean), .groups = "drop")

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  dplyr::select(Shelf, Year, Reef)

# Step 3: Species matrix
PIT_matrix_filtered <- site_by_PIT %>%
  dplyr::select(
    Acroporidae:Halophila
  )
```

```{r}
# create a distance matrix
PIT_dist <- vegdist(PIT_matrix_filtered, method = "bray")

# With respect to Shelf
betadisper(PIT_dist, metadata_PIT$Shelf)
anova(betadisper(PIT_dist, metadata_PIT$Shelf)) # Significant was 0.01143 * ***
```


Not appropriate for composition analysis.
```{r}
# # create a distance matrix
# PIT_dist <- vegdist(PIT_matrix_filtered, method = "hellinger")
# 
# # With respect to Shelf
# betadisper(PIT_dist, metadata_PIT$Shelf)
# anova(betadisper(PIT_dist, metadata_PIT$Shelf)) # Significant was 0.006932 * ***
```

```{r}
PIT <- PIT |> 
  mutate(
    Total_HC_Cover = Acroporidae+ Pocilloporidae + Poritidae + `Other Hard coral`
  )
```


##4.2. Spatio-temporal variability in brood stock

### 4.2.1 Total Hard Coral cover %

```{r}
PIT <- PIT |> 
  mutate(
    Total_HC_Cover = Acroporidae+ Pocilloporidae + Poritidae + `Other Hard coral`
  )

nbin_HC_adult_int <- glmmTMB(Total_HC_Cover ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)
```


### 4.2.2. Acroporidae Hard coral cover

```{r}
nbin_acro_adult_int <- glmmTMB(Acroporidae ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)
```


### 4.2.3. Pocilloporidae Hard Coral Cover %

```{r}
pois_poci_adult_int <- glmmTMB(Pocilloporidae ~ Shelf * Year + (1 | Reef/Site), 
      family = "poisson", data = PIT)
nbin_poci_adult_int <- glmmTMB(Pocilloporidae ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)
```

### 4.2.4 Poritidae Hard coral cover %
```{r}
nbin_por_adult_int <- glmmTMB(Poritidae ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)

nbin_por_adult <- glmmTMB(Poritidae ~ Shelf +  Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)
```

```{r}
emm_shelf <- emmeans(nbin_por_adult, ~Shelf, type = "response")

pairs(emm_shelf)
```

```{r}
emm_year_response <- emmeans(nbin_por_adult, ~Year, type = "response")
pairs(emm_year_response)
```



# 5. Brood stock (PIT) and Recruitment (Tiles)

```{r}
st_join <- Coral_settlement_nGBR |> 
  group_by(Site, Year) |> 
  summarise(
    acro_mean_st = mean(Acroporidae),
    poci_mean_st = mean(Pocilloporidae),
    por_mean_st = mean(Poritid),
    acro_sd_st = sd(Acroporidae),
    poci_sd_st = sd(Pocilloporidae),
    por_sd_st = sd(Poritid),
    acro_median_st = median(Acroporidae),
    poci_median_st = median(Pocilloporidae),
    por_median_st = median(Poritid),
  )

join_pit <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    PIT = n()
  )

combined_averages <- left_join(join_pit, st_join)
```

Now these are both averages, it can't follow an integar and would need to follow a different family of error distribution.

## 4.1. Acroporidae

Despite not being insteresting in across the shelf differences, as shelf has only three levels there are two few levels for a stable randm effect
Despite not being interested in across the shelf differences, as shelf has only three levels there are two few levels for a stable randm effect
```{r}
Acro <- combined_averages |> 
  filter(acro_mean_st > 0)
```

```{r}
gamma_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Reef/Site) + (1 | Year), 
         family = Gamma(link = "log"), data = Acro)
# sim <- simulateResiduals(gamma_avg_Acroporidae_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

ln_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Reef/Site) + (1 | Year), 
        family = "lognormal", data = Acro)
# sim <- simulateResiduals(ln_avg_Acroporidae_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim)

gau_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Reef/Site) + (1 | Year), 
        family = "gaussian", data = Acro)
# sim <- simulateResiduals(gau_avg_Acroporidae_fixed)
# testResiduals(sim) #fails
# testZeroInflation(sim) #passes

gamma_avg_Acroporidae_nested <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Shelf/Reef/Site), 
         family = Gamma(link = "log"), data = Acro)
# sim <- simulateResiduals(gamma_avg_Acroporidae_nested)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

ln_avg_Acroporidae_nested <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Shelf/Reef/Site), 
        family = "lognormal", data = Acro)
# sim <- simulateResiduals(ln_avg_Acroporidae_nested)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

gaus_avg_Acroporidae_nested <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Shelf/Reef/Site), 
        family = "gaussian", data = Acro)
# sim <- simulateResiduals(gaus_avg_Acroporidae_nested)
# testResiduals(sim) #fails miserably
# testZeroInflation(sim)

#interaction with Shelf
gamma_avg_Acroporidae_int <- glmmTMB(acro_mean_st ~ mean_aco*Shelf +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Acro)
# sim <- simulateResiduals(gamma_avg_Acroporidae_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

ln_avg_Acroporidae_int <- glmmTMB(acro_mean_st ~ mean_aco*Shelf +(1 | Reef/Site), 
        family = "lognormal", data = Acro)
# sim <- simulateResiduals(ln_avg_Acroporidae_int)
# testResiduals(sim) #passes
# testZeroInflation(sim) #passes

gaus_avg_Acroporidae_int <- glmmTMB(acro_mean_st ~ mean_aco*Shelf +(1 | Reef/Site), 
        family = "gaussian", data = Acro)
# sim <- simulateResiduals(gaus_avg_Acroporidae_int)
# testResiduals(sim) #terrible, truly.
# testZeroInflation(sim)
```
## 5.1. Acroporidae
Model selection Table
```{r}
model.sel(gamma_avg_Acroporidae_fixed, ln_avg_Acroporidae_fixed, gamma_avg_Acroporidae_nested, ln_avg_Acroporidae_nested, gamma_avg_Acroporidae_int, ln_avg_Acroporidae_int)
```

Appendix summary of model
```{r}

gamma_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ scale(mean_aco, scale = FALSE) +(1 | Reef/Site) + (1 | Year), 
         family = Gamma(link = "log"), data = Acro)

summary(gamma_avg_Acroporidae_fixed)
```

## 5.2. Pocilloporidae
```{r}
## Pocilloporidae
Poci <- combined_averages |> 
  filter(poci_mean_st > 0)

ln_avg_poci_fixed <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Reef/Site) + (1 | Year), 
        family = "lognormal", data = Poci)
ln_avg_poci_fixed2 <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Reef/Site), 
        family = "lognormal", data = Poci)

gaus_avg_poci_fixed <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Reef/Site) + (1 | Year), 
        family = "gaussian", data = Poci)

gamma_avg_Pocilloporidae_fixed <- glmmTMB(poci_mean_st ~ mean_poci + (1 | Reef/Site) + (1 | Year),
        family = Gamma(link = "log"), data = Poci)
gamma_avg_Pocilloporidae_fixed2 <- glmmTMB(poci_mean_st ~ mean_poci + (1 | Reef/Site),
        family = Gamma(link = "log"), data = Poci)

ln_avg_poci_nested <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Shelf/Reef/Site), 
        family = "lognormal", data = Poci)

gaus_avg_poci_nested <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Shelf/Reef/Site), 
        family = "gaussian", data = Poci)

gamma_avg_Pocilloporidae_nested <- glmmTMB(poci_mean_st ~ mean_poci + (1 | Shelf/Reef/Site),
        family = Gamma(link = "log"), data = Poci)

ln_avg_Pocilloporidae_int <- glmmTMB(poci_mean_st ~ mean_poci * Shelf +(1 | Reef/Site), 
         family = "lognormal", data = Poci)

gau_avg_Pocilloporidae_int <- glmmTMB(poci_mean_st ~ mean_poci * Shelf +(1 | Reef/Site), 
         family = "gaussian", data = Poci)

gamma_avg_Pocilloporidae_int <- glmmTMB(poci_mean_st ~ mean_poci * Shelf +(1 | Reef/Site), 
         family = Gamma(link = "log"), data = Poci)
```

```{r}
# sim14.1 <- simulateResiduals(gamma_avg_Pocilloporidae_fixed) #passed
# testResiduals(sim14.1)
# testZeroInflation(sim14.1)
# sim14.1 <- simulateResiduals(gamma_avg_Pocilloporidae_fixed2) #passed
# testResiduals(sim14.1)
# testZeroInflation(sim14.1)
# sim14.2 <- simulateResiduals(ln_avg_poci_fixed) # very good
# testResiduals(sim14.2)
# testZeroInflation(sim14.2)
# 
# sim14.3 <- simulateResiduals(gaus_avg_poci_fixed)
# testResiduals(sim14.3) # definitelty not this one
# testZeroInflation(sim14.3)
# 
# sim14.4 <- simulateResiduals(gamma_avg_Pocilloporidae_nested) #passed
# testResiduals(sim14.4)
# testZeroInflation(sim14.4)
# 
# sim14.5 <- simulateResiduals(ln_avg_poci_nested) # very good
# testResiduals(sim14.5)
# testZeroInflation(sim14.5)
# 
# sim14.6 <- simulateResiduals(gaus_avg_poci_fixed)
# testResiduals(sim14.6) # Outliers
# testZeroInflation(sim14.6)
# 
# sim14.7 <- simulateResiduals(ln_avg_Pocilloporidae_int)
# testResiduals(sim14.7) # passes
# testZeroInflation(sim14.7) #passes
# 
# sim14.8 <- simulateResiduals(gamma_avg_Pocilloporidae_int)
# testResiduals(sim14.8) # passes
# testZeroInflation(sim14.8) #passes
# 
# sim14.9 <- simulateResiduals(gau_avg_Pocilloporidae_int)
# testResiduals(sim14.9) # passes
# testZeroInflation(sim14.9) #passes
```


Model selection Table
```{r}
model.sel(gamma_avg_Pocilloporidae_fixed, gamma_avg_Pocilloporidae_nested, ln_avg_poci_fixed, ln_avg_poci_nested, gamma_avg_Pocilloporidae_int, gau_avg_Pocilloporidae_int, ln_avg_Pocilloporidae_int, gamma_avg_Pocilloporidae_fixed2, ln_avg_poci_fixed2)
```

Summary of best performing model for appendix

Without year performs slightly better
```{r}
gamma_avg_Pocilloporidae_fixed <- glmmTMB(poci_mean_st ~ scale(mean_poci, scale = FALSE) + (1 | Reef/Site) + (1 | Year),
        family = Gamma(link = "log"), data = Poci)

summary(gamma_avg_Pocilloporidae_fixed)
```

## 5.3. Poritidae
```{r}
## Poritidae
Por <- combined_averages |> 
  filter(por_mean_st > 0)

gamma_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Reef/Site) + (1 | Year),
        family = Gamma(link = "log"), data = Por)
ln_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Reef/Site) + (1 | Year), 
        family = "lognormal", data = Por)
gaus_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Reef/Site) + (1 | Year), 
        family = "gaussian", data = Por)

```

```{r}
# sim15.1 <- simulateResiduals(gamma_avg_Poritid_fixed) #passed
# testResiduals(sim15.1)
# testZeroInflation(sim15.1)
# sim15.2 <- simulateResiduals(ln_avg_Poritid_fixed) # very good
# testResiduals(sim15.2)
# testZeroInflation(sim15.2)
# 
# sim15.3 <- simulateResiduals(gaus_avg_Poritid_fixed)
# testResiduals(sim15.3) # definitely not this one
# testZeroInflation(sim15.3)
```
Model selection Table
```{r}
model.sel(gamma_avg_Poritid_fixed, ln_avg_Poritid_fixed)
```

```{r}
ln_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ scale(mean_por, scale = FALSE) + (1 | Reef/Site) + (1 | Year), 
        family = "lognormal", data = Por)
summary(ln_avg_Poritid_fixed)
```

# 6. Larval Supply and Juveniles

```{r}
combine_belt_st <- left_join(join_belt, Join_Coral)

# Only including years in which both data is included (once lagging has happened)
combine_belt_st <- combine_belt_st |> 
  mutate(Year_start = as.numeric(substr(Year, 1, 4)))

gropd_df2 <- combine_belt_st %>%
  arrange(Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_mean_acro_lag1 = lag(st_mean_acro, 1),
    st_mean_poci_lag1 = lag(st_mean_poci, 1),
    st_mean_por_lag1 = lag(st_mean_por, 1),
  ) %>%
  ungroup()
```

## 6.1. Spatial patterns

### 6.1.1 Acroporidae
Model selection
Remove first year row
```{r}
model_df_acro <- gropd_df2 %>%
  filter(!is.na(mean_acro))

ln_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + Shelf + Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)


gau_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 +Shelf + Year + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)


#without year
ln_acro_juv_st_int_wo_yr <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)


gau_acro_juv_st_int_wo_yr<- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)


ln_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf * Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)

gau_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf * Year + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)
```

```{r}
model.sel(gau_acro_juv_st_int, gau_acro_juv_st_fixed, ln_acro_juv_st_fixed, gau_acro_juv_st_int_wo_yr, ln_acro_juv_st_int_wo_yr)

Acro_juv_st_df <- model.sel(gau_acro_juv_st_int, gau_acro_juv_st_fixed, ln_acro_juv_st_fixed, gau_acro_juv_st_int_wo_yr, ln_acro_juv_st_int_wo_yr) |> as.data.frame()
```
```{r}
ln_acro_juv_st_int <- glmmTMB(mean_acro ~ scale(st_mean_acro_lag1, scale = FALSE) * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)

summary(model_df_acro$st_mean_acro_lag1)

summary(ln_acro_juv_st_int)
```




Tukey HSD among Shelf
```{r}
emm_shelf <- emmeans(ln_acro_juv_st_int_wo_yr, ~Shelf, type = "response")

pairs(emm_shelf)
```

### 6.1.2. Pocilloporidae

Model selection
```{r}
model_df_poci <- gropd_df2 %>%
  filter(!is.na(mean_poci))

ln_poci_juv_st_fixed <- glmmTMB(mean_poci ~ st_mean_poci_lag1 + Shelf + Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)

gau_poci_juv_st_fixed <- glmmTMB(mean_poci ~ st_mean_poci_lag1+Shelf + Year+ (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)

#without year
ln_poci_juv_st_int_wo_yr <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)

gau_poci_juv_st_int_wo_yr<- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)

ln_poci_juv_st_int <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf * Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)

gau_poci_juv_st_int <- glmmTMB(mean_poci ~ st_mean_poci_lag1 * Shelf * Year + (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)
```


```{r}
model.sel(gau_poci_juv_st_fixed, gau_poci_juv_st_int, ln_poci_juv_st_fixed, ln_poci_juv_st_int, gau_poci_juv_st_int_wo_yr, ln_poci_juv_st_int_wo_yr)

Poci_juv_st_df <- model.sel(gau_poci_juv_st_fixed, gau_poci_juv_st_int, ln_poci_juv_st_fixed, ln_poci_juv_st_int, gau_poci_juv_st_int_wo_yr, ln_poci_juv_st_int_wo_yr) |> as.data.frame()
```

```{r}
summary(ln_poci_juv_st_int_wo_yr)
```

Tukey HSD among Shelf
```{r}
emm_shelf <- emmeans(ln_poci_juv_st_int_wo_yr, ~Shelf, type = "response")

pairs(emm_shelf)
```

### 6.1.3. Poritidae
Model selection
```{r}
model_df_por <- gropd_df2 %>%
  filter(!is.na(mean_por))

model_df_por <- model_df_por |> 
  filter(mean_por > 0)
```

```{r}
ln_por_juv_st_fixed <- glmmTMB(mean_por ~ st_mean_por_lag1 + Shelf + Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_por)

gau_por_juv_st_fixed <- glmmTMB(mean_por ~ st_mean_por_lag1 + Shelf + Year + (1|Reef/Site),
              family = "gaussian",
           data = model_df_por)
#without year
ln_por_juv_st_int_wo_yr <- glmmTMB(mean_por ~ st_mean_por_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_por)
gau_por_juv_st_int_wo_yr<- glmmTMB(mean_por ~ st_mean_por_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_por)

ln_por_juv_st_int <- glmmTMB(mean_por ~ st_mean_por_lag1 * Shelf * Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_por)

gau_por_juv_st_int<- glmmTMB(mean_por ~ st_mean_por_lag1 * Shelf * Year + (1|Reef/Site),
              family = "gaussian",
           data = model_df_por)
```


Tukey HSD among Shelf
```{r}
emm_shelf <- emmeans(ln_por_juv_st_int_wo_yr, ~Shelf, type = "response")

pairs(emm_shelf)
```

# 7. Exporting the data to excel spreadsheets to make my Tukey tables

Note that in my emmeans the Tukey-adjusted p-values are within each year (the 3 comparisons), whereas Benjamin Hochberg would adjust across all comparisons you include in the p.adjust call.

```{r}
library(openxlsx)

extract_emm_results <- function(model, filename) {
  ## ==============================
  ## Shelf within Year
  ## ==============================

  emm_shelf <- emmeans(model, ~ Shelf | Year, type = "response")

  emm_shelf_df <- as.data.frame(summary(emm_shelf)) %>%
    dplyr::rename(
      Mean = response,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL
    )

  contrast_shelf_df <- as.data.frame(
    summary(pairs(emm_shelf), infer = TRUE)
  ) %>%
    dplyr::rename(
      Ratio = ratio,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL,
      p_Tukey = p.value
    )

  contrast_shelf_df$p_BH <- p.adjust(
    contrast_shelf_df$p_Tukey,
    method = "BH"
  )


  ## ==============================
  ## Year within Shelf
  ## ==============================

  emm_year <- emmeans(model, ~ Year | Shelf, type = "response")

  emm_year_df <- as.data.frame(summary(emm_year)) %>%
    dplyr::rename(
      Mean = response,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL
    )

  contrast_year_df <- as.data.frame(
    summary(pairs(emm_year), infer = TRUE)
  ) %>%
    dplyr::rename(
      Ratio = ratio,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL,
      p_Tukey = p.value
    )

  contrast_year_df$p_BH <- p.adjust(
    contrast_year_df$p_Tukey,
    method = "BH"
  )


  ## ==============================
  ## Export to Excel
  ## ==============================

  openxlsx::write.xlsx(
    list(
      "EMM_Shelf_within_Year" = emm_shelf_df,
      "Shelf_Contrasts" = contrast_shelf_df,
      "EMM_Year_within_Shelf" = emm_year_df,
      "Year_Contrasts" = contrast_year_df
    ),
    file = filename
  )
}

# nbin_por_adult <- glmmTMB(Poritidae ~ Shelf +  Year +(1 | Reef/Site), 
#         family = "nbinom2", data = PIT_adult)
# extract_emm_results(nbin_total_yr2, "output/Total_Recruits.xlsx")
# extract_emm_results(nbin_Acroporidae_interaction, "output/Acroporidae_recruits.xlsx")
# extract_emm_results(nbin_Pocilloporidae_interaction, "output/Pocilloporidae_recruits.xlsx")
# extract_emm_results(nbin_Poritid_interaction, "output/Poritidae_recruits.xlsx")
# extract_emm_results(nbin_total_juv, "output/Total_juveniles.xlsx")
# extract_emm_results(nbin_acro_juv, "output/Acroporidae_juveniles.xlsx")
# extract_emm_results(nbin_poci_juv, "output/Pocilloporidae_juveniles.xlsx")
# extract_emm_results(nbin_por_juv, "output/Poritidae_juveniles.xlsx")
# 
# extract_emm_results(nbin_HC_adult_int, "output/Adult_HC.xlsx")
# extract_emm_results(nbin_acro_adult_int, "output/Adult_acro.xlsx")
extract_emm_results(nbin_poci_adult_int, "output/Adult_poci.xlsx")
```

```{r}
extract_emm_results_pois <- function(model, filename) {
  ## ==============================
  ## Shelf within Year
  ## ==============================

  emm_shelf <- emmeans(model, ~ Shelf | Year, type = "response")

  emm_shelf_df <- as.data.frame(summary(emm_shelf)) %>%
    dplyr::rename(
      Mean = rate,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL
    )

  contrast_shelf_df <- as.data.frame(
    summary(pairs(emm_shelf), infer = TRUE)
  ) %>%
    dplyr::rename(
      Ratio = ratio,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL,
      p_Tukey = p.value
    )

  contrast_shelf_df$p_BH <- p.adjust(
    contrast_shelf_df$p_Tukey,
    method = "BH"
  )


  ## ==============================
  ## Year within Shelf
  ## ==============================

  emm_year <- emmeans(model, ~ Year | Shelf, type = "response")

  emm_year_df <- as.data.frame(summary(emm_year)) %>%
    dplyr::rename(
      Mean = rate,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL
    )

  contrast_year_df <- as.data.frame(
    summary(pairs(emm_year), infer = TRUE)
  ) %>%
    dplyr::rename(
      Ratio = ratio,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL,
      p_Tukey = p.value
    )

  contrast_year_df$p_BH <- p.adjust(
    contrast_year_df$p_Tukey,
    method = "BH"
  )


  ## ==============================
  ## Export to Excel
  ## ==============================

  openxlsx::write.xlsx(
    list(
      "EMM_Shelf_within_Year" = emm_shelf_df,
      "Shelf_Contrasts" = contrast_shelf_df,
      "EMM_Year_within_Shelf" = emm_year_df,
      "Year_Contrasts" = contrast_year_df
    ),
    file = filename
  )
}

extract_emm_results_pois(pois_poci_adult_int, "output/Adult_poci.xlsx")
```


Extracting with the 23/24 year as the main one

Extracting juveniles with the outer shelf as the dominant 

Extractions of models where there is no interaction term
```{r}
library(openxlsx)

extract_emm_results <- function(model, filename) {
  ## ==============================
  ## Shelf within Year
  ## ==============================

  emm_shelf <- emmeans(model, ~ Shelf, type = "response")

  emm_shelf_df <- as.data.frame(summary(emm_shelf)) %>%
    dplyr::rename(
      Mean = response,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL
    )

  contrast_shelf_df <- as.data.frame(
    summary(pairs(emm_shelf), infer = TRUE)
  ) %>%
    dplyr::rename(
      Ratio = ratio,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL,
      p_Tukey = p.value
    )

  contrast_shelf_df$p_BH <- p.adjust(
    contrast_shelf_df$p_Tukey,
    method = "BH"
  )


  ## ==============================
  ## Year within Shelf
  ## ==============================

  emm_year <- emmeans(model, ~ Year, type = "response")

  emm_year_df <- as.data.frame(summary(emm_year)) %>%
    dplyr::rename(
      Mean = response,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL
    )

  contrast_year_df <- as.data.frame(
    summary(pairs(emm_year), infer = TRUE)
  ) %>%
    dplyr::rename(
      Ratio = ratio,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL,
      p_Tukey = p.value
    )

  contrast_year_df$p_BH <- p.adjust(
    contrast_year_df$p_Tukey,
    method = "BH"
  )


  ## ==============================
  ## Export to Excel
  ## ==============================

  openxlsx::write.xlsx(
    list(
      "EMM_Shelf" = emm_shelf_df,
      "Shelf_Contrasts" = contrast_shelf_df,
      "EMM_Year" = emm_year_df,
      "Year_Contrasts" = contrast_year_df
    ),
    file = filename
  )
}

nbin_por_adult <- glmmTMB(Poritidae ~ Shelf +  Year +(1 | Reef/Site), 
        family = "nbinom2", data = PIT)
Poc <- Belt_10m |> 
  dplyr::select(Year, Shelf, Reef, Site, Pocilloporidae)
nbin_poci_juv <- glmmTMB(Pocilloporidae ~ Year + Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Poc)

# extract_emm_results(nbin_por_adult, "output/Adult_Por_fixed.xlsx")
# extract_emm_results(nbin_poci_juv, "output/Juv_Poci_fixed.xlsx")
```

