---
title: "Summary Statistics"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)


# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")

Pocilloporidae <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, 
         Reef, Site, Tile, 
         Pocilloporidae, Top_Pocil, 
         Under_Pocil, Side_Pocil, total)

# Pocilloporidae_orientation <- Pocilloporidae %>% 
#   rename(Top = Top_Pocil,
#                    Bottom = Under_Pocil,
#                    Side = Side_Pocil,
#          Recruits_per_tile = total)
# 
# Pocilloporidae_orientation <- Pocilloporidae_orientation %>% 
#   pivot_longer(cols = 7:9,
#                names_to = "Tile_Surface",
#                values_to = "Recruits")

Acroporidae <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, 
         Reef, Site, Tile, 
         Acroporidae, Top_Acr, 
         Under_Acr, Side_Acr, total)

# Acroporidae_orientation <- Acroporidae %>% 
#   rename(Top = Top_Acr,
#                    Bottom = Under_Acr,
#                    Side = Side_Acr,
#          Recruits_per_tile = total)
# 
# Acroporidae_orientation <- Acroporidae_orientation %>% 
#   pivot_longer(cols = 7:9,
#                names_to = "Tile_Surface",
#                values_to = "Recruits")

Poritid <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, 
         Reef, Site, Tile, 
         Poritid, Top_Por, Under_Por, 
         Side_Por, total)

# Poritidae_orientation <- Poritid %>% 
#   rename(Top = Top_Por,
#                    Bottom = Under_Por,
#                    Side = Side_Por,
#          Recruits_per_tile = total) 
# 
# Poritidae_orientation <- Poritidae_orientation %>% 
#   pivot_longer(cols = 7:9,
#                names_to = "Tile_Surface",
#                values_to = "Recruits")


# Benthic Data

# Benthic Data
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0)))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    )  )



PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))


PIT <- PIT %>%
  mutate(
    Acroporidae = rowSums(select(., `Acropora - branching`:`Montipora`), na.rm = TRUE),
    
    Pocilloporidae = rowSums(select(., Pocillopora:Stylophora), na.rm = TRUE),
    
    Poritidae = rowSums(select(., `Porites - branching`:Goniopora), na.rm = TRUE),
    
    `Other Hard coral` = rowSums(select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE),
    
    `Soft coral` = rowSums(select(., Lobophytum:`Soft-other`), na.rm = TRUE),
    `Brown MA` = rowSums(select(., Sargassum:Lobophora), na.rm = TRUE),
    `Red MA` = rowSums(select(., Asparagopsis:Jania), na.rm = TRUE),
    `Green MA`= rowSums(select(., Halimeda:Valonia), na.rm = TRUE),
    `Sand and Rubble` = rowSums(select(., Sand:Rubble), na.rm = TRUE),
    Turf = rowSums(select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE),
    Other = rowSums(select(., Sponge:Clam), na.rm = TRUE),
    Zoantharian = rowSums(select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  select(Year, Shelf, Reef, Site, Transect, Acroporidae:Zoantharian,CCA, Cyanobacteria, Halophila)

PIT <- PIT |> 
  mutate(
    sum = rowSums(across(Acroporidae:Halophila), na.rm = TRUE)
  )

summary(PIT$sum)

PIT <- PIT |> 
  mutate(
    across(Acroporidae:Halophila, ~ .x * 2)
  )
PIT <- PIT |> 
  mutate(
    Total_HC_Cover = Acroporidae+ Pocilloporidae + Poritidae + `Other Hard coral`
  )
# Combined <- left_join(Coral_settlement_nGBR, join)

## Total recuits
total_recruits <- Coral_settlement_nGBR %>%
  select(Year_Collected, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 5:8, names_to = "Family", values_to = "Total")

#Order so that Other is last in Family
total_recruits <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, Reef, Site, Obs, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 6:9, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))

total_recruits$Year_Collected <- as.factor(total_recruits$Year_Collected)

## Position on tile

position_total_recruits <- Coral_settlement_nGBR %>%
  select(2:9, 15:29, 32) %>% 
  pivot_longer(cols = 21:23, names_to = "Pos", values_to = "Total")

position_total_recruits <- position_total_recruits |> 
  mutate(Pos = case_when(
      Pos == "Top.Total" ~ "Top",
      Pos == "Side.Total" ~ "Side",
      Pos == "Under.Total" ~ "Bottom"
  ))
 
position_total_recruits <- position_total_recruits |>                           
mutate(Pos = factor(Pos, levels = c("Bottom", "Side", "Top")))

# Juvenile data
Belt_10m <- read_csv("Belt_10m_final.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    ),
    Site = factor(Site, levels = levels(PIT$Site))
  )

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))

Belt_10m <- Belt_10m |>
  mutate(
    Other_HC = Favia + Favites + Goniastrea + Astrea +
                 Echinopora + Galaxea + Turbinaria +
                 Psammacora + Pavona
  )

## Family recruitment data
family_total_recruits <- Coral_settlement_nGBR %>%
  select(2:13) %>% 
  pivot_longer(cols = 9:12, names_to = "Family", values_to = "Total") |> 
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))
```

Summary statistics for results text in thesis.

# 1. Coral Recruitment

## 1.1. Families
```{r echo=FALSE}
Coral_settlement_nGBR |> 
  summarise(total_tiles = n(),
            total_recruits = sum(total),
            total_acro = sum(Acroporidae),
            total_poci = sum(Pocilloporidae),
            total_poritidae = sum(Poritid),
            total_other = sum(Other),
            total_bottom = sum(Under.Total),
            total_side = sum(Side.Total),
            total_top = sum(Top.Total)) |> 
  knitr::kable()
1839/21168*100

```

Relative abundance of recruits

```{r}
Coral_settlement_nGBR |> 
  group_by(Shelf, Year) |> 
  summarise(
            total_acro = sum(Acroporidae),
            total_poci = sum(Pocilloporidae),
            total_poritidae = sum(Poritid),
            total_other = sum(Other),
            total_recruits = sum(total),
            total_tiles = n()
            )

Coral_settlement_nGBR |> 
  group_by(Shelf, Reef, Year) |> 
  summarise(
            total_acro = sum(Acroporidae),
            total_poci = sum(Pocilloporidae),
            total_poritidae = sum(Poritid),
            total_other = sum(Other),
            total_recruits = sum(total),
            total_tiles = n()
            ) |> 
  write.csv(file = "output/relative_prop_recruits_reef.csv")


```


Range of recruits to tiles

```{r}
Coral_settlement_nGBR |> 
  summarise(
    min_recruits = min(total),
    max_recruits = max(total)
  )
```

compare to hughes that found the max no reecruits at Lizard was 366
```{r}
Coral_settlement_nGBR |> 
  filter(Reef == "Lizard") |> 
  summarise(
    min_recruits = min(total),
    max_recruits = max(total),
    n_replicates = n()
  )

Coral_settlement_nGBR |> 
  filter(Reef == "MacGillivray") |> 
  summarise(
    min_recruits = min(total),
    max_recruits = max(total),
    n_replicates = n()
  )
Coral_settlement_nGBR |> 
  filter(Reef == "MacGillivray") |> 
  group_by(Year) |> 
  summarise(
    mean_recruits = mean(total),
    se_recruits = sd(total)/n()^0.5,
    n_panels = n()
  )

```


Relative proportion
```{r}
Coral_settlement_nGBR |> 
  group_by(Year, Reef, Site, Shelf) |> 
  summarise(
    Acro_prop = sum(Acroporidae)/sum(total),
    Poci_prop =  sum(Pocilloporidae)/sum(total),
    por_prop =  sum(Poritid)/sum(total),
    Tiles = n()
  ) 
```

```{r}
Coral_settlement_nGBR |> 
  group_by(Year, Reef, Site, Shelf) |> 
  summarise(
    Acro_prop = sum(Acroporidae)/sum(total),
    Poci_prop =  sum(Pocilloporidae)/sum(total),
    por_prop =  sum(Poritid)/sum(total),
    Tiles = n()
  ) |> 
  ungroup() |> 
  pivot_longer(cols = 5:7, names_to = "Family", values_to = "Avg Relative abundance" ) |> 
  group_by(Shelf, Year, Reef, Family) |> 
  summarise(
    mean = mean(`Avg Relative abundance`),
    min = min(`Avg Relative abundance`),
    max = max(`Avg Relative abundance`)
  )
  
```

### 1.1.1. Acroporidae

```{r}
Coral_settlement_nGBR |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Acroporidae),
    se = sd(Acroporidae)/(n()^0.5),
    n= n()
  )

# inner ver mid
47.754167/16.299107
# inner vs outer
47.754167/2.214953
```

```{r}
Coral_settlement_nGBR |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Acroporidae),
    se = sd(Acroporidae)/(n()^0.5),
    n= n()
  )

#2022 and 2023
22.295302/7.393064

#2024 and 2022
47.145251/22.295302

#2024 and 2023
47.145251/7.393064

#2025 and 2024
47.145251/14.384181
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2022") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Acroporidae),
    se = sd(Acroporidae)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2023") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Acroporidae),
    se = sd(Acroporidae)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2024") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Acroporidae),
    se = sd(Acroporidae)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2025") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Acroporidae),
    se = sd(Acroporidae)/(n()^0.5),
    n= n()
  )
```

Shelf
Inner
```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Inner") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Acroporidae),
    se = sd(Acroporidae)/(n()^0.5),
    n= n()
  )

#22 and 23
48.65000/10.46667

#2022 and 2024
100.41667/48.65000
```
Mid

```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Mid") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Acroporidae),
    se = sd(Acroporidae)/(n()^0.5),
    n= n()
  )
```

Outer

```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Outer") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Acroporidae),
    se = sd(Acroporidae)/(n()^0.5),
    n= n()
  )
```

### 1.1.2. Pocilloporidae

```{r}
Coral_settlement_nGBR |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Pocilloporidae),
    se = sd(Pocilloporidae)/(n()^0.5),
    n= n()
  )
```



```{r}
Coral_settlement_nGBR |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Pocilloporidae),
    se = sd(Pocilloporidae)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2022") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Pocilloporidae),
    se = sd(Pocilloporidae)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2023") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Pocilloporidae),
    se = sd(Pocilloporidae)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2024") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Pocilloporidae),
    se = sd(Pocilloporidae)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2025") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Pocilloporidae),
    se = sd(Pocilloporidae)/(n()^0.5),
    n= n()
  )
```

Shelf
Inner
```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Inner") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Pocilloporidae),
    se = sd(Pocilloporidae)/(n()^0.5),
    n= n()
  )
```

Mid

```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Mid") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Pocilloporidae),
    se = sd(Pocilloporidae)/(n()^0.5),
    n= n()
  )
```

Outer

```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Outer") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Pocilloporidae),
    se = sd(Pocilloporidae)/(n()^0.5),
    n= n()
  )
```

### 1.1.3. Poritidae

```{r}
Coral_settlement_nGBR |> 
  # group_by(Shelf) |> 
  summarise(
    mean = mean(Poritid),
    se = sd(Poritid)/(n()^0.5),
    min = min(Poritid),
    max = max(Poritid),
    n= n()
  )
Coral_settlement_nGBR |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Poritid),
    se = sd(Poritid)/(n()^0.5),
    min = min(Poritid),
    max = max(Poritid),
    n= n()
  )
```


## 1.2.Spatiotemporal

Temporal statistics

```{r echo=FALSE}
Coral_settlement_nGBR |> 
  group_by(Year_Collected) |> 
  summarise(total_tiles = n(),
            total_recruits = sum(total),
            total_acro = sum(Acroporidae),
            total_poci = sum(Pocilloporidae),
            total_poritidae = sum(Poritid),
            total_bottom = sum(Under.Total),
            total_side = sum(Side.Total),
            total_top = sum(Top.Total)) |> 
  knitr::kable()
```

```{r}
Coral_settlement_nGBR |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )

Coral_settlement_nGBR |> 
  group_by(Year) |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )

#21/22 and 23/24
59.24022/27.15976

#22/23 and 23/24
59.24022/13.52571

#23/24 and 24/25
59.24022/20.56497


```

By Shelf Position

```{r}
Coral_settlement_nGBR |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )

55.520833/24.259740
55.520833/9.921397
```

Inner 

```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Inner") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )

```

Mid

```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Mid") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )
```

Outer

```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Outer") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )
```

Year and Shelf
2022
```{r}
Coral_settlement_nGBR |> 
  filter(Year_Collected == "2022") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )
```

2023 

```{r}
Coral_settlement_nGBR |> 
  filter(Year_Collected == "2023") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )
```

2024 

```{r}
Coral_settlement_nGBR |> 
  filter(Year_Collected == "2024") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )
```

2025

```{r}
Coral_settlement_nGBR |> 
  filter(Year_Collected == "2025") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )
```


## 1.3. Position on tile

```{r}
position_total_recruits |> 
  group_by(Pos) |> 
  summarise(
    mean = mean(Total),
    se = sd(Total)/(n()^0.5),
    n= n()
  )
```

```{r}
position_total_recruits |> 
  group_by(Pos, Shelf) |> 
  summarise(
    mean = mean(Total),
    se = sd(Total)/(n()^0.5),
    n= n()
  )
```

```{r}
position_total_recruits |> 
  group_by(Pos, Year_Collected) |> 
  summarise(
    mean = mean(Total),
    se = sd(Total)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Top.Total),
    se = sd(Top.Total)/(n()^0.5),
    n= n()
  )
```



```{r}
Coral_settlement_nGBR |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Top.Total),
    se = sd(Top.Total)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2022") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Top.Total),
    se = sd(Top.Total)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2023") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Top.Total),
    se = sd(Top.Total)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2024") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Top.Total),
    se = sd(Top.Total)/(n()^0.5),
    n= n()
  )
```

```{r}
Coral_settlement_nGBR |>
  filter(Year_Collected == "2025") |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(Top.Total),
    se = sd(Top.Total)/(n()^0.5),
    n= n()
  )
```

Shelf
Inner
```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Inner") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Top.Total),
    se = sd(Top.Total)/(n()^0.5),
    n= n()
  )
```

Mid

```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Mid") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Top.Total),
    se = sd(Top.Total)/(n()^0.5),
    n= n()
  )
```

Outer

```{r}
Coral_settlement_nGBR |> 
  filter(Shelf == "Outer") |> 
  group_by(Year_Collected) |> 
  summarise(
    mean = mean(Top.Total),
    se = sd(Top.Total)/(n()^0.5),
    n= n()
  )
```

# 2. Juveniles
```{r}
Belt_10m <- read_csv("Belt_10m.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))

Belt_10m <- Belt_10m |>
  mutate(
    Other_HC_R = Favia + Favites + Goniastrea + Astrea +
                 Echinopora + Galaxea + Turbinaria +
                 Psammacora + Pavona
  )

Belt_10m |> 
  summarise(
    juveniles = sum(Total_juveniles)
  )

#how many belt transects were there no juveniles
Belt_10m |> 
  filter(Total_juveniles == 0) |> 
  summarise(
    belt_transects_wo_juveniles = n()
  )

#how many belt transects were there no juveniles
Belt_10m |> 
  filter(Total_juveniles == 39) |> 
  summarise(
    belt_transects_max_juveniles = n()
  )



```

Mean no juveniles
```{r}
Belt_10m |> 
  summarise(
    mean_juv = mean(Total_juveniles),
    se_juv = sd(Total_juveniles)/n()^0.5
  )
```

```{r}
Belt_10m |> 
  mutate(juveniles_metre = Total_juveniles/10) |> 
  summarise(
    mean_juv = mean(juveniles_metre),
    se_juv = sd(juveniles_metre)/n()^0.5
  )
 
```

Cross shelf averages
```{r}
Belt_10m |> 
  mutate(juveniles_metre = Total_juveniles/10) |>
  group_by(Shelf) |> 
  summarise(
    mean_juv = mean(juveniles_metre),
    se_juv = sd(juveniles_metre)/n()^0.5
  )
```


## 2.1. Families

```{r}
Belt_10m |> 
  summarise(
    sum = sum(Total_juveniles),
    sum_acro = sum(Acroporidae),
    sum_poci = sum(Pocilloporidae),
    sum_pori = sum(Poritidae),
    sum_oth = sum(Other_HC_R),
    n = n()
  )

Belt_10m |> 
  group_by(Shelf) |> 
  summarise(
    sum = sum(Total_juveniles),
    sum_acro = sum(Acroporidae),
    sum_poci = sum(Pocilloporidae),
    sum_pori = sum(Poritidae),
    sum_oth = sum(Other_HC_R),
    sum_Favites = sum(Favites),
    sum_favia = sum(Favia),
    sum_goniastrea = sum(Goniastrea),
    sum_tub = sum(Turbinaria),
    sum_Galaxea = sum(Galaxea),
    n = n()
  )
```

```{r}
Belt_10m |> 
  summarise(
    min_total = min(Total_juveniles),
    max_total = max(Total_juveniles),
    min_acro = min(Acroporidae),
    max_acro = max(Acroporidae),
    min_poci = min(Pocilloporidae),
    max_poci = max(Pocilloporidae),
    min_por = min(Poritidae),
    max_por = max(Poritidae)
  )
```

Relative abundance of coral juveniles across the shelf
```{r}
Belt_10m |>
  group_by(Shelf, Year) |> 
  summarise(
    prop_acro = sum(Acroporidae)/sum(Total_juveniles),
    prop_poci = sum(Pocilloporidae)/sum(Total_juveniles),
    prop_pori = sum(Poritidae)/sum(Total_juveniles),
    prop_oth = sum(Other_HC_R)/sum(Total_juveniles),
    n = n()
  ) 
```

```{r}
Belt_10m |>
  group_by(Shelf, Reef, Year) |> 
  summarise(
    sum_Acroporidae = sum(Acroporidae),
    sum_Pocilloporidae = sum(Pocilloporidae),
    sum_Poritidae = sum(Poritidae),
    sum_other = sum(Other_HC_R),
    sum_Total = sum(Total_juveniles),
    n = n()
  ) |> 
  write.csv(file = "output/relative_prop_juveniles.csv")
```

Favia + Favites + Goniastrea + Astrea +
                 Echinopora + Galaxea + Turbinaria +
                 Psammacora + Pavona
```{r}
Belt_10m |> 
  group_by(Shelf, Reef) |> 
  summarise(
    sum = sum(Total_juveniles),
    sum_acro = sum(Acroporidae),
    sum_acropora = sum(Acropora),
    sum_isopora = sum(Isopora),
    sum_montipora = sum(Montipora),
    sum_poci = sum(Pocilloporidae),
    sum_pocillopora = sum(Pocillopora),
    sum_seriatopora = sum(Seriatopora),
    sum_stylophora = sum(Stylophora),
    sum_pori = sum(Poritidae),
    sum_porites = sum(Porites),
    sum_oth = sum(Other_HC_R),
    sum_Favites = sum(Favites),
    sum_favia = sum(Favia),
    sum_goniastrea = sum(Goniastrea),
    sum_tub = sum(Turbinaria),
    sum_echinopora = sum(Echinopora),
    sum_Galaxea = sum(Galaxea),
    sum_Pavona = sum(Pavona),
    sum_psammacora = sum(Psammacora),
    n = n()
  ) |> write.csv(file = "output/juvenile_reef_taxa.csv")
```


###2.1.2. Other Hard corals

```{r}
Other_HC_juv <- Belt_10m |> select(2:7, 15:23)

Other_HC_juv_long <- Other_HC_juv |> 
 pivot_longer(cols = 7:15, names_to = "Benthos", values_to = "Points")

juv_oth_summary <- Other_HC_juv_long |> 
  group_by(Benthos,Year, Shelf, Reef, Site) |> 
  summarise(
    mean = mean(Points),
    sd = sd(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points),
    belts = n()
  )
juv_oth_summary <- juv_oth_summary |> filter(max > 0)

juv_oth_summary |> 
  group_by(Benthos) |> 
  summarise(
    sites_where_juveniles_family_present_at_least_once = n()
  )
```

Need to explore this and how the other families observed differ across the shelf
```{r}
Belt_10m |> 
  group_by(Shelf, Reef, Year) |> 
  summarise(
    sum_oth_HC = sum(Other_HC),
    sum_Favites = sum(Favites),
    sum_favia = sum(Favia),
    sum_goniastrea = sum(Goniastrea),
    sum_tub = sum(Turbinaria),
    sum_Galaxea = sum(Galaxea),
    belt_transect = n()
  )
```

## 2.2.Spatiotemporal

```{r}
Belt_10m |> 
  group_by(Year) |> 
  summarise(
    min_total = min(Total_juveniles),
    max_total = max(Total_juveniles),
    min_acro = min(Acroporidae),
    max_acro = max(Acroporidae),
    min_poci = min(Pocilloporidae),
    max_poci = max(Pocilloporidae),
    min_por = min(Poritidae),
    max_por = max(Poritidae)
  )
```

```{r}
Belt_10m |> 
  group_by(Shelf) |> 
  summarise(
    min_total = min(Total_juveniles),
    max_total = max(Total_juveniles),
    min_acro = min(Acroporidae),
    max_acro = max(Acroporidae),
    min_poci = min(Pocilloporidae),
    max_poci = max(Pocilloporidae),
    min_por = min(Poritidae),
    max_por = max(Poritidae),
    mean_acro = mean(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    mean_pori = mean(Poritidae),
    n_transects = n()
  )
```

# 3. Benthic Cover
```{r}
PIT |> 
  group_by(Shelf) |> 
  summarise(
    n_PIT = n()
  )
# PIT <- PIT |> 
#   mutate(
#     Total_HC_Cover = Acroporidae+ Pocilloporidae + Poritidae + `Other Hard coral`
#   )
```


```{r}
PIT_long <- PIT |> 
  pivot_longer(cols = c(6:20, Total_HC_Cover), names_to = "Benthos", values_to = "Points")

```

## 3.1. Families

```{r}
Summary_table_PIT <- PIT_long |> 
  group_by(Benthos, Site, Year, Shelf, Reef) |> 
  summarise(
    mean = mean(Points),
    sd = sd(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points),
    PITs = n()
  )

Summary_table_PIT

PIT_long |> 
  group_by(Shelf, Reef, Year, Benthos) |> 
  # filter(Benthos == "Total_HC_Cover") |> 
  summarise(
    mean = mean(Points),
    # sd = sd(Points),
    se = sd(Points)/(n()^0.5),
    # min = min(Points),
    # max = max(Points),
    PITs = n()
  ) |> write.csv(file = "coral_cover_reef_year_mean.csv")
```

```{r}
PIT_long |> 
  group_by(Shelf, Reef, Year, Benthos) |> 
  filter(Benthos == "Acroporidae", Shelf == "Mid") |> 
  summarise(
    mean = mean(Points),
    # sd = sd(Points),
    se = sd(Points)/(n()^0.5),
    # min = min(Points),
    # max = max(Points),
    PITs = n()
  ) #|> write.csv(file = "Acroporidae_mean_se_reef_year.csv")
```

```{r}
PIT_long |> 
  group_by(Shelf, Reef, Year, Benthos) |> 
  filter(Benthos %in% c("Acroporidae", "Pocilloporidae", "Poritidae", "Total_HC_Cover")) |> 
  summarise(
    mean = mean(Points),
    # sd = sd(Points),
    se = sd(Points)/(n()^0.5),
    # min = min(Points),
    # max = max(Points),
    PITs = n()
  ) |> 
  ggplot(aes(y=mean, x=Reef, fill = Shelf))+
  geom_point()+
  geom_errorbar(aes(y=se, x=Reef, fill = Shelf))

PIT_long |> 
  group_by(Shelf, Reef, Year, Benthos) |> 
  filter(Benthos %in% c("Acroporidae", "Pocilloporidae", "Poritidae", "Total_HC_Cover")) |> 
  summarise(
    mean = mean(Points),
    se = sd(Points)/(n()^0.5),
    PITs = n(),
    .groups = "drop"
  ) |> 
  ggplot(aes(x = Year, y = mean, fill = Shelf, colour = Shelf)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(
    aes(ymin = mean - se, ymax = mean + se),
    width = 0.2,
    position = position_dodge(width = 0.5)
  )+
  facet_grid(Benthos~Reef, scale = "free_y")+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  theme(panel.grid = element_blank(),
        text = element_text(size =12)
        )+
  theme_bw()+
  labs(x = "Year", y = "Percent cover (mean ± SE)")+
  scale_x_discrete(
    "Year",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )
```


```{r}
PIT_long |> 
  group_by(Benthos, Shelf, Reef) |> 
  summarise(
    mean = mean(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points),
    PITs = n()
  ) |> 
write.csv(file = "output/PIT_long_reef_total_HC.csv")
```
```{r}
#brown macroalgae
PIT_long |> 
  filter(Benthos == "Brown MA") |> 
  group_by(Benthos, Site, 
           #Year, 
           Shelf, Reef) |> 
  summarise(
    mean = mean(Points),
    sd = sd(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points),
    PITs = n()
  )

PIT_long |> 
  filter(Benthos == "Brown MA",
         Points > 25)
# Brown macroaglae > 50% in 53 transects

53/241 * 100
PIT_long |> 
  filter(Benthos == "Brown MA",
         Points > 25) |> 
  group_by(Site) |> 
  summarise(
    n = n()
  )
```

CCA

```{r}
PIT_long |> 
  filter(Benthos == "CCA",) |> 
  group_by(Site) |> 
  summarise(
    n = n()
  )

```

```{r}
#Turf
PIT_long |> 
  filter(Benthos == "Turf") |> 
  group_by(Benthos, Site, 
           #Year, 
           Shelf, Reef) |> 
  summarise(
    mean = mean(Points),
    sd = sd(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points),
    PITs = n()
  )
```

```{r}
#present / absent
PIT_long |> 
  filter(Points > 0) |> 
  group_by(Site, Benthos, Shelf) |> 
  summarise(
    mean = mean(Points),
    sd = sd(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points),
    PITs = n()
  ) |> 
  filter(
    Benthos == "Soft coral"
    #Benthos == "Sand and Rubble"
    #Benthos == "Green MA"
    #Benthos == "Poritidae"
    #Benthos == "Pocilloporidae"
    #Benthos == "Acroporidae"
    #Benthos == "Brown MA" #all inner shelf sites, 4 mid shelf sites (3 of which are around Lizard)
    #Benthos == "CCA" #CCA was not present at any Inner shelf sites
    )
```


```{r}
Summary_table_PIT |> 
  group_by(Benthos, Site) |> 
  summarise(
    min_mean = min(mean),
    max_mean = max(mean)
  )
```


###3.1.1 Acroporidae

```{r}
PIT_long |> 
  group_by(Benthos,Year, Shelf, Reef, Site) |> 
  summarise(
    mean = mean(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points)
  ) |> 
  filter(Benthos == "Acroporidae") |> 
  arrange(desc(mean))
```

```{r}
PIT_long |> 
  group_by(Benthos,Year, Shelf, Reef, Site) |> 
  summarise(
    mean = mean(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points)
  ) |> 
  filter(Benthos == "Acroporidae") |> 
  arrange(-desc(mean))
```



### 3.1.2. Pocilloporidae
```{r}
PIT_long |> 
  group_by(Benthos,Year, Shelf, Reef, Site) |> 
  summarise(
    mean = mean(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points)
  ) |> 
  filter(Benthos == "Pocilloporidae") |> 
  arrange(desc(mean))
```

```{r}
PIT_long |> 
  group_by(Benthos,Year, Shelf, Reef, Site) |> 
  summarise(
    mean = mean(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points)
  ) |> 
  filter(Benthos == "Pocilloporidae") |> 
  arrange(-desc(mean))
```

###3.1.3. Poritidae

```{r}
PIT_long |> 
  group_by(Benthos,Year, Shelf, Reef, Site) |> 
  summarise(
    mean = mean(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points)
  ) |> 
  filter(Benthos == "Poritidae") |> 
  arrange(desc(mean))
```

```{r}
PIT_long |> 
  group_by(Benthos,Year, Shelf, Reef, Site) |> 
  summarise(
    mean = mean(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points)
  ) |> 
  filter(Benthos == "Poritidae") |> 
  arrange(-desc(mean))
```


##3.2. Other Benthic Groupds
Macroalgae brown

```{r}
# PIT_brown_macro_long <- PIT |> 
#   pivot_longer(cols = 42:56, names_to = "Benthos", values_to = "Points") |> 
#   select(1:7, Benthos, Points)
```

```{r}
# Summary_table_brown_macroalgae <- PIT_brown_macro_long |> 
#   group_by(Benthos,Year, Shelf, Reef, Site) |> 
#   summarise(
#     mean = mean(Points),
#     sd = sd(Points),
#     se = sd(Points)/(n()^0.5),
#     min = min(Points),
#     max = max(Points),
#     PITs = n()
#   )
```

```{r}
# PIT_Coral_Genera <- PIT |> 
#   pivot_longer(cols = 8:37, names_to = "Benthos", values_to = "Points") |> 
#   select(1:7, Benthos, Points)
```

```{r}
Summary_table_coral <- PIT_Coral_Genera |> 
  group_by(Benthos,Year, Shelf, Reef, Site) |> 
  summarise(
    mean = mean(Points),
    sd = sd(Points),
    se = sd(Points)/(n()^0.5),
    min = min(Points),
    max = max(Points),
    PITs = n()
  )
Summary_table_coral
```


## 3.2.Spatiotemporal



