---
title: "06_Benthic Composition"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(tidyr)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)

# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))
```

Models of Benthic Composition

```{r}
PIT_plot <- PIT |> 
  select(Shelf, Year, Reef, Site, Acroporidae, Pocilloporidae, Poritidae, Macroalgae,
         Rubble_and_Sand, `Hard Coral`,
         #`Brown Macroalgae`, `Red Macroalgae`, `Green Macroalgae`, 
         Turf, `Soft Coral`)

PIT_long <- PIT_plot |> pivot_longer(cols = 5:12, names_to = "Benthic Cover", values_to = "Cover")

PIT_long <- PIT_long |> 
  mutate(`Benthic Cover` = factor(`Benthic Cover`, levels = c("Hard Coral", "Acroporidae", "Pocilloporidae","Poritidae", "Rubble_and_Sand", "Macroalgae", "Turf", "Soft Coral")))

PIT_long <- PIT_long|> 
  mutate(
    "percent_cover" = Cover/100
    )
```

Reorder so that Acroporidae, Pocilloporidae, Poritidae, and then Macroalgae, Turf, and then Soft Coral

```{r}
ggplot(data = PIT_long)+
  geom_jitter(aes(x = Year, y = Cover, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.6),
    size = 0.5,
    alpha = 0.5,
    colour = "grey"
  )+
   # Mean points
  stat_summary(aes(x = Year, y = Cover, colour = Shelf),
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.6),
    size = 2.25
  ) +
  # Error bars (mean ± SE)
  stat_summary(aes(x = Year, y = Cover, colour = Shelf),
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.6),
    width = 0.45
  ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  theme_bw()+
  theme(text =element_text(size =8),
        title = element_text(size = 8),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        legend.position = "bottom")+
  labs(y = "Percent Cover (%)")+
  facet_wrap(~`Benthic Cover`, scale = "free_y", ncol = 2)+
  scale_y_log10()
```

## Macroalgae Model

```{r}
nbin_macro_i <- glmmTMB(Macroalgae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_macro_i)
testResiduals(sim1) # fails KS test (deviation significant), and has significant outliers

nbin_macro_f <- glmmTMB(Macroalgae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_macro_f)
testResiduals(sim1) # Fits the data well

zbin_macro_i <- glmmTMB(Macroalgae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", ziformula = ~1, data = PIT_plot)
sim1 <- simulateResiduals(zbin_macro_i) # now passes KS test
testResiduals(sim1)
```
```{r}
model.sel(nbin_macro_f, zbin_macro_i)
```
```{r}
emm_shelf <- emmeans(zbin_macro_i, ~Shelf | Year, type = "response")
pairs(emm_shelf)
```


```{r}
emm_year_response <- emmeans(zbin_macro_i, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

### Figure
```{r}
macro_model_emm <- zbin_macro_i|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()
```

```{r echo=FALSE}
  ggplot()+
  geom_pointrange(data = macro_model_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(y = expression("Predicted macroalgae cover ( % / " * 10~m^{2} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1),
        text = element_text(size =9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_y_log10(
    breaks = scales::breaks_log(n = 4),
    labels = scales::label_number()
  )
```

## Rubble and Sand model

```{r}
nbin_sand_i <- glmmTMB(Rubble_and_Sand ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_sand_i)
testResiduals(sim1) #passes DHARMA

nbin_sand_f <- glmmTMB(Rubble_and_Sand ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_sand_f)
testResiduals(sim1) # passes DHARMA
```

```{r}
model.sel(nbin_sand_f, nbin_sand_i)
```


## Turf Model

```{r}
nbin_turf_i <- glmmTMB(Turf ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_turf_i) # passes DHARMA
testResiduals(sim1)

nbin_turf_f <- glmmTMB(Turf ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_turf_f) #passes DHARMA
testResiduals(sim1)
```

```{r}
model.sel(nbin_turf_f, nbin_turf_i)
```

## Acroporidae cover

```{r}
#interaction
nbin_Acro_i <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_Acro_i) #passes DHARMA
testResiduals(sim1)

#fixed
nbin_Acro_f <- glmmTMB(Acroporidae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_Acro_f) #passes DHARMA
testResiduals(sim1)
```

```{r}
model.sel(nbin_Acro_f, nbin_Acro_i)
```

### Tukey HSD tests

```{r}
emm_shelf <- emmeans(nbin_Acro_i, ~Shelf | Year, type = "response")
pairs(emm_shelf)
```

### Figure
```{r}
Acro_model_emm <- nbin_Acro_i|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()
```
```{r echo=FALSE}
  ggplot()+
  geom_pointrange(data = Acro_model_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(y = expression("Predicted Acroporidae cover ( % / " * 10~m^{2} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1),
        text = element_text(size =9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_y_log10(
    breaks = scales::breaks_log(n = 4),
    labels = scales::label_number()
  )
```

## Pocilloporidae cover

```{r}
#Interaction
nbin_poci_i <- glmmTMB(Pocilloporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_poci_i) # passes
testResiduals(sim1)

#Fixed
nbin_poci_f <- glmmTMB(Pocilloporidae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_poci_f) #passes
testResiduals(sim1)

#Interaction
pois_poci_i <- glmmTMB(Pocilloporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "poisson", data = PIT_plot)
sim1 <- simulateResiduals(pois_poci_i) # passes
testResiduals(sim1)


#Fixed
pois_poci_f <- glmmTMB(Pocilloporidae ~ Shelf+Year + (1 | Reef/Site), 
        family = "poisson", data = PIT_plot)
sim1 <- simulateResiduals(pois_poci_f) #passes
testResiduals(sim1)
```

```{r}
model.sel(nbin_poci_f, nbin_poci_i, pois_poci_f, pois_poci_i)
```
Not sure why the nbin don't have AICc values for Pocilloporidae.

### Tukey HSD

```{r}
emm_shelf <- emmeans(pois_poci_i, ~Shelf | Year, type = "response")
pairs(emm_shelf)
```

### Figure

```{r}
poci_model_emm <- pois_poci_i|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()
```

Note in the poisson output the repsonse is replaced with rate?

```{r echo=FALSE}
  ggplot()+
  geom_pointrange(data = poci_model_emm, aes(x=Year,
             y=rate, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(y = expression("Predicted Pocilloporidae cover ( % / " * 10~m^{2} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1),
        text = element_text(size =9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_y_log10(
    breaks = scales::breaks_log(n = 4),
    labels = scales::label_number()
  )
```

## Poritidae cover

```{r}
nbin_pori_i <- glmmTMB(Poritidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)

sim1 <- simulateResiduals(nbin_pori_i)
testResiduals(sim1)

nbin_pori_f <- glmmTMB(Poritidae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_pori_f)
testResiduals(sim1)
```
```{r}
model.sel(nbin_pori_f, nbin_pori_i)
```
```{r}
summary(nbin_pori_f)
```


## Soft Coral

```{r}
nbin_sc_i <- glmmTMB(`Soft Coral` ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)

sim1 <- simulateResiduals(nbin_sc_i)
testResiduals(sim1) #passes tests

nbin_sc_f <- glmmTMB(`Soft Coral` ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_sc_f)
testResiduals(sim1) #passes tests
```

```{r}
model.sel(nbin_sc_f, nbin_sc_i)
```

```{r}
summary(nbin_sc_f)
```

Figure
```{r}
sc_model_emm <- nbin_sc_f|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()
```
```{r echo=FALSE}
  ggplot()+
  geom_pointrange(data = sc_model_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(y = expression("Predicted soft coral cover ( % / " * 10~m^{2} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1),
        text = element_text(size =9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  scale_y_log10(
    breaks = scales::breaks_log(n = 4),
    labels = scales::label_number()
  )
```

## NMDS and PERMANOVA
```{r}
# Step 1: Create Matrix
site_by_PIT <- PIT %>%
  select(
    Year, Reef, Shelf, Site, Transect, 107:118
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

site_by_PIT <- site_by_PIT |>                                                         # ensure only vegetated tiles included 
  group_by(Reef, Year, Shelf) %>%               # this was missing in your code
  summarise(across(`Hard Coral`:Rubble_and_Sand, mean), .groups = "drop")

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  select(Shelf, Year, Reef)

#create species matrix
PIT_matrix_filtered <- site_by_PIT |> 
  select(4:15)

```



Normal distance matrix
```{r}
# create a distance matrix
PIT_dist <- vegdist(PIT_matrix_filtered, method = "bray")

# With respect to Shelf
betadisper(PIT_dist, metadata_PIT$Shelf)
anova(betadisper(PIT_dist, metadata_PIT$Shelf)) # Significant was 1.561e-05 ***
```

```{r}
PIT_hellinger <- decostand(PIT_matrix_filtered, method = "hellinger")
```


Check dispersion
```{r}
# create a distance matrix
PIT_dist <- vegdist(PIT_hellinger, method = "euclidean")

# With respect to Shelf
betadisper(PIT_dist, metadata_PIT$Shelf)
anova(betadisper(PIT_dist, metadata_PIT$Shelf)) # oh no, thats not great. Significance was 1.08e-07 ***
```

```{r}
set.seed(999)     # for reproducibility

nmds <- metaMDS(PIT_matrix_filtered, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, metadata_PIT[, c("Year", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year == "2024-25" ~ "24/25",
      Year == "2023-24" ~ "23/24",
      Year == "2022-23" ~ "22/23",
      Year == "2021-22" ~ "21/22")
  )

ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 1.5, alpha = 0.8) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "top",
         text = element_text(size = 8),
        title = element_text(size = 8)) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
             x = min(scores$NMDS1),
             y = min(scores$NMDS2),              
             label = paste0("Stress = ", round(nmds$stress, 3)),
              hjust = -2, vjust = 0.5, size = 3)+
  scale_x_continuous(limits = c(-1, 1.45))+
  scale_y_continuous(limits = c(-1.3, 1.1))
```

```{r}
# Creating Vectors for driving the patterns
Benthic_envfit <- envfit(
  nmds,
  PIT_matrix_filtered,
  permutations = 999
) 

#Saving the vectors as a data frame
Benthic_envfit <- as.data.frame(scores(Benthic_envfit, display = "vectors"))

Benthic_envfit <- Benthic_envfit |> 
  rownames_to_column(var = "label_category")
```

```{r}
ggplot()+
  geom_segment(data = Benthic_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Benthic_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_category),
                  inherit.aes = FALSE,
                  size = 3,
                  box.padding = 0.5, 
                  min.segment.length = Inf,
                  nudge_x = -0.1,
                  nudge_y = 0.1)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year"
    )
```

```{r}
permanova_result <- adonis2(PIT_matrix_filtered ~ Shelf,
  data = site_by_PIT,
  method = "bray",
  by = "margin",
  permutations = 999)

print(permanova_result)
```