---
title: "06_Benthic Composition"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(tidyr)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)

# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))
```

Models of Benthic Composition

```{r}
PIT_plot <- PIT |> 
  select(Shelf, Year, Reef, Site, Acroporidae, Pocilloporidae, Poritidae, Macroalgae,
         Rubble_and_Sand, `Hard Coral`,
         #`Brown Macroalgae`, `Red Macroalgae`, `Green Macroalgae`, 
         Turf, `Soft Coral`)

PIT_long <- PIT_plot |> pivot_longer(cols = 5:12, names_to = "Benthic Cover", values_to = "Cover")

PIT_long <- PIT_long |> 
  mutate(`Benthic Cover` = factor(`Benthic Cover`, levels = c("Hard Coral", "Acroporidae", "Pocilloporidae","Poritidae", "Rubble_and_Sand", "Macroalgae", "Turf", "Soft Coral")))

PIT_long <- PIT_long|> 
  mutate(
    "percent_cover" = Cover/100
    )
```

Reorder so that Acroporidae, Pocilloporidae, Poritidae, and then Macroalgae, Turf, and then Soft Coral

```{r}
ggplot(data = PIT_long)+
  geom_jitter(aes(x = Year, y = Cover, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.6),
    size = 0.5,
    alpha = 0.5,
    colour = "grey"
  )+
   # Mean points
  stat_summary(aes(x = Year, y = Cover, colour = Shelf),
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.6),
    size = 2.25
  ) +
  # Error bars (mean ± SE)
  stat_summary(aes(x = Year, y = Cover, colour = Shelf),
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.6),
    width = 0.45
  ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  theme_bw()+
  theme(text =element_text(size =8),
        title = element_text(size = 8),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        legend.position = "bottom")+
  labs(y = "Percent Cover (%)")+
  facet_wrap(~`Benthic Cover`, scale = "free_y", ncol = 2)
```

## Macroalgae Model

```{r}
nbin_macro_i <- glmmTMB(Macroalgae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_macro_i)
testResiduals(sim1) # fails KS test (deviation significant), and has significant outliers

nbin_macro_f <- glmmTMB(Macroalgae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_macro_f)
testResiduals(sim1) # Fits the data well

zbin_macro_i <- glmmTMB(Macroalgae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", ziformula = ~1, data = PIT_plot)
sim1 <- simulateResiduals(zbin_macro_i) # now passes KS test
testResiduals(sim1)
```
```{r}
model.sel(nbin_macro_f, zbin_macro_i)
```
```{r}
emm_shelf <- emmeans(zbin_macro_i, ~Shelf | Year, type = "response")
pairs(emm_shelf)
```


```{r}
emm_year_response <- emmeans(zbin_macro_i, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

### Figure
```{r}
macro_model_emm <- zbin_macro_i|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()
```

```{r echo=FALSE}
  ggplot()+
  geom_pointrange(data = macro_model_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(y = expression("Predicted macroalgae cover ( % / " * 10~m^{2} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1),
        text = element_text(size =9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
```

## Rubble and Sand model

```{r}
nbin_sand_i <- glmmTMB(Rubble_and_Sand ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_sand_i)
testResiduals(sim1) #passes DHARMA

nbin_sand_f <- glmmTMB(Rubble_and_Sand ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_sand_f)
testResiduals(sim1) # passes DHARMA
```

```{r}
model.sel(nbin_sand_f, nbin_sand_i)
```


## Turf Model

```{r}
nbin_turf_i <- glmmTMB(Turf ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_turf_i) # passes DHARMA
testResiduals(sim1)

nbin_turf_f <- glmmTMB(Turf ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_turf_f) #passes DHARMA
testResiduals(sim1)
```

```{r}
model.sel(nbin_turf_f, nbin_turf_i)
```

## Acroporidae cover

```{r}
#interaction
nbin_Acro_i <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_Acro_i) #passes DHARMA
testResiduals(sim1)

#fixed
nbin_Acro_f <- glmmTMB(Acroporidae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_Acro_f) #passes DHARMA
testResiduals(sim1)
```

```{r}
model.sel(nbin_Acro_f, nbin_Acro_i)
```

### Tukey HSD tests

```{r}
emm_shelf <- emmeans(nbin_Acro_i, ~Shelf | Year, type = "response")
pairs(emm_shelf)
```

### Figure
```{r}
Acro_model_emm <- nbin_Acro_i|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()
```
```{r echo=FALSE}
  ggplot()+
  geom_pointrange(data = Acro_model_emm,aes(x=Year,
             y=response, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(y = expression("Predicted Acroporidae cover ( % / " * 10~m^{2} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1),
        text = element_text(size =9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
```

## Pocilloporidae cover

```{r}
#Interaction
nbin_poci_i <- glmmTMB(Pocilloporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_poci_i) # passes
testResiduals(sim1)

#Fixed
nbin_poci_f <- glmmTMB(Pocilloporidae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_poci_f) #passes
testResiduals(sim1)

#Interaction
pois_poci_i <- glmmTMB(Pocilloporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "poisson", data = PIT_plot)
sim1 <- simulateResiduals(pois_poci_i) # passes
testResiduals(sim1)


#Fixed
pois_poci_f <- glmmTMB(Pocilloporidae ~ Shelf+Year + (1 | Reef/Site), 
        family = "poisson", data = PIT_plot)
sim1 <- simulateResiduals(pois_poci_f) #passes
testResiduals(sim1)
```

```{r}
model.sel(nbin_poci_f, nbin_poci_i, pois_poci_f, pois_poci_i)
```
Not sure why the nbin don't have AICc values for Pocilloporidae.

### Tukey HSD

```{r}
emm_shelf <- emmeans(pois_poci_i, ~Shelf | Year, type = "response")
pairs(emm_shelf)
```

### Figure

```{r}
poci_model_emm <- pois_poci_i|> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame()
```

Note in the poisson output the repsonse is replaced with rate?

```{r echo=FALSE}
  ggplot()+
  geom_pointrange(data = poci_model_emm, aes(x=Year,
             y=rate, 
             color=Shelf,ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.8))+
  theme_classic()+
  labs(y = expression("Predicted Pocilloporidae cover ( % / " * 10~m^{2} * ")"))+
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1),
        text = element_text(size =9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
```

## Poritidae cover

```{r}
nbin_pori_i <- glmmTMB(Poritidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)

sim1 <- simulateResiduals(nbin_pori_i)
testResiduals(sim1)

nbin_pori_f <- glmmTMB(Poritidae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_pori_f)
testResiduals(sim1)
```
```{r}
model.sel(nbin_pori_f, nbin_pori_i)
```
```{r}
summary(nbin_pori_f)
```


## Soft Coral

```{r}
nbin_sc_i <- glmmTMB(`Soft Coral` ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)

sim1 <- simulateResiduals(nbin_sc_i)
testResiduals(sim1)

nbin_sc_f <- glmmTMB(`Soft Coral` ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_sc_f)
testResiduals(sim1)
```

```{r}
model.sel(nbin_sc_f, nbin_sc_i)
```

```{r}
summary(nbin_sc_f)
```


