---
title: "Benthic Survey Multivariate analysis"
author: "Monique White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)

# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(prop_acro),
    sd_aco = sd(prop_acro),
    mean_poci = mean(prop_poci),
    sd_poci = sd(prop_poci),
    mean_por = mean(prop_pori),
    sd_por = sd(prop_pori),
    mean_macroalgae = mean(Prop_macroalgae),
    sd_macroalgae = sd(Prop_macroalgae),
    mean_sand =mean(Rubble_and_Sand),
    sd_sand = sd(Rubble_and_Sand),
    PIT = n()
  )

Combined <- left_join(Coral_settlement_nGBR, join)

## Total recuits
total_recruits <- Coral_settlement_nGBR %>%
  select(Year_Collected, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 5:8, names_to = "Family", values_to = "Total")

#Order so that Other is last in Family
total_recruits <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, Reef, Site, Obs, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 6:9, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))

total_recruits$Year_Collected <- as.factor(total_recruits$Year_Collected)



## Family recruitment data
family_total_recruits <- Coral_settlement_nGBR %>%
  select(2:13) %>% 
  pivot_longer(cols = 9:12, names_to = "Family", values_to = "Total") |> 
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))

```



```{r}
# Step 1: Create site-by-species matrix with Site_ID
site_by_PIT <- PIT %>%
  select(
    Year, Reef, Shelf, Site, Transect, 107:118
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  select(Site_ID, Shelf, Year, Reef, Site)

# Step 3: Species matrix
PIT_matrix <- site_by_PIT %>%
  select(
    Site_ID,
    6:17
  )
```

```{r}
PIT_matrix <- PIT_matrix |> 
  select(Site_ID, Acroporidae, Pocilloporidae, Poritidae, `Brown Macroalgae`, `Red Macroalgae`, `Green Macroalgae`, Turf, `Soft Coral`, Rubble_and_Sand)

# Step 4: Remove all-zero rows BEFORE rownames
PIT_matrix_filtered <- PIT_matrix %>%
  filter(rowSums(across(-Site_ID)) > 0) %>%
  column_to_rownames("Site_ID")

# Step 5: Match metadata
PIT_metadata_filtered <- metadata_PIT %>%
  filter(Site_ID %in% rownames(PIT_matrix_filtered)) %>%
  column_to_rownames("Site_ID")
```

```{r}
# Step 6: PERMANOVA
PIT_adonis_result1 <- adonis2(
  PIT_matrix_filtered ~ Shelf + Year, #where shelf and year are fixed effects but without interaction
  data = PIT_metadata_filtered,
  method = "bray",
  by = "term",
  permutations = 999
)
print(PIT_adonis_result1)

PIT_adonis_result <- adonis2(
  PIT_matrix_filtered ~ Shelf * Year, #includes interaction term
  data = PIT_metadata_filtered,
  method = "bray",
  by = "term",
  permutations = 999
)
print(PIT_adonis_result)

# interaction term was significant so i think it should be included.

set.seed(999)  # reproducibility

PIT_nmds <- metaMDS(
  PIT_matrix_filtered,
  distance = "bray",
  k = 2,            # 2D solution
  trymax = 100,
  autotransform = FALSE
)
```

After PERMANOVA, run ANOSIM
Check dispersion is correct for this

```{r}
# create a distance matrix
PIT_dist <- vegdist(PIT_matrix_filtered, method = "bray")

# With respect to Shelf
betadisper(PIT_dist, metadata_PIT$Shelf)
anova(betadisper(PIT_dist, metadata_PIT$Shelf))
```

Well this doesn't meet the assumptions for either shelf or for year.

```{r}
# with respect to year
betadisper(PIT_dist, metadata_PIT$Year)
anova(betadisper(PIT_dist, metadata_PIT$Year))

```


This article notes that ANOSIM is fully non-parametric and cannot test interactions; instead it “provides a robust, comparable and globally interpretable measure of magnitude of overall community change associated with each factor, having excised any possible effect from the factor(s) it is crossed with, irrespective of whether the factors interact or not” (p. 911). https://uw.pressbooks.pub/appliedmultivariatestatistics/chapter/anosim/#:~:text=Analysis%20of%20similarities%20(ANOSIM)%20is%20a%20non%2Dparametric,significant%20differences%20between%20two%20or%20more%20groups.


```{r}
PIT_nmds$stress
```
Stress is 0.1409

```{r}
stressplot(PIT_nmds)
```

Extract Scores

```{r}
nmds_scores_PIT <- as.data.frame(scores(PIT_nmds, display = "sites"))

nmds_scores_PIT$Site_ID <- rownames(nmds_scores_PIT)

nmds_scores_PIT <- nmds_scores_PIT %>%
  left_join(
    PIT_metadata_filtered %>% 
      rownames_to_column("Site_ID"),
    by = "Site_ID"
  )
nmds_scores_PIT <- nmds_scores_PIT |> 
mutate(
    Deployment =
    case_when(
      Year == "2024-25" ~ "24/25",
      Year == "2023-24" ~ "23/24",
      Year == "2022-23" ~ "22/23",
      Year == "2021-22" ~ "21/22")
  )
```

```{r}

Figure6A <- ggplot() +
  geom_point(data = nmds_scores_PIT, aes(NMDS1, NMDS2, colour = Shelf), size = 2, alpha = 0.8) +
  stat_ellipse(data = nmds_scores_PIT, aes(NMDS1, NMDS2, colour = Shelf), type = "t") +
  theme_bw() +
  coord_equal() +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  theme(text =element_text(size =8),
        title = element_text(size = 8),
        legend.position = "bottom",
        panel.grid = element_blank())+
  facet_wrap(~Deployment, ncol =2)+
  scale_x_continuous(limits = c(-1.35, 2.5))+
  scale_y_continuous(limits = c(-1.55, 1.5))
  #labs(x = "")+
  # scale_shape_discrete(
  #   "Deployment",
  #   labels = c("2022" = "21/22",
  #              "2023" = "22/23",
  #              "2024" = "23/24",
  #              "2025" = "24/25",
  #              "2026" = "25/26"))

Figure6A
```

```{r}
Figure6a <- nmds_scores_PIT |> filter(Year == "2021-22") |> 
  ggplot() +
  geom_point(aes(NMDS1, NMDS2, colour = Shelf), size = 2, alpha = 0.8) +
  stat_ellipse(aes(NMDS1, NMDS2, colour = Shelf), type = "t") +
  theme_classic() +
  coord_equal() +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  theme(text =element_text(size =8),
        title = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.position = "bottom")+
  labs(x = "",, title = "21/22")+
  scale_x_continuous(limits = c(-1.1, 2.32))+
  scale_y_continuous(limits = c(-1.33, 1.63))
Figure6a
  #labs(x = "")

Figure6b <- nmds_scores_PIT |> filter(Year == "2022-23") |> 
  ggplot() +
  geom_point(aes(NMDS1, NMDS2, colour = Shelf), size = 2, alpha = 0.8) +
  stat_ellipse(aes(NMDS1, NMDS2, colour = Shelf), type = "t") +
  theme_classic() +
  coord_equal() +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  theme(text =element_text(size =8),
        title = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.position = "bottom")+
  labs(x = "", 
       #y = "",
       title = "22/23")+
  scale_x_continuous(limits = c(-1.1, 2.32))+
  scale_y_continuous(limits = c(-1.33, 1.63))
Figure6b

Figure6c <- nmds_scores_PIT |> filter(Year == "2023-24") |> 
  ggplot() +
  geom_point(aes(NMDS1, NMDS2, colour = Shelf), size = 2, alpha = 0.8) +
  stat_ellipse(aes(NMDS1, NMDS2, colour = Shelf), type = "t") +
  theme_classic() +
  coord_equal() +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  theme(text =element_text(size =8),
        title = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.position = "bottom")+
  labs(title = "23/24",
       x= ""
       )+
  scale_x_continuous(limits = c(-1.1, 2.32))+
  scale_y_continuous(limits = c(-1.33, 1.63))
Figure6c

Figure6d <- nmds_scores_PIT |> filter(Year == "2024-25") |> 
  ggplot() +
  geom_point(aes(NMDS1, NMDS2, colour = Shelf), size = 2, alpha = 0.8) +
  stat_ellipse(aes(NMDS1, NMDS2, colour = Shelf), type = "t") +
  theme_classic() +
  coord_equal() +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  theme(text =element_text(size =8),
        title = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.position = "bottom")+
  labs(y = "", title = "24/25")+
  scale_x_continuous(limits = c(-1.1, 2.32))+
  scale_y_continuous(limits = c(-1.33, 1.63))
Figure6d
```


```{r}
PIT_envfit <- envfit(
  PIT_nmds,
  PIT_matrix_filtered,
  permutations = 999
) 

PIT_envfit <- as.data.frame(scores(PIT_envfit, display = "vectors"))


PIT_envfit <- PIT_envfit |> 
  rownames_to_column(var = "label_Benthos")


Figure6B <- ggplot()+
  geom_text_repel(
  data = PIT_envfit,
  aes(x = NMDS1, y = NMDS2, label = label_Benthos),
  inherit.aes = FALSE,
  size = 3,
  nudge_x = ifelse(PIT_envfit$NMDS1 > 0, 0.15, -0.2),
  nudge_y = ifelse(PIT_envfit$NMDS2 > 0, 0.15, -0.1),
  force = 1,
  direction = "both"
)+
  geom_segment(data = PIT_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(1/3, "picas")),
               linetype = "solid",
               linewidth = 1)+
   theme_classic() +
  coord_equal() +
  theme(text =element_text(size =8),
        title = element_text(size = 8),
        legend.position = "right")+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 8),
        title = element_text(size = 8)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2"
    ) +
  # labs(y = "")+
   scale_x_continuous(limits = c(-1.35, 2.5))+
  scale_y_continuous(limits = c(-1.55, 1.5))#+
    # annotate("text",
    #          x = min(scores_hellinger$NMDS1),
    #          y = min(scores_hellinger$NMDS2),              
             #label = paste0("Stress = ", round(nmds_hellinger$stress, 3)),
              #hjust = 0, vjust = 0, size = 4
             # )+
  # scale_x_continuous(limits = c(-1.75, 3.75))+
  # scale_y_continuous(limits = c(-1.5, 1.75))

Figure6A
Figure6B

Right <- ggarrange(Figure6a, Figure6b, Figure6c, Figure6d, ncol = 2, 
                   nrow = 2,
          labels = c("C", "D", "E", "F"),
          #heights = c(4, 1),
          common.legend = TRUE, legend = "right"
          )
Left <- ggarrange(Figure6A, Figure6B, ncol = 1, 
          labels = c("A", "B"),
          #heights = c(4, 1),
          common.legend = TRUE, legend = "none"
          )

ggarrange(Left, Right, widths = c(1, 1.5), 
          common.legend = TRUE, legend = "right"
          ) |> ggsave(file = "Figure6.png",
                      height = 200,
                      width = 169,
                      unit = "mm")

ggarrange(Figure6A, Figure6B, ncol = 1, 
          labels = c("A", "B"),
          heights = c(1.1, 1),
          common.legend = TRUE, legend = "right"
          )|> ggsave(file = "Figure6.png",
                      height = 220,
                      width = 150,
                      unit = "mm")

ggarrange(Right, Figure6B,  ncol =1, nrow = 2, 
          common.legend = TRUE, legend = "right"
          ) |> ggsave(file = "Figure61.png",
                      height = 225,
                      width = 169,
                      unit = "mm")
```

```{r}
PIT_plot <- PIT |> 
  select(Shelf, Year, Reef, Site, Acroporidae, Pocilloporidae, Poritidae, Macroalgae,
         Rubble_and_Sand, `Hard Coral`,
         #`Brown Macroalgae`, `Red Macroalgae`, `Green Macroalgae`, 
         Turf, `Soft Coral`)

PIT_long <- PIT_plot |> pivot_longer(cols = 5:12, names_to = "Benthic Cover", values_to = "Cover")

PIT_long <- PIT_long |> 
  mutate(`Benthic Cover` = factor(`Benthic Cover`, levels = c("Hard Coral", "Acroporidae", "Pocilloporidae","Poritidae", "Rubble_and_Sand", "Macroalgae", "Turf", "Soft Coral")))

PIT_long <- PIT_long|> 
  mutate(
    "percent_cover" = Cover/100
    )
```

Reorder so that Acroporidae, Pocilloporidae, Poritidae, and then Macroalgae, Turf, and then Soft Coral

```{r}
ggplot(data = PIT_long)+
  geom_jitter(aes(x = Year, y = Cover, group = Shelf),
    position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.6),
    size = 0.5,
    alpha = 0.5,
    colour = "grey"
  )+
   # Mean points
  stat_summary(aes(x = Year, y = Cover, colour = Shelf),
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.6),
    size = 2.25
  ) +
  # Error bars (mean ± SE)
  stat_summary(aes(x = Year, y = Cover, colour = Shelf),
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.6),
    width = 0.45
  ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  theme_bw()+
  theme(text =element_text(size =8),
        title = element_text(size = 8),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        legend.position = "bottom")+
  labs(y = "Percent Cover (%)")+
  facet_wrap(~`Benthic Cover`, scale = "free_y", ncol = 2)
```

Macroalgae Model
```{r}
nbin_macro_i <- glmmTMB(Macroalgae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_macro_i)
testResiduals(sim1)

nbin_macro_f <- glmmTMB(Macroalgae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_macro_f)
testResiduals(sim1)
```
```{r}
model.sel(nbin_macro_f, nbin_macro_i)
```

Rubble and Sand model
```{r}
nbin_sand_i <- glmmTMB(Rubble_and_Sand ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_sand_i)
testResiduals(sim1)

nbin_sand_f <- glmmTMB(Rubble_and_Sand ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_sand_f)
testResiduals(sim1)
```

```{r}
model.sel(nbin_sand_f, nbin_sand_i)
```


Turf Model
```{r}
nbin_turf_i <- glmmTMB(Turf ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_turf_i)
testResiduals(sim1)

nbin_turf_f <- glmmTMB(Turf ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_turf_f)
testResiduals(sim1)
```

```{r}
model.sel(nbin_turf_f, nbin_turf_i)
```

Acroporidae cover
```{r}
nbin_Acro_i <- glmmTMB(Acroporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)

sim1 <- simulateResiduals(nbin_Acro_i)
testResiduals(sim1)

nbin_Acro_f <- glmmTMB(Acroporidae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_Acro_f)
testResiduals(sim1)
```

```{r}
model.sel(nbin_Acro_f, nbin_Acro_i)
```

Pocilloporidae cover

```{r}
nbin_poci_i <- glmmTMB(Pocilloporidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)

sim1 <- simulateResiduals(nbin_poci_i)
testResiduals(sim1)

nbin_poci_f <- glmmTMB(Pocilloporidae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_poci_f)
testResiduals(sim1)
```

```{r}
model.sel(nbin_poci_f, nbin_poci_i)
```

Poritidae cover

```{r}
nbin_pori_i <- glmmTMB(Poritidae ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)

sim1 <- simulateResiduals(nbin_pori_i)
testResiduals(sim1)

nbin_pori_f <- glmmTMB(Poritidae ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_pori_f)
testResiduals(sim1)
```
```{r}
model.sel(nbin_pori_f, nbin_pori_i)
```
```{r}
summary(nbin_pori_f)
```


Soft Coral

```{r}
nbin_sc_i <- glmmTMB(`Soft Coral` ~ Shelf*Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)

sim1 <- simulateResiduals(nbin_sc_i)
testResiduals(sim1)

nbin_sc_f <- glmmTMB(`Soft Coral` ~ Shelf+Year + (1 | Reef/Site), 
        family = "nbinom2", data = PIT_plot)
sim1 <- simulateResiduals(nbin_sc_f)
testResiduals(sim1)
```

```{r}
model.sel(nbin_sc_f, nbin_sc_i)
```
```{r}
summary(nbin_sc_f)
```

