---
title: "Data_Analysis_Thesis"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
```

Clean data

```{r message=FALSE}
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "dislodged",
    TRUE ~ notes
  ))

```

# Determining if loose or dettached tiles are significantly different to other tiles within a site.

```{r}
Coral_settlement_nGBR |> 
  group_by(Reef,notes) |> summarise(n= n())

Coral_settlement_nGBR |> 
  ggplot(aes(x=Year, y=total, colour = notes))+
  geom_jitter()+
  geom_boxplot()+
  facet_wrap(~Site, scale = "free_y", ncol = 3)
```

```{r}
dettached_mod <- glmmTMB(total ~ notes + Year_Collected + (1 | Reef/Site), 
      family = "nbinom2", data = Coral_settlement_nGBR)

sim_mod <- simulateResiduals(dettached_mod)
testResiduals(sim_mod)

summary(dettached_mod)
```

Does appear that dislodged tiles are significantly lower than other tiles when reef and site are a random effect.. maybe they need to be a fixed effect.

The ones 
```{r}
#remove dislodged
Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(is.na(notes) | trimws(notes) == "") |>
  group_by(Year..Jan., Reef, Site) |>
  mutate(
    Tile = row_number()
  ) |>
  filter(Tile <= 10) |>
  ungroup()

# When I do this and summarise I have 71 sites, not 72. Unsure if one whole site was completely dislodged.

```


```{r echo=FALSE}
sample_design <- Coral_settlement_nGBR |> 
  group_by(Year, Reef, Site) |> 
  summarise(n = n()) |> 
  arrange(desc(n)) 
```


```{r echo=FALSE, message=FALSE}
total_recruits <- Coral_settlement_nGBR %>%
  select(Year_Collected, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 5:8, names_to = "Family", values_to = "Total")

#Order so that Other is last in Family
total_recruits <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, Reef, Site, Obs, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 6:9, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))
```

```{r}
total_recruits$Year_Collected <- as.factor(total_recruits$Year_Collected)
```

## Exploratory analysis

```{r}
Fig.1<-  Coral_settlement_nGBR |>  
  ggplot(aes(x=Year_Collected, y=total, colour = Shelf))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Recruits (" * tile^{-1}*")" ),
  x = "Deployment",
  colour = "Shelf position")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(~Shelf)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )
```

Fig 1

```{r}
Coral_settlement_nGBR |>  
  ggplot(aes(x=Year_Collected, y=total, colour = notes))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Recruits (" * tile^{-1}*")" ),
  x = "Deployment",
  colour = "Notes")+
  facet_wrap(~Shelf)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )
```


```{r Detect Outliers, echo=FALSE, fig.height=8, fig.width=6.5}
p.1 <- total_recruits %>% 
  group_by(Year_Collected, Reef, Family, Shelf)  %>% 
  ggplot(aes(x=Family, y=Total, colour = Shelf))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Recruits (" * Reef^{-1} ~ Year^{-1} * ")"),
  x = "Family",
  colour = "Shelf Position")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))

p.2 <- total_recruits %>% 
  group_by(Year_Collected, Reef, Family, Shelf) %>%
  summarise(average_no_recruits = mean(Total)) %>% 
  ggplot(aes(x=Family, y=(average_no_recruits), colour = Shelf))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Mean Recruits (" * Reef^{-1} ~ Year^{-1} * ")"),
  x = "Family",
  colour = "Shelf Position"
) +
scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))

ggarrange(p.1, p.2, ncol = 1)
```

```{r}
total_recruits %>% 
  group_by(Year_Collected, Reef, Family, Shelf)  %>% 
  ggplot(aes(x=Family, y=Total, colour = Shelf))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Recruits (" *Tile^{-1} ~ Year^{-1} * ")"),
  x = "Family",
  colour = "Shelf Position")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(~Year_Collected, scales = "free_y")
```



```{r message=FALSE, echo=FALSE}
total_recruits %>% 
  group_by(Family, Shelf, Year_Collected, Reef) %>%
  summarise(average = mean(Total),
            se = sd(Total)/ sqrt(n()),
            max = max(Total),
            min = min(Total)) %>% 
  arrange(desc(average)) %>% 
  print(n=20)
```


# Models

## Question one. is the total number of recruits different across the shelf?

### Use GLMM for total recruits

Run generalised linear models with the Poisson, normal binomail and the zero inflated normal bionomial.
Total is the dependent variable (the count of coral recruits at a given sampling unit)
Shelf position is a fixed effect predictor, it is used to test if the average coral settlement differs between shelf positions
Random effects: 
a nested random effect, modelling the variations between reefs, and variation between sites nested within reefs

***Check the levels of grouping variables***

```{r check levels of grouping variables, echo=FALSE}
Coral_settlement_nGBR %>% count(Reef, Site)
```

All reefs have multiple sites, so nesting should work.

```{r warning=FALSE, echo=TRUE}
pois_total_yr <- glmmTMB(total ~ Shelf + Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = Coral_settlement_nGBR)

nbin_total_yr <- glmmTMB(total ~ Shelf + Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

zbin_total_yr <- glmmTMB(total ~ Shelf + Year_Collected + (1 | Reef/Site),
                      ziformula = ~1,
        family = "nbinom2", data = Coral_settlement_nGBR)

AIC(pois_total_yr, nbin_total_yr, zbin_total_yr) #nbin has the lowest AIC
```


normal binomial model has the lowest AIC so I will get the summary of that model.
```{r}
Coral_settlement_nGBR |> 
group_by(Reef, Site) |> 
  summarise(count = n())
```


```{r echo=FALSE}
#GLMM negative binomial for total recruits
summary(nbin_total_yr)
```

2023 and 2024 are significantly different to 2022.  2025 is not significantly different to 2022.  Outer shelf is significantly different to inner shelf (p<0.01),   Mid shelf is significantly different to inner shelf reefs (p<0.05).  Intercept is also significant: not quite sure what that means in a biological framework.  When making a figure should I then be using summary statistics or raw count data? Because the large number of zeros will make it difficult to do a boxplot of the raw data. 

```{r}
library(parameters)
print(model_parameters(nbin_total_yr,
                       exponentiate = TRUE #required for log or logit function
                       ))
```


```{r}
# Specify 'exponentiate = TRUE' to interpret coefficients as ratios
nbin_total_yr.results <- parameters::model_parameters(nbin_total_yr, exponentiate = TRUE) 
```




### Simulations of the model 

Simulating model -> need to check the reason people do this... I think it is to see if the observed values are explained by the model or not.


```{r}
sim1 <- simulateResiduals(nbin_total_yr)
testResiduals(sim1)
```

**Key take aways:**
- The uniformity indicates the scaled residuals are uniformly distributed. The p value > 0.05 suggests there is no evidence the residuals deviate from uniformity.
- Dispersion is close to 1 and p value is >0.05, suggesting that there is no overdispersion or under dispersion.
- Outliers: suggesting that the number of outliers is not significantly different to what would be expected.

Because the simulation suggests not overdispersed, that means I can either do Wald Z or chi squared test (Bolker, Brooks, Clark, Geange, Poulsen, Stevens, White; 2008 http://dx.doi.org/10.1016/j.tree.2008.10.008)

Attempting without seperating families and using family as a fixed effect

```{r}
pois_Family_yr <- glmmTMB(Total ~ Family * Shelf + Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = total_recruits)

sim2 <- simulateResiduals(pois_Family_yr)
testResiduals(sim2)
```
Not a good fit

```{r}
nbin_Family_yr <- glmmTMB(Total ~ 1+ Family * Shelf + Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = total_recruits)

sim3 <- simulateResiduals(nbin_Family_yr)
testResiduals(sim3)

#seems to be a good model


zbin_Family_yr <- glmmTMB(Total ~ 1+Family * Shelf + Year_Collected + (1 | Reef/Site),
                      ziformula = ~1,
        family = "nbinom2", data = total_recruits)
sim4 <- simulateResiduals(zbin_Family_yr)
testResiduals(sim4)
```

```{r}
AIC(nbin_Family_yr, zbin_Family_yr)
```

```{r}
summary(nbin_Family_yr)
```
I guess my issue is what to make the baseline that everything is different too... should it be inner:Acroporidae? I don't think so, but I don't know what it should be.

# Q2 is the composition of the recruits different across the shelf?

PERMANOVA - variation within a group can be calculated directly from a distance matrix. 


Assess the fit of a NMDS by plotting the original dissimilarities (diss) against the Euclidean ordination distances (dist). Shepard plot
```{r}
plot(NMDS$diss, NMDS$dist)
stressplot(object = NMDS)
```

NMDS ordinations are often plotted without units or labels on the axes so that people do not try to interpret them.

```{r}
ordiplot(NMDS, type = "n")
orditorp(NMDS, display = "species", col = "blue", air = 0.01)
orditorp(NMDS, display = "sites", cex = 0.5, air = 0.01)
```

### PERMANOVA


```{r}
library(vegan)
library(dplyr)
library(tidyr)

# Step 1: Create site-by-species matrix and "Site" column
site_by_species <- Coral_settlement_nGBR %>%
  select(Year_Collected, Reef, Shelf, Site, Tile, Acroporidae, Pocilloporidae, Poritid, Other) 

# Add a unique Site ID *after* extracting metadata
site_by_species <- site_by_species %>%
  mutate(Site_ID = paste(Year_Collected, Reef, Site, Tile, sep = "_"))

# Step 2: Separate metadata and species data
metadata <- site_by_species %>%
  select(Site, Shelf, Year_Collected)

species_matrix <- site_by_species %>%
  select(Site, Acroporidae, Pocilloporidae, Poritid, Other)

# Step 3: Set row names and remove species rows with all zeros
species_matrix_filtered <- species_matrix %>%
  column_to_rownames("Site_ID") %>%
  filter(rowSums(.) > 0)

# Step 4: Match metadata rows to filtered species matrix
metadata_filtered <- metadata %>%
  filter(Site_ID %in% rownames(species_matrix_filtered)) %>%
  column_to_rownames("Site_ID")

# Step 5: PERMANOVA
adonis_result <- vegan::adonis2(species_matrix_filtered ~ Shelf+ Year_Collected,
                                data = metadata_filtered,
                                method = "bray",
                                by = "term",
                                permutations = 999)

print(adonis_result)

```
The PERMANOVA suggests that shelf position and year do significantly affect the coral community composition.
Shelf position explains 17.4% of the variation (strong and significant)
Year collected explains 5% of the variation (suggesting significant temporal differences)
77.5% of the variation in composition of coral recruits remains unexplained.


# Redundancy Analysis
https://pmassicotte.github.io/stats-denmark-2019/07_rda.html#/a-concrete-example-3

Suggests using the Hellinger-transform method (Borcard, Gillet, and Legendre 2011; Legendre and Gallagher 2001) that is suited to species abundance data, 

- gives low weights to variables with low counts and many zeros.
Divide each value in a data matrix by its row sum, and taking the square root of the quotient

```{r}
Recruits <- Coral_settlement_nGBR %>%
  select(Acroporidae, Pocilloporidae, Poritid, Other)
recruit_hellinger <- decostand(Recruits, method = "hellinger")

environment_data <- Coral_settlement_nGBR %>%
  select(Year_Collected, Site, Reef, Shelf) #worried because environment data is nested here.
```

```{r}
my_rda <- rda(recruit_hellinger~Year_Collected + Shelf+ (Reef/Site), data = environment_data)

summary(my_rda)
```
```{r}
plot(my_rda)
```
Explained variance
```{r}
RsquareAdj(my_rda)
```
Tests of RDA results
For canonical analysis such as RDA, model significance relies on permutation tests.

The principle of a permutation test is to generate a reference distribution of the chosen statistic under the null hypothesis H0 by randomly permuting appropriate elements of the data a large number of times and recomputing the statistic each time. Then, one compares the true value of the statistic to this reference distribution. The p value is computed as the proportion of the permuted values equal to or larger than the true (unpermuted) value of the statistic for a one-tailed test in the upper tail, like the F test used in RDA (Borcard, Gillet, and Legendre 2011).

H0: Observed results can be produced by random chance.

The null hypothesis is rejected if this p
value is equal to or smaller than the predefined significance level α (ex.: α=0.05).

Test for 
1. Global RDA significance
2. Axis significance
3. Terms (explanatory variables) significance

### Global significance test
```{r}
anova.cca(my_rda, permutations = 999)
```
Since p = 0.001, reject H0, meaning that the RDA model is significant and not produced by random chance

### Axis significance
```{r}
anova.cca(my_rda, by = "axis")
```

### Terms significance
```{r}
anova.cca(my_rda, by = "terms")
```

### Check for multicollinearity
Use the vif.cca function
```{r}
sqrt(vif.cca(my_rda))
```
Not sure what the go is with some being NA. "As a rule of thumb, if sqrt vif >2, collinarity is considered high and some of the explanatory cariables are redundant.

I will remove reef and site.
```{r}
my_rda2 <- rda(recruit_hellinger~Year_Collected + Shelf, data = environment_data)

summary(my_rda2)
```

```{r}
plot(my_rda2)

RsquareAdj(my_rda2)

anova.cca(my_rda2, permutations = 999)
```
```{r}
anova.cca(my_rda2, by = "axis")
anova.cca(my_rda2, by = "terms")

sqrt(vif.cca(my_rda2))
```

###Unexplained Variance
Kaiser Guttman criterion

```{r}
my_rda2$CA$eig[my_rda2$CA$eig > mean(my_rda2$CA$eig)]
```

## Partial RDA
It is possible to run an RDA of a matrix Y, explained by a matrix variables X, in the presence of co-variable(s) W.

This analysis allows to display the patterns of the response data (Y) uniquely explained by a linear model of the explanatory variables (X) when the effect of other covariates (W) is held constant (Borcard, Gillet, and Legendre 2011).
### Hold Year constant
Model the effect of shelf position on coral recruit family abundances once the effect of year is removed

```{r}
partial_rda <- rda(recruit_hellinger~ Shelf + Condition(Year_Collected), data = environment_data)
```

```{r}
summary(partial_rda)
```
Conditioned gives the amount of variance that has been explained by the covariables and removed
Constrained gives the amount of variance uniquely explained by the explanatory variables
```{r}
RsquareAdj(partial_rda)
```

### Hold Shelf Position Constant
Model the effect of year on coral recruit family abundances once the effect of shelf position is removed

```{r}
partial_rda2 <- rda(recruit_hellinger~ Year_Collected + Condition(Shelf), data = environment_data)
```
```{r}
summary(partial_rda2)
```

```{r}
RsquareAdj(partial_rda2)
```
Test significance of partial RDA
```{r}
anova.cca(partial_rda2)
```
Since p≤0.05 we reject H0 and can conclude that the result was not produced randomly.

```{r}
plot(partial_rda2, scaling = 1)
```

### Variance partitioning
```{r}
env_shelf <- environment_data[,"Shelf"]

env_year <- environment_data[,"Year_Collected"]
vp <- varpart(recruit_hellinger, env_shelf, env_year)
vp

plot(vp)
```

# Multivariate WorkFlow

### Need to do test of significance of fractions

```{r}
anova.cca(rda(recruit_hellinger, env_shelf))

anova.cca(rda(recruit_hellinger, env_year))

#[a + b+c]
anova.cca(rda(recruit_hellinger, environment_data))

# [a]
anova.cca(rda(recruit_hellinger, env_shelf, env_year))

# [c]
anova.cca(rda(recruit_hellinger, env_year, env_shelf))
```

# Benthic Data

Importing benthic data
```{r}
PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    )
  )

Belt_10m <- read_csv("Belt_10m.csv") %>%
  mutate(across(8:25, ~replace_na(.x, 0))) # Read in data
Belt_10m <- Belt_10m %>%
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))
```

```{r}
sample_design2 <- PIT |> 
  group_by(Year, Reef, Site) |> 
  summarise(PIT = n()) |> 
  arrange(desc(PIT)) 
# check names of sites in PIT and recruitment data

# site_pit <- PIT |> 
#   mutate("PIT" = "PIT")|> 
#   distinct(Site, PIT)

# sites_recruits <- Coral_settlement_nGBR |>
#   mutate("Recruits" = "recruits")|> 
#   distinct(Site, Recruits)

# sites_clean <- full_join(site_pit, sites_recruits)


# reef_pit <- PIT |> 
#    mutate("PIT" = "PIT")|> 
#    distinct(Reef, PIT)
# 
# reef_recruits <- Coral_settlement_nGBR |>
#    mutate("Recruits" = "recruits")|> 
#    distinct(Reef, Recruits)
# 
# reef_clean <- full_join(reef_pit, reef_recruits)
```

***Note to self - perhaps calculate a proportion of soft coral, given that they are known to have a negative effect on coral recruitment***

```{r}
PIT |> 
  group_by(Year, Site) |> 
  summarise(
    count = n_distinct(Transect)
  ) #the number of transects per year differs among sites, so not sure whether calculating mean and standard error is appropriate or not.


```

## is the total number of recruits related to the number proportion of coral in PIT?


```{r}
library(mvabund)

PIT_abund <- mvabund(PIT[,95:103])
boxplot(PIT[, 95:103], horizontal = TRUE, las = 2, main = "Abundance")

pit_long <- PIT |> 
  pivot_longer(cols = 95:103, names_to = "Benthos", values_to = "Proportion")

pit_long |> 
  ggplot(aes(x=Year, y=Proportion, colour = Site))+
  geom_boxplot()+
  facet_wrap(~Benthos)

pit_long |> 
  group_by(Year, Site, `Depth/Zone`, Benthos) |> 
  summarise(
    mean = mean(Proportion),
    se = sd(Proportion)/(n()^0.5),
    mad = mad(Proportion),
    min = min(Proportion),
    max = max(Proportion),
    n = n(),
    range = max(Proportion)- min(Proportion)
  )
```

I am going to have to make a decision if I just keep the January or October, and then will need to change the year to facilitate an inner join.

```{r}
#PIT_and_recruits <- inner_join(Coral_settlement_nGBR, PIT_50m)
```

```{r}
PIT_long <- PIT |> 
  select(Year, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritidae, Prop_macroalgae, prop_coral) %>% 
  pivot_longer(cols = 5:9, names_to = "Family", values_to = "Proportion")
```

```{r}
PIT_long |> 
ggplot(aes(x=Year, y= Proportion, colour = Shelf))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Proportion cover (" * PIT^{-1}*")" ),
  x = "Benthos",
  colour = "Shelf position")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(~Family)

```

Mid shelf has had variable acroporidae cover, 

```{r}
PIT_long |> 
ggplot(aes(x=Year, y= Proportion, colour = Site))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Proportion cover (" * Transect^{-1}*")" ),
  x = "Benthos",
  colour = "Shelf position")+
  # scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_grid(Reef~Family)
```

Two takeaways:
1. Explicit temporal trend in benthic cover in Macgillvary reef (increasing then decreased final year); and
2. Significant variation in prop_coral and prop_macroalgae in Turtle group reefs.
# Bayesian Analysis 

Model Diagnostics Based on Murray Feedback
```{r}
Coral_settlement_nGBR_CS <- read_csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(`Shelf posn` != "Coral Sea", #remove coral sea data
        Obs != c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= `Shelf posn`,
         "Year_Collected"= `Year (Jan)`,
         "Site_#"= `Site #`,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected),
         Shelf = factor(Shelf))
```

```{r include=FALSE, eval=FALSE}
library(tidyverse)  #for data wrangling etc
library(cmdstanr)   #for cmdstan
library(brms)       #for fitting models in STAN
library(standist)   #for exploring distributions
library(HDInterval) #for HPD intervals
library(posterior)  #for posterior draws
library(coda)       #for diagnostics
library(bayesplot)  #for diagnostics
library(rstan)      #for interfacing with STAN
library(DHARMa)     #for residual diagnostics
library(emmeans)    #for marginal means etc
library(broom)      #for tidying outputs
library(broom.mixed) #for tidying MCMC outputs
library(tidybayes)  #for more tidying outputs
library(ggeffects)  #for partial plots
library(patchwork)  #for multiple figures
library(bayestestR) #for ROPE
library(ggridges)   #for ridge plots
library(easystats)     #framework for stats, modelling and visualisation
#source('helperFunctions.R')

ggplot(Coral_settlement_nGBR, aes(y = total, #response
                                  x=Reef))+ #categorical variable
  geom_boxplot()+
  facet_wrap(~Site) 

ggplot(Coral_settlement_nGBR, aes(y = total, #response
                                  x=Shelf))+ #categorical variable
  geom_boxplot()+
  facet_wrap(~Year_Collected) 
```
Suggests that the variance is not equal.

The 0 count skewed
So we would do poisson or negative binomial

```{r include=FALSE, eval=FALSE}
formula.coral1 <- bf(total~(1|Reef) + Year_Collected,
                     family = negbinomial(link = "log"))

get_prior(formula.coral1, data = Coral_settlement_nGBR)
```
Priors would be beta, intercept, sd, and "shape"?

```{r determining priors, include=FALSE, eval=FALSE}
Coral_settlement_nGBR |> 
  group_by(Year_Collected, Reef) |> 
  summarise(median(total),
            mad(total))
```

```{r}
sample_joined <- left_join(sample_design2, sample_design)
```

