---
title: "Data_Analysis_Thesis"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict

# Clean data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "dislodged",
    TRUE ~ notes
  ))

```


# Determining if loose or dettached tiles are significantly different to other tiles within a site.

```{r}
Coral_settlement_nGBR |> 
  group_by(Reef,notes) |> summarise(n= n())

Coral_settlement_nGBR |> 
  ggplot(aes(x=Year, y=total, colour = notes))+
  geom_jitter()+
  geom_boxplot()+
  facet_wrap(~Site, scale = "free_y", ncol = 3)
```


```{r}
Coral_settlement_nGBR |>  
  ggplot(aes(x=Year_Collected, y=total, colour = notes))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Recruits (" * tile^{-1}*")" ),
  x = "Deployment",
  colour = "Notes")+
  facet_wrap(~Shelf)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )
```

```{r}
dettached_mod <- glmmTMB(total ~ notes + Year_Collected + (1 | Reef/Site), 
      family = "nbinom2", data = Coral_settlement_nGBR)

sim_mod <- simulateResiduals(dettached_mod)
testResiduals(sim_mod)

summary(dettached_mod)
```

Does appear that dislodged tiles are significantly lower than other tiles when reef and site are a random effect.. maybe they need to be a fixed effect.

```{r}
plot(ggpredict(dettached_mod, terms = "notes"))+
  labs(y= expression("Predicted no. recruits (" * tile^{-1}*")" ), x= "Notes")+
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        text = element_text(size=8),
        title = element_blank())

d_tukey <- emmeans(dettached_mod, ~notes)
pairs(d_tukey, adjust = "tukey")
```
The dislodged tiles have 2.5 recruits less than attached tiles, when reef and site are held constant as a random effect, and this difference is significant at P<0.05.
The tiles with lots of tube worms were slightly lower than those without lots of tube worms. p<0.1

*also run with year as a random variable*
```{r}
dettached_mod <- glmmTMB(total ~ notes + (1 | Reef/Site) + (1| Year_Collected), 
      family = "nbinom2", data = Coral_settlement_nGBR)
```

```{r}
sim_mod <- simulateResiduals(dettached_mod)
testResiduals(sim_mod)

plot(ggpredict(dettached_mod, terms = "notes"))

d_tukey <- emmeans(dettached_mod, ~notes)
pairs(d_tukey, adjust = "tukey")
```

# Remove dislodged tiles

```{r}
Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "dislodged")
```

# Summary Statistics

```{r echo=FALSE}
Coral_settlement_nGBR |> 
  summarise(total_tiles = n(),
            total_recruits = sum(total),
            total_acro = sum(Acroporidae),
            total_poci = sum(Pocilloporidae),
            total_poritidae = sum(Poritid),
            total_other = sum(Other),
            total_bottom = sum(Under.Total),
            total_side = sum(Side.Total),
            total_top = sum(Top.Total)) |> 
  knitr::kable()
```

```{r echo=FALSE}
Coral_settlement_nGBR |> 
  group_by(Year_Collected,Reef, Site) |> 
  summarise(total_tiles = n(),
            total_recruits = sum(total))
```

Temporal statistics

```{r echo=FALSE}
Coral_settlement_nGBR |> 
  group_by(Year_Collected) |> 
  summarise(total_tiles = n(),
            total_recruits = sum(total),
            total_acro = sum(Acroporidae),
            total_poci = sum(Pocilloporidae),
            total_poritidae = sum(Poritid),
            total_bottom = sum(Under.Total),
            total_side = sum(Side.Total),
            total_top = sum(Top.Total)) |> 
  knitr::kable()
```



## Summary statistics
```{r}
Coral_settlement_nGBR |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )

Coral_settlement_nGBR |> 
  group_by(Year) |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )
```

### By Shelf Position

```{r}
Coral_settlement_nGBR |> 
  group_by(Shelf) |> 
  summarise(
    mean = mean(total),
    se = sd(total)/(n()^0.5),
    n= n()
  )
```



```{r echo=FALSE}
sample_design <- Coral_settlement_nGBR |> 
  group_by(Year, Reef, Site) |> 
  summarise(n = n()) |> 
  arrange(desc(n)) 
```


```{r echo=FALSE, message=FALSE}
total_recruits <- Coral_settlement_nGBR %>%
  select(Year_Collected, Shelf, Reef, Site, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 5:8, names_to = "Family", values_to = "Total")

#Order so that Other is last in Family
total_recruits <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, Reef, Site, Obs, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  pivot_longer(cols = 6:9, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritid","Pocilloporidae", "Other")))

total_recruits$Year_Collected <- as.factor(total_recruits$Year_Collected)
```

#Figure 1

```{r}
Coral_settlement_nGBR |>  
  ggplot(aes(x=Year_Collected, y=total, colour = Shelf))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Total recruits (" * tile^{-1}*")" ),
  x = "Deployment",
  colour = "Shelf")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(~Shelf)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  geom_jitter(size = 0.35, alpha = 0.5)

Fig.1 <- Coral_settlement_nGBR |>  
  ggplot(aes(x=Year_Collected, y=total, colour = Shelf))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Recruits (" * tile^{-1}*")" ),
  x = "Year",
  colour = "Shelf")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  #facet_wrap(~Shelf)+
  theme(legend.position = c(1,1), legend.justification = c(1,1))+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25")
  )+
  geom_jitter(size = 0.35, alpha = 0.5)

# Fig.1 <- Coral_settlement_nGBR |>
#   ggplot(aes(x = Year_Collected, y = total, colour = Shelf)) +
#   geom_boxplot(position = position_dodge(width = 0.75)) +
#   geom_jitter(
#     aes(group = Shelf),
#     position = position_jitterdodge(
#       jitter.width = 0.35,
#       dodge.width  = 0.75
#     ),
#     size = 0.35,
#     alpha = 0.5
#   ) +
#   theme_classic() +
#   labs(
#     y = expression("Recruits (" * tile^{-1} * ")"),
#     x = "Year",
#     colour = "Shelf position"
#   ) +
#   scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366")) +
#   theme(legend.position = c(1,1), legend.justification = c(1,1)) +
#   scale_x_discrete(
#     "Deployment",
#     labels = c(
#       "2022" = "21/22",
#       "2023" = "22/23",
#       "2024" = "23/24",
#       "2025" = "24/25"
#     )
#   )


Fig.1
```

```{r}
Fig1_mean <- Coral_settlement_nGBR |>  
  ggplot(aes(x = Year_Collected, y = total, colour = Shelf)) +
  # Mean points
  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.6),
    size = 2.5
  ) +
  # Error bars (mean ± SE)
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.6),
    width = 0.2
  ) +
  # Raw data
  geom_jitter(
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.6),
    size = 0.35,
    alpha = 0.5
  ) +
  theme_classic() +
  labs(
    y = expression("Recruits (" * tile^{-1} * ")"),
    x = "Year",
    colour = "Shelf"
  ) +
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366")) +
  theme(
    legend.position = c(1, 1),
    legend.justification = c(1, 1)
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c(
      "2022" = "21/22",
      "2023" = "22/23",
      "2024" = "23/24",
      "2025" = "24/25"
    )
  )+
  theme(text = element_text(size =8))

```

Fig 1
```{r echo=FALSE, eval=FALSE}
ggsave(filename = "Figure_1.png",
       plot = Fig.1,
       width = 169,
       height = 120,
       units = "mm")

ggsave(filename = "Figure_1mean.png",
       plot = Fig1_mean,
       width = 169,
       height = 120,
       units = "mm")
```


# Figure 2


```{r}
Figure.2 <- total_recruits |> 
group_by(Year_Collected, Reef, Family, Shelf)  %>% 
  ggplot(aes(x=Year_Collected, y=Total, colour = Shelf))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Recruits (" *tile^{-1} * ")"),
  x = "Year",
  colour = "Shelf")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(~Family, scales = "free_y", ncol =1)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  geom_jitter(
    aes(group = Shelf),
    position = position_jitterdodge(
      jitter.width = 0.35,
      dodge.width  = 0.75
    ),
    size = 0.35,
    alpha = 0.5
  )
```


```{r eval = FALSE, echo=FALSE}
ggsave(filename = "Figure 2.png",
       plot = Figure.2,
       height = 200,
       width = 125,
       unit = "mm")
```

```{r}
Fig2_mean <- total_recruits |> 
  ggplot(aes(x=Year_Collected, y=Total, colour = Shelf))+
stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.6),
    size = 2
  ) +
  # Error bars (mean ± SE)
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.6),
    width = 0.2
  ) +
  # Raw data
  geom_jitter(
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.6),
    size = 0.35,
    alpha = 0.3
  ) +
  theme_classic() +
  labs(
    y = expression("Recruits (" * tile^{-1} * ")"),
    x = "Year",
    colour = "Shelf"
  ) +
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366")) +
  theme(
    # legend.position = c(1, 0.4),
    # legend.justification = c(1, 1)
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c(
      "2022" = "21/22",
      "2023" = "22/23",
      "2024" = "23/24",
      "2025" = "24/25"
    )
  )+
  facet_wrap(~Family, scales = "free_y", ncol =2)+
  theme(text = element_text(size =8),
        legend.position = "bottom")
```

```{r}
ggsave(filename = "Fig2_mean.png",
       plot = Fig2_mean,
       height = 175,
       width = 160,
       unit = "mm")
```


## Position on tiles

```{r}
position_total_recruits <- Coral_settlement_nGBR %>%
  select(2:9, 15:29) %>% 
  pivot_longer(cols = 21:23, names_to = "Pos", values_to = "Total")

position_total_recruits <- position_total_recruits |> 
  mutate(Pos = case_when(
      Pos == "Top.Total" ~ "Top",
      Pos == "Side.Total" ~ "Side",
      Pos == "Under.Total" ~ "Bottom"
  ))
 
position_total_recruits <- position_total_recruits |>                           
mutate(Pos = factor(Pos, levels = c("Bottom", "Side", "Top")))
```



```{r message=FALSE, echo=FALSE}
total_recruits %>% 
  group_by(Family, Shelf, Year_Collected, Reef) %>%
  summarise(average = mean(Total),
            se = sd(Total)/ sqrt(n()),
            max = max(Total),
            min = min(Total)) %>% 
  arrange(desc(average)) %>% 
  print(n=20)
```

```{r}
Figure.3 <- position_total_recruits |> 
group_by(Year_Collected, Reef, Pos, Shelf)  %>% 
  ggplot(aes(x=Year_Collected, y=Total, colour = Shelf))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Recruits (" *tile^{-1} * ")"),
  x = "Year",
  colour = "Shelf")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(~Pos, scales = "free_y", ncol =1)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  geom_jitter(
    aes(group = Shelf),
    position = position_jitterdodge(
      jitter.width = 0.35,
      dodge.width  = 0.75
    ),
    size = 0.35,
    alpha = 0.5
  )

Figure.3

```
```{r echo=FALSE, eval=FALSE}
ggsave(filename = "Figure 3.png",
       plot = Figure.3,
       height = 275,
       width = 125,
       unit = "mm")
```

```{r}
Fig3_mean <- position_total_recruits |> 
group_by(Year_Collected, Reef, Pos, Shelf)  %>% 
  ggplot(aes(x=Year_Collected, y=Total, colour = Shelf))+
stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.6),
    size = 2
  ) +
  # Error bars (mean ± SE)
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.6),
    width = 0.2
  ) +
  # Raw data
  geom_jitter(
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.6),
    size = 0.35,
    alpha = 0.3
  ) +
  theme_classic() +
  labs(
    y = expression("Recruits (" * tile^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366")) +
  theme(
    # legend.position = c(1, 0.4),
    # legend.justification = c(1, 1)
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c(
      "2022" = "21/22",
      "2023" = "22/23",
      "2024" = "23/24",
      "2025" = "24/25"
    )
  )+
  facet_wrap(~Pos, scales = "free_y", ncol =1)+
  theme(text = element_text(size =8),
        legend.position = "right")
```

```{r echo=FALSE, eval=FALSE}
ggsave(filename = "Fig3_mean.png",
       plot = Fig3_mean,
       height = 275,
       width = 125,
       unit = "mm")
```

# Composition on tile

```{r}
position_family_recruits <- Coral_settlement_nGBR %>%
  select(2:9, 15:29) %>% 
  pivot_longer(cols = 9:20, names_to = "Pos", values_to = "Total")

position_family_recruits <- position_family_recruits |> 
  mutate(Family = case_when(
      Pos %in% c("Top_Acr", "Side_Acr", "Under_Acr") ~ "Acroporidae",
      Pos %in% c("Top_Pocil", "Side_Pocil", "Under_Pocil") ~ "Pocilloporidae",
      Pos %in% c("Top_Por", "Side_Por", "Under_Por") ~ "Poritidae",
     Pos %in% c("Top_other", "Side_other", "Under_oth") ~ "Other"
  ))

position_family_recruits <- position_family_recruits |> 
  mutate(Pos = case_when(
      Pos %in% c("Top_Acr","Top_Pocil", "Top_Por", "Top_other") ~ "Top",
      Pos %in% c("Side_Pocil", "Side_Acr", "Side_Por","Side_other") ~ "Side",
      Pos %in% c("Under_Acr", "Under_Pocil", "Under_Por", "Under_oth") ~ "Bottom"
  ))
 
position_family_recruits <- position_family_recruits |>                           
mutate(Pos = factor(Pos, levels = c("Bottom", "Side", "Top")),
       Family = factor(Family, levels = c("Acroporidae", "Pocilloporidae","Poritidae", "Other")))
```

```{r}
Figure.4 <- position_family_recruits |> 
  ggplot(aes(x=Year_Collected, y=Total, colour = Shelf))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Recruits (" * surface^{-1}~tile^{-1} * ")"),
  x = "Year",
  colour = "Shelf")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(Family~Pos, scales = "free_y", ncol =3)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  geom_jitter(
    aes(group = Shelf),
    position = position_jitterdodge(
      jitter.width = 0.35,
      dodge.width  = 0.75
    ),
    size = 0.35,
    alpha = 0.5
  )+
  theme(text =element_text(size =8),
        title = element_text(size = 8))

Figure.4
```

```{r}
Acro_pos <- position_family_recruits |> 
  filter(Family == "Acroporidae") |> 
  ggplot(aes(x=Year_Collected, y=Total, colour = Shelf))+
  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.6),
    size = 2
  ) +
  # Error bars (mean ± SE)
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.6),
    width = 0.2
  ) +
  # Raw data
  geom_jitter(
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.6),
    size = 0.35,
    alpha = 0.3
  ) +
  theme_classic()+
  labs(
  y = expression("Acroporidae recruits (" * ~tile~surface^{-1} * ")"),
  x = "Year",
  colour = "Shelf")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(~Pos, scales = "free_y", ncol =3)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  geom_jitter(
    aes(group = Shelf),
    position = position_jitterdodge(
      jitter.width = 0.35,
      dodge.width  = 0.75
    ),
    size = 0.35,
    alpha = 0.5
  )+
  theme(text =element_text(size =8),
        title = element_text(size = 8))

Acro_pos
```

```{r}
Poci_pos <- position_family_recruits |> 
  filter(Family == "Pocilloporidae") |> 
  ggplot(aes(x=Year_Collected, y=Total, colour = Shelf))+
  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.6),
    size = 2
  ) +
  # Error bars (mean ± SE)
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.6),
    width = 0.2
  ) +
  # Raw data
  geom_jitter(
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.6),
    size = 0.35,
    alpha = 0.3
  ) +
  theme_classic()+
  labs(
  y = expression("Pocilloporidae recruits (" * ~tile~surface^{-1} * ")"),
  x = "Year",
  colour = "Shelf")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(~Pos, scales = "free_y", ncol =3)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  geom_jitter(
    aes(group = Shelf),
    position = position_jitterdodge(
      jitter.width = 0.4,
      dodge.width  = 0.75
    ),
    size = 0.35,
    alpha = 0.4
  )+
  theme(text =element_text(size =8),
        title = element_text(size = 8))

Poci_pos
```

```{r}
Por_pos <- position_family_recruits |> 
  filter(Family == "Poritidae") |> 
  ggplot(aes(x=Year_Collected, y=Total, colour = Shelf))+
  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.6),
    size = 2
  ) +
  # Error bars (mean ± SE)
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.6),
    width = 0.2
  ) +
  # Raw data
  geom_jitter(
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.6),
    size = 0.35,
    alpha = 0.3
  ) +
  theme_classic()+
  labs(
  y = expression("Poritidae recruits (" * ~tile~surface^{-1} * ")"),
  x = "Year",
  colour = "Shelf")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(~Pos, scales = "free_y", ncol =3)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  geom_jitter(
    aes(group = Shelf),
    position = position_jitterdodge(
      jitter.width = 0.4,
      dodge.width  = 0.75
    ),
    size = 0.35,
    alpha = 0.4
  )+
  theme(text =element_text(size =8),
        title = element_text(size = 8))

Por_pos
```

```{r}
Oth_pos <- position_family_recruits |> 
  filter(Family == "Other") |> 
  ggplot(aes(x=Year_Collected, y=Total, colour = Shelf))+
  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.6),
    size = 2
  ) +
  # Error bars (mean ± SE)
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.6),
    width = 0.2
  ) +
  # Raw data
  geom_jitter(
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.6),
    size = 0.35,
    alpha = 0.3
  ) +
  theme_classic()+
  labs(
  y = expression("Other recruits (" * ~tile~surface^{-1} * ")"),
  x = "Year",
  colour = "Shelf")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(~Pos, scales = "free_y", ncol =3)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  geom_jitter(
    aes(group = Shelf),
    position = position_jitterdodge(
      jitter.width = 0.4,
      dodge.width  = 0.75
    ),
    size = 0.35,
    alpha = 0.4
  )+
  theme(text =element_text(size =8),
        title = element_text(size = 8))

Oth_pos
```

```{r}
ggarrange(Acro_pos, Poci_pos, Por_pos, Oth_pos, ncol =1, common.legend = TRUE)
```

```{r}
library(ggh4x)


Family_pos <- position_family_recruits |> 
  ggplot(aes(x=Year_Collected, y=Total, colour = Shelf))+
  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.6),
    size = 2
  ) +
  # Error bars (mean ± SE)
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.6),
    width = 0.2
  ) +
  # Raw data
  geom_jitter(
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.6),
    size = 0.35,
    alpha = 0.3
  ) +
  theme_classic()+
  labs(
  y = expression("Recruits (" * ~tile~surface^{-1} * ")"),
  x = "Year",
  colour = "Shelf")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_grid2(
   Pos~Family,
  scales = "free_y",
  independent = "y",
  axes = "y"
)+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  geom_jitter(
    aes(group = Shelf),
    position = position_jitterdodge(
      jitter.width = 0.35,
      dodge.width  = 0.75
    ),
    size = 0.35,
    alpha = 0.5
  )+
  theme(text =element_text(size =8),
        title = element_text(size = 8),
        legend.position = "bottom")

Family_pos
```

```{r eval=FALSE, echo=FALSE}
ggsave(filename = "Fig4.png",
       plot = Family_pos,
       height = 175,
       width = 200,
       unit = "mm")
```


```{r eval=FALSE, echo=FALSE}
ggsave(filename = "Figure 4.png",
       plot = Figure.4,
       height = 170,
       width = 155,
       unit = "mm")
```

# Models

## Question one. is the total number of recruits different across the shelf?

### Use GLMM for total recruits

Run generalised linear models with the Poisson, normal binomail and the zero inflated normal bionomial.
Total is the dependent variable (the count of coral recruits at a given sampling unit)
Shelf position is a fixed effect predictor, it is used to test if the average coral settlement differs between shelf positions
Random effects: 
a nested random effect, modelling the variations between reefs, and variation between sites nested within reefs


Run generalised linear models with the Poisson, normal binomail and the zero inflated normal bionomial. Total is the dependent variable (the count of coral recruits at a given sampling unit) Shelf position is a fixed effect predicator, it is used to test if the average coral settlement differs between shelf positions Random effects: a nested random effect, modelling the variations between reefs, and variation between sites nested within reefs


Victor: I recommend including Year as a fixed effect since you are interested in differences in recruitment across years, and the levels are not interchangeable. In contrast, a variable like Site would be more appropriate as a random effect in your analysis.

*Should I be scaling or centring the data?*

*Should I be doing a liklihood ratio test here to determine whether the model is better with or without year?*

```{r warning=FALSE, message=FALSE}
pois_total_yr <- glmmTMB(total ~ Shelf + Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = Coral_settlement_nGBR)

nbin_total_yr <- glmmTMB(total ~ Shelf +  Year_Collected +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

zbin_total_yr <- glmmTMB(total ~ Shelf + Year_Collected + (1 | Reef/Site),
                      ziformula = ~1,
        family = "nbinom2", data = Coral_settlement_nGBR)

pois_total_yr2 <- glmmTMB(total ~ Shelf * Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = Coral_settlement_nGBR)

nbin_total_yr2 <- glmmTMB(total ~ Shelf * Year_Collected +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

zbin_total_yr2 <- glmmTMB(total ~ Shelf * Year_Collected + (1 | Reef/Site),
                      ziformula = ~1,
        family = "nbinom2", data = Coral_settlement_nGBR)
```

Use DHARMA to assess residuals

```{r}
sim1.1 <- simulateResiduals(pois_total_yr) 
testResiduals(sim1.1) #not good fit

sim1.2 <- simulateResiduals(pois_total_yr2)
testResiduals(sim1.2) #not good fit

sim1.3 <- simulateResiduals(nbin_total_yr)
testResiduals(sim1.3) #good

sim1.4 <- simulateResiduals(nbin_total_yr2)
testResiduals(sim1.4) #good

sim1.5 <- simulateResiduals(zbin_total_yr)
testResiduals(sim1.5) #good

sim1.6 <- simulateResiduals(zbin_total_yr2)
testResiduals(sim1.6) #good
```

Poisson model residuals were bad, therefore will only compare the four models that testResiduals were good for.

```{r}
AIC(nbin_total_yr, zbin_total_yr, nbin_total_yr2, zbin_total_yr2) |>  #nbin has the lowest AIC with interaction terms
  knitr::kable()
```

negative binomial model has the lowest AIC so I will get the summary of that model.

```{r echo=FALSE}
summary(nbin_total_yr2)
```
2023 and 2024 are significantly different to 2022. 2025 is not significantly different to 2022. Outer shelf is significantly different to inner shelf (p<0.01), Mid shelf is significantly different to inner shelf reefs (p<0.05). Intercept is also significant: not quite sure what that means in a biological framework.

When making a figure should I then be using summary statistics or raw count data? Because the large number of zeros will make it difficult to do a boxplot of the raw data.


### Tukey tests

```{r}
emm_shelf <- emmeans(nbin_total_yr2, ~Shelf)

emm_shelf

pairs(emm_shelf, adjust = "tukey")

```

```{r}
emm_year <- emmeans(nbin_total_yr2, ~Year_Collected)

emm_year

pairs(emm_year, adjust = "tukey")
```


```{r}
emm_year_shelf <- emmeans(nbin_total_yr2, ~Year_Collected*Shelf)

emm_year_shelf

pairs(emm_year_shelf, adjust = "tukey", 
      #type = "response"
      )
```

```{r}
nbin_total_yr2|> 
  emmeans(~Shelf| Year_Collected, type = "response") |> 
  as.data.frame() |> 
  ggplot(aes(x=Year_Collected,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(y= "Estimated marginal mean total\nrecruits (response scale)")+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1))+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))
```

```{r}
# Caterpillar plot
nbin_total_yr2|> 
  emmeans(~Shelf| Year_Collected, type = "response") |> 
  as.data.frame() |>
  ggplot(aes(x = response, y = Shelf, color=Shelf)) +
  geom_point() +
  geom_errorbarh(aes(xmin =asymp.LCL, xmax = asymp.UCL), height = 0.2) +
  facet_wrap(~Year_Collected, scales = "free_y") +
  labs(x = "Estimated marginal mean (response scale)",
       y = "Shelf Position",
       title = "Caterpillar plot of EMMs by Year") +
  theme_minimal()+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))
```

# Q2 is the composition of the recruits different across the shelf?

Isolating Data for a single Family of coral

```{r echo=FALSE}
Pocilloporidae <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, 
         Reef, Site, Tile, 
         Pocilloporidae, Top_Pocil, 
         Under_Pocil, Side_Pocil, total)

Pocilloporidae_orientation <- Pocilloporidae %>% 
  rename(Top = Top_Pocil,
                   Bottom = Under_Pocil,
                   Side = Side_Pocil,
         Recruits_per_tile = total)

Pocilloporidae_orientation <- Pocilloporidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Acroporidae <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, 
         Reef, Site, Tile, 
         Acroporidae, Top_Acr, 
         Under_Acr, Side_Acr, total)

Acroporidae_orientation <- Acroporidae %>% 
  rename(Top = Top_Acr,
                   Bottom = Under_Acr,
                   Side = Side_Acr,
         Recruits_per_tile = total)

Acroporidae_orientation <- Acroporidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Poritid <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf, 
         Reef, Site, Tile, 
         Poritid, Top_Por, Under_Por, 
         Side_Por, total)

Poritid_orientation <- Poritid %>% 
  rename(Top = Top_Por,
                   Bottom = Under_Por,
                   Side = Side_Por,
         Recruits_per_tile = total) 

Poritid_orientation <- Poritid_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")
```

GLMM with Total Recruits for a given coral family, per tile.

To analyze whether there's a difference in the number of recruits of a given coral family (e.g., Poritids) across shelf positions, while accounting for random effects due to site nested within reef and controlling for year, the model you choose depends on:

Whether the response is aggregated or split by tile facet (side/top/bottom),

The structure of your random effects (nested site within reef),

Whether Year_Collected should be a fixed or random effect.

```{r poisson of total Recruits, message=FALSE}
pois_Poritid_fixed <- glmmTMB(Poritid ~ Shelf + Year_Collected+(1 | Reef/Site) , 
      family = "poisson", data = Poritid) 

pois_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = Poritid)

pois_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + Year_Collected+ (1 | Reef/Site), 
      family = "poisson", data = Acroporidae)

pois_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = Acroporidae)

pois_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf + Year_Collected  + (1 | Reef/Site), 
      family = "poisson", data = Pocilloporidae)

pois_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf*Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = Pocilloporidae)
```

Negative Binomial of Total Recruits (for a family)

```{r}
## Fixed
nbin_Poritid_fixed <- glmmTMB(Poritid ~ Shelf + (1 | Reef/Site) + Year_Collected, 
        family = "nbinom2", data = Poritid)

nbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf + (1 | Reef/Site) + Year_Collected, 
        family = "nbinom2", data = Pocilloporidae)

nbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + (1 | Reef/Site) + Year_Collected, 
        family = "nbinom2", data = Acroporidae)

## Interaction
nbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = Poritid)

nbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf * Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = Pocilloporidae)

nbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = Acroporidae)
```


Negative Binomial Zero Inflated of Total recruits (per family)

```{r}
#Fixed
# Fit model with zero-inflated negative binomial distribution 

zbin_Poritid_fixed <- glmmTMB(Poritid ~ Shelf+ Year_Collected + (1 | Reef/Site) ,
                        ziformula = ~1,
                        family = "nbinom2",
                        data = Poritid)
zbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf * Year_Collected + (1 | Reef/Site),
                        ziformula = ~1,
                        family = "nbinom2",
                        data = Poritid)

zbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf+ Year_Collected + (1 | Reef/Site),
                               ziformula = ~1,
                        family = "nbinom2",
                        data = Pocilloporidae)

zbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf*Year_Collected + (1 | Reef/Site),
                               ziformula = ~1,
                        family = "nbinom2",
                        data = Pocilloporidae)

zbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + Year_Collected+ (1 | Reef/Site),
                            ziformula = ~1,
                        family = "nbinom2",
                        data = Acroporidae)
zbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*Year_Collected + (1 | Reef/Site),
                            ziformula = ~1,
                        family = "nbinom2",
                        data = Acroporidae)

```

### Acroporidae

```{r}
Acroporidae |> 
  ggplot(aes(x = Year_Collected, y = Acroporidae, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")
```

Simulate residuals for models

```{r}
sim2.1 <- simulateResiduals(nbin_Acroporidae_fixed)
testResiduals(sim2.1)

sim2.2 <- simulateResiduals(nbin_Acroporidae_interaction)
testResiduals(sim2.2)

sim2.3 <- simulateResiduals(zbin_Acroporidae_fixed)
testResiduals(sim2.3)

sim2.4 <- simulateResiduals(zbin_Acroporidae_interaction)
testResiduals(sim2.4)

sim2.5 <- simulateResiduals(pois_Acroporidae_fixed) #not a good fit
testResiduals(sim2.5)

sim2.6 <- simulateResiduals(pois_Acroporidae_interaction) #not a good fit
testResiduals(sim2.6)

```


Compare AIC of those models

```{r}
AIC( nbin_Acroporidae_fixed, zbin_Acroporidae_fixed, nbin_Acroporidae_interaction, zbin_Acroporidae_interaction)
```


```{r echo=FALSE}
nbin_Acroporidae_interaction %>% 
  summary()
```

Acroporidae abundance significantly different in the outer reefs than inshore reefs.

```{r}
emm_acro_shelf <- emmeans(nbin_Acroporidae_interaction, ~Shelf)

emm_acro_shelf

pairs(emm_acro_shelf, adjust = "tukey")
```

```{r}
nbin_Acroporidae_interaction|> 
  emmeans(~Shelf | Year_Collected, type = "response") |> 
  as.data.frame() |> 
  ggplot(aes(x=Year_Collected,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(y= "Estimated marginal mean Acroporidae\nrecruits (response scale)")+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1))+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))
```


### Pocillopora

```{r}
Pocilloporidae |> 
  ggplot(aes(x = Year_Collected, y = Pocilloporidae, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")
```


Test Residuals of models for Pocillopora

```{r}
sim3.1 <- simulateResiduals(pois_Pocilloporidae_fixed)
testResiduals(sim3.1)

sim3.2 <- simulateResiduals(pois_Pocilloporidae_interaction)
testResiduals(sim3.2)

sim3.3 <- simulateResiduals(nbin_Pocilloporidae_interaction)
testResiduals(sim3.3)

sim3.4 <- simulateResiduals(nbin_Pocilloporidae_fixed)
testResiduals(sim3.4)

sim3.5 <- simulateResiduals(zbin_Pocilloporidae_fixed)
testResiduals(sim3.5)

sim3.6 <- simulateResiduals(zbin_Pocilloporidae_interaction)
testResiduals(sim3.6)
```

```{r}
AIC(pois_Pocilloporidae_fixed, nbin_Pocilloporidae_fixed, zbin_Pocilloporidae_fixed, nbin_Pocilloporidae_interaction)
```


```{r echo=FALSE}
nbin_Pocilloporidae_interaction %>% 
  summary()
```

```{r}
emm_poci_shelf <- emmeans(nbin_Pocilloporidae_interaction, ~Shelf)

emm_poci_shelf

pairs(emm_poci_shelf, adjust = "tukey")
```

Significant difference in Pocilloporidae abundance in mid shore reefs compared to inshore reefs

```{r}
nbin_Pocilloporidae_interaction|> 
  emmeans(~Shelf | Year_Collected, type = "response") |> 
  as.data.frame() |> 
  ggplot(aes(x=Year_Collected,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(y= "Estimated marginal mean Acroporidae\nrecruits (response scale)")+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1))+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))
```


### Poritid

```{r}
Poritid |> 
  ggplot(aes(x = Year_Collected, y = Poritid, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")

Poritid |> 
  filter(Poritid>60)
```

Test residuals of models

```{r}
sim4.1 <- simulateResiduals(pois_Poritid_fixed)
testResiduals(sim4.1) # bad fit

sim4.2 <- simulateResiduals(pois_Poritid_interaction)
testResiduals(sim4.2) # bad fit

sim4.3 <- simulateResiduals(nbin_Poritid_interaction)
testResiduals(sim4.3)

sim4.4 <- simulateResiduals(nbin_Poritid_fixed)
testResiduals(sim4.4)

sim4.5 <- simulateResiduals(zbin_Poritid_fixed)
testResiduals(sim4.5)

sim4.6 <- simulateResiduals(zbin_Poritid_interaction)
testResiduals(sim4.6)
```

Both nbin_interaction, and zbin_interaction for Portidae have significant dispersion.
Don't see much of a difference in Poritid across shelf position. Although slightly less in mid shore reefs (p<0.1)

```{r echo=FALSE}
AIC(nbin_Poritid_fixed, zbin_Poritid_fixed, nbin_Poritid_interaction, zbin_Poritid_interaction)
```


```{r}
nbin_Poritid_interaction|> 
  emmeans(~Shelf | Year_Collected, type = "response") |> 
  as.data.frame() |> 
  ggplot(aes(x=Year_Collected,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(y= "Estimated marginal mean Acroporidae\nrecruits (response scale)")+
  scale_x_discrete(
    "Deployment",
    labels = c("2022" = "21/22",
               "2023" = "22/23",
               "2024" = "23/24",
               "2025" = "24/25",
               "2026" = "25/26")
    )+
  theme(legend.position = c(1,1), legend.justification = c(1,1))+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))
```

# Composition and Orientation

```{r}
nb_Orientation_Acroporidae <- glmmTMB(Recruits ~ 0+Shelf* Tile_Surface + (1 | Reef/Site), 
        family = "nbinom2", data = Acroporidae_orientation)

pois_Orientation_Acro <- glmmTMB(Recruits ~ 0+Shelf* Tile_Surface + (1 | Reef/Site), 
        family = "poisson", data = Acroporidae_orientation)

zi_orientation_Acro <- glmmTMB(Recruits ~ 0+Shelf* Tile_Surface + (1 | Reef/Site),
                               ziformula = ~1,
        family = "nbinom2", data = Acroporidae_orientation)
```

AIC to determine model of best fit
```{r echo=FALSE}
AIC(nb_Orientation_Acroporidae, pois_Orientation_Acro,zi_orientation_Acro)
```


```{r echo=FALSE}
nb_Orientation_Acroporidae |> 
  summary()
```

```{r echo=FALSE}
sim5 <- simulateResiduals(nb_Orientation_Acroporidae)
testResiduals(sim5)
```

```{r}
sim5.1 <- simulateResiduals(pois_Orientation_Acro)
testResiduals(sim5.1)
```

# Brood Stock Analysis

Check whether benthic composition is behaving consistently between Sites in a reef
```{r}
PIT_50m <- read_csv("PIT_50m.csv", col_types = cols(Shelf = col_character()))

PIT_50m <- PIT_50m |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site)
         ) |> 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))
```

```{r echo=FALSE}
ggplot(PIT_50m, aes(x=Year, y=prop_hard_coral, colour = Reef))+
  geom_boxplot(aes(group = interaction(Year, Site))) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")

ggplot(PIT_50m, aes(x=Year, y=Prop_macroalgae, colour = Reef))+
  geom_boxplot(aes(group = interaction(Year, Site))) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")
```

### Acroporidae

```{r echo=FALSE}
ggplot(PIT_50m, aes(x=Year, y=prop_acro, colour = Reef))+
  geom_boxplot(aes(group = interaction(Year, Site))) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")
```

### Pocilloporidae

```{r echo=FALSE}
ggplot(PIT_50m, aes(x=Year, y=prop_poci, colour = Reef))+
  geom_boxplot(aes(group = interaction(Year, Site))) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")
```

### Poritidae

```{r echo=FALSE} 
ggplot(PIT_50m, aes(x=Year, y=prop_pori, colour = Reef))+
  geom_boxplot(aes(group = interaction(Year, Site))) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")
```

Victor Email: Regarding the PERMANOVA: you start with a community matrix (sites x coral taxa). A PERMANOVA tests whether community composition differs among groups, using a distance matrix derived from your multivariate raw data. The inputs are a dissimilarity (distance) matrix, and a grouping factor (i.e., a categorical variable you want to test, like “reef”, “shelf.position”, or “year”). In your case, I'd suggest calculating a Bray-Curtis distance matrix and then running a PERMANOVA with the vegan package. Here’s a quick example you can run in R:
```{r}
library(vegan)
abund <- PIT_50m |> 
  select(Acroporidae, Pocilloporidae, Poritidae)

prop <- PIT_50m |> 
  select(prop_hard_coral, Prop_macroalgae)
```

Some of the rows do not have Acropora, Pocillopora or Porites present... should I then remove them? Because the distance matrix won't work otherwise.

```{r}
metadata <- PIT_50m |> 
  select(Site, Shelf, Reef, Year) |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site)
  )
```


```{r echo=FALSE}
dist.mat <- vegdist(prop, method = "bray")                                                       # Step 1: Calculate dissimilarity matrix (Bray-Curtis)
```

```{r}
PERMANOVA_prop <- adonis2(dist.mat ~ Reef:Shelf, data = metadata, permutations = 999, by = "terms")   # Step 2: run PERMANOVA

print(PERMANOVA_prop)

PERMANOVA_prop2 <- adonis2(dist.mat ~ Year+Shelf+Year:Shelf, data = metadata, permutations = 999, by = "terms")

print(PERMANOVA_prop2)
```

formula – model formula such as Y ~ A + B * C.  See the ‘Complex Models‘ chapter for more details of how to specify a model formula.  For now, the key point is that we are specifying that we want to analyze the response (Y) as a function of (~) one or more explanatory variables (A, B, C).  The class of Y determines what is done:
If Y is a dissimilarity matrix (i.e., output from dist() or vegdist()), the pseudo-F statistic is calculated directly.
If Y is a data frame or matrix containing data, vegdist() is applied to calculate the distance matrix first and then the pseudo-F statistic is calculated.

- 60% os total community variation (in dissimilarity matrix) explained by Year, SHelf, and their interaction.
Shelf contributes 52.2% of total variation in community composition (proportion macro algae and proportion coral)

### Try to do for abundances, but will need to remove rows where it is 0
```{r}
data_for_abundance <- PIT_50m |>
  mutate(clean = rowSums(across(c(Acroporidae, Pocilloporidae, Poritidae)), na.rm = TRUE))

data_for_abundance <- data_for_abundance |> 
  filter(clean > 0)

abund <- data_for_abundance |> 
  select(Acroporidae, Pocilloporidae, Poritidae)

abund_std <- abund |> 
  wisconsin()

abund_bray <- vegdist(abund_std, method = "bray")

meta_abund <- data_for_abundance |> 
  select(Shelf, Year, Reef, Site)
```


```{r}
PERMANOVA_abund <- adonis2(abund_bray ~ Year+Shelf+ Year:Shelf, data = meta_abund, permutations = 999, by = "terms")   # Step 2: run PERMANOVA
```

```{r echo=FALSE}
PERMANOVA_abund
```

Abundance is the community matrix data, with 237 sample units of 3 species.

```{r}
set.seed(42)
z <- metaMDS(comm = abund_std,
             autotransform = FALSE,
             distance = "bray",
             engine = "monoMDS",
             k=2,
             weakties = TRUE,
             model = "global",
             maxit = 300,
             try = 40,
             trymax = 100)

print(z)
# Extract NMDS site scores
scores_nmds <- as.data.frame(scores(z, display = "sites"))

# Join to metadata (assuming same row order)
scores_nmds <- scores_nmds |> 
  bind_cols(meta_abund)


NMDS <- ggplot(scores_nmds, aes(x = NMDS1, y = NMDS2, colour = Shelf)) +
  geom_point(size = 3) +
  theme_minimal() +
  labs(
    title = "NMDS (Bray–Curtis)",
    colour = "Shelf Position"
  )
```

Stress <0.05 therefore excellent representation with no prospect of misinterpretation.

https://uw.pressbooks.pub/appliedmultivariatestatistics/chapter/nmds/

```{r}
gof <- goodness(object = z)
plot(z, display = "sites", type = "none")
points(z, display = "sites", cex = 2*gof/mean(gof))
```

Evaluating fit
plot original dissimilarities (z$diss) against Euclidean ordination distances (z$dist).

```{r}
plot(z$diss, z$dist)
```

```{r}
z.points <- data.frame(z$points,
                       Acroporidae = abund_std$Acroporidae,
                       Pocilloporidae = abund_std$Pocilloporidae,
                       Poritid = abund_std$Poritidae)

p <- ggplot(data = z.points, aes(x = MDS1, y = MDS2)) +
theme_bw() +
theme(axis.title = element_blank(),
  axis.ticks = element_blank(),
  axis.text = element_blank())

p + geom_point()
```



## Acroporidae

```{r echo=FALSE}
Acroporidae_orientation |> 
  ggplot(aes(x = Year_Collected, y = Recruits, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_grid(Shelf~Tile_Surface, scale = "free_y")
```

## Pocilloporidae

```{r echo=FALSE}
Pocilloporidae_orientation |> 
  ggplot(aes(x = Year_Collected, y = Recruits, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_grid(Shelf~Tile_Surface, scale = "free_y")+
  theme_bw()
```


## Poritidae

```{r}
Poritid_orientation |> 
  ggplot(aes(x = Year_Collected, y = Recruits, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_grid(Shelf~Tile_Surface, scale = "free_y")+
  theme_bw()
```

***Check the levels of grouping variables***


```{r}
library(parameters)
print(model_parameters(nbin_total_yr,
                       exponentiate = TRUE #required for log or logit function
                       ))
```


```{r}
# Specify 'exponentiate = TRUE' to interpret coefficients as ratios
nbin_total_yr.results <- parameters::model_parameters(nbin_total_yr, exponentiate = TRUE) 
```



Because the simulation suggests not overdispersed, that means I can either do Wald Z or chi squared test (Bolker, Brooks, Clark, Geange, Poulsen, Stevens, White; 2008 http://dx.doi.org/10.1016/j.tree.2008.10.008)


```{r}
nbin_Family_yr <- glmmTMB(Total ~ 1+ Family * Shelf + Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = total_recruits)

sim3 <- simulateResiduals(nbin_Family_yr)
testResiduals(sim3)

#seems to be a good model


zbin_Family_yr <- glmmTMB(Total ~ 1+Family * Shelf + Year_Collected + (1 | Reef/Site),
                      ziformula = ~1,
        family = "nbinom2", data = total_recruits)
sim4 <- simulateResiduals(zbin_Family_yr)
testResiduals(sim4)
```

```{r}
AIC(nbin_Family_yr, zbin_Family_yr)
```

```{r}
summary(nbin_Family_yr)
```

  I guess my issue is what to make the baseline that everything is different too... should it be inner:Acroporidae? I don't think so, but I don't know what it should be.

# Q2 is the composition of the recruits different across the shelf?

PERMANOVA - variation within a group can be calculated directly from a distance matrix. 

Assess the fit of a NMDS by plotting the original dissimilarities (diss) against the Euclidean ordination distances (dist). Shepard plot
```{r}
# plot(NMDS$diss, NMDS$dist)
# stressplot(object = NMDS)
```

NMDS ordinations are often plotted without units or labels on the axes so that people do not try to interpret them.

```{r}
# ordiplot(NMDS, type = "n")
# orditorp(NMDS, display = "species", col = "blue", air = 0.01)
# orditorp(NMDS, display = "sites", cex = 0.5, air = 0.01)
```

### PERMANOVA

```{r}
library(vegan)
library(dplyr)
library(tidyr)

# Step 1: Create site-by-species matrix and "Site" column
site_by_species <- Coral_settlement_nGBR %>%
  select(Year_Collected, Reef, Shelf, Site, Tile, Acroporidae, Pocilloporidae, Poritid, Other) 

# Add a unique Site ID *after* extracting metadata
site_by_species <- site_by_species %>%
  mutate(Site_ID = paste(Year_Collected, Reef, Site, Tile, sep = "_"))

# Step 2: Separate metadata and species data
metadata <- site_by_species %>%
  select(Site_ID, Shelf, Year_Collected)

species_matrix <- site_by_species %>%
  select(Site_ID, Acroporidae, Pocilloporidae, Poritid, Other)

# Step 3: Set row names and remove species rows with all zeros
species_matrix_filtered <- species_matrix %>%
  column_to_rownames("Site_ID") %>%
  filter(rowSums(.) > 0)

# Step 4: Match metadata rows to filtered species matrix
metadata_filtered <- metadata %>%
  filter(Site_ID %in% rownames(species_matrix_filtered)) %>%
  column_to_rownames("Site_ID")

# Step 5: PERMANOVA
adonis_result <- vegan::adonis2(species_matrix_filtered ~ Shelf+ Year_Collected,
                                data = metadata_filtered,
                                method = "bray",
                                by = "term",
                                permutations = 999)

print(adonis_result)

```
The PERMANOVA suggests that shelf position and year do significantly affect the coral community composition.
Shelf position explains 17.4% of the variation (strong and significant)
Year collected explains 5% of the variation (suggesting significant temporal differences)
77.5% of the variation in composition of coral recruits remains unexplained.


# Redundancy Analysis
https://pmassicotte.github.io/stats-denmark-2019/07_rda.html#/a-concrete-example-3

Suggests using the Hellinger-transform method (Borcard, Gillet, and Legendre 2011; Legendre and Gallagher 2001) that is suited to species abundance data, 

- gives low weights to variables with low counts and many zeros.
Divide each value in a data matrix by its row sum, and taking the square root of the quotient

```{r}
Recruits <- Coral_settlement_nGBR %>%
  select(Acroporidae, Pocilloporidae, Poritid, Other)
recruit_hellinger <- decostand(Recruits, method = "hellinger")

environment_data <- Coral_settlement_nGBR %>%
  select(Year_Collected, Site, Reef, Shelf) #worried because environment data is nested here.
```

```{r}
my_rda <- rda(recruit_hellinger~Year_Collected + Shelf+ (Reef/Site), data = environment_data)

summary(my_rda)
```
```{r}
plot(my_rda)
```
Explained variance
```{r}
RsquareAdj(my_rda)
```
Tests of RDA results
For canonical analysis such as RDA, model significance relies on permutation tests.

The principle of a permutation test is to generate a reference distribution of the chosen statistic under the null hypothesis H0 by randomly permuting appropriate elements of the data a large number of times and recomputing the statistic each time. Then, one compares the true value of the statistic to this reference distribution. The p value is computed as the proportion of the permuted values equal to or larger than the true (unpermuted) value of the statistic for a one-tailed test in the upper tail, like the F test used in RDA (Borcard, Gillet, and Legendre 2011).

H0: Observed results can be produced by random chance.

The null hypothesis is rejected if this p
value is equal to or smaller than the predefined significance level alpha (ex.: alpha=0.05).

Test for 
1. Global RDA significance
2. Axis significance
3. Terms (explanatory variables) significance

### Global significance test
```{r}
anova.cca(my_rda, permutations = 999)
```
Since p = 0.001, reject H0, meaning that the RDA model is significant and not produced by random chance

### Axis significance
```{r}
anova.cca(my_rda, by = "axis")
```

### Terms significance
```{r}
anova.cca(my_rda, by = "terms")
```

### Check for multicollinearity
Use the vif.cca function
```{r}
sqrt(vif.cca(my_rda))
```
Not sure what the go is with some being NA. "As a rule of thumb, if sqrt vif >2, collinarity is considered high and some of the explanatory cariables are redundant.

I will remove reef and site.
```{r}
my_rda2 <- rda(recruit_hellinger~Year_Collected + Shelf, data = environment_data)

summary(my_rda2)
```

```{r}
plot(my_rda2)

RsquareAdj(my_rda2)

anova.cca(my_rda2, permutations = 999)
```
```{r}
anova.cca(my_rda2, by = "axis")
anova.cca(my_rda2, by = "terms")

sqrt(vif.cca(my_rda2))
```

###Unexplained Variance
Kaiser Guttman criterion

```{r}
my_rda2$CA$eig[my_rda2$CA$eig > mean(my_rda2$CA$eig)]
```

## Partial RDA
It is possible to run an RDA of a matrix Y, explained by a matrix variables X, in the presence of co-variable(s) W.

This analysis allows to display the patterns of the response data (Y) uniquely explained by a linear model of the explanatory variables (X) when the effect of other covariates (W) is held constant (Borcard, Gillet, and Legendre 2011).
### Hold Year constant
Model the effect of shelf position on coral recruit family abundances once the effect of year is removed

```{r}
partial_rda <- rda(recruit_hellinger~ Shelf + Condition(Year_Collected), data = environment_data)
```

```{r}
summary(partial_rda)
```
Conditioned gives the amount of variance that has been explained by the covariables and removed
Constrained gives the amount of variance uniquely explained by the explanatory variables
```{r}
RsquareAdj(partial_rda)
```

### Hold Shelf Position Constant
Model the effect of year on coral recruit family abundances once the effect of shelf position is removed

```{r}
partial_rda2 <- rda(recruit_hellinger~ Year_Collected + Condition(Shelf), data = environment_data)
```
```{r}
summary(partial_rda2)
```

```{r}
RsquareAdj(partial_rda2)
```
Test significance of partial RDA
```{r}
anova.cca(partial_rda2)
```
Since p< =0.05 we reject H0 and can conclude that the result was not produced randomly.

```{r}
plot(partial_rda2, scaling = 1)
```

### Variance partitioning
```{r}
env_shelf <- environment_data[,"Shelf"]

env_year <- environment_data[,"Year_Collected"]
# vp <- varpart(recruit_hellinger, env_shelf, env_year)
# vp
# 
# plot(vp)
```

# Multivariate WorkFlow

### Need to do test of significance of fractions

```{r eval = FALSE}
anova.cca(rda(recruit_hellinger, env_shelf))

anova.cca(rda(recruit_hellinger, env_year))

#[a + b+c]
anova.cca(rda(recruit_hellinger, environment_data))

# [a]
anova.cca(rda(recruit_hellinger, env_shelf, env_year))

# [c]
anova.cca(rda(recruit_hellinger, env_year, env_shelf))
```

# Benthic Data

Importing benthic data
```{r}
PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    )
  )
```

```{r}
sample_design2 <- PIT |> 
  group_by(Year, Reef, Site) |> 
  summarise(PIT = n()) |> 
  arrange(desc(PIT)) 
```

***Note to self - perhaps calculate a proportion of soft coral, given that they are known to have a negative effect on coral recruitment***

```{r}
PIT |> 
  group_by(Year, Site) |> 
  summarise(
    count = n_distinct(Transect)
  ) #the number of transects per year differs among sites, so not sure whether calculating mean and standard error is appropriate or not.


```

## is the total number of recruits related to the number proportion of coral in PIT?

```{r}
# Step 1: Create site-by-species matrix with Site_ID
site_by_PIT <- PIT %>%
  select(
    Year, Reef, Shelf, Site, Transect,
    prop_acro, prop_hard_coral, prop_poci, prop_pori,
    prop_soft_coral, prop_other_hc,
    Prop_DC, Prop_Other, Prop_macroalgae
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  select(Site_ID, Shelf, Year)

# Step 3: Species matrix
PIT_matrix <- site_by_PIT %>%
  select(
    Site_ID,
    prop_acro, prop_hard_coral, prop_poci, prop_pori,
    prop_soft_coral, prop_other_hc,
    Prop_DC, Prop_Other, Prop_macroalgae
  )

# Step 4: Remove all-zero rows BEFORE rownames
PIT_matrix_filtered <- PIT_matrix %>%
  filter(rowSums(across(-Site_ID)) > 0) %>%
  column_to_rownames("Site_ID")

# Step 5: Match metadata
PIT_metadata_filtered <- metadata_PIT %>%
  filter(Site_ID %in% rownames(PIT_matrix_filtered)) %>%
  column_to_rownames("Site_ID")

# Step 6: PERMANOVA
PIT_adonis_result <- adonis2(
  PIT_matrix_filtered ~ Shelf + Year,
  data = PIT_metadata_filtered,
  method = "bray",
  by = "term",
  permutations = 999
)

print(PIT_adonis_result)

set.seed(999)  # reproducibility

PIT_nmds <- metaMDS(
  PIT_matrix_filtered,
  distance = "bray",
  k = 2,            # 2D solution
  trymax = 100,
  autotransform = FALSE
)
```

```{r}
PIT_nmds$stress
```
 Stress isn't great, but usable.
 
```{r}
nmds_scores_PIT <- as.data.frame(scores(PIT_nmds, display = "sites"))

nmds_scores_PIT$Site_ID <- rownames(nmds_scores_PIT)

nmds_scores_PIT <- nmds_scores_PIT %>%
  left_join(
    PIT_metadata_filtered %>% 
      rownames_to_column("Site_ID"),
    by = "Site_ID"
  )

ggplot(nmds_scores_PIT, aes(NMDS1, NMDS2, colour = Shelf)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(type = "t") +
  theme_classic() +
  coord_equal() +
  labs(title = "NMDS of PIT benthic composition")
```
 
```{r}
PIT_envfit <- envfit(
  PIT_nmds,
  PIT_matrix_filtered,
  permutations = 999
)

PIT_envfit

plot(PIT_nmds)
plot(PIT_envfit, p.max = 0.05, col = "black")

```

 
 
```{r}
binomial_acro_benthic <- glmmTMB(data = PIT, prop_acro~Year * Shelf+(1|Reef/Site), family = "binomial")

sim_prop_acro <- simulateResiduals(binomial_acro_benthic)
testResiduals(sim_prop_acro)

summary(binomial_acro_benthic)


binomial_poci_benthic <- glmmTMB(data = PIT, prop_poci~Year * Shelf+(1|Reef/Site), family = "binomial")
sim_prop_poci <- simulateResiduals(binomial_poci_benthic)
testResiduals(sim_prop_poci)
summary(binomial_poci_benthic)

PIT_matrix <- PIT[,95:103]
```


```{r}
library(mvabund)

PIT_abund <- mvabund(PIT[,95:103])
boxplot(PIT[, 95:103], horizontal = TRUE, las = 2, main = "Abundance")

pit_long <- PIT |> 
  pivot_longer(cols = 95:103, names_to = "Benthos", values_to = "Proportion")

pit_long |> 
  ggplot(aes(x=Year, y=Proportion, colour = Site))+
  geom_boxplot()+
  facet_wrap(~Benthos)

pit_long |> 
  group_by(Year, Site, `Depth/Zone`, Benthos) |> 
  summarise(
    mean = mean(Proportion),
    se = sd(Proportion)/(n()^0.5),
    mad = mad(Proportion),
    min = min(Proportion),
    max = max(Proportion),
    n = n(),
    range = max(Proportion)- min(Proportion)
  )
```

I am going to have to make a decision if I just keep the January or October, and then will need to change the year to facilitate an inner join.

```{r}
PIT_long <- PIT |> 
  select(Year, Shelf, Reef, Site, prop_hard_coral, prop_soft_coral, Prop_macroalgae, prop_acro, prop_poci, prop_pori) %>% 
  pivot_longer(cols = 5:10, names_to = "Family", values_to = "Proportion")
```

```{r}
PIT_long |> 
ggplot(aes(x=Year, y= Proportion, colour = Shelf))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Proportion cover (" * PIT^{-1}*")" ),
  x = "Benthos",
  colour = "Shelf position")+
  scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_wrap(~Family)
```

Mid shelf has had variable Acroporidae cover, 

```{r}
PIT_long |> 
ggplot(aes(x=Year, y= Proportion, colour = Site))+
  geom_boxplot(position = "dodge")+
  theme_classic()+
  labs(
  y = expression("Proportion cover (" * Transect^{-1}*")" ),
  x = "Benthos",
  colour = "Shelf position")+
  # scale_colour_manual(values = c("#99CCFF", "#3399FF", "#003366"))+
  facet_grid(Family~Reef, scales = "free_y")
```

Two takeaways:
1. Explicit temporal trend in benthic cover in Macgillvary reef (increasing then decreased final year); and
2. Significant variation in prop_coral and prop_macroalgae in Turtle group reefs.

# Trying something new, to hopefully be able to include the proportion of a family in a model?
```{r}
join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    PIT = n()
  )

Combined <- left_join(Coral_settlement_nGBR, join)

```

```{r}
## Fixed
nbin_Poritid_fixed <- glmmTMB(Poritid ~ Shelf + mean_aco + (1 | Reef/Site) + Year_Collected, 
        family = "nbinom2", data = Combined)

nbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf+mean_poci + (1 | Reef/Site) + Year_Collected, 
        family = "nbinom2", data = Combined)

nbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf + mean_por +(1 | Reef/Site) + Year_Collected, 
        family = "nbinom2", data = Combined)

## Interaction
nbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf*mean_por*Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = Combined)

nbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf *mean_poci* Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = Combined)

nbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf*mean_aco*Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = Combined)
```

```{r}
sim_aco <- simulateResiduals(nbin_Acroporidae_interaction)
testResiduals(sim_aco)

summary(nbin_Acroporidae_interaction) #suspect that I may need to do a LRT test to determine if the model fit is better with or without the average cover of acro
```

```{r}
sim_poci <- simulateResiduals(nbin_Pocilloporidae_interaction)
testResiduals(sim_poci)

summary(nbin_Pocilloporidae_interaction) #Doesn't seem to add to the model much tbh
```


```{r}
sim_por <- simulateResiduals(nbin_Poritid_interaction)
testResiduals(sim_por)

summary(nbin_Pocilloporidae_interaction) #Doesn't seem to add to the model much tbh

```



# Bayesian Analysis 

Model Diagnostics Based on Murray Feedback


```{r include=FALSE, eval=FALSE}
library(tidyverse)  #for data wrangling etc
library(cmdstanr)   #for cmdstan
library(brms)       #for fitting models in STAN
library(standist)   #for exploring distributions
library(HDInterval) #for HPD intervals
library(posterior)  #for posterior draws
library(coda)       #for diagnostics
library(bayesplot)  #for diagnostics
library(rstan)      #for interfacing with STAN
library(DHARMa)     #for residual diagnostics
library(emmeans)    #for marginal means etc
library(broom)      #for tidying outputs
library(broom.mixed) #for tidying MCMC outputs
library(tidybayes)  #for more tidying outputs
library(ggeffects)  #for partial plots
library(patchwork)  #for multiple figures
library(bayestestR) #for ROPE
library(ggridges)   #for ridge plots
library(easystats)     #framework for stats, modelling and visualisation
#source('helperFunctions.R')

ggplot(Coral_settlement_nGBR, aes(y = total, #response
                                  x=Reef))+ #categorical variable
  geom_boxplot()+
  facet_wrap(~Site) 

ggplot(Coral_settlement_nGBR, aes(y = total, #response
                                  x=Shelf))+ #categorical variable
  geom_boxplot()+
  facet_wrap(~Year_Collected) 
```
Suggests that the variance is not equal.

The 0 count skewed
So we would do poisson or negative binomial

```{r include=FALSE, eval=FALSE}
formula.coral1 <- bf(total~(1|Reef) + Year_Collected,
                     family = negbinomial(link = "log"))

get_prior(formula.coral1, data = Coral_settlement_nGBR)
```
Priors would be beta, intercept, sd, and "shape"?

```{r determining priors, include=FALSE, eval=FALSE}
Coral_settlement_nGBR |> 
  group_by(Year_Collected, Reef) |> 
  summarise(median(total),
            mad(total))
```


