---
title: "Dislodged Tiles"
author: "Monique White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
```

Clean data

```{r message=FALSE}
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "dislodged",
    notes == "tile loose on sand, lots of tube worms" ~ "tile loose on sand,\nlots of tube worms",
    TRUE ~ notes
  ))

```

# Determining if loose or dettached tiles are significantly different to other tiles within a site.

```{r}
Coral_settlement_nGBR |> 
  group_by(Reef,notes) |> summarise(n= n())
```

```{r fig.width=10, fig.height=20}
Coral_settlement_nGBR |> 
  ggplot(aes(x=Year, y=total, colour = notes))+
  geom_jitter()+
  geom_boxplot()+
  facet_wrap(~interaction(Site,Reef), scale = "free_y", ncol = 3)+
  theme_bw()
```

Here I want to determine whether the dislodged tiles are significantly different to the tiles in the one position, as I would assume that if they are getting moved in surge, the larvae are less likely to settle on them.

Choose notes and year as fixed factors. Not shelf position because we aren't really interested in that. 


```{r}
dettached_mod <- glmmTMB(total ~ notes + Year_Collected + (1 | Reef/Site), 
      family = "nbinom2", data = Coral_settlement_nGBR)
```

```{r}
sim_mod <- simulateResiduals(dettached_mod)
testResiduals(sim_mod)
```

To me the fit of the residuals seems to fit well. And there were no significant p values for outliers, dispersion, or uniformity. (p>0.05)

```{r}
summary(dettached_mod)
```

What this output says is that when site and reef are random, on average, the tiles that are attached have 2.98 recruits, and that it is significant.

The dislodged tiles have 2.5 recruits less than attached tiles, when reef and site are held constant as a random effect, and this difference is significant at P<0.05.
The tiles with lots of tube worms were slightly lower than those without lots of tube worms. p<0.1

*also run with year as a random variable*
```{r}
dettached_mod <- glmmTMB(total ~ notes + (1 | Reef/Site) + (1| Year_Collected), 
      family = "nbinom2", data = Coral_settlement_nGBR)
```

```{r}
sim_mod <- simulateResiduals(dettached_mod)
testResiduals(sim_mod)
```

```{r}
summary(dettached_mod)
```


Thoughts on this:

  1. Perhaps I should remove the tiles that are dislodged, or include the "notes" variable as a random factor or fixed factor in my models - because it does appear to have a significant effect.