---
title: "Dislodged Tiles"
author: "Monique White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects)
```

Clean data

```{r message=FALSE, echo=FALSE}
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  ),
  Deployment = case_when(
      Year_Collected == "2022" ~ "21/22",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2026" ~ "25/26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes == "Dislodged" ~ "Attached to baseplate",
    TRUE ~ notes
  ))

```

# Determining if loose or dettached tiles are significantly different to other tiles within a site.

```{r}
Coral_settlement_nGBR |> 
  # group_by(notes, Reef,) |> 
  summarise(n= n()) |> 
  summarise(sum = sum(n))

Coral_settlement_nGBR |> 
  group_by(notes, Reef,) |> 
  summarise(n= n()) |> 
  summarise(sum = sum(n))
```

```{r fig.width=10, fig.height=20}
Coral_settlement_nGBR |> 
  ggplot(aes(x=Year, y=total, colour = notes))+
  geom_jitter()+
  geom_boxplot()+
  facet_wrap(~interaction(Site,Reef), scale = "free_y", ncol = 3)+
  theme_bw()
```

Here I want to determine whether the dislodged tiles are significantly different to the tiles in the one position, as I would assume that if they are getting moved in surge, the larvae are less likely to settle on them.

Choose notes and year as fixed factors. Not shelf position because we aren't really interested in that. 

```{r}
dettached_mod <- glmmTMB(total ~ notes + Year_Collected + (1 | Reef/Site), 
      family = "nbinom2", data = Coral_settlement_nGBR)
```

```{r}
sim_mod <- simulateResiduals(dettached_mod)
testResiduals(sim_mod)
```

To me the fit of the residuals seems to fit well. And there were no significant p values for outliers, dispersion, or uniformity. (p>0.05)

```{r}
summary(dettached_mod)
```

What this output says is that when site and reef are random, on average, the tiles that are attached have 2.98 recruits, and that it is significant.

The dislodged tiles have 2.5 recruits less than attached tiles, when reef and site are held constant as a random effect, and this difference is significant at P<0.05.
The tiles with lots of tube worms were slightly lower than those without lots of tube worms. p<0.1

*also run with year as a random variable*
```{r}
dettached_mod2 <- glmmTMB(total ~ notes + (1 | Reef/Site) + (1| Year_Collected), 
      family = "nbinom2", data = Coral_settlement_nGBR)
```

```{r}
sim_mod2 <- simulateResiduals(dettached_mod2)
testResiduals(sim_mod2)
```

```{r}
plot(ggpredict(dettached_mod, terms = "notes"))+
  labs(y= expression("Predicted no. recruits (" * tile^{-1}*")" ), x= "Status of tile at collection")+
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        text = element_text(size=8),
        title = element_blank())

AIC(dettached_mod, dettached_mod2)
# model.sel(dettached_mod, dettached_mod2)
#detached model behaved better
```

```{r}
dislodged_emm <- dettached_mod|> 
  emmeans(~notes, type = "response") |> 
  as.data.frame()

ggplot()+
  geom_pointrange(data = dislodged_emm,aes(x=notes,
             y=response,ymin = asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.85))+
  theme_classic()+
  labs(y = expression("Predicted recruits ( " * tile^{-1} * ")"))+
  scale_x_discrete(
    "Tile at collection",
    labels = c("No notes" = "Attached",
               "Dislodged" = "Dislodged")
    )

d_tukey <- emmeans(dettached_mod, ~notes,  type = "response")
d_tukey

pairs(d_tukey)
# |> 
  # regrid() |> #back transforms 
  # pairs(reverse = FALSE) |> 
  # as.data.frame() |> 
  # mutate(
  #   lower  = estimate - 1.96*SE,
  #   upper  = estimate + 1.96*SE,
  #   CI = 1.96*SE
  # )
d_tukey
```

```{r}
summary(dettached_mod)
```

Thoughts on this:

  1. Perhaps I should remove the tiles that are dislodged, or include the "notes" variable as a random factor or fixed factor in my models - because it does appear to have a significant effect.
  
# Check that site or reef are not driving the trends
Likelihood ratio test in anova?

Doesn't the nesting and having it as a random effect account for this? I just want to ensure I have answered Jenny questions.


# Benthic Data question


## is the total number of recruits related to the number proportion of coral in PIT?

```{r}
# Step 1: Create site-by-species matrix with Site_ID
site_by_PIT <- PIT %>%
  select(
    Year, Reef, Shelf, Site, Transect,
    prop_acro, prop_hard_coral, prop_poci, prop_pori,
    prop_soft_coral, prop_other_hc,
    Prop_DC, Prop_Other, Prop_macroalgae
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  select(Site_ID, Shelf, Year)

# Step 3: Species matrix
PIT_matrix <- site_by_PIT %>%
  select(
    Site_ID,
    prop_acro, prop_hard_coral, prop_poci, prop_pori,
    prop_soft_coral, prop_other_hc,
    Prop_DC, Prop_Other, Prop_macroalgae
  )

# Step 4: Remove all-zero rows BEFORE rownames
PIT_matrix_filtered <- PIT_matrix %>%
  filter(rowSums(across(-Site_ID)) > 0) %>%
  column_to_rownames("Site_ID")

# Step 5: Match metadata
PIT_metadata_filtered <- metadata_PIT %>%
  filter(Site_ID %in% rownames(PIT_matrix_filtered)) %>%
  column_to_rownames("Site_ID")

# Step 6: PERMANOVA
PIT_adonis_result <- adonis2(
  PIT_matrix_filtered ~ Shelf + Year,
  data = PIT_metadata_filtered,
  method = "bray",
  by = "term",
  permutations = 999
)

print(PIT_adonis_result)

set.seed(999)  # reproducibility

PIT_nmds <- metaMDS(
  PIT_matrix_filtered,
  distance = "bray",
  k = 2,            # 2D solution
  trymax = 100,
  autotransform = FALSE
)
```

```{r}
PIT_nmds$stress
```
 Stress isn't great, but usable.
 
```{r}
nmds_scores_PIT <- as.data.frame(scores(PIT_nmds, display = "sites"))

nmds_scores_PIT$Site_ID <- rownames(nmds_scores_PIT)

nmds_scores_PIT <- nmds_scores_PIT %>%
  left_join(
    PIT_metadata_filtered %>% 
      rownames_to_column("Site_ID"),
    by = "Site_ID"
  )

ggplot(nmds_scores_PIT, aes(NMDS1, NMDS2, colour = Shelf)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(type = "t") +
  theme_classic() +
  coord_equal() +
  labs(title = "NMDS of PIT benthic composition")
```

```{r}
PIT_envfit <- envfit(
  PIT_nmds,
  PIT_matrix_filtered,
  permutations = 999
)

PIT_envfit

plot(PIT_nmds)
plot(PIT_envfit, p.max = 0.05, col = "black")
```

*I thought about potentially doing a model of the prop_X, and then saying that the results could be due to that.*
However the issue is that proportional data is [0,1], so therefore I needed to use a binomial model. And the result was terrible

```{r}
binomial_acro_benthic <- glmmTMB(data = PIT, prop_acro~Year * Shelf+(1|Reef/Site), family = "binomial")

sim_prop_acro <- simulateResiduals(binomial_acro_benthic)
testResiduals(sim_prop_acro)

summary(binomial_acro_benthic)


binomial_poci_benthic <- glmmTMB(data = PIT, prop_poci~Year * Shelf+(1|Reef/Site), family = "binomial")
sim_prop_poci <- simulateResiduals(binomial_poci_benthic)
testResiduals(sim_prop_poci)
summary(binomial_poci_benthic)

```