---
title: "Belt Transects"
author: "Monique White"
date: '`r Sys.Date()`'
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function


# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))
Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")

# Mean or median no. tiles per site / year
Join_Coral <- Coral_settlement_nGBR |> 
  group_by(Shelf, Year, Site, Reef) |> 
  summarise(
    st_mean_acro = mean(Acroporidae),
    st_sd_acro = sd(Acroporidae),
    st_median_acro = median(Acroporidae),
    st_mean_poci = mean(Pocilloporidae),
    st_sd_poci = sd(Pocilloporidae),
    st_median_poci = median(Pocilloporidae),
    st_mean_por = mean(Poritid),
    st_sd_por = sd(Poritid),
    st_median_por = median(Poritid),
    st_mean_total = mean(total),
    st_sd_total = sd(total),
    Tiles = n()
  )

# Juvenile data
Belt_10m <- read_csv("Belt_10m.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(total_juveniles),
    sd_total = sd(total_juveniles),
    Belt = n()
  )

Combined1 <- left_join(Coral_settlement_nGBR, join_belt)
# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_hc_oth = mean(prop_other_hc)*100,
    sd_hc_oth = sd(prop_other_hc)*100,
    mean_b_macroalgae = mean(`Brown Macroalgae`),
    sd_b_macroalgae = sd(`Brown Macroalgae`),
    mean_r_macroalgae = mean(`Red Macroalgae`),
    sd_r_macroalgae = sd(`Red Macroalgae`),
    mean_g_macroalgae = mean(`Green Macroalgae`),
    sd_g_macroalgae = sd(`Green Macroalgae`),
    mean_turf = mean(Turf),
    sd_turf = sd(Turf),
    PIT = n()
  )

Combined <- left_join(Coral_settlement_nGBR, join)
```

```{r}
Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    transect = n()
  )

# appears that PIT and belt have the same sample size / no. repeats

```

Relationship between Juveniles and recruits

```{r}
ggplot()+
  geom_point(data = Combined1, aes(x=mean_acro, y= Acroporidae, colour = Shelf))
```

```{r}
ggplot()+
  geom_point(data = Combined1, aes(x=mean_poci, y= Pocilloporidae, colour = Shelf))
```

```{r}
ggplot()+
  geom_point(data = Combined1, aes(x=mean_por, y= Poritid, colour = Shelf))
```

Inner shelf is always translating to lower abundances of the juveniles.

Is there a significant difference in the number of Juveniles across the shelf?

Mean is not a integer, so would need to look at a different model family, such as log.normal, gamma dist, or normal dist.
I also have a feeling that I need to create the variable year + 1 for juveniles, as that would suggest the larval supply from the previous year translated to increased juveniles.
```{r}
ggplot()+
  geom_bar(data = Combined1, aes(x=mean_total))+
  facet_wrap(~Year)
```


Or, hypothesis could be based on 23/24 high larval supply, there was an increase in juveniles relative to the previous years?

Is it driven by larval supply?


```{r}
Model <- glmmTMB(mean_total~Shelf*Year_Collected + (1| Reef/Site), data = Combined1)

sim <- simulateResiduals(Model)
testResiduals(sim) # okay so this default family isn't a good fit, not sure what to go though.
```

```{r}
pois <- glmmTMB(total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Belt_10m)

nbin <- glmmTMB(total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_10m)

```

```{r}
sim <- simulateResiduals(pois)
testResiduals(pois)
```

```{r}
sim <- simulateResiduals(nbin)
testResiduals(nbin) #not the best looking model ever but not significant deviation
```

```{r}
summary(nbin)
```
Interaction terms significant so going to look at pairwise

```{r}
emm_shelf <- emmeans(nbin, ~Shelf | Year)
pairs(emm_shelf, adjust = "tukey")#has estimate whereas the response has ratio

emm_shelf_response <- emmeans(nbin, ~Shelf| Year, type = "response")
#create dataframe
emm_total_df <- as.data.frame(emm_shelf_response) |> 
  rename(predicted_recruits = response)
# Contrast Dataframe
emm_total_pair_df <- as.data.frame(pairs(emm_shelf_response, adjust = "tukey"))

#Calcualte additive differences in total recruits
difference_total <- emm_total_df |> 
  select(Year, Shelf, predicted_recruits) |> 
  pivot_wider(names_from = Shelf, values_from = predicted_recruits) |> 
  mutate(
    `Inner / Mid`   = Inner - Mid,
    `Inner / Outer` = Inner - Outer,
    `Mid / Outer`   = Mid - Outer,
    pct_change_Mid_Inner   = 100 * (Mid - Inner)/Inner,
    pct_change_Outer_Inner = 100 * (Outer - Inner)/Inner,
    pct_change_Outer_Mid   = 100 * (Outer - Mid)/Mid
  )

difference_total

long_total_shelf <- emm_total_pair_df |> 
  select(Year, contrast, ratio, SE, z.ratio, p.value) 

long_diff_total_shelf <- difference_total |> 
  select(Year, starts_with("Inner"), starts_with("Mid")) |> 
  pivot_longer(-Year, names_to = "contrast", values_to = "additive_difference")

final_shelf_total <- left_join(long_total_shelf, long_diff_total_shelf, by = c("Year", "contrast"))

final_shelf_total
```

```{r}
emm_year_response <- emmeans(nbin, ~Year|Shelf, type = "response")
#create dataframe
emm_totalyr_df <- as.data.frame(emm_year_response) |> 
  rename(predicted_recruits = response)
# Contrast Dataframe
emm_totalyr_pair_df <- as.data.frame(pairs(emm_year_response, adjust = "tukey"))

#Calcualte additive differences in total recruits
difference_totalyr <- emm_totalyr_df |> 
  select(Year, Shelf, predicted_recruits) |> 
  pivot_wider(names_from = Year, values_from = predicted_recruits) |> 
  mutate(
    `(2021-22) / (2022-23)` = `2021-22` - `2022-23`,
    `(2021-22) / (2023-24)` = `2021-22` - `2023-24`,
    `(2021-22) / (2024-25)` = `2021-22` - `2024-25`,
    `(2022-23) / (2023-24)` = `2022-23` - `2023-24`,
    `(2022-23) / (2024-25)` = `2022-23` - `2024-25`,
    `(2023-24) / (2024-25)` = `2023-24` - `2024-25`,
    
    pct_change_22_23 = 100 * (`2021-22` - `2022-23`) / `2021-22`,
    pct_change_22_24 = 100 * (`2021-22` - `2023-24`) / `2021-22`,
    pct_change_22_25 = 100 * (`2021-22` - `2024-25`) / `2021-22`,
    pct_change_23_24 = 100 * (`2022-23` - `2023-24`) / `2022-23`,
    pct_change_23_25 = 100 * (`2022-23` - `2024-25`) / `2022-23`,
    pct_change_24_25 = 100 * (`2023-24` - `2024-25`) / `2023-24`
  )


difference_totalyr

long_total_yr <- emm_totalyr_pair_df |> 
  select(Shelf, contrast, ratio, SE, z.ratio, p.value)  

long_diff_total_year <- difference_totalyr |> 
  select(Shelf, starts_with("(2021-22)"), starts_with("(2022-23)"), starts_with("(2023-24)")) |> 
  pivot_longer(-Shelf, names_to = "contrast", values_to = "additive_difference")

final_yr_total <- left_join(long_total_yr, long_diff_total_year, by = c("Shelf", "contrast"))
```

Figure of total juveniles
```{r}
nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
regrid() |> #back transforms 
  pairs(reverse = TRUE) 

total_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=lower, ymax = upper),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted juvenile abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
total_plot
```

```{r}
total_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted juvenile abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
```


Pivot longer
```{r}
Belt_long <- Belt_10m %>% 
  select(Year, Shelf, Reef, Site, Transect, Acroporidae, Pocilloporidae, Poritidae) %>% 
  pivot_longer(cols = 6:8, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritidae","Pocilloporidae")))
```

```{r}
pois <- glmmTMB(Total ~ Family*Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Belt_long)

nbin <- glmmTMB(Total ~ Family*Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_long)
```

```{r}
sim <- simulateResiduals(pois)
testResiduals(pois)
```

```{r}
sim <- simulateResiduals(nbin)
testResiduals(nbin) #fails outlier test
```

Will have to do individual families.
```{r}
ggplot()+
  geom_boxplot(data = Belt_long, aes(x=Year, y=Total, colour = Shelf))+
  facet_wrap(~Family, ncol =1)
```
Mid and outer have higher juvenile abundances of all families typically... This is the inverse of what we see in recruits.

```{r}
Acro <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Acroporidae)

pois <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Acro)

nbin <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Acro)
```

```{r}
sim <- simulateResiduals(pois)
testResiduals(pois)
```

```{r}
sim <- simulateResiduals(nbin)
testResiduals(nbin) #fails outlier test
```
Both meet assumptions
```{r}
model.sel(pois, nbin) #nbin behaves better
```

```{r}
nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
regrid() |> #back transforms 
  pairs(reverse = TRUE) 

Acro_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=lower, ymax = upper),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted juvenile Acroporidae abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

Acro_plot
```

```{r}
Acro_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted juvenile Acroporidae abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

Acro_plot

```


Pocilloporidae
```{r}
Poc <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Pocilloporidae)

pois <- glmmTMB(Pocilloporidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Poc)

nbin <- glmmTMB(Pocilloporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Poc)
```

```{r}
sim <- simulateResiduals(pois)
testResiduals(pois)
```

```{r}
sim <- simulateResiduals(nbin)
testResiduals(nbin) 
```
```{r}
model.sel(pois, nbin) #nbin behaves better
```

```{r}
Poci_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=lower, ymax = upper),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted Pocilloporidae juvenile abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

Poci_plot
```

Asymp instead of the calculated 1.96 x SE
```{r}
Poci_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted Pocilloporidae juvenile abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

Poci_plot
```


Poritidae
```{r}
Por <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Poritidae)

pois <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Por)

nbin <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Por)
```

```{r}
sim <- simulateResiduals(pois)
testResiduals(pois) #outlier significant
```

```{r}
sim <- simulateResiduals(nbin)
testResiduals(nbin) 
```

Both models have outliers as significant.
```{r}
model.sel(pois, nbin) #nbin behaves better
```

```{r}
Poritidae_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=lower, ymax = upper),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted Poritidae juvenile abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

Poritidae_plot+geom_jitter(data = Por, aes(x=Year, y=Poritidae, group = Shelf),
                           position = position_dodge(width = 0.5),
                           colour = "grey")
```

```{r}
Poritidae_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted Poritidae juvenile abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

```


```{r}
ggarrange(total_plot, Acro_plot, Poci_plot, Poritidae_plot, 
          ncol = 1,
          common.legend = TRUE, 
          legend="right",
          labels = c("A", "B", "C", "D")
          ) |> 
ggsave(file = "Juvenile_analysis.png",
       height = 200,
       width = 189,
       units = "mm")
```

```{r}
## ----emm_table------------------------------------------------------------------
emm_table <- function(fit_emm){
  table_out <- data.table(summary(fit_emm))
  if("response" %in% colnames(table_out)){
    # if the lm with log(y) or glm with log link, use " / "
    setnames(table_out,
             old = c("response"),
             new = c("emmean"))
  }
  if("asymp.LCL" %in% colnames(table_out)){
    # if the lm with log(y) or glm with log link, use " / "
    setnames(table_out,
             old = c("asymp.LCL", "asymp.UCL"),
             new = c("lower.CL", "upper.CL"))
  }
  return(table_out)
}
## ----effects_table--------------------------------------------------------------
effects_table <- function(fit_pairs,
                          digits = 2,
                          accuracy = 1e-04,
                          add.p = FALSE){
  if(is.data.frame(fit_pairs) == TRUE){
    pairs_dt <- data.table(fit_pairs)
  }else{
    pairs_dt <- summary(fit_pairs) %>%
      data.table()
  }
 
  if("asymp.LCL" %in% colnames(pairs_dt)){
    # if the lm with log(y) or glm with log link, use " / "
    setnames(pairs_dt,
             old = c("asymp.LCL", "asymp.UCL"),
             new = c("lower.CL", "upper.CL"))
  }
  
  # for simple = each,
  if(which(colnames(pairs_dt)=="contrast") == 3){
    # replace "." with blank
    g_col <- colnames(fit_pairs)[1]
    x_col <- colnames(fit_pairs)[2]
    pairs_dt[get(g_col) == ".", 
             (1) := ""]
    pairs_dt[get(x_col) == ".", 
             (2) := ""]

    # beautify contrast column and group1, group2
    pairs_dt[, contrast := paste0(get(names(pairs_dt)[1]),
                                   get(names(pairs_dt)[2]),
                                   ": ",
                                   contrast)]
  }
  
  # create a column of nicely formatted p-values for display.
  # pairs_dt[, p := pvalString(p.value)]
  pairs_dt[, p := p_format(p.value,
                           digits = digits,
                           accuracy = accuracy,
                           add.p = add.p)]
  pairs_dt[, contrast := factor(contrast, contrast)]
  return(pairs_dt)
}

## ----pvalue_table---------------------------------------------------------------
pvalue_table <- function(fit_pairs){
  if(is.data.frame(fit_pairs) == TRUE){
    pairs_dt <- data.table(fit_pairs)
  }else{
    pairs_dt <- summary(fit_pairs) %>%
      data.table()
  }
  
  if("ratio" %in% colnames(pairs_dt)){
    # if the lm with log(y) or glm with log link, use " / "
    groups <- unlist(str_split(pairs_dt$contrast, " / "))
    setnames(pairs_dt, old = "ratio", new = "estimate")
  }else{
    # if lm use " - "
    groups <- unlist(str_split(pairs_dt$contrast, " - "))
  }

  # remove parentheses if they exist from group cols
  groups <- lapply(groups, remove_parentheses) %>%
    unlist
  
  # add the group1 and group 2 columns
  pairs_dt[, group1 := groups[odd(1:length(groups))]]
  pairs_dt[, group2 := groups[even(1:length(groups))]]
  
  # match for simple and remove parentheses if they exist
  if(names(pairs_dt)[1] != "contrast"){
    x_col <- names(pairs_dt)[2]
    g_col <- names(pairs_dt)[1]
    x_col_levels <- unique(pairs_dt[, get(x_col)])
    x_col_levels <- x_col_levels[x_col_levels != "."]
    g_col_levels <- unique(pairs_dt[, get(g_col)])
    g_col_levels <- g_col_levels[g_col_levels != "."]
    repair <- pairs_dt[, group1] %in% c(x_col_levels, g_col_levels)
    pairs_dt[, group1 := ifelse(repair,
                                group1,
                                gsub(")", "", group1, fixed = TRUE))]
    pairs_dt[, group1 := ifelse(repair,
                                group1,
                                gsub("(", "", group1, fixed = TRUE))]
    repair <- pairs_dt[, group2] %in% c(x_col_levels, g_col_levels)
    pairs_dt[, group2 := ifelse(repair,
                                group2,
                                gsub(")", "", group2, fixed = TRUE))]
    pairs_dt[, group2 := ifelse(repair,
                                group2,
                                gsub("(", "", group2, fixed = TRUE))]
  }
  
  
  if("asymp.LCL" %in% colnames(pairs_dt)){
    # if the lm with log(y) or glm with log link, use " / "
    setnames(pairs_dt,
             old = c("asymp.LCL", "asymp.UCL"),
             new = c("lower.CL", "upper.CL"))
  }
  
  # for simple = each,
  if(which(colnames(pairs_dt)=="contrast") == 3){
    # replace "." with blank
    pairs_dt[get(g_col) == ".", 
             (1) := ""]
    pairs_dt[get(x_col) == ".", 
             (2) := ""]

    # beautify contrast column and group1, group2
    pairs_dt[, contrast := paste0(get(names(pairs_dt)[1]),
                                   get(names(pairs_dt)[2]),
                                   ": ",
                                   contrast)]
    pairs_dt[get(g_col) == "",
             group1 := paste0(get(names(pairs_dt)[1]),
                              get(names(pairs_dt)[2]),
                              ",",
                              group1)]
    pairs_dt[get(g_col) == "",
             group2 := paste0(get(names(pairs_dt)[1]),
                              get(names(pairs_dt)[2]),
                              ",",
                              group2)]
    pairs_dt[get(x_col) == "",
             group1 := paste0(group1,
                              ",",
                              get(names(pairs_dt)[1]),
                              get(names(pairs_dt)[2]))]
    pairs_dt[get(x_col) == "",
             group2 := paste0(group2,
                              ",",
                              get(names(pairs_dt)[1]),
                              get(names(pairs_dt)[2]))]
  }
  
  # create a column of nicely formatted p-values for display.
  pairs_dt[, p := pvalString(p.value)]
  pairs_dt[, contrast := factor(contrast, contrast)]
  return(pairs_dt)
}

## ----response-plot--------------------------------------------------------------
response_plot <- function(
  fit, # model fit from lm, lmer, nlme, glmmTMB
  fit_emm, # data frame with means, error
  x_label = "none",
  y_label = "Response (units)",
  g_label = NULL,
  dots = "sina",
  dodge_width = 0.8,
  adjust = 0.5,
  palette = pal_okabe_ito,
  legend_position = "top",
  flip_horizontal = FALSE,
  group_lines = FALSE
){
   dt <- model.frame(fit) %>% # correctly subsets
    data.table()
  y_col <- insight::find_response(fit)
  
  fit_emm_dt <- emm_table(fit_emm)
  
  # get number of pre emmean/ratio/rate columns
  if("emmean" %in% names(fit_emm_dt)){response <- "emmean"}
  if("rate" %in% names(fit_emm_dt)){response <- "rate"}
  if("ratio" %in% names(fit_emm_dt)){response <- "ratio"}
  
  if(which(names(fit_emm_dt) == response) == 2){
    x_col <- names(fit_emm_dt)[1]
    g_col <- NA
  }else{
    x_col <- names(fit_emm_dt)[1]
    g_col <- names(fit_emm_dt)[2]
  }
    
  rand_int <- find_random(fit)
  
  if(is.na(g_col) & group_lines==TRUE){
    group_lines <- FALSE
  }
  if(is.factor(dt[, get(x_col)]) == FALSE){
    dt[, (x_col) := factor(get(x_col))]
  }
  if(!is.na(g_col)){
    if(is.factor(dt[, get(g_col)]) == FALSE){
      dt[, (g_col) := factor(get(g_col))]
    }
  } 
  
  if(is.na(g_col)){
    pd <- position_dodge(width = 0)
  }else{
    pd <- position_dodge(width = dodge_width)
  }
  
  if(is.na(g_col)){
    color_col <- x_col
  }else{
    color_col <- g_col
  }
  
  gg <- ggplot(data=dt, aes(x = get(x_col),
                            y = get(y_col),
                            color = get(color_col)))
  # plot points
  if(dots == "sina"){
    gg <- gg + geom_sina(alpha = 0.5,
                         position = pd,
                         adjust = adjust)
  }
  if(dots == "jitter"){
    gg <- gg + geom_point(alpha = 0.5,
                          position = position_jitter(seed = 1))
  }
  if(dots == "dotplot"){
    gg <- gg + geom_dotplot(binaxis='y',
                            stackdir='center',
                            alpha = 0.5,
                            position = pd)
  }
  
  # plot means and CI
  gg <- gg +
    geom_errorbar(data = fit_emm_dt, aes(y = get(response),
                                         ymin = lower.CL,
                                         ymax = upper.CL,
                                         color = get(color_col)),
                  width = 0,
                  position = pd
    ) +
    geom_point(data = fit_emm_dt, aes(y = get(response),
                                      color = get(color_col)),
               size = 3,
               position = pd
    ) +
    
    # aesthetics
    ylab(y_label) +
    scale_color_manual(values = palette,
                       name = color_col) +
    theme_pubr() +
    theme(legend.position = legend_position) +
    
    NULL
  
  # if(group_lines == TRUE | group_lines == g_col){
  #   gg <- gg +
  #     geom_line(data = fit_emm_dt,
  #               aes(y = emmean,
  #                   group = get(g_col)),
  #               position = pd)
  # }
  # if(group_lines == x_col){
  #   gg <- gg +
  #     geom_line(data = fit_emm_dt,
  #               aes(y = emmean,
  #                   group = get(x_col)))
  # }
  
  if(is.na(g_col)){
    gg <- gg + theme(legend.position="none")
  }
  
  if(is.null(g_label)){
    gg <- gg + guides(color = guide_legend(title=g_col))
  }else{
    if(g_label == "none"){
      gg <- gg + guides(color = guide_legend(title=NULL))
    }else{
      gg <- gg + guides(color = guide_legend(title=g_label))
    }
  }
  
  # remove x axis title
  if(x_label == "none"){
    gg <- gg + theme(axis.title.x = element_blank())
  }else{
    gg <- gg + xlab(x_label)}
  
  if(flip_horizontal == TRUE){
    gg <- gg + coord_flip() +
      theme(legend.position = legend_position)
    gg <- gg + theme(axis.title.x = element_text())
    if(x_label == "none"){
      gg <- gg + theme(axis.title.y = element_blank())
    }else{
      gg <- gg + ylab(x_label)
    }
  }
  
  # gg
  
  return(gg) 
}
## ----ggplot_the_response--------------------------------------------------------
ggplot_the_response <- function(
 fit, # model fit from lm, lmer, nlme, glmmTMB
  fit_emm,
  fit_pairs,
  wrap_col = NULL,
  x_label = "none",
  y_label = "Response (units)",
  g_label = NULL,
  dots = "sina",
  dodge_width = 0.8,
  adjust = 0.5,
  contrast_rows = "all",
  y_pos = NULL,
  palette = pal_okabe_ito,
  legend_position = "top",
  flip_horizontal = FALSE,
  group_lines = FALSE  
){
  
  # base plot
  gg <- response_plot(
    fit = fit,
    fit_emm = fit_emm,
    x_label = x_label,
    y_label = y_label,
    g_label = g_label,
    dots = dots,
    dodge_width = dodge_width,
    adjust = adjust,
    palette = palette,
    legend_position = legend_position,
    flip_horizontal = flip_horizontal,
    group_lines = group_lines    
  )
  
  # add p-values
  if(contrast_rows[1] != "none" |
     contrast_rows[1] == 0 ){
    gg <- plot_pvalues(
      gg = gg,
      fit_emm = fit_emm,
      fit_pairs = fit_pairs,
      contrast_rows = contrast_rows,
      y_pos = y_pos
    )
  }
  
  gg
  return(gg)
}
```
From the tutorial https://www.middleprofessor.com/files/applied-biostatistics_bookdown/_book/plotting-models

```{r}
nbin_emm <- nbin |> 
  emmeans(~Shelf | Year, type = "response") 
nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
regrid() |> #back transforms 
  pairs(reverse = TRUE) 

nbin_simple <- contrast(nbin_emm,
                        method = "revpairwise") |> 
  summary(infer = TRUE)

nbin_simple
  
ggplot_the_response(
  fit =nbin,
  fit_emm = nbin_emm,
  fit_pairs = nbin_simple
                    )
```


```{r}
combine_belt_st <- left_join(join_belt, Join_Coral)
```

```{r}
combine_belt_st <- combine_belt_st |> 
  mutate(Year_start = as.numeric(substr(Year, 1, 4)))
```

```{r}
gropd_df2 <- combine_belt_st %>%
  arrange(Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_median_acro_lag1 = lag(st_median_acro, 1)
  ) %>%
  ungroup()
```

Remove first year rows
```{r}
model_df <- gropd_df2 %>%
  filter(!is.na(st_median_acro_lag1))
```

Fit model. check if mean_acro is approximately normal
```{r}
m1 <- lmer(mean_acro ~ st_median_acro_lag1 + (1|Reef/Site),
           data = model_df)

m1
summary(m1)
```
```{r}
ggplot()+
  geom_point(data = gropd_df2, aes(x=st_median_acro_lag1, y = mean_acro, colour = Shelf))
```


# Total recruits related to total juveniles

```{r}

total_group <- combine_belt_st %>%
  arrange(Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_mean_total_lag1 = lag(st_mean_total, 1)
  ) %>%
  ungroup()
```

Remove first year rows
```{r}
model_df <- total_group %>%
  filter(!is.na(st_mean_total_lag1))
```

```{r}
ggplot()+
  geom_point(data = model_df, aes(x=st_mean_total_lag1, y = mean_total, colour = Shelf))+
  labs(
    x= "Mean number of recruits per tile (t-1)",
    y= "Mean number of juvenile colonies per 10m2 (t)"
  )
```

We would expect that when larval supply is high (e.g., x is high), that the number of juveniles the following year will also be high (i.e., y high). 

```{r}
ggplot()+
  geom_point(data = model_df, aes(x=st_mean_total_lag1, y = mean_total, colour = Shelf))+
  labs(
    x= "Mean number of recruits per tile (t-1)",
    y= "Mean number of juvenile colonies per 10m2 (t)"
  )+
  theme_bw()+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  facet_wrap(~Shelf, ncol = 1)+
  theme(panel.grid.minor =  element_blank(),
        panel.grid.major.x = element_blank())
```

This relationship doesn't seem to hold. Likely reflecting the fact that 
particularly in inner and outer shelf positions. Outer shelf not many larval settle on tile so unlikely to be a good representation of larval supply

Inner shelf doesn't translate into juveniles due to high competition and settlement tile perhaps acting as a better substrate than the actual substrate available.
