---
title: "Belt Transects"
author: "Monique White"
date: '`r Sys.Date()`'
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function


# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))
Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")

# Juvenile data
Belt_10m <- read_csv("Belt_10m.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(total_juveniles),
    sd_total = sd(total_juveniles),
    Belt = n()
  )

Combined1 <- left_join(Coral_settlement_nGBR, join_belt)
# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_hc_oth = mean(prop_other_hc)*100,
    sd_hc_oth = sd(prop_other_hc)*100,
    mean_b_macroalgae = mean(`Brown Macroalgae`),
    sd_b_macroalgae = sd(`Brown Macroalgae`),
    mean_r_macroalgae = mean(`Red Macroalgae`),
    sd_r_macroalgae = sd(`Red Macroalgae`),
    mean_g_macroalgae = mean(`Green Macroalgae`),
    sd_g_macroalgae = sd(`Green Macroalgae`),
    mean_turf = mean(Turf),
    sd_turf = sd(Turf),
    PIT = n()
  )

Combined <- left_join(Coral_settlement_nGBR, join)
```

Relationship between Juveniles and recruits

```{r}
ggplot()+
  geom_point(data = Combined1, aes(x=mean_acro, y= Acroporidae, colour = Shelf))
```

```{r}
ggplot()+
  geom_point(data = Combined1, aes(x=mean_poci, y= Pocilloporidae, colour = Shelf))
```

```{r}
ggplot()+
  geom_point(data = Combined1, aes(x=mean_por, y= Poritid, colour = Shelf))
```

Inner shelf is always translating to lower abundances of the juveniles.

Is there a significant difference in the number of Juveniles across the shelf?

Mean is not a integer, so would need to look at a different model family, such as log.normal, gamma dist, or normal dist.
I also have a feeling that I need to create the variable year + 1 for juveniles, as that would suggest the larval supply from the previous year translated to increased juveniles.
```{r}
ggplot()+
  geom_bar(data = Combined1, aes(x=mean_total))+
  facet_wrap(~Year)
```


Or, hypothesis could be based on 23/24 high larval supply, there was an increase in juveniles relative to the previous years?

Is it driven by larval supply?


```{r}
Model <- glmmTMB(mean_total~Shelf*Year_Collected + (1| Reef/Site), data = Combined1)

sim <- simulateResiduals(Model)
testResiduals(sim) # okay so this default family isn't a good fit, not sure what to go though.
```

```{r}
pois <- glmmTMB(total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Belt_10m)

nbin <- glmmTMB(total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_10m)

```

```{r}
sim <- simulateResiduals(pois)
testResiduals(pois)
```

```{r}
sim <- simulateResiduals(nbin)
testResiduals(nbin) #not the best looking model ever but not significant deviation
```

```{r}
summary(nbin)
```
Interaction terms significant so going to look at pairwise

```{r}
emm_shelf <- emmeans(nbin, ~Shelf | Year)
pairs(emm_shelf, adjust = "tukey")#has estimate whereas the response has ratio

emm_shelf_response <- emmeans(nbin, ~Shelf| Year, type = "response")
#create dataframe
emm_total_df <- as.data.frame(emm_shelf_response) |> 
  rename(predicted_recruits = response)
# Contrast Dataframe
emm_total_pair_df <- as.data.frame(pairs(emm_shelf_response, adjust = "tukey"))

#Calcualte additive differences in total recruits
difference_total <- emm_total_df |> 
  select(Year, Shelf, predicted_recruits) |> 
  pivot_wider(names_from = Shelf, values_from = predicted_recruits) |> 
  mutate(
    `Inner / Mid`   = Inner - Mid,
    `Inner / Outer` = Inner - Outer,
    `Mid / Outer`   = Mid - Outer,
    pct_change_Mid_Inner   = 100 * (Mid - Inner)/Inner,
    pct_change_Outer_Inner = 100 * (Outer - Inner)/Inner,
    pct_change_Outer_Mid   = 100 * (Outer - Mid)/Mid
  )

difference_total

long_total_shelf <- emm_total_pair_df |> 
  select(Year, contrast, ratio, SE, z.ratio, p.value) 

long_diff_total_shelf <- difference_total |> 
  select(Year, starts_with("Inner"), starts_with("Mid")) |> 
  pivot_longer(-Year, names_to = "contrast", values_to = "additive_difference")

final_shelf_total <- left_join(long_total_shelf, long_diff_total_shelf, by = c("Year", "contrast"))

final_shelf_total
```

```{r}
emm_year_response <- emmeans(nbin, ~Year|Shelf, type = "response")
#create dataframe
emm_totalyr_df <- as.data.frame(emm_year_response) |> 
  rename(predicted_recruits = response)
# Contrast Dataframe
emm_totalyr_pair_df <- as.data.frame(pairs(emm_year_response, adjust = "tukey"))

#Calcualte additive differences in total recruits
difference_totalyr <- emm_totalyr_df |> 
  select(Year, Shelf, predicted_recruits) |> 
  pivot_wider(names_from = Year, values_from = predicted_recruits) |> 
  mutate(
    `(2021-22) / (2022-23)` = `2021-22` - `2022-23`,
    `(2021-22) / (2023-24)` = `2021-22` - `2023-24`,
    `(2021-22) / (2024-25)` = `2021-22` - `2024-25`,
    `(2022-23) / (2023-24)` = `2022-23` - `2023-24`,
    `(2022-23) / (2024-25)` = `2022-23` - `2024-25`,
    `(2023-24) / (2024-25)` = `2023-24` - `2024-25`,
    
    pct_change_22_23 = 100 * (`2021-22` - `2022-23`) / `2021-22`,
    pct_change_22_24 = 100 * (`2021-22` - `2023-24`) / `2021-22`,
    pct_change_22_25 = 100 * (`2021-22` - `2024-25`) / `2021-22`,
    pct_change_23_24 = 100 * (`2022-23` - `2023-24`) / `2022-23`,
    pct_change_23_25 = 100 * (`2022-23` - `2024-25`) / `2022-23`,
    pct_change_24_25 = 100 * (`2023-24` - `2024-25`) / `2023-24`
  )


difference_totalyr

long_total_yr <- emm_totalyr_pair_df |> 
  select(Shelf, contrast, ratio, SE, z.ratio, p.value)  

long_diff_total_year <- difference_totalyr |> 
  select(Shelf, starts_with("(2021-22)"), starts_with("(2022-23)"), starts_with("(2023-24)")) |> 
  pivot_longer(-Shelf, names_to = "contrast", values_to = "additive_difference")

final_yr_total <- left_join(long_total_yr, long_diff_total_year, by = c("Shelf", "contrast"))
```


Pivot longer
```{r}
Belt_long <- Belt_10m %>% 
  select(Year, Shelf, Reef, Site, Transect, Acroporidae, Pocilloporidae, Poritidae) %>% 
  pivot_longer(cols = 6:8, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritidae","Pocilloporidae")))
```

```{r}
pois <- glmmTMB(Total ~ Family*Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Belt_long)

nbin <- glmmTMB(Total ~ Family*Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_long)
```

```{r}
sim <- simulateResiduals(pois)
testResiduals(pois)
```

```{r}
sim <- simulateResiduals(nbin)
testResiduals(nbin) #fails outlier test
```

Will have to do individual families.
```{r}
ggplot()+
  geom_boxplot(data = Belt_long, aes(x=Year, y=Total, colour = Shelf))+
  facet_wrap(~Family, ncol =1)
```
Mid and outer have higher juvenile abundances of all families typically... This is the inverse of what we see in recruits.
