---
title: "Belt Transects"
author: "Monique White"
date: '`r Sys.Date()`'
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function


# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))
Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")

# Mean or median no. tiles per site / year
Join_Coral <- Coral_settlement_nGBR |> 
  group_by(Shelf, Year, Site, Reef) |> 
  summarise(
    st_mean_acro = mean(Acroporidae),
    st_sd_acro = sd(Acroporidae),
    st_median_acro = median(Acroporidae),
    st_mean_poci = mean(Pocilloporidae),
    st_sd_poci = sd(Pocilloporidae),
    st_median_poci = median(Pocilloporidae),
    st_mean_por = mean(Poritid),
    st_sd_por = sd(Poritid),
    st_median_por = median(Poritid),
    st_mean_total = mean(total),
    st_sd_total = sd(total),
    Tiles = n()
  )

# Juvenile data
Belt_10m <- read_csv("Belt_10m.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(total_juveniles),
    sd_total = sd(total_juveniles),
    Belt = n()
  )

Combined1 <- left_join(Coral_settlement_nGBR, join_belt)
# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_hc_oth = mean(prop_other_hc)*100,
    sd_hc_oth = sd(prop_other_hc)*100,
    mean_b_macroalgae = mean(`Brown Macroalgae`),
    sd_b_macroalgae = sd(`Brown Macroalgae`),
    mean_r_macroalgae = mean(`Red Macroalgae`),
    sd_r_macroalgae = sd(`Red Macroalgae`),
    mean_g_macroalgae = mean(`Green Macroalgae`),
    sd_g_macroalgae = sd(`Green Macroalgae`),
    mean_turf = mean(Turf),
    sd_turf = sd(Turf),
    PIT = n()
  )

Combined <- left_join(Coral_settlement_nGBR, join)
```

```{r eval=FALSE, include=FALSE}
Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    transect = n()
  )

# appears that PIT and belt have the same sample size / no. repeats

```

Relationship between Juveniles and recruits

```{r include=FALSE}
ggplot()+
  geom_point(data = Combined1, aes(x=mean_acro, y= Acroporidae, colour = Shelf))
```

```{r include=FALSE}
ggplot()+
  geom_point(data = Combined1, aes(x=mean_poci, y= Pocilloporidae, colour = Shelf))
```

```{r include=FALSE}
ggplot()+
  geom_point(data = Combined1, aes(x=mean_por, y= Poritid, colour = Shelf))
```

Inner shelf is always translating to lower abundances of the juveniles.

Is there a significant difference in the number of Juveniles across the shelf?

Mean is not a integer, so would need to look at a different model family, such as log.normal, gamma dist, or normal dist.
I also have a feeling that I need to create the variable year + 1 for juveniles, as that would suggest the larval supply from the previous year translated to increased juveniles.
```{r include=FALSE}
ggplot()+
  geom_bar(data = Combined1, aes(x=mean_total))+
  facet_wrap(~Year)
```

Or, hypothesis could be based on 23/24 high larval supply, there was an increase in juveniles relative to the previous years?

Is it driven by larval supply?


```{r}
pois <- glmmTMB(total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Belt_10m)

nbin <- glmmTMB(total_juveniles ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_10m)

```

```{r}
sim <- simulateResiduals(pois)
testResiduals(pois)
```

```{r}
sim <- simulateResiduals(nbin)
testResiduals(nbin) #not the best looking model ever but not significant deviation
```

```{r}
summary(nbin)
```
Interaction terms significant so going to look at pairwise

```{r}
emm_shelf <- emmeans(nbin, ~Shelf | Year, type = "response")
pairs(emm_shelf)
```

```{r}
emm_year_response <- emmeans(nbin, ~Year|Shelf, type = "response")
pairs(emm_year_response)
```

Figure of total juveniles
```{r echo=FALSE}
nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
regrid() |> #back transforms 
  pairs(reverse = TRUE) 

total_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted juvenile abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
total_plot
```

```{r echo = FALSE}
total_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted juvenile abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))
```


Pivot longer
```{r}
Belt_long <- Belt_10m %>% 
  select(Year, Shelf, Reef, Site, Transect, Acroporidae, Pocilloporidae, Poritidae) %>% 
  pivot_longer(cols = 6:8, names_to = "Family", values_to = "Total") %>%
  mutate(Family = factor(Family, levels = c("Acroporidae", "Poritidae","Pocilloporidae")))
```

Models for the total don't meet assumptions. Poisson has both KS deviation and significant outliers.
negative binomial has significant outliers. Therefore, I will just model families instead.
```{r}
pois <- glmmTMB(Total ~ Family*Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Belt_long)

nbin <- glmmTMB(Total ~ Family*Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Belt_long)
```

```{r eval=FALSE}
sim <- simulateResiduals(pois) # KS deviation significant and outliers significant
testResiduals(pois)
```

```{r eval=FALSE}
sim <- simulateResiduals(nbin)
testResiduals(nbin) #fails outlier test
```

Will have to do individual families. Just using a boxplot to see how the data typically looks.
```{r echo=FALSE}
ggplot()+
  geom_boxplot(data = Belt_long, aes(x=Year, y=Total, colour = Shelf))+
  facet_wrap(~Family, ncol =1)
```
Mid and outer have higher juvenile abundances of all families typically... This is the inverse of what we see in recruits.

```{r}
Acro <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Acroporidae)

pois <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Acro)

nbin <- glmmTMB(Acroporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Acro)
```

```{r}
sim <- simulateResiduals(pois)
testResiduals(pois)
```

```{r}
sim <- simulateResiduals(nbin)
testResiduals(nbin) #fails outlier test
```
Both meet assumptions
```{r}
model.sel(pois, nbin) #nbin behaves better
```

```{r echo=FALSE}
Acro_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted juvenile Acroporidae abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

Acro_plot
```

```{r echo=FALSE}
Acro_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted juvenile Acroporidae abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

Acro_plot

```


Pocilloporidae
```{r}
Poc <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Pocilloporidae)

pois <- glmmTMB(Pocilloporidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Poc)

nbin <- glmmTMB(Pocilloporidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Poc)
```

```{r}
sim <- simulateResiduals(pois)
testResiduals(pois)
```

```{r}
sim <- simulateResiduals(nbin)
testResiduals(nbin) 
```
```{r}
model.sel(pois, nbin) #nbin behaves better
```

```{r eval = FALSE, echo=FALSE}
Poci_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted Pocilloporidae juvenile abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

Poci_plot
```

Asymp instead of the calculated 1.96 x SE
```{r echo=FALSE}
Poci_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted Pocilloporidae juvenile abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

Poci_plot
```


Poritidae
```{r}
Por <- Belt_10m |> 
  select(Year, Shelf, Reef, Site, Poritidae)

pois <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
      family = "poisson", data = Por)

nbin <- glmmTMB(Poritidae ~ Year*Shelf+(1 | Reef/Site), 
        family = "nbinom2", data = Por)
```

```{r}
sim <- simulateResiduals(pois)
testResiduals(pois) #outlier significant
```

```{r}
sim <- simulateResiduals(nbin)
testResiduals(nbin) 
```

Both models have outliers as significant.
```{r}
model.sel(pois, nbin) #nbin behaves better
```

```{r echo=FALSE}
Poritidae_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted Poritidae juvenile abundance (" * 10~m^{2} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

Poritidae_plot+geom_jitter(data = Por, aes(x=Year, y=Poritidae, group = Shelf),
                           position = position_dodge(width = 0.5),
                           colour = "grey")
```

```{r eval=FALSE, echo=FALSE}
Poritidae_plot <-  nbin |> 
  emmeans(~Shelf | Year, type = "response") |> 
  as.data.frame() |> 
  mutate(
    lower = response - 1.96*SE,
    upper = response + 1.96*SE
  ) |>
  ggplot(aes(x=Year,
             y=response, 
             color=Shelf))+
  geom_pointrange(aes(ymin=asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(width = 0.5))+
  theme_classic()+
  labs(
    y = expression("Predicted Poritidae juvenile abundance (" * belt~transect^{-1} * ")"),,
    colour = "Shelf"
  ) +
  scale_x_discrete(
    "Deployment",
    labels = c("2021-22" = "21/22",
               "2022-23" = "22/23",
               "2023-24" = "23/24",
               "2024-25" = "24/25",
               "2025-26" = "25/26")
    )+
  #facet_wrap(~Family, ncol = 1, scales = "free_y")+
  theme(legend.position = "bottom",
        text = element_text(size = 9))+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))

```


```{r}
ggarrange(total_plot, Acro_plot, Poci_plot, Poritidae_plot, 
          ncol = 1,
          common.legend = TRUE, 
          legend="right",
          labels = c("A", "B", "C", "D")
          ) |> 
ggsave(file = "Juvenile_analysis.png",
       height = 200,
       width = 189,
       units = "mm")
```


```{r}
combine_belt_st <- left_join(join_belt, Join_Coral)
```

```{r}
combine_belt_st <- combine_belt_st |> 
  mutate(Year_start = as.numeric(substr(Year, 1, 4)))
```

```{r}
gropd_df2 <- combine_belt_st %>%
  arrange(Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_median_acro_lag1 = lag(st_median_acro, 1)
  ) %>%
  ungroup()
```

Remove first year rows
```{r}
model_df <- gropd_df2 %>%
  filter(!is.na(st_median_acro_lag1))
```

Fit model. check if mean_acro is approximately normal
```{r}
m1 <- lmer(mean_acro ~ st_median_acro_lag1 + (1|Reef/Site),
           data = model_df)

m1
summary(m1)
```
```{r}
ggplot()+
  geom_point(data = gropd_df2, aes(x=st_median_acro_lag1, y = mean_acro, colour = Shelf))
```


# Total recruits related to total juveniles

```{r}

total_group <- combine_belt_st %>%
  arrange(Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_mean_total_lag1 = lag(st_mean_total, 1)
  ) %>%
  ungroup()
```

Remove first year rows
```{r}
model_df <- total_group %>%
  filter(!is.na(st_mean_total_lag1))
```

```{r}
ggplot()+
  geom_point(data = model_df, aes(x=st_mean_total_lag1, y = mean_total, colour = Shelf))+
  labs(
    x= "Mean number of recruits per tile (t-1)",
    y= "Mean number of juvenile colonies per 10m2 (t)"
  )
```

We would expect that when larval supply is high (e.g., x is high), that the number of juveniles the following year will also be high (i.e., y high). 

```{r}
ggplot()+
  geom_point(data = model_df, aes(x=st_mean_total_lag1, y = mean_total, colour = Shelf))+
  labs(
    x= "Mean number of recruits per tile (t-1)",
    y= "Mean number of juvenile colonies per 10m2 (t)"
  )+
  theme_bw()+
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  facet_wrap(~Shelf, ncol = 1)+
  theme(panel.grid.minor =  element_blank(),
        panel.grid.major.x = element_blank())
```

This relationship doesn't seem to hold. Likely reflecting the fact that 
particularly in inner and outer shelf positions. Outer shelf not many larval settle on tile so unlikely to be a good representation of larval supply

Inner shelf doesn't translate into juveniles due to high competition and settlement tile perhaps acting as a better substrate than the actual substrate available.
