---
title: "Models_10Sep2025"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval=TRUE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan)
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
```

```{r echo=FALSE, message=FALSE}
Coral_settlement_nGBR_CS <- read_csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(`Shelf posn` != "Coral Sea", #remove coral sea data
        Obs != c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf_Position"= `Shelf posn`,
         "Year_Collected"= `Year (Jan)`,
         "Site_#"= `Site #`,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected),
         Shelf_Position = factor(Shelf_Position))

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf_Position != "Inner" | (Shelf_Position %in% c( "Inner", "Mid") & Tile %in% as.character(1:10)))
```

```{r echo=FALSE, include=FALSE, eval=FALSE}
#Coral_settlement_nGBR |> 
  ggplot(aes(x=Year_Collected, y=total, colour = Site))+
  geom_boxplot()+
  facet_wrap(~Shelf_Position, ncol  =1)
```

```{r echo=FALSE}
Coral_settlement_nGBR |> 
  ggplot(aes(x = Year_Collected, y = total, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_wrap(~Shelf_Position, ncol = 1, scale = "free_y")

```

# Summary Statistics

```{r echo=FALSE}
Coral_settlement_nGBR |> 
  summarise(total_tiles = n(),
            total_recruits = sum(total),
            total_acro = sum(Acroporidae),
            total_poci = sum(Pocilloporidae),
            total_poritidae = sum(Poritid),
            total_bottom = sum(`Under Total`),
            total_side = sum(`Side Total`),
            total_top = sum(`Top Total`)) |> 
  knitr::kable()
```
Temporal statistics

```{r echo=FALSE}
Coral_settlement_nGBR |> 
  group_by(Year_Collected) |> 
  summarise(total_tiles = n(),
            total_recruits = sum(total),
            total_acro = sum(Acroporidae),
            total_poci = sum(Pocilloporidae),
            total_poritidae = sum(Poritid),
            total_bottom = sum(`Under Total`),
            total_side = sum(`Side Total`),
            total_top = sum(`Top Total`)) |> 
  knitr::kable()
```


Run generalised linear models with the Poisson, normal binomail and the zero inflated normal bionomial. Total is the dependent variable (the count of coral recruits at a given sampling unit) Shelf position is a fixed effect predicator, it is used to test if the average coral settlement differs between shelf positions Random effects: a nested random effect, modelling the variations between reefs, and variation between sites nested within reefs


Victor: I recommend including Year as a fixed effect since you are interested in differences in recruitment across years, and the levels are not interchangeable. In contrast, a variable like Site would be more appropriate as a random effect in your analysis.

*Should I be scaling or centring the data?*

```{r warning=FALSE, message=FALSE}
pois_total_yr <- glmmTMB(total ~ Shelf_Position + Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = Coral_settlement_nGBR)

nbin_total_yr <- glmmTMB(total ~ Shelf_Position +  Year_Collected +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

zbin_total_yr <- glmmTMB(total ~ Shelf_Position + Year_Collected + (1 | Reef/Site),
                      ziformula = ~1,
        family = "nbinom2", data = Coral_settlement_nGBR)

pois_total_yr2 <- glmmTMB(total ~ Shelf_Position * Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = Coral_settlement_nGBR)

nbin_total_yr2 <- glmmTMB(total ~ Shelf_Position * Year_Collected +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

zbin_total_yr2 <- glmmTMB(total ~ Shelf_Position* Year_Collected + (1 | Reef/Site),
                      ziformula = ~1,
        family = "nbinom2", data = Coral_settlement_nGBR)
```

Use DHARMA to assess residuals

```{r}
sim1.1 <- simulateResiduals(pois_total_yr) 
testResiduals(sim1.1) #not good fit

sim1.2 <- simulateResiduals(pois_total_yr2)
testResiduals(sim1.2) #not good fit

sim1.3 <- simulateResiduals(nbin_total_yr)
testResiduals(sim1.3) #good

sim1.4 <- simulateResiduals(nbin_total_yr2)
testResiduals(sim1.4) #good

sim1.5 <- simulateResiduals(zbin_total_yr)
testResiduals(sim1.5) #good

sim1.6 <- simulateResiduals(zbin_total_yr2)
testResiduals(sim1.6) #good
```

Poisson model residuals were bad, therefore will only compare the four models that testResiduals were good for.

```{r}
AIC(nbin_total_yr, zbin_total_yr, nbin_total_yr2, zbin_total_yr2) |>  #nbin has the lowest AIC with interaction terms
  knitr::kable()
```

negative binomial model has the lowest AIC so I will get the summary of that model.

```{r echo=FALSE}
summary(nbin_total_yr2)
```
2023 and 2024 are significantly different to 2022. 2025 is not significantly different to 2022. Outer shelf is significantly different to inner shelf (p<0.01), Mid shelf is significantly different to inner shelf reefs (p<0.05). Intercept is also significant: not quite sure what that means in a biological framework.

When making a figure should I then be using summary statistics or raw count data? Because the large number of zeros will make it difficult to do a boxplot of the raw data.


# Q2 is the composition of the recruits different across the shelf?

Isolating Data for a single Family of coral

```{r echo=FALSE}
Pocilloporidae <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf_Position, 
         Reef, Site, Tile, 
         Pocilloporidae, Top_Pocil, 
         Under_Pocil, Side_Pocil, total)

Pocilloporidae_orientation <- Pocilloporidae %>% 
  rename(Top = Top_Pocil,
                   Bottom = Under_Pocil,
                   Side = Side_Pocil,
         Recruits_per_tile = total)

Pocilloporidae_orientation <- Pocilloporidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Acroporidae <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf_Position, 
         Reef, Site, Tile, 
         Acroporidae, Top_Acr, 
         Under_Acr, Side_Acr, total)

Acroporidae_orientation <- Acroporidae %>% 
  rename(Top = Top_Acr,
                   Bottom = Under_Acr,
                   Side = Side_Acr,
         Recruits_per_tile = total)

Acroporidae_orientation <- Acroporidae_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")

Poritid <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Shelf_Position, 
         Reef, Site, Tile, 
         Poritid, Top_Por, Under_Por, 
         Side_Por, total)

Poritid_orientation <- Poritid %>% 
  rename(Top = Top_Por,
                   Bottom = Under_Por,
                   Side = Side_Por,
         Recruits_per_tile = total) 

Poritid_orientation <- Poritid_orientation %>% 
  pivot_longer(cols = 7:9,
               names_to = "Tile_Surface",
               values_to = "Recruits")
```

GLMM with Total Recruits for a given coral family, per tile.

To analyze whether there's a difference in the number of recruits of a given coral family (e.g., Poritids) across shelf positions, while accounting for random effects due to site nested within reef and controlling for year, the model you choose depends on:

Whether the response is aggregated or split by tile facet (side/top/bottom),

The structure of your random effects (nested site within reef),

Whether Year_Collected should be a fixed or random effect.

```{r poisson of total Recruits, message=FALSE}
pois_Poritid_fixed <- glmmTMB(Poritid ~ Shelf_Position + Year_Collected+(1 | Reef/Site) , 
      family = "poisson", data = Poritid) 

pois_Poritid_interaction <- glmmTMB(Poritid ~ Shelf_Position*Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = Poritid)

pois_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf_Position + Year_Collected+ (1 | Reef/Site), 
      family = "poisson", data = Acroporidae)

pois_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf_Position*Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = Acroporidae)

pois_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf_Position + Year_Collected  + (1 | Reef/Site), 
      family = "poisson", data = Pocilloporidae)

pois_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf_Position*Year_Collected + (1 | Reef/Site), 
      family = "poisson", data = Pocilloporidae)
```

Negative Binomial of Total Recruits (for a family)

```{r}
## Fixed
nbin_Poritid_fixed <- glmmTMB(Poritid ~ Shelf_Position + (1 | Reef/Site) + Year_Collected, 
        family = "nbinom2", data = Poritid)

nbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf_Position + (1 | Reef/Site) + Year_Collected, 
        family = "nbinom2", data = Pocilloporidae)

nbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf_Position + (1 | Reef/Site) + Year_Collected, 
        family = "nbinom2", data = Acroporidae)

## Interaction
nbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf_Position*Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = Poritid)

nbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf_Position * Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = Pocilloporidae)

nbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf_Position*Year_Collected + (1 | Reef/Site), 
        family = "nbinom2", data = Acroporidae)
```


Negative Binomial Zero Inflated of Total recruits (per family)

```{r}
#Fixed
# Fit model with zero-inflated negative binomial distribution 

zbin_Poritid_fixed <- glmmTMB(Poritid ~ Shelf_Position+ Year_Collected + (1 | Reef/Site) ,
                        ziformula = ~1,
                        family = "nbinom2",
                        data = Poritid)
zbin_Poritid_interaction <- glmmTMB(Poritid ~ Shelf_Position * Year_Collected + (1 | Reef/Site),
                        ziformula = ~1,
                        family = "nbinom2",
                        data = Poritid)

zbin_Pocilloporidae_fixed <- glmmTMB(Pocilloporidae ~ Shelf_Position+ Year_Collected + (1 | Reef/Site),
                               ziformula = ~1,
                        family = "nbinom2",
                        data = Pocilloporidae)

zbin_Pocilloporidae_interaction <- glmmTMB(Pocilloporidae ~ Shelf_Position*Year_Collected + (1 | Reef/Site),
                               ziformula = ~1,
                        family = "nbinom2",
                        data = Pocilloporidae)

zbin_Acroporidae_fixed <- glmmTMB(Acroporidae ~ Shelf_Position + Year_Collected+ (1 | Reef/Site),
                            ziformula = ~1,
                        family = "nbinom2",
                        data = Acroporidae)
zbin_Acroporidae_interaction <- glmmTMB(Acroporidae ~ Shelf_Position*Year_Collected + (1 | Reef/Site),
                            ziformula = ~1,
                        family = "nbinom2",
                        data = Acroporidae)

```

### Acroporidae
```{r}
Acroporidae |> 
  ggplot(aes(x = Year_Collected, y = Acroporidae, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_wrap(~Shelf_Position, ncol = 1, scale = "free_y")
```

Simulate residuals for models

```{r}
sim2.1 <- simulateResiduals(nbin_Acroporidae_fixed)
testResiduals(sim2.1)

sim2.2 <- simulateResiduals(nbin_Acroporidae_interaction)
testResiduals(sim2.2)

sim2.3 <- simulateResiduals(zbin_Acroporidae_fixed)
testResiduals(sim2.3)

sim2.4 <- simulateResiduals(zbin_Acroporidae_interaction)
testResiduals(sim2.4)

sim2.5 <- simulateResiduals(pois_Acroporidae_fixed) #not a good fit
testResiduals(sim2.5)

sim2.6 <- simulateResiduals(pois_Acroporidae_interaction) #not a goodf fit
testResiduals(sim2.6)

```


Compare AIC of those models

```{r}
AIC( nbin_Acroporidae_fixed, zbin_Acroporidae_fixed, nbin_Acroporidae_interaction, zbin_Acroporidae_interaction)
```


```{r echo=FALSE}
nbin_Acroporidae_interaction %>% 
  summary()
```

Acroporidae abundance significantly different in the outer reefs than inshore reefs.

### Pocillopora

```{r}
Pocilloporidae |> 
  ggplot(aes(x = Year_Collected, y = Pocilloporidae, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_wrap(~Shelf_Position, ncol = 1, scale = "free_y")
```


Test Residuals of models for Pocillopora

```{r}
sim3.1 <- simulateResiduals(pois_Pocilloporidae_fixed)
testResiduals(sim3.1)

sim3.2 <- simulateResiduals(pois_Pocilloporidae_interaction)
testResiduals(sim3.2)

sim3.3 <- simulateResiduals(nbin_Pocilloporidae_interaction)
testResiduals(sim3.3)

sim3.4 <- simulateResiduals(nbin_Pocilloporidae_fixed)
testResiduals(sim3.4)

sim3.5 <- simulateResiduals(zbin_Pocilloporidae_fixed)
testResiduals(sim3.5)

sim3.6 <- simulateResiduals(zbin_Pocilloporidae_interaction)
testResiduals(sim3.6)
```

```{r}
AIC(pois_Pocilloporidae_fixed, nbin_Pocilloporidae_fixed, zbin_Pocilloporidae_fixed, nbin_Pocilloporidae_interaction)
```


```{r echo=FALSE}
nbin_Pocilloporidae_interaction %>% 
  summary()
```



Significant difference in Pocilloporidae abundance in mid shore reefs compared to inshore reefs

### Poritid

```{r}
Poritid |> 
  ggplot(aes(x = Year_Collected, y = Poritid, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_wrap(~Shelf_Position, ncol = 1, scale = "free_y")

Poritid |> 
  filter(Poritid>60)
```

Test residuals of models

```{r}
sim4.1 <- simulateResiduals(pois_Poritid_fixed)
testResiduals(sim4.1) # bad model

sim4.2 <- simulateResiduals(pois_Poritid_interaction)
testResiduals(sim4.2) # bad model

sim4.3 <- simulateResiduals(nbin_Poritid_interaction)
testResiduals(sim4.3)

sim4.4 <- simulateResiduals(nbin_Poritid_fixed)
testResiduals(sim4.4)

sim4.5 <- simulateResiduals(zbin_Poritid_fixed)
testResiduals(sim4.5)

sim4.6 <- simulateResiduals(zbin_Poritid_interaction)
testResiduals(sim4.6)
```


```{r echo=FALSE}
AIC(pois_Poritid_fixed, nbin_Poritid_fixed, zbin_Poritid_fixed, nbin_Poritid_interaction, zbin_Poritid_interaction, pois_Poritid_interaction)
```


```{r echo=FALSE}
nbin_Poritid_interaction %>% 
  summary() 
```

Both nbin_interaction, and zbin_interaction for Portidae have significant dispersion.
Don't see much of a difference in Poritid across shelf position. Although slightly less in mid shore reefs (p<0.1)




```{r}
nb_Orientation_Acroporidae <- glmmTMB(Recruits ~ 0+Shelf_Position* Tile_Surface + (1 | Reef/Site), 
        family = "nbinom2", data = Acroporidae_orientation)

pois_Orientation_Acro <- glmmTMB(Recruits ~ 0+Shelf_Position* Tile_Surface + (1 | Reef/Site), 
        family = "poisson", data = Acroporidae_orientation)

zi_orientation_Acro <- glmmTMB(Recruits ~ 0+Shelf_Position* Tile_Surface + (1 | Reef/Site),
                               ziformula = ~1,
        family = "nbinom2", data = Acroporidae_orientation)
```

AIC to determine model of best fit
```{r echo=FALSE}
AIC(nb_Orientation_Acroporidae, pois_Orientation_Acro,zi_orientation_Acro)
```


```{r echo=FALSE}
nb_Orientation_Acroporidae |> 
  summary()
```

```{r echo=FALSE}
sim5 <- simulateResiduals(nb_Orientation_Acroporidae)
testResiduals(sim5)
```



```{r}
sim5.1 <- simulateResiduals(pois_Orientation_Acro)
testResiduals(sim5.1)
```



# Brood Stock Analysis

Check whether benthic composition is behaving consistently between Sites in a reef
```{r}
PIT_50m <- read_csv("PIT_50m.csv", col_types = cols(Shelf = col_character()))

PIT_50m <- PIT_50m |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site)
         ) |> 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))



ggplot(PIT_50m, aes(x=Year, y=prop_coral, colour = Reef))+
  geom_boxplot(aes(group = interaction(Year, Site))) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")

ggplot(PIT_50m, aes(x=Year, y=Prop_macroalgae, colour = Reef))+
  geom_boxplot(aes(group = interaction(Year, Site))) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")


```

### Acroporidae

```{r}
ggplot(PIT_50m, aes(x=Year, y=Acroporidae, colour = Reef))+
  geom_boxplot(aes(group = interaction(Year, Site))) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")
```

### Pocilloporidae

```{r}
ggplot(PIT_50m, aes(x=Year, y=Pocilloporidae, colour = Reef))+
  geom_boxplot(aes(group = interaction(Year, Site))) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")
```

### Poritidae

```{r}
ggplot(PIT_50m, aes(x=Year, y=Poritidae, colour = Reef))+
  geom_boxplot(aes(group = interaction(Year, Site))) +
  facet_wrap(~Shelf, ncol = 1, scale = "free_y")
```

Why is there not benthic surveys at Turtle S in the past two years?

Victor Email: Regarding the PERMANOVA: you start with a community matrix (sites x coral taxa). A PERMANOVA tests whether community composition differs among groups, using a distance matrix derived from your multivariate raw data. The inputs are a dissimilarity (distance) matrix, and a grouping factor (i.e., a categorical variable you want to test, like “reef”, “shelf.position”, or “year”). In your case, I'd suggest calculating a Bray-Curtis distance matrix and then running a PERMANOVA with the vegan package. Here’s a quick example you can run in R:
```{r}
library(vegan)
abund <- PIT_50m |> 
  select(Acroporidae, Pocilloporidae, Poritidae)

prop <- PIT_50m |> 
  select(prop_coral, Prop_macroalgae)
```

Some of the rows do not have Acropora, Pocillopora or Porites present... should I then remove them? Because the distance matrix won't work otherwise.

```{r}
metadata <- PIT_50m |> 
  select(Site, Shelf, Reef, Year) |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site)
  )
```



```{r echo=FALSE}
dist.mat <- vegdist(prop, method = "bray")                                                       # Step 1: Calculate dissimilarity matrix (Bray-Curtis)
```

```{r}
PERMANOVA_prop <- adonis2(dist.mat ~ Reef:Shelf, data = metadata, permutations = 999, by = "terms")   # Step 2: run PERMANOVA

print(PERMANOVA_prop)

PERMANOVA_prop2 <- adonis2(dist.mat ~ Year+Shelf+Year:Shelf, data = metadata, permutations = 999, by = "terms")

print(PERMANOVA_prop2)
```

formula – model formula such as Y ~ A + B * C.  See the ‘Complex Models‘ chapter for more details of how to specify a model formula.  For now, the key point is that we are specifying that we want to analyze the response (Y) as a function of (~) one or more explanatory variables (A, B, C).  The class of Y determines what is done:
If Y is a dissimilarity matrix (i.e., output from dist() or vegdist()), the pseudo-F statistic is calculated directly.
If Y is a data frame or matrix containing data, vegdist() is applied to calculate the distance matrix first and then the pseudo-F statistic is calculated.

- 60% os total community variation (in dissimilarity matrix) explained by Year, SHelf, and their interaction.
Shelf contributes 52.2% of total variation in community composition (proportion macro algae and proportion coral)

### Try to do for abundances, but will need to remove rows where it is 0
```{r}
data_for_abundance <- PIT_50m |>
  mutate(clean = rowSums(across(c(Acroporidae, Pocilloporidae, Poritidae)), na.rm = TRUE))

data_for_abundance <- data_for_abundance |> 
  filter(clean > 0)


abund <- data_for_abundance |> 
  select(Acroporidae, Pocilloporidae, Poritidae) |> 
  mutate(
    Acroporidae = Acroporidae*50,
    Pocilloporidae = Pocilloporidae*50,
    Poritidae = Poritidae*50
  )

abund_std <- abund |> 
  wisconsin()



abund_bray <- vegdist(abund_std, method = "bray")

meta_abund <- data_for_abundance |> 
  select(Shelf, Year, Reef, Site)
```



```{r}
PERMANOVA_abund <- adonis2(abund_bray ~ Year+Shelf+ Year:Shelf, data = meta_abund, permutations = 999, by = "terms")   # Step 2: run PERMANOVA
```

```{r echo=FALSE}
PERMANOVA_abund
```

Abundance is the community matrix data, with 237 sample units of 3 species.

```{r}
set.seed(42)
z <- metaMDS(comm = abund_std,
             autotransform = FALSE,
             distance = "bray",
             engine = "monoMDS",
             k=2,
             weakties = TRUE,
             model = "global",
             maxit = 300,
             try = 40,
             trymax = 100)

print(z)
# Extract NMDS site scores
scores_nmds <- as.data.frame(scores(z, display = "sites"))

# Join to metadata (assuming same row order)
scores_nmds <- scores_nmds |> 
  bind_cols(meta_abund)


ggplot(scores_nmds, aes(x = NMDS1, y = NMDS2, colour = Shelf)) +
  geom_point(size = 3) +
  theme_minimal() +
  labs(
    title = "NMDS (Bray–Curtis)",
    colour = "Shelf Position"
  )

```

Stress <0.05 therefore excellent representation with no prospect of misinterpretation.
https://uw.pressbooks.pub/appliedmultivariatestatistics/chapter/nmds/
```{r}
gof <- goodness(object = z)
plot(z, display = "sites", type = "none")
points(z, display = "sites", cex = 2*gof/mean(gof))
```

Evaluating fit
plot original dissimilarities (z$diss) against Euclidean ordination distances (z$dist).

```{r}
plot(z$diss, z$dist)
```
```{r}
z.points <- data.frame(z$points,
                       Acroporidae = abund_std$Acroporidae,
                       Pocilloporidae = abund_std$Pocilloporidae,
                       Poritid = abund_std$Poritidae)

p <- ggplot(data = z.points, aes(x = MDS1, y = MDS2)) +
theme_bw() +
theme(axis.title = element_blank(),
  axis.ticks = element_blank(),
  axis.text = element_blank())

p + geom_point()


```


# Correlation between brood stock and corals

### Total hard coral cover and total recruits

```{r}
PIT_50m |> 
  group_by(Site, Year, Reef, Shelf) |> 
  summarise(n = n())

```

Would need to do average or something to have a single value for PIT and single value for tiles for a given combination of site, year, and recruits/brood stock.

### Total Acroporidae cover and total acroproidae recruits

### Total Pocilloporidae cover and total Pocillopordiae recruits

# Position on tile (IGNORE)

## Acroporidae

```{r echo=FALSE}
Acroporidae_orientation |> 
  ggplot(aes(x = Year_Collected, y = Recruits, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_grid(Shelf_Position~Tile_Surface, scale = "free_y")
```

## Pocilloporidae

```{r echo=FALSE}
Pocilloporidae_orientation |> 
  ggplot(aes(x = Year_Collected, y = Recruits, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_grid(Shelf_Position~Tile_Surface, scale = "free_y")+
  theme_bw()
```


## Poritidae

```{r}
Poritid_orientation |> 
  ggplot(aes(x = Year_Collected, y = Recruits, fill = Reef)) +
  geom_boxplot(aes(group = interaction(Year_Collected, Site)), 
               position = position_dodge(width = 0.8),
               alpha = 0.6) +
  facet_grid(Shelf_Position~Tile_Surface, scale = "free_y")+
  theme_bw()
```
