---
title: "Multivariate_Analysis_Thesis"
author: "Monique White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r eval=TRUE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
```

Clean data

```{r message=FALSE}
Coral_settlement_nGBR_CS <- read_csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(`Shelf posn` != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR |> 
  filter(notes != "NA")

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf_Position"= `Shelf posn`,
         "Year_Collected"= `Year (Jan)`,
         "Site_#"= `Site #`,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf_Position == "Outer" | (Shelf_Position %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)
```

Notes from Victor email: Regarding the PERMANOVA: you start with a community matrix (sites x coral taxa). A PERMANOVA tests whether community composition differs among groups, using a distance matrix derived from your multivariate raw data. The inputs are a dissimilarity (distance) matrix, and a grouping factor (i.e., a categorical variable you want to test, like “reef”, “shelf.position”, or “year”). In your case, I'd suggest calculating a Bray-Curtis distance matrix and then running a PERMANOVA with the vegan package. 



```{r}
library(data.table)
Recruits <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  select(Acroporidae, Pocilloporidae, Poritid, Other)
recruit_hellinger <- decostand(Recruits, method = "hellinger")



environment_data <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  select(Year_Collected, Site, Reef, Shelf_Position) #worried because environment data is nested here.
```

## PERMANOVA
```{r}
# Filtered recruit data
Recruits_sqrt <- sqrt(Recruits)

# Bray–Curtis PERMANOVA
adonis2(Recruits_sqrt ~ Year_Collected * Shelf_Position, data = environment_data, method = "bray")

#Nested PERMANOVA
adonis2(Recruits_sqrt ~ Year_Collected * Shelf_Position, data = environment_data, method = "bray",  strata = environment_data$Reef )
```

Example: test effects with full nested structure explicitly:

```{r}
adonis2(
  Recruits_sqrt ~ Year_Collected + Shelf_Position + Reef %in% Shelf_Position + Site %in% Reef,
  data = environment_data,
  method = "bray"
)

```

## inspect dispersion (PERMDISP)
```{r}
bray <- vegdist(Recruits_sqrt, method = "bray")

betadisper(bray, environment_data$Shelf_Position)
anova(betadisper(bray, environment_data$Shelf_Position))
```

```{r}
library(vegan)
bray <- vegdist(Recruits_sqrt, method = "bray")

set.seed(999)     # for reproducibility

nmds <- metaMDS(Recruits_sqrt, distance = "bray", k = 2, trymax = 200)

#Check the Stress Value
nmds$stress

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, environment_data[, c("Year_Collected", "Shelf_Position", "Reef", "Site")])
```

Your stress = 0.154 sits comfortably in the:

“Acceptable / good for ecological community data”
“Reliable representation of multivariate structure”

Especially because:

Coral recruit data are sparse, zero-heavy, and nonlinear

Bray–Curtis on square-root transformed data is typical

nMDS often struggles with noisy ecological count data

So 0.15 is strong.

It means:

The 2D ordination is capturing the major differences in assemblages

Patterns you see between groups (Shelf, Year, Reef) are trustworthy

```{r}
p <- ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf_Position,
                         shape = Year_Collected)) +
     geom_point(size = 3, alpha = 0.9) +
    stat_ellipse(aes(group = Shelf_Position, colour = Shelf_Position),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         axis.title = element_text(face = "bold")
     ) +
     labs(
         x = "nMDS1",
        y = "nMDS2",
         title = "nMDS of Coral Recruit Assemblages (Bray–Curtis)",
        colour = "Shelf Position",
        shape = "Year Collected"
    ) +
    annotate("text",
             x = min(scores$MDS1),
             y = min(scores$MDS2),              label = paste0("Stress = ", round(nmds$stress, 3)),
              hjust = 0, vjust = 0, size = 5)
 p

```

## NMDS

Asking the question: how does the coral recruitment composition differ in vegetated tiles...

```{r}
#Create a site-by-species matrix which includes year and reef as the "site"

# Step 1: Create Site column and keep only species data
site_by_species <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Reef, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  unite("Site", Year_Collected, Reef, sep = "_")

# Step 2: Store species-only matrix separately
species_matrix <- site_by_species %>%
  select(Acroporidae, Pocilloporidae, Poritid, Other)

# Remove rows where all species counts are 0
species_matrix_filtered <- species_matrix[rowSums(species_matrix) > 0, ] #this step removes 43 tiles that have no recruits.

# Step 3: Run NMDS
NMDS <- metaMDS(species_matrix_filtered, distance = "bray", autotransform = FALSE, k = 2, trymax = 100)

#NMDS_3 <- metaMDS(species_matrix_filtered, distance = "bray", autotransform = FALSE, k = 3, trymax = 100)


print(NMDS)
#print(NMDS_3)

NMDS$stress
#NMDS_3$stress
#the stress score is 0.118
```

## Zero-adjusted Bray-curtis following methids in Harrison and dela crus, 2024

```{r}
za_recruits <- Coral_settlement_nGBR |> 
  select(Acroporidae, Pocilloporidae, Poritid, Other) |> 
  mutate(
    adj = 1
  )

za_sqrt <- sqrt(za_recruits)

za_env <- Coral_settlement_nGBR |> 
  select(Year_Collected, Site, Reef, Shelf_Position)

```

Transformation
```{r}
# Bray–Curtis PERMANOVA
adonis2(za_sqrt ~ Year_Collected * Shelf_Position, data = za_env, method = "bray")
```

```{r}
set.seed(999)     # for reproducibility

za_nmds <- metaMDS(za_sqrt, distance = "bray", k = 2, trymax = 100)

```

```{r}
print(za_nmds)


za_nmds$stress

za_site_scores <- as.data.frame(vegan::scores(za_nmds, display = "sites"))
za_scores <- cbind(za_site_scores, za_env[, c("Year_Collected", "Shelf_Position", "Reef", "Site")])
```

```{r}
za_nmds_plot <- ggplot(za_scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf_Position,
                         shape = Year_Collected)) +
     geom_point(size = 3, alpha = 0.9) +
    stat_ellipse(aes(group = Shelf_Position, colour = Shelf_Position),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         axis.title = element_text(face = "bold")
     ) +
     labs(
         x = "nMDS1",
        y = "nMDS2",
         title = "nMDS of Coral Recruit Assemblages (Bray–Curtis)",
        colour = "Shelf Position",
        shape = "Year Collected"
    ) +
    annotate("text",
             x = min(scores$MDS1),
             y = min(scores$MDS2),              label = paste0("Stress = ", round(nmds$stress, 3)),
              hjust = 0, vjust = 0, size = 5)
 za_nmds_plot
```

Oh dear oh dear, when I do the zero-adjust and keep in the tiles that had no recruits, the solution did not repeat and the nmds shows more overlap

## Trying to weight by dispersion... 
So I will need to group by site x reef x year for a given family, calculate the var and mean. (based on https://www.int-res.com/articles/meps2006/320/m320p011.pdf)

```{r}
dispersion_weighting <- Coral_settlement_nGBR |> 
  group_by(Year_Collected, Shelf_Position, Reef, Site) |> 
  summarise(
    acro_var = var(Acroporidae),
    acro_mean = mean(Acroporidae),
    poci_var = var(Pocilloporidae),
    poci_mean = mean(Pocilloporidae),
    pori_var = var(Poritid),
    pori_mean = mean(Poritid),
    other_var = var(Other),
    other_mean = mean(Other)
  )

dispersion_weighting <- dispersion_weighting |> 
  mutate(
    "acro_dispersion_index" = acro_var/acro_mean,
    "poci_dispersion_index" = poci_var/poci_mean,
    "pori_dispersion_index" = pori_var/pori_mean
  )

dw_recruits <- left_join(Coral_settlement_nGBR, dispersion_weighting, by = c("Year_Collected", "Shelf_Position" , "Reef", "Site"))

#XiDW = Xij/D actually d with the hat, which is caculated as (x^2/df); and I am not sure what the degrtee of freedom is here
dw_recruits <- dw_recruits |> 
  mutate(
    "acro_value_over_dispersion_index" = Acroporidae/acro_dispersion_index,
    "poci_value_over_dispersion_index" = Pocilloporidae/poci_dispersion_index,
    "pori_value_over_dispersion_index" = Poritid/pori_dispersion_index
  )
```


# Redundancy Analysis
https://pmassicotte.github.io/stats-denmark-2019/07_rda.html#/a-concrete-example-3

Suggests using the Hellinger-transform method (Borcard, Gillet, and Legendre 2011; Legendre and Gallagher 2001) that is suited to species abundance data, 

- gives low weights to variables with low counts and many zeros.
Divide each value in a data matrix by its row sum, and taking the square root of the quotient

```{r}
Recruits <- Coral_settlement_nGBR %>%
  select(Acroporidae, Pocilloporidae, Poritid, Other)
recruit_hellinger <- decostand(Recruits, method = "hellinger")

environment_data <- Coral_settlement_nGBR %>%
  select(Year_Collected, Site, Reef, Shelf_Position) #worried because environment data is nested here.
```

```{r}
my_rda <- rda(recruit_hellinger~Year_Collected + Shelf_Position+ (Reef/Site), data = environment_data)

summary(my_rda)
```
```{r}
plot(my_rda)
```
Explained variance
```{r}
RsquareAdj(my_rda)
```

Tests of RDA results
For canonical analysis such as RDA, model significance relies on permutation tests.

The principle of a permutation test is to generate a reference distribution of the chosen statistic under the null hypothesis H0 by randomly permuting appropriate elements of the data a large number of times and recomputing the statistic each time. Then, one compares the true value of the statistic to this reference distribution. The p value is computed as the proportion of the permuted values equal to or larger than the true (unpermuted) value of the statistic for a one-tailed test in the upper tail, like the F test used in RDA (Borcard, Gillet, and Legendre 2011).

H0: Observed results can be produced by random chance.

The null hypothesis is rejected if this p
value is equal to or smaller than the predefined significance level α (ex.: α=0.05).

Test for 
1. Global RDA significance
2. Axis significance
3. Terms (explanatory variables) significance

### Global significance test
```{r}
anova.cca(my_rda, permutations = 999)
```

Since p = 0.001, reject H0, meaning that the RDA model is significant and not produced by random chance

### Axis significance

```{r}
anova.cca(my_rda, by = "axis")
```

### Terms significance

```{r}
anova.cca(my_rda, by = "terms")
```

### Check for multicollinearity
Use the vif.cca function
```{r}
sqrt(vif.cca(my_rda))
```

Not sure what the go is with some being NA. "As a rule of thumb, if sqrt vif >2, collinarity is considered high and some of the explanatory cariables are redundant.

I will remove reef and site.
```{r}
my_rda2 <- rda(recruit_hellinger~Year_Collected + Shelf_Position, data = environment_data)

summary(my_rda2)
```

```{r}
plot(my_rda2)

RsquareAdj(my_rda2)

anova.cca(my_rda2, permutations = 999)
```

```{r}
anova.cca(my_rda2, by = "axis")
anova.cca(my_rda2, by = "terms")

sqrt(vif.cca(my_rda2))
```

### Unexplained Variance
Kaiser Guttman criterion

```{r}
my_rda2$CA$eig[my_rda2$CA$eig > mean(my_rda2$CA$eig)]
```

# Partial RDA
It is possible to run an RDA of a matrix Y, explained by a matrix variables X, in the presence of co-variable(s) W.

This analysis allows to display the patterns of the response data (Y) uniquely explained by a linear model of the explanatory variables (X) when the effect of other covariates (W) is held constant (Borcard, Gillet, and Legendre 2011).
## Hold Year constant
Model the effect of shelf position on coral recruit family abundances once the effect of year is removed

```{r}
partial_rda <- rda(recruit_hellinger~ Shelf_Position + Condition(Year_Collected), data = environment_data)
```

```{r}
summary(partial_rda)
```

Conditioned gives the amount of variance that has been explained by the covariables and removed
Constrained gives the amount of variance uniquely explained by the explanatory variables
```{r}
RsquareAdj(partial_rda)
```

## Hold Shelf Position Constant

Model the effect of year on coral recruit family abundances once the effect of shelf position is removed

```{r}
partial_rda2 <- rda(recruit_hellinger~ Year_Collected + Condition(Shelf_Position), data = environment_data)
```

```{r}
summary(partial_rda2)
```

```{r}
RsquareAdj(partial_rda2)
```
Test significance of partial RDA
```{r}
anova.cca(partial_rda2)
```

Since p < =0.05 we reject H0 and can conclude that the result was not produced randomly.

```{r}
plot(partial_rda2, scaling = 1)
```

## Variance partitioning
```{r}
env_shelf <- environment_data[,"Shelf_Position"]

env_year <- environment_data[,"Year_Collected"]
vp <- varpart(recruit_hellinger, env_shelf, env_year)
vp

plot(vp)
```

### Need to do test of significance of fractions

```{r}
anova.cca(rda(recruit_hellinger, env_shelf))

anova.cca(rda(recruit_hellinger, env_year))

#[a + b+c]
anova.cca(rda(recruit_hellinger, environment_data))

# [a]
anova.cca(rda(recruit_hellinger, env_shelf, env_year))

# [c]
anova.cca(rda(recruit_hellinger, env_year, env_shelf))
```