---
title: "Multivariate_Analysis_Thesis"
author: "Monique White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
#library(lme4)
#library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(data.table) #not sure if required at this point
library(plotly) #interactive nmds

#import and clean data
Coral_settlement_nGBR_CS <- read_csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(`Shelf posn` != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR |> 
  filter(notes != "NA")

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= `Shelf posn`,
         "Year_Collected"= `Year (Jan)`,
         "Site_#"= `Site #`,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)
Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  )) |> 
  filter(notes != "Dislodged")

```

Notes from Victor email: Regarding the PERMANOVA: you start with a community matrix (sites x coral taxa). A PERMANOVA tests whether community composition differs among groups, using a distance matrix derived from your multivariate raw data. The inputs are a dissimilarity (distance) matrix, and a grouping factor (i.e., a categorical variable you want to test, like “reef”, “shelf.position”, or “year”). In your case, I'd suggest calculating a Bray-Curtis distance matrix and then running a PERMANOVA with the vegan package. 


```{r}
Recruits <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  select(Acroporidae, Pocilloporidae, Poritid, Other)
recruit_hellinger <- decostand(Recruits, method = "hellinger")

environment_data <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  select(Year_Collected, Site, Reef, Shelf) #worried because environment data is nested here.
```

# PERMANOVA

## Sqrt

```{r eval =FALSE}
# Filtered recruit data
Recruits_sqrt <- sqrt(Recruits)

# Bray–Curtis PERMANOVA
adonis2(Recruits_sqrt ~ Year_Collected * Shelf, data = environment_data, method = "bray")

#Nested PERMANOVA
adonis2(Recruits_sqrt ~ Year_Collected * Shelf, data = environment_data, method = "bray",  strata = environment_data$Reef) |> 
print()
```

Example: test effects with full nested structure explicitly:

```{r eval = FALSE}
adonis2(
  Recruits_sqrt ~ Year_Collected + Shelf + Reef %in% Shelf + Site %in% Reef,
  data = environment_data,
  method = "bray"
)
```

## inspect dispersion (PERMDISP)

```{r eval =FALSE}
bray <- vegdist(Recruits_sqrt, method = "bray")

betadisper(bray, environment_data$Shelf)
anova(betadisper(bray, environment_data$Shelf))
```

```{r eval = FALSE}
bray <- vegdist(Recruits_sqrt, method = "bray")

set.seed(999)     # for reproducibility

nmds <- metaMDS(Recruits_sqrt, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, environment_data[, c("Year_Collected", "Shelf", "Reef", "Site")])
```

Your stress = 0.149 sits comfortably in the:

“Acceptable / good for ecological community data”
“Reliable representation of multivariate structure”

Especially because:

Coral recruit data are sparse, zero-heavy, and nonlinear Bray–Curtis on square-root transformed data is typical.
nMDS often struggles with noisy ecological count data

So 0.15 is strong.

It means:

The 2D ordination is capturing the major differences in assemblages

Patterns you see between groups (Shelf, Year, Reef) are trustworthy
*(ChatGPT output - further investigation required)*

```{r eval = FALSE}
p <- ggplot(data =scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf,
                         shape = Year_Collected)) +
     geom_point(size = 3, alpha = 0.9) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         axis.title = element_text(face = "bold")
     ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
     labs(
         x = "nMDS1",
        y = "nMDS2",
         colour = "Shelf Position",
        shape = "Year Collected"
    ) +
    annotate("text",
             x = min(scores$MDS1),
             y = min(scores$MDS2),              label = paste0("Stress = ", round(nmds$stress, 3)),
              hjust = 0, vjust = 0, size = 5)
 p
```
 
```{r}
 # ggplotly(p)
```

## Hellinger transformed

```{r}
# Bray–Curtis PERMANOVA
adonis2(recruit_hellinger ~ Year_Collected * Shelf, data = environment_data, method = "bray", by = "term", permutations = 999) |> 
  print()

# #Nested PERMANOVA
# adonis2(recruit_hellinger ~ Year_Collected * Shelf, data = environment_data, method = "bray",  strata = environment_data$Reef, by= "term", permutations = 999 ) |> 
#   print()
```

Question: I don't understand why these outputs are the same, if one is nested and one is not?


### inspect dispersion (PERMDISP) on Hellinger transformed

```{r}
bray_hellinger <- vegdist(recruit_hellinger, method = "bray")

# With respect to Shelf
betadisper(bray_hellinger, environment_data$Shelf)
anova(betadisper(bray_hellinger, environment_data$Shelf))

# with respect to year
betadisper(bray_hellinger, environment_data$Year_Collected)
anova(betadisper(bray_hellinger, environment_data$Year_Collected))
```

```{r}
set.seed(999)     # for reproducibility

nmds_hellinger <- metaMDS(recruit_hellinger, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds_hellinger$stress

# Extract NMDS for plot
site_scores_hellinger <- as.data.frame(vegan::scores(nmds_hellinger, display = "sites"))
scores_hellinger <- cbind(site_scores_hellinger, environment_data[, c("Year_Collected", "Shelf", "Reef", "Site")])
```

```{r}
Recruits_tile <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  select(Tile, Site, Acroporidae, Pocilloporidae, Poritid, Other)

scores_hellinger <- left_join(scores_hellinger, Recruits_tile)
scores_hellinger <- scores_hellinger |> 
  mutate(
    Deployment =
    case_when(
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2022" ~ "21/22")
  )

p_h <- ggplot(scores_hellinger, aes(x = NMDS1, y = NMDS2,                          colour = Shelf,
                         shape = Year_Collected)) +
     geom_point(size = 3, alpha = 0.9) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         axis.title = element_text(face = "bold")
     ) +
     labs(
         x = "nMDS1",
        y = "nMDS2",
         colour = "Shelf",
        shape = "Year Collected"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
             x = min(scores_hellinger$NMDS1),
             y = min(scores_hellinger$NMDS2),              
             label = paste0("Stress = ", round(nmds_hellinger$stress, 3)),
              hjust = 0, vjust = 0, size = 5)
 p_h
 
 
Figure5A1 <-  ggplot(scores_hellinger, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 3, alpha = 0.9) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)) +
     labs(
         x = "nMDS1",
        y = "nMDS2",
         colour = "Shelf",
        shape = "Year Collected"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
             x = min(scores_hellinger$NMDS1),
             y = min(scores_hellinger$NMDS2),              
             label = paste0("Stress = ", round(nmds_hellinger$stress, 3)),
              hjust = 0, vjust = 0, size = 5)+
   facet_wrap(~ Deployment, ncol = 1)+
  scale_x_continuous(limits = c(-1, 1.6))+
  scale_y_continuous(limits = c(-1.5, 1))
```

```{r}
p_h <- ggplot(scores_hellinger,
              aes(x = NMDS1, y = NMDS2,
                  colour = Shelf,
                  shape = Year_Collected,
                  text = paste(
                    "Tile:", Tile,
                    "<br>Acroporidae:", Acroporidae,
                    "<br>Pocilloporidae:", Pocilloporidae,
                    "<br>Poritid:", Poritid,
                    "<br>Other:", Other
                  ))) +
  geom_point(size = 3, alpha = 0.9) +
  stat_ellipse(aes(group = Shelf, colour = Shelf),
               level = 0.75, linewidth = 1) +
  theme_bw(base_size = 14) +
 theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)) +
  labs(
    x = "nMDS1",
    y = "nMDS2",
    colour = "Shelf",
    shape = "Year Collected"
  ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
  annotate("text",
           x = min(scores_hellinger$NMDS1),
           y = min(scores_hellinger$NMDS2),
           label = paste0("Stress = ", round(nmds_hellinger$stress, 3)),
           hjust = 0, vjust = 0, size = 5)

# ggplotly(p_h, tooltip = "text")

```

```{r}
Recruits_envfit <- envfit(
  nmds_hellinger,
  recruit_hellinger,
  permutations = 999
) 

Recruits_envfit <- as.data.frame(scores(Recruits_envfit, display = "vectors"))

Recruits_envfit <- Recruits_envfit |> 
  rownames_to_column(var = "label_family")

 #check p values to determine whether to include

# Prop other not significant at p<0.05

# plot(PIT_nmds)
# plot(PIT_envfit, p.max = 0.05, col = "black")
```

```{r}
library(ggrepel)

# Decide limits of x and y axis, so that both plots will be the same
scores_hellinger |> 
  summarise(
    x_min = min(NMDS1),
    x_max = max(NMDS1),
    y_min = min(NMDS2),
    y_max = max(NMDS2)
  )

ggplot() +
     geom_point(data =scores_hellinger, aes(x = NMDS1, y = NMDS2,
                                      colour = Shelf,
                         shape = Year_Collected),
                size = 3, alpha = 0.9) +
    stat_ellipse(data =scores_hellinger, aes(x = NMDS1, y = NMDS2,
                                      group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         axis.title = element_text(face = "bold")
     ) +
     labs(
         x = "nMDS1",
        y = "nMDS2",
         colour = "Shelf Position",
        shape = "Year Collected"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
             x = min(scores_hellinger$NMDS1),
             y = min(scores_hellinger$NMDS2),              
             label = paste0("Stress = ", round(nmds_hellinger$stress, 3)),
              hjust = 0, vjust = 0, size = 5)+
geom_segment(data = Recruits_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Recruits_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_family),
                  inherit.aes = FALSE,
                  size = 6,box.padding = 0.5, min.segment.length = Inf) +
  scale_x_continuous(limits = c(-1, 1.6))+
  scale_y_continuous(limits = c(-1.5, 1))
```

The issue with the hellinger transformed nMDS is that there is like an outline of triangle. Not sure what that means.

# Composition of vegetated tiles, hellinger transformed

```{r}
Fig5.A <- ggplot() +
     geom_point(data =scores_hellinger, aes(x = NMDS1, y = NMDS2,
                                      colour = Shelf,
                         shape = Year_Collected),
                size = 3, alpha = 0.9) +
    stat_ellipse(data =scores_hellinger, aes(x = NMDS1, y = NMDS2,
                                      group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
     labs(
         x = "nMDS1",
        y = "nMDS2",
         colour = "Shelf",
        shape = "Deployment"
    ) +
   scale_shape_discrete(
    "Deployment",
    labels = c(
      "2022" = "21/22",
      "2023" = "22/23",
      "2024" = "23/24",
      "2025" = "24/25"
    ))+
    annotate("text",
             x = min(scores_hellinger$NMDS1),
             y = min(scores_hellinger$NMDS2),              
             label = paste0("Stress = ", round(nmds_hellinger$stress, 3)),
              hjust = 0, vjust = 0, size = 4)+
  scale_x_continuous(limits = c(-1, 1.6))+
  scale_y_continuous(limits = c(-1.35, 0.85))

Fig5.A

Fig5.B <- ggplot()+
  geom_segment(data = Recruits_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Recruits_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_family),
                  inherit.aes = FALSE,
                  size = 5,
                  box.padding = 0.5, 
                  min.segment.length = Inf)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "nMDS1",
        y = "nMDS2",
         colour = "Shelf",
        shape = "Year Collected"
    ) +
  scale_x_continuous(limits = c(-1, 1.6))+
  scale_y_continuous(limits = c(-1.35, 0.85))

Figure5 <- ggpubr::ggarrange(Fig5.A, Fig5.B, ncol =1, common.legend = TRUE, labels = c("A", "B"), legend = "right")

#Figure5 <- ggpubr::ggarrange(Figure5A1, Fig5.B, ncol =1, common.legend = TRUE, labels = c("A", "B"), legend = "right", heights = c(1.75, 0.5))


ggsave(filename = "Fig5.png",
       plot = Figure5,
       height = 250,
       width = 169,
       unit = "mm")
```


## Determining what each of the lines represents

```{r}
scores_hellinger |> 
  filter(NMDS1 > 0.5) # outer shelf ones

scores_hellinger |> 
  filter(NMDS2 < 0.5,
         NMDS2 > 0,
         NMDS1 < 0.5)

scores_hellinger |> 
  filter(NMDS2 < 0,
         NMDS1 < 0) #Bottom half of triangle
```


## Zero_adj hellinger
```{r}
Recruit_adj <- Coral_settlement_nGBR |> 
   select(Acroporidae, Pocilloporidae, Poritid, Other) #|> 
  # mutate(
  #   adj = 1
  # )
recruit_h_adj <- decostand(Recruit_adj, method = "hellinger")

recruit_h_adj <- recruit_h_adj |> 
  mutate(
    adj = 0.01
  )

environment_data <- Coral_settlement_nGBR|> 
  select(Year_Collected, Site, Reef, Shelf)
```

```{r}
# Bray–Curtis PERMANOVA
adonis2(recruit_h_adj ~ Year_Collected * Shelf, data = environment_data, method = "bray")

#Nested PERMANOVA
adonis2(recruit_h_adj ~ Year_Collected * Shelf, data = environment_data, method = "bray",  strata = environment_data$Reef )
```

### inspect dispersion (PERMDISP) on Hellinger transformed
```{r}
bray_h_adj <- vegdist(recruit_h_adj, method = "bray")

betadisper(bray_h_adj, environment_data$Shelf)
anova(betadisper(bray_h_adj, environment_data$Shelf))
```

```{r}
set.seed(999)     # for reproducibility

nmds_h_adj <- metaMDS(recruit_h_adj, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds_h_adj$stress

# Extract NMDS for plot
site_scores_h_adj <- as.data.frame(vegan::scores(nmds_h_adj, display = "sites"))
scores_h_adj <- cbind(site_scores_h_adj, environment_data[, c("Year_Collected", "Shelf", "Reef", "Site")])
```

```{r}
p_h_adj <- ggplot(scores_h_adj, aes(x = NMDS1, y = NMDS2,                          colour = Shelf,
                         shape = Year_Collected)) +
     geom_point(size = 3, alpha = 0.9) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         axis.title = element_text(face = "bold")
     ) +
     labs(
         x = "nMDS1",
        y = "nMDS2",
         colour = "Shelf Position",
        shape = "Year Collected"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
             x = min(scores_h_adj$MDS1),
             y = min(scores_h_adj$MDS2),              label = paste0("Stress = ", round(nmds_h_adj$stress, 3)),
              hjust = 0, vjust = 0, size = 5)
 p_h_adj
```

I am going to do the same thing, but do the mutate prior to the transformation.
```{r}
Recruit_adj <- Coral_settlement_nGBR |> 
   select(Acroporidae, Pocilloporidae, Poritid, Other) |> 
  mutate(
    adj = 1
  )
recruit_h_adj <- decostand(Recruit_adj, method = "hellinger")


environment_data <- Coral_settlement_nGBR|> 
  select(Year_Collected, Site, Reef, Shelf)
```

```{r}
# Bray–Curtis PERMANOVA
adonis2(recruit_h_adj ~ Year_Collected * Shelf, data = environment_data, method = "bray")

#Nested PERMANOVA
adonis2(recruit_h_adj ~ Year_Collected * Shelf, data = environment_data, method = "bray",  strata = environment_data$Reef )
```

```{r}
bray_h_adj <- vegdist(recruit_h_adj, method = "bray")

betadisper(bray_h_adj, environment_data$Shelf)
anova(betadisper(bray_h_adj, environment_data$Shelf))
```

```{r}
set.seed(999)     # for reproducibility

nmds_h_adj <- metaMDS(recruit_h_adj, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds_h_adj$stress

# Extract NMDS for plot
site_scores_h_adj <- as.data.frame(vegan::scores(nmds_h_adj, display = "sites"))
scores_h_adj <- cbind(site_scores_h_adj, environment_data[, c("Year_Collected", "Shelf", "Reef", "Site")])
```
```{r}
p_h_adj <- ggplot(scores_h_adj, aes(x = NMDS1, y = NMDS2,                          colour = Shelf,
                         shape = Year_Collected)) +
     geom_point(size = 3, alpha = 0.9) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         axis.title = element_text(face = "bold")
     ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
     labs(
         x = "nMDS1",
        y = "nMDS2",
         colour = "Shelf Position",
        shape = "Year Collected"
    ) +
    annotate("text",
             x = min(scores_h_adj$MDS1),
             y = min(scores_h_adj$MDS2),              label = paste0("Stress = ", round(nmds_h_adj$stress, 3)),
              hjust = 0, vjust = 0, size = 5)
 p_h_adj
```

# NMDS

Asking the question: how does the coral recruitment composition differ in vegetated tiles...

```{r}
#Create a site-by-species matrix which includes year and reef as the "site"

# Step 1: Create Site column and keep only species data
site_by_species <- Coral_settlement_nGBR %>% 
  select(Year_Collected, Reef, Acroporidae, Pocilloporidae, Poritid, Other) %>% 
  unite("Site", Year_Collected, Reef, sep = "_")

# Step 2: Store species-only matrix separately
species_matrix <- site_by_species %>%
  select(Acroporidae, Pocilloporidae, Poritid, Other)

# Remove rows where all species counts are 0
species_matrix_filtered <- species_matrix[rowSums(species_matrix) > 0, ] #this step removes 43 tiles that have no recruits.

# Step 3: Run NMDS
NMDS <- metaMDS(species_matrix_filtered, distance = "bray", autotransform = FALSE, k = 2, trymax = 100)

#NMDS_3 <- metaMDS(species_matrix_filtered, distance = "bray", autotransform = FALSE, k = 3, trymax = 100)

print(NMDS)
#print(NMDS_3)

NMDS$stress
#NMDS_3$stress
#the stress score is 0.118
```

## Zero-adjusted Bray-curtis following methods in Harrison and dela cruz, 2024

```{r}
za_recruits <- Coral_settlement_nGBR |> 
  select(Acroporidae, Pocilloporidae, Poritid, Other) |> 
  mutate(
    adj = 1
  )

za_sqrt <- sqrt(za_recruits)

za_env <- Coral_settlement_nGBR |> 
  select(Year_Collected, Site, Reef, Shelf)

```

Transformation
```{r}
# Bray–Curtis PERMANOVA
adonis2(za_sqrt ~ Year_Collected * Shelf, data = za_env, method = "bray")
```

```{r}
set.seed(999)     # for reproducibility

za_nmds <- metaMDS(za_sqrt, distance = "bray", k = 2, trymax = 100)

```

```{r}
print(za_nmds)

za_nmds$stress

za_site_scores <- as.data.frame(vegan::scores(za_nmds, display = "sites"))
za_scores <- cbind(za_site_scores, za_env[, c("Year_Collected", "Shelf", "Reef", "Site")])
```

```{r}
za_nmds_plot <- ggplot(za_scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf,
                         shape = Year_Collected)) +
     geom_point(size = 3, alpha = 0.9) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         axis.title = element_text(face = "bold")
     ) +
     labs(
         x = "nMDS1",
        y = "nMDS2",
         title = "nMDS of Coral Recruit Assemblages (Bray–Curtis)",
        colour = "Shelf Position",
        shape = "Year Collected"
    ) +
    annotate("text",
             x = min(za_scores$NMDS1),
             y = min(za_scores$NMDS2),              label = paste0("Stress = ", round(za_nmds$stress, 3)),
              hjust = 0, vjust = 0, size = 5)
 za_nmds_plot
```

Oh dear oh dear, when I do the zero-adjust and keep in the tiles that had no recruits, the solution did not repeat and the nmds shows more overlap

### 0.01
Including all tiles.
```{r}
za2_recruits <- Coral_settlement_nGBR |> 
  select(Acroporidae, Pocilloporidae, Poritid, Other) |> 
  mutate(
    adj = 0.01
  )

za2_sqrt <- sqrt(za2_recruits)

za2_env <- Coral_settlement_nGBR |> 
  select(Year_Collected, Site, Reef, Shelf)

```

Transformation
```{r}
# Bray–Curtis PERMANOVA
adonis2(za2_sqrt ~ Year_Collected * Shelf, data = za2_env, method = "bray")
#not sure why i do sqrt here, need to determine that.

adonis2(za2_recruits ~ Year_Collected * Shelf, data = za2_env, method = "bray")
```

```{r}
set.seed(999)     # for reproducibility

za2_nmds <- metaMDS(za2_sqrt, distance = "bray", k = 2, trymax = 100)

za2_nmds <- metaMDS(za2_recruits, distance = "bray", k = 2, trymax = 100)

```

```{r}
print(za2_nmds)


za2_nmds$stress

za2_site_scores <- as.data.frame(vegan::scores(za2_nmds, display = "sites"))
za2_scores <- cbind(za2_site_scores, za2_env[, c("Year_Collected", "Shelf", "Reef", "Site")])
```

```{r}
za2_nmds_plot <- ggplot(za2_scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf,
                         shape = Year_Collected)) +
     geom_point(size = 3, alpha = 0.9) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         axis.title = element_text(face = "bold")
     ) +
     labs(
         x = "nMDS1",
        y = "nMDS2",
         title = "nMDS of Coral Recruit Assemblages (Bray–Curtis)",
        colour = "Shelf Position",
        shape = "Year Collected"
    ) +
    annotate("text",
             x = min(za2_scores$NMDS1),
             y = min(za2_scores$NMDS2),              label = paste0("Stress = ", round(za2_nmds$stress, 3)),
              hjust = 0, vjust = 0, size = 5)

za2_nmds_plot
```


## Trying to weight by dispersion... 
So I will need to group by site x reef x year for a given family, calculate the var and mean. (based on https://www.int-res.com/articles/meps2006/320/m320p011.pdf)

```{r}
dispersion_weighting <- Coral_settlement_nGBR |> 
  group_by(Year_Collected, Shelf, Reef, Site) |> 
  summarise(
    acro_var = var(Acroporidae),
    acro_mean = mean(Acroporidae),
    poci_var = var(Pocilloporidae),
    poci_mean = mean(Pocilloporidae),
    pori_var = var(Poritid),
    pori_mean = mean(Poritid),
    other_var = var(Other),
    other_mean = mean(Other)
  )

dispersion_weighting <- dispersion_weighting |> 
  mutate(
    "acro_dispersion_index" = acro_var/acro_mean,
    "poci_dispersion_index" = poci_var/poci_mean,
    "pori_dispersion_index" = pori_var/pori_mean
  )

dw_recruits <- left_join(Coral_settlement_nGBR, dispersion_weighting, by = c("Year_Collected", "Shelf" , "Reef", "Site"))

#XiDW = Xij/D actually d with the hat, which is caculated as (x^2/df); and I am not sure what the degrtee of freedom is here
dw_recruits <- dw_recruits |> 
  mutate(
    "acro_value_over_dispersion_index" = Acroporidae/acro_dispersion_index,
    "poci_value_over_dispersion_index" = Pocilloporidae/poci_dispersion_index,
    "pori_value_over_dispersion_index" = Poritid/pori_dispersion_index
  )
```


# Benthic data PERMANOVA

```{r}
PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    )
  )
```


```{r}
environment_data <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  select(Year_Collected, Site, Reef, Shelf) #worried because environment data is nested here.
# Bray–Curtis PERMANOVA
adonis2(Recruits_sqrt ~ Year_Collected * Shelf, data = environment_data, method = "bray")

#Nested PERMANOVA
adonis2(Recruits_sqrt ~ Year_Collected * Shelf, data = environment_data, method = "bray",  strata = environment_data$Reef )
```

