---
title: "Murray Meeting 3rd of March"
author: "Monique White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)

#Coral Recruits
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)
Coral_settlement_nGBR$Shelf <- factor(Coral_settlement_nGBR$Shelf)


Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  ),
  Deployment = case_when(
      Year_Collected == "2022" ~ "21/22",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2026" ~ "25/26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")

# Mean or mean no. tiles per site / year
Join_Coral <- Coral_settlement_nGBR |> 
  group_by(Shelf, Year, Site, Reef) |> 
  summarise(
    st_mean_acro = mean(Acroporidae),
    st_sd_acro = sd(Acroporidae),
    st_mean_poci = mean(Pocilloporidae),
    st_sd_poci = sd(Pocilloporidae),
    st_mean_por = mean(Poritid),
    st_sd_por = sd(Poritid),
    st_mean_total = mean(total),
    st_sd_total = sd(total),
    Tiles = n()
  )
# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:109, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

join_pit <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    PIT = n()
  )

# Juvenile data
Belt_10m <- read_csv("Belt_10m.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))
Belt_10m <- Belt_10m |>
  mutate(
    Other_HC = Favia + Favites + Goniastrea + Astrea +
                 Echinopora + Galaxea + Turbinaria +
                 Psammacora + Pavona
  )

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(Total_juveniles),
    sd_total = sd(Total_juveniles),
    Belt = n()
  )
```

# 1. Multiple comparison correction

when the pairs function, I get the Tukey-adjusted p values.
But those P-values are Tukey adjusted within a given year (when ~ Shelf | Year), or in a given Shelf (when the formula is ~ Year | Shelf); right?
I saw some papers then do a Benjamin Hochberg adjustment, and I wonder if I apply that to all p-values finally, to get the significant pairwise contrasts? But even when I do that, if I am applying the BH correction at the point of the finished model, but not both models together at the same time (e.g., the following BH adjustment
```{r}
# emm_shelf <- emmeans(model, ~ Shelf | Year, type = "response")
# contrast_shelf_df <- as.data.frame(
#     summary(pairs(emm_shelf), infer = TRUE)
#   ) %>%
#     dplyr::rename(
#       Ratio = ratio,
#       Lower_CI = asymp.LCL,
#       Upper_CI = asymp.UCL,
#       p_Tukey = p.value )
# 
#   contrast_shelf_df$p_BH <- p.adjust(
#     contrast_shelf_df$p_Tukey,
#     method = "BH" )
```

does not account for the other comparison, (emm_year <- emmeans(model, ~ Year | Shelf, type = "response")).

## 1.1. Example for Total Recruits

```{r}
nbin_total_yr2 <- glmmTMB(total ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)
```

```{r}
# tuk1 <- emmeans:::tukey.emmc(unique(Coral_settlement_nGBR$Shelf))
# tuk2 <- emmeans:::tukey.emmc(unique(Coral_settlement_nGBR$Year))
```

Example with shelf year
```{r}
Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    shelfyear = paste(Shelf, Year)
  )
unique(Coral_settlement_nGBR$shelfyear)
nbin_total_yr2 <- glmmTMB(total ~ shelfyear +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)

summary(nbin_total_yr2)

cmat <- cbind(
  "Inner v Mid 21/22" = c(1, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
  "Inner v Out 21/22" = c(1, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0),
  "Mid v Out 21/22" = c(0, 1, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0),
  "Inner v Mid 22/23" = c(0, 0, 0, 1, -1, 0, 0, 0, 0, 0, 0, 0),
  "Inner v Out 22/23" = c(0, 0, 0, 1, 0, -1, 0, 0, 0, 0, 0, 0),
  "Mid v Out 22/23" = c(0, 0, 0, 0, 1, -1, 0, 0, 0, 0, 0, 0),
  "Inner v Mid 23/24" = c(0, 0, 0, 0, 0, 0, 1, -1, 0, 0, 0, 0),
  "Inner v Out 23/24" = c(0, 0, 0, 0, 0, 0, 1, 0, -1, 0, 0, 0),
  "Mid v Out 23/24" = c(0, 0, 0, 0, 0, 0, 0, 1, -1, 0, 0, 0),
  "Inner v Mid 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, -1, 0),
  "Inner v Out 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, -1),
  "Mid v Out 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, -1),
  "Inner 21/22 v 22/23" = c(1, 0, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0),
  "Inner 21/22 v 23/24" = c(1, 0, 0, 0, 0, 0, -1, 0, 0, 0, 0, 0),
  "Inner 21/22 v 24/25" = c(1, 0, 0, 0, 0, 0, 0, 0, 0, -1, 0, 0),
  "Inner 22/23 v 23/24" = c(0, 0, 0, 1, 0, 0, -1, 0, 0, 0, 0, 0),
  "Inner 22/23 v 24/25" = c(0, 0, 0, 1, 0, 0, 0, 0, 0, -1, 0, 0),
  "Inner 23/24 v 24/25" = c(0, 0, 0, 0, 0, 0, 1, 0, 0, -1, 0, 0),
  "Mid 21/22 v 22/23" = c(0, 1, 0, 0, -1, 0, 0, 0, 0, 0, 0, 0),
  "Mid 21/22 v 23/24" = c(0, 1, 0, 0, 0, 0, 0, -1, 0, 0, 0, 0),
  "Mid 21/22 v 24/25" = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, -1, 0),
  "Mid 22/23 v 23/24" = c(0, 0, 0, 0, 1, 0, 0, -1, 0, 0, 0, 0),
  "Mid 22/23 v 24/25" = c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, -1, 0),
  "Mid 23/24 v 24/25" = c(0, 0, 0, 0, 0, 0, 0, 1, 0, 0, -1, 0),
  "Outer 21/22 v 22/23" = c(0, 0, 1, 0, 0, -1, 0, 0, 0, 0, 0, 0),
  "Outer 21/22 v 23/24" = c(0, 0, 1, 0, 0, 0, 0, 0, -1, 0, 0, 0),
  "Outer 21/22 v 24/25" = c(0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, -1),
  "Outer 22/23 v 23/24" = c(0, 0, 0, 0, 0, 1, 0, 0, -1, 0, 0, 0),
  "Outer 22/23 v 24/25" = c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, -1),
  "Outer 23/24 v 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, -1)
)
nbin_total_yr2 |> emmeans(~shelfyear, type="response") |> 
  contrast(method=list(shelfyear=cmat), adjust= "tukey")
```

```{r}
Coral_settlement_nGBR$shelfyear <- factor(Coral_settlement_nGBR$shelfyear,
                                          levels = c(
    "Inner 2021-22","Mid 2021-22","Outer 2021-22",
    "Inner 2022-23","Mid 2022-23","Outer 2022-23",
    "Inner 2023-24","Mid 2023-24","Outer 2023-24",
    "Inner 2024-25","Mid 2024-25","Outer 2024-25"
  )
)
emm <- emmeans(nbin_total_yr2, ~ shelfyear, type = "response")

levels(emm)
contrast_list <- as.list(as.data.frame(cmat))

contrast(
  emm,
  method = contrast_list,
  adjust = "holm"   # recommended
)
```


```{r}

nbin_total_yr2 <- glmmTMB(total ~ Shelf * Year +(1 | Reef/Site), 
        family = "nbinom2", data = Coral_settlement_nGBR)
## Shelf within Year
  emm_shelf <- emmeans(nbin_total_yr2, ~ Shelf | Year, type = "response")

  emm_shelf_df <- as.data.frame(summary(emm_shelf)) %>%
    dplyr::rename(
      Mean = response,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL
    )

  contrast_shelf_df <- as.data.frame(
    summary(pairs(emm_shelf), infer = TRUE)
  ) %>%
    dplyr::rename(
      Ratio = ratio,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL,
      p_Tukey = p.value
    )

  contrast_shelf_df$p_BH <- p.adjust(
    contrast_shelf_df$p_Tukey,
    method = "BH"
  )
  
contrast_shelf_df

## Year within Shelf

  emm_year <- emmeans(nbin_total_yr2, ~ Year | Shelf, type = "response")

  emm_year_df <- as.data.frame(summary(emm_year)) %>%
    dplyr::rename(
      Mean = response,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL
    )

  contrast_year_df <- as.data.frame(
    summary(pairs(emm_year), infer = TRUE)
  ) %>%
    dplyr::rename(
      Ratio = ratio,
      Lower_CI = asymp.LCL,
      Upper_CI = asymp.UCL,
      p_Tukey = p.value
    )

  contrast_year_df$p_BH <- p.adjust(
    contrast_year_df$p_Tukey,
    method = "BH"
  )

contrast_year_df
```


# 2. Brood stock (PIT) and Recruitment (Tiles)

```{r}
st_join <- Coral_settlement_nGBR |> 
  group_by(Site, Year) |> 
  summarise(
    acro_mean_st = mean(Acroporidae),
    poci_mean_st = mean(Pocilloporidae),
    por_mean_st = mean(Poritid),
    acro_sd_st = sd(Acroporidae),
    poci_sd_st = sd(Pocilloporidae),
    por_sd_st = sd(Poritid),
    acro_median_st = median(Acroporidae),
    poci_median_st = median(Pocilloporidae),
    por_median_st = median(Poritid),
  )

combined_averages <- left_join(join_pit, st_join)
```

Now these are both averages, it can't follow an integer and would need to follow a different family of error distribution.

## 2.1. Acroporidae

Despite not being interested in across the shelf differences, as shelf has only three levels there are two few levels for a stable random effect. (Is that correct?)
Or should I be including Shelf as a random effect too?

*Note that I could not get the Gamma family ones to run in glmmTMB, despite filtering so that the response variable (acro_mean_st) was not 0*
Maybe centre data to make the intercept make more sense
scale(mean_aco, scale = FALSE) only centreing
scale(mean_aco) centre and scale (mean of 0 and SD 1)

```{r}
Acro <- combined_averages |> 
  filter(acro_mean_st > 0)

gamma_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ scale(mean_aco, scale = FALSE) +(1 | Reef/Site) + (1 | Year),
        family = Gamma(link = "log"), data = Acro)

ln_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Reef/Site) + (1 | Year), 
        family = "lognormal", data = Acro)

gaus_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Reef/Site) + (1 | Year), 
        family = "gaussian", data = Acro)

```

```{r}
sim13.1 <- simulateResiduals(gamma_avg_Acroporidae_fixed) #didn't work
testResiduals(sim13.1)
testZeroInflation(sim13.1)
# sim13.2 <- simulateResiduals(ln_avg_Acroporidae_fixed)
# testResiduals(sim13.2)
# testZeroInflation(sim13.2)
# 
# sim13.3 <- simulateResiduals(gaus_avg_Acroporidae_fixed)
# testResiduals(sim13.3) # definitelty not this one
# testZeroInflation(sim13.3)
```

Log normal models worked best for all of these
```{r}
summary(ln_avg_Acroporidae_fixed)
```
When I am reporting this output, the mean cover is not a significant predictor of recruitment, and the estimate (slope) is nearly 0.
Just want to confirm I am reporting the correct thing.
Otherwise I would use a Spearmans correlation that may also make more sense and be easier to report.

# 3. Recruitment and Juveniles
These models will need to be lagged. 

```{r}
combine_belt_st <- left_join(join_belt, Join_Coral)

# Only including years in which both data is included (once lagging has happened)
combine_belt_st <- combine_belt_st |> 
  mutate(Year_start = as.numeric(substr(Year, 1, 4)))

gropd_df2 <- combine_belt_st %>%
  arrange(Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_mean_acro_lag1 = lag(st_mean_acro, 1),
    st_mean_poci_lag1 = lag(st_mean_poci, 1),
    st_mean_por_lag1 = lag(st_mean_por, 1),
  ) %>%
  ungroup()
```


## 3.1. Acroporidae

Remove first year row
```{r}
model_df_acro <- gropd_df2 %>%
  filter(!is.na(mean_acro))
```


Fit model. check if mean_acro is approximately normal. *Note this may be the wrong approach. I am not sure if this method of shopping around with variables is problematic, and that I should have a model in mind before fitting... The reason I sometimes left out variables was to see if they were contributing enough to justify the penalty of having them...*


```{r}
ln_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + Shelf + Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)
# sim <- simulateResiduals(ln_acro_juv_st_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim)

gau_acro_juv_st_fixed <- glmmTMB(mean_acro ~ st_mean_acro_lag1 +Shelf + Year + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)
# sim <- simulateResiduals(gau_acro_juv_st_fixed)
# testResiduals(sim) #passes
# testZeroInflation(sim)

#without year
ln_acro_juv_st_int_wo_yr <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)
# sim <- simulateResiduals(ln_acro_juv_st_int_wo_yr)
# testResiduals(sim) # passed
# testZeroInflation(sim)

gau_acro_juv_st_int_wo_yr<- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)
# sim <- simulateResiduals(gau_acro_juv_st_int_wo_yr)
# testResiduals(sim) # passed but doesn't look great
# testZeroInflation(sim)

ln_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf * Year + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)
# sim <- simulateResiduals(ln_acro_juv_st_int)
# testResiduals(sim) #outlier test significant
# testZeroInflation(sim)

gau_acro_juv_st_int <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf * Year + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)
# sim <- simulateResiduals(gau_acro_juv_st_int)
# testResiduals(sim) #passes
# testZeroInflation(sim)

# year as random variable
ln_acro_juv_st_int_yr <- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site) + (1 | Year),
              family = "lognormal",
           data = model_df_acro)
# sim <- simulateResiduals(ln_acro_juv_st_int_yr)
# testResiduals(sim) # passed
# testZeroInflation(sim)

gau_acro_juv_st_int_yr<- glmmTMB(mean_acro ~ st_mean_acro_lag1 * Shelf + (1|Reef/Site) + (1 | Year),
              family = "gaussian",
           data = model_df_acro)
# sim <- simulateResiduals(gau_acro_juv_st_int_yr)
# testResiduals(sim) # passed
# testZeroInflation(sim)
```

```{r}
model.sel(gau_acro_juv_st_int, gau_acro_juv_st_fixed, ln_acro_juv_st_fixed, gau_acro_juv_st_int_wo_yr, ln_acro_juv_st_int_wo_yr, gau_acro_juv_st_int_yr, ln_acro_juv_st_int_yr)
```

```{r echo=FALSE}
model.sel(gau_acro_juv_st_int, gau_acro_juv_st_fixed, ln_acro_juv_st_fixed, gau_acro_juv_st_int_wo_yr, ln_acro_juv_st_int_wo_yr, gau_acro_juv_st_int_yr, ln_acro_juv_st_int_yr) |> knitr::kable()
```

No inferential justification for having interaction term
```{r}
library(multcomp)
```

```{r}
Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    shelfyear = paste(Shelf, Year)
  )
Coral_settlement_nGBR$shelfyear <- factor(Coral_settlement_nGBR$shelfyear,
                                          levels = c(
    "Inner 2021-22","Mid 2021-22","Outer 2021-22",
    "Inner 2022-23","Mid 2022-23","Outer 2022-23",
    "Inner 2023-24","Mid 2023-24","Outer 2023-24",
    "Inner 2024-25","Mid 2024-25","Outer 2024-25"
  )
)
nbin_total_yr2 <- glmmTMB(
  total ~ 0 + shelfyear + (1 | Reef/Site),
  family = "nbinom2",
  data = Coral_settlement_nGBR
)

summary(nbin_total_yr2)

cmat <- cbind(
  "Inner v Mid 21/22" = c(1, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
  "Inner v Out 21/22" = c(1, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0),
  "Mid v Out 21/22" = c(0, 1, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0),
  "Inner v Mid 22/23" = c(0, 0, 0, 1, -1, 0, 0, 0, 0, 0, 0, 0),
  "Inner v Out 22/23" = c(0, 0, 0, 1, 0, -1, 0, 0, 0, 0, 0, 0),
  "Mid v Out 22/23" = c(0, 0, 0, 0, 1, -1, 0, 0, 0, 0, 0, 0),
  "Inner v Mid 23/24" = c(0, 0, 0, 0, 0, 0, 1, -1, 0, 0, 0, 0),
  "Inner v Out 23/24" = c(0, 0, 0, 0, 0, 0, 1, 0, -1, 0, 0, 0),
  "Mid v Out 23/24" = c(0, 0, 0, 0, 0, 0, 0, 1, -1, 0, 0, 0),
  "Inner v Mid 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, -1, 0),
  "Inner v Out 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, -1),
  "Mid v Out 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, -1),
  "Inner 21/22 v 22/23" = c(1, 0, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0),
  "Inner 21/22 v 23/24" = c(1, 0, 0, 0, 0, 0, -1, 0, 0, 0, 0, 0),
  "Inner 21/22 v 24/25" = c(1, 0, 0, 0, 0, 0, 0, 0, 0, -1, 0, 0),
  "Inner 22/23 v 23/24" = c(0, 0, 0, 1, 0, 0, -1, 0, 0, 0, 0, 0),
  "Inner 22/23 v 24/25" = c(0, 0, 0, 1, 0, 0, 0, 0, 0, -1, 0, 0),
  "Inner 23/24 v 24/25" = c(0, 0, 0, 0, 0, 0, 1, 0, 0, -1, 0, 0),
  "Mid 21/22 v 22/23" = c(0, 1, 0, 0, -1, 0, 0, 0, 0, 0, 0, 0),
  "Mid 21/22 v 23/24" = c(0, 1, 0, 0, 0, 0, 0, -1, 0, 0, 0, 0),
  "Mid 21/22 v 24/25" = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, -1, 0),
  "Mid 22/23 v 23/24" = c(0, 0, 0, 0, 1, 0, 0, -1, 0, 0, 0, 0),
  "Mid 22/23 v 24/25" = c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, -1, 0),
  "Mid 23/24 v 24/25" = c(0, 0, 0, 0, 0, 0, 0, 1, 0, 0, -1, 0),
  "Outer 21/22 v 22/23" = c(0, 0, 1, 0, 0, -1, 0, 0, 0, 0, 0, 0),
  "Outer 21/22 v 23/24" = c(0, 0, 1, 0, 0, 0, 0, 0, -1, 0, 0, 0),
  "Outer 21/22 v 24/25" = c(0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, -1),
  "Outer 22/23 v 23/24" = c(0, 0, 0, 0, 0, 1, 0, 0, -1, 0, 0, 0),
  "Outer 22/23 v 24/25" = c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, -1),
  "Outer 23/24 v 24/25" = c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, -1)
)
cmat <- t(cmat)
#glht(nbin_total_yr2, linfct = cmat(period = "Tukey"))
glht_res <- glht(nbin_total_yr2, linfct = mcp(shelfyear = cmat), type = "tukey")

glht_res <- glht(nbin_total_yr2, linfct = mcp(shelfyear = "Tukey"))
summary(glht_res)
```

