---
title: "PERMANOVA NMDS vegetated tiles, 26th Feb Meeting"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)

#import and clean data
Coral_settlement_nGBR_CS <- read_csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(`Shelf posn` != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR |> 
  filter(notes != "NA")

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= `Shelf posn`,
         "Year_Collected"= `Year (Jan)`,
         "Site_#"= `Site #`,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)
Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  )) |> 
  filter(notes != "Dislodged")

# Mean or median no. tiles per site / year
Join_Coral <- Coral_settlement_nGBR |> 
  group_by(Shelf, Year, Site, Reef) |> 
  summarise(
    st_mean_acro = mean(Acroporidae),
    st_sd_acro = sd(Acroporidae),
    st_median_acro = median(Acroporidae),
    st_mean_poci = mean(Pocilloporidae),
    st_sd_poci = sd(Pocilloporidae),
    st_median_poci = median(Pocilloporidae),
    st_mean_por = mean(Poritid),
    st_sd_por = sd(Poritid),
    st_median_por = median(Poritid),
    st_mean_total = mean(total),
    st_sd_total = sd(total),
    Tiles = n()
  )

# Benthic Data

PIT <- read_csv("PIT_50m.csv") %>%
  mutate(across(8:91, ~replace_na(.x, 0)))# Read in data

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))

PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_macroalgae = mean(Macroalgae),
    sd_macroalgae = sd(Macroalgae),
    mean_sand =mean(Rubble_and_Sand),
    sd_sand = sd(Rubble_and_Sand),
    mean_turf = mean(Turf),
    sd_turf = sd(Turf),
    PIT = n()
  )

Combined <- left_join(Coral_settlement_nGBR, join)

# Juvenile data
Belt_10m <- read_csv("Belt_10m.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(Total_juveniles),
    sd_total = sd(Total_juveniles),
    Belt = n()
  )
```

```{r}
Recruits <- Coral_settlement_nGBR %>%
  filter(total > 0) %>%                                                          # ensure only vegetated tiles included 
  group_by(Reef, Year_Collected, Shelf) %>%               # this was missing in your code
  summarise(across(Acroporidae:Other, mean), .groups = "drop")

recruit_hellinger <- decostand(Recruits %>% select(Acroporidae, Pocilloporidae, Poritid, Other),
                                                         method = "hellinger")

# reef year?
reef_year <- Recruits |> 
  select(Reef, Year_Collected, Shelf)

permanova_result <- adonis2(recruit_hellinger ~ Year_Collected * Shelf,
  data = Recruits,
  method = "euclidean",    # the Hellinger transformation is designed for Euclidean distnaces. Bray-Curtis is not “wrong” but it’s redundant if you’ve already done a Hellinger transformation.
  by = "margin",
  permutations = 999)

print(permanova_result)


euc_hellinger <- vegdist(recruit_hellinger, method = "euclidean")
# With respect to Shelf
betadisper(euc_hellinger, Recruits$Shelf)
anova(betadisper(euc_hellinger, Recruits$Shelf))
```

Using this method, we can see that the model explains only 37.58 percent of the data, and there is a large amount of residual...
ANd there is no significant effect of year and shelf

```{r}
set.seed(999)     # for reproducibility

nmds <- metaMDS(recruit_hellinger, distance = "euclidean", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, reef_year[, c("Year_Collected", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2022" ~ "21/22")
  )

ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 1.5, alpha = 0.8) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(
        panel.grid = element_blank(),
         legend.position = "top",
         text = element_text(size = 8),
        title = element_text(size = 8)) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
             x = min(scores$NMDS1),
             y = min(scores$NMDS2),              
             label = paste0("Stress = ", round(nmds$stress, 3)),
              hjust = -2, vjust = 0.5, size = 3)+
  scale_x_continuous(limits = c(-1, 1.45))+
  scale_y_continuous(limits = c(-1.3, 1.1))
```

```{r}
# Creating Vectors for driving the patterns
Recruits_envfit <- envfit(
  nmds,
  recruit_hellinger,
  permutations = 999
) 

#Saving the vectors as a data frame
Recruits_envfit <- as.data.frame(scores(Recruits_envfit, display = "vectors"))

Recruits_envfit <- Recruits_envfit |> 
  rownames_to_column(var = "label_family")
```

```{r}
ggplot()+
  geom_segment(data = Recruits_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Recruits_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_family),
                  inherit.aes = FALSE,
                  size = 3,
                  box.padding = 0.5, 
                  min.segment.length = Inf,
                  nudge_x = -0.1,
                  nudge_y = 0.1)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year Collected"
    )
```

The fact that there isn't significant differences to me suggests perhaps not having year collected in this model.

```{r}
permanova_result <- adonis2(recruit_hellinger ~ Shelf,
  data = Recruits,
  method = "euclidean",    # the Hellinger transformation is designed for Euclidean distnaces. Bray-Curtis is not “wrong” but it’s redundant if you’ve already done a Hellinger transformation.
  by = "margin",
  permutations = 999)

print(permanova_result)
```
Okay, Shelf itself is better model and I might just use that?
Shelf position explains 68.7% of the total variation observed in the mean counts of each family on a tile, for a given reef and year.

```{r}
p1 <- ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 1.5, alpha = 0.8) +
    stat_ellipse(aes(group = Shelf, colour = Shelf),
                 level = 0.75, linewidth = 1) +
     theme_bw(base_size = 14) +
     theme(legend.position = c(1,1), legend.justification = c(1,1),
        panel.grid = element_blank(),
         #legend.position = "top",
         text = element_text(size = 9),
        title = element_text(size = 9)) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
             x = min(scores$NMDS1),
             y = min(scores$NMDS2),              
             label = paste0("Stress = ", round(nmds$stress, 3)),
              hjust = -2, vjust = 0.5, size = 3)+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))

p2 <- ggplot()+
  geom_segment(data = Recruits_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Recruits_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_family),
                  inherit.aes = FALSE,
                  size = 3,
                  box.padding = 0.5, 
                  min.segment.length = Inf,
                  nudge_x = -0.1,
                  nudge_y = 0.1)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year Collected"
    )+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))

ggarrange(p1, p2, ncol = 1, heights = c(1, 1))
```

Is it okay that the sample size looks so small?

# Lag juvenile Models

```{r}
combine_belt_st <- left_join(join_belt, Join_Coral)
```

```{r}
combine_belt_st <- combine_belt_st |> 
  mutate(Year_start = as.numeric(substr(Year, 1, 4)))
```

```{r}
gropd_df2 <- combine_belt_st %>%
  arrange(Reef, Site, Year_start) %>%   # make sure data are ordered
  group_by(Reef, Site) %>%              # or group_by(Reef) if Site not needed
  mutate(
    st_median_acro_lag1 = lag(st_median_acro, 1),
    st_mean_acro_lag1 = lag(st_mean_acro, 1),
    st_median_poci_lag1 = lag(st_median_poci, 1),
    st_mean_poci_lag1 = lag(st_mean_poci, 1),
    st_median_por_lag1 = lag(st_median_por, 1),
    st_mean_por_lag1 = lag(st_mean_por, 1),
  ) %>%
  ungroup()
```

Remove first year rows
```{r}
model_df_acro <- gropd_df2 %>%
  filter(!is.na(mean_acro))

model_df_poci <- gropd_df2 %>%
  filter(!is.na(mean_poci))

model_df_por <- gropd_df2 %>%
  filter(!is.na(mean_por))
```

Fit model. check if mean_acro is approximately normal
```{r}
m1 <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + (1|Reef/Site),
              family = "lognormal",
           data = model_df_acro)

sim <- simulateResiduals(m1)
testResiduals(sim)
testZeroInflation(sim)

m2 <- glmmTMB(mean_acro ~ st_mean_acro_lag1 + (1|Reef/Site),
              family = "gaussian",
           data = model_df_acro)

sim <- simulateResiduals(m2)
testResiduals(sim)
testZeroInflation(sim)

testResiduals(sim)
testZeroInflation(sim)
model.sel(m1, m2)
```


```{r}
m1 <- glmmTMB(mean_poci ~ st_mean_poci_lag1 + (1|Reef/Site),
              family = "lognormal",
           data = model_df_poci)

sim <- simulateResiduals(m1)
testResiduals(sim)
testZeroInflation(sim)


m2 <- glmmTMB(mean_poci ~ st_mean_poci_lag1*Shelf + (1|Reef/Site),
              family = "gaussian",
           data = model_df_poci)

sim <- simulateResiduals(m2)
testResiduals(sim)
testZeroInflation(sim)
model.sel(m1, m2)
```
```{r}
summary(m1)
```


```{r}
model_df_por <- model_df_por |> 
  filter(mean_por > 0)
m1 <- glmmTMB(mean_por ~ st_mean_por_lag1 + (1|Reef/Site),
              family = "lognormal",
           data = model_df_por)

sim <- simulateResiduals(m1)
testResiduals(sim)
testZeroInflation(sim)

m2 <- glmmTMB(mean_por ~ st_mean_por_lag1 + (1|Reef/Site),
              family = "gaussian",
           data = model_df_por)

sim <- simulateResiduals(m2)
testResiduals(sim)
testZeroInflation(sim)

model.sel(m1, m2)
```

```{r}
summary(m1)
```

No significant relationship between poritidae 

## Averaging both tiles and mean cover

```{r}
st_join <- Coral_settlement_nGBR |> 
  group_by(Site, Year) |> 
  summarise(
    acro_mean_st = mean(Acroporidae),
    poci_mean_st = mean(Pocilloporidae),
    por_mean_st = mean(Poritid),
    acro_sd_st = sd(Acroporidae),
    poci_sd_st = sd(Pocilloporidae),
    por_sd_st = sd(Poritid),
    acro_median_st = median(Acroporidae),
    poci_median_st = median(Pocilloporidae),
    por_median_st = median(Poritid),
  )
```

```{r}
combined_averages <- left_join(join, st_join)
```

Now these are both averages, it can't follow an intergar and would need to follow a different family of error distribution.

## Acroporidae

```{r echo=FALSE}
ggplot()+
  geom_point(data = combined_averages, aes(x=mean_aco, y=acro_mean_st, colour = Shelf))
```

## Correlation

```{r}
# Test for normality
shapiro.test(combined_averages$acro_mean_st)
hist(combined_averages$acro_mean_st)
shapiro.test(combined_averages$mean_aco)
```

```{r}
cor.test(combined_averages$acro_mean_st, combined_averages$mean_aco, method = "spearman")
```

### Models

```{r}
Acro <- combined_averages |> 
  filter(acro_mean_st > 0)

# gamma_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Shelf/Site) + (1 | Year),
#         family = "Gamma", data = Acro)

ln_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Shelf/Site) + (1 | Year), 
        family = "lognormal", data = Acro)

gaus_avg_Acroporidae_fixed <- glmmTMB(acro_mean_st ~ mean_aco +(1 | Shelf/Site) + (1 | Year), 
        family = "gaussian", data = Acro)

```

```{r}
# sim13.1 <- simulateResiduals(gamma_avg_Acroporidae_fixed) #didn't work
# testResiduals(sim13.1) 
# testZeroInflation(sim13.1)
sim13.2 <- simulateResiduals(ln_avg_Acroporidae_fixed)
testResiduals(sim13.2)

sim13.3 <- simulateResiduals(gaus_avg_Acroporidae_fixed)
testResiduals(sim13.3) # definitelty not this one
```

Log normal models worked best for all of these
```{r}
summary(ln_avg_Acroporidae_fixed)
```

## Pocilloporidae

```{r echo=FALSE}
ggplot()+
  geom_point(data = combined_averages, aes(x=mean_poci, y=poci_mean_st, colour = Shelf))+
  facet_wrap(~Year)
```

### Correlation

```{r}
# Test for normality
shapiro.test(combined_averages$poci_mean_st)
hist(combined_averages$poci_mean_st)

shapiro.test(combined_averages$mean_poci)
hist(combined_averages$mean_poci)
hist(combined_averages$mean_poci)
```

```{r}
cor.test(combined_averages$poci_mean_st, combined_averages$mean_poci, method = "spearman")
```

### Model

```{r}
## Pocilloporidae
Poci <- combined_averages |> 
  filter(poci_mean_st > 0)
ln_avg_poci_fixed <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Shelf/Site) + (1 | Year), 
        family = "lognormal", data = Poci)

gaus_avg_poci_fixed <- glmmTMB(poci_mean_st ~ mean_poci +(1 | Shelf/Site) + (1 | Year), 
        family = "gaussian", data = Poci)
# gamma_avg_Pocilloporidae_fixed <- glmmTMB(poci_mean_st ~ mean_poci + (1 | Shelf/Site) + (1 | Year), 
#         family = "Gamma", data = Poci)
```

```{r}
# sim14.1 <- simulateResiduals(gamma_avg_poci_fixed) #didn't work
# testResiduals(sim14.1) 
# testZeroInflation(sim14.1)
sim14.2 <- simulateResiduals(ln_avg_poci_fixed) # very good
testResiduals(sim14.2)
testZeroInflation(sim14.2)
sim14.3 <- simulateResiduals(gaus_avg_poci_fixed)
testResiduals(sim14.3) # definitelty not this one
testZeroInflation(sim14.3)
```

```{r}
summary(ln_avg_poci_fixed)
```


## Poritidae

### Correlation

```{r}
# Test for normality
shapiro.test(combined_averages$por_mean_st)
shapiro.test(combined_averages$mean_por)
```

```{r}
cor.test(combined_averages$por_mean_st, combined_averages$mean_por, method = "spearman")
```

### Model

```{r}
## Poritidae
Por <- combined_averages |> 
  filter(por_mean_st > 0)
# gamma_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Shelf/Site) + (1 | Year), 
#         family = "Gamma", data = Por)
ln_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Shelf/Site) + (1 | Year), 
        family = "lognormal", data = Por)
gaus_avg_Poritid_fixed <- glmmTMB(por_mean_st ~ mean_por + (1 | Shelf/Site) + (1 | Year), 
        family = "gaussian", data = Por)

```

```{r}
# sim15.1 <- simulateResiduals(gamma_avg_Poritid_fixed) #didn't work
# testResiduals(sim15.1) 
# testZeroInflation(sim15.1)
sim15.2 <- simulateResiduals(ln_avg_Poritid_fixed) # very good
testResiduals(sim15.2)
testZeroInflation(sim15.2)

sim15.3 <- simulateResiduals(gaus_avg_Poritid_fixed)
testResiduals(sim15.3) # definitely not this one
testZeroInflation(sim15.3)
```

```{r}
summary(ln_avg_Poritid_fixed)
```

