---
title: "PERMANOVA_WF"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)
library(ggordiplots) #nmds and ordination plots

# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  ),
  Deployment = case_when(
      Year_Collected == "2022" ~ "21/22",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2026" ~ "25/26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")


# Mean or mean no. tiles per site / year
Join_Coral <- Coral_settlement_nGBR |> 
  group_by(Shelf, Year, Site, Reef) |> 
  summarise(
    st_mean_acro = mean(Acroporidae),
    st_sd_acro = sd(Acroporidae),
    st_median_acro = median(Acroporidae),
    st_mean_poci = mean(Pocilloporidae),
    st_sd_poci = sd(Pocilloporidae),
    st_median_poci = median(Pocilloporidae),
    st_mean_por = mean(Poritid),
    st_sd_por = sd(Poritid),
    st_median_por = median(Poritid),
    st_mean_total = mean(total),
    st_sd_total = sd(total),
    Tiles = n()
  )

# Benthic Data

# Benthic Data
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    )
  )


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

PIT_detailed <- PIT |> 
  dplyr::select(1:93)

PIT <- PIT %>%
  mutate(
    Acroporidae = rowSums(select(., `Acropora - branching`:`Montipora`), na.rm = TRUE),
    
    Pocilloporidae = rowSums(select(., Pocillopora:Stylophora), na.rm = TRUE),
    
    Poritidae = rowSums(select(., `Porites - branching`:Goniopora), na.rm = TRUE),
    
    `Other Hard coral` = rowSums(select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE),
    
    `Soft coral` = rowSums(select(., Lobophytum:`Soft-other`), na.rm = TRUE),
    `Brown MA` = rowSums(select(., Sargassum:Lobophora), na.rm = TRUE),
    `Red MA` = rowSums(select(., Asparagopsis:Jania), na.rm = TRUE),
    `Green MA`= rowSums(select(., Halimeda:Valonia), na.rm = TRUE),
    `Sand and Rubble` = rowSums(select(., Sand:Rubble), na.rm = TRUE),
    Turf = rowSums(select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE),
    Other = rowSums(select(., Sponge:Clam), na.rm = TRUE),
    Zoantharian = rowSums(select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  select(Year, Shelf, Reef, Site, Transect, Acroporidae:Zoantharian,CCA, Cyanobacteria, Halophila)

PIT <- PIT |> 
  mutate(
    sum = rowSums(across(Acroporidae:Halophila), na.rm = TRUE)
  )

summary(PIT$sum)

PIT <- PIT |> 
  mutate(
    across(Acroporidae:Halophila, ~ .x * 2)
  )

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae)
  )

Combined <- left_join(Coral_settlement_nGBR, join)
# Juvenile data
Belt_10m <- read_csv("Belt_10m_final.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
 mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))
Belt_10m <- Belt_10m |>
  mutate(
    Other_HC = Favia + Favites + Goniastrea + Astrea +
                 Echinopora + Galaxea + Turbinaria +
                 Psammacora + Pavona
  )

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(Total_juveniles),
    sd_total = sd(Total_juveniles),
    Belt = n()
  )

Combined1 <- left_join(Coral_settlement_nGBR, join_belt)
```

# 1. Recruits
## 1.1 Composition of Vegetated tiles

```{r echo=FALSE}
# Coral_settlement_nGBR <- Coral_settlement_nGBR |>
#   rename("Other Hard Corals" = Other)
```

```{r echo=FALSE}
# Vegetated tiles 
Recruits <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  dplyr::select(Acroporidae, Pocilloporidae, Poritid, Other)

environment_data <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  dplyr::select(Year_Collected, Site, Reef, Shelf)

# environment_data |> count(Shelf, Year_Collected)
```

```{r}
# Filtered recruit data
Recruits_sqrt <- sqrt(Recruits) #Squareroot transformed because Acroporidae is more abundant than the other taxa

recruits_sqrt_enviro <- bind_cols(environment_data, Recruits_sqrt)
#write.csv(Recruits, file = "Matrices_for_PRIMER/Recruits_raw.csv")
# write.csv(Recruits_sqrt, file = "Matrices_for_PRIMER/Recruits_sqrt.csv")
# write.csv(environment_data, file = "Matrices_for_PRIMER/recruit_enviro.csv")
#write.csv(recruits_sqrt_enviro, file = "Matrices_for_PRIMER/recruit_sqrt_enviro.csv")
# To stratify by year (this asks "do shelf differences exist after accounting for year structure?") 
h <- how(blocks = environment_data$Reef)                                                   # restrict permutations within years

#Bray curtis PERMANOVA
adonis2(Recruits_sqrt ~ Shelf*Year_Collected, 
        data = environment_data, 
        permutations = h,
        method = "bray", by = "margin")   # PERMANOVA, where main effects are not estimable independently of the interaction, 

adonis2(Recruits_sqrt ~ Shelf*Year_Collected, 
        data = environment_data, 
        permutations = h,
        method = "bray", by = "terms") #



```

```{r}
mod_unrestricted <- adonis2(Recruits_sqrt ~ Shelf * Year_Collected, 
                           data = environment_data, 
                           permutations = 999,
                           method = "bray",
                           by = "margin")
mod_restricted <- adonis2(Recruits_sqrt ~ Shelf * Year_Collected, 
                         data = environment_data, 
                         permutations = h,
                         method = "bray",
                         by = "margin")

mod_restricted #permutations restricted to reef, which there are 2 levels
mod_unrestricted #permutations were not restricted to reef

adonis2(Recruits_sqrt ~ Shelf + Year_Collected, 
        data = environment_data, 
        permutations = h,
        method = "bray",
        by = "margin")
adonis2(Recruits_sqrt ~ Shelf * Year_Collected, 
        data = environment_data, 
        permutations = h,
        method = "bray",
        by = "terms")
```
1. What your PERMANOVA is telling you
From the correct model (restricted, by = "margin"):
Shelf × Year:
R² approximately equal to 0.049 (~5%)
p = 0.005

So:
There is a significant interaction
But from the additive model:
Shelf: R² ≈ 0.224 (strong)
Year: R² ≈ 0.045 (moderate)

👉 This is crucial:

Unlike your belt data, Shelf remains strong even without the interaction.


##1.2. PERMDISP for Recruits
What that means for PERMDISP strategy

You’ve got:

Strong main effect (Shelf)
Moderate Year effect
Moderate interaction (~5%)

So you need to test both scales:
```{r}
#PERMDISP
# Distance matrix
bray_sqrt_recruits<- vegdist(Recruits_sqrt, method = "bray")

bd_shelf <- betadisper(bray_sqrt_recruits, environment_data$Shelf)
permutest(bd_shelf, permutations = h)

bd_int <- betadisper(bray_sqrt_recruits, 
                   interaction(environment_data$Shelf, environment_data$Year_Collected))
permutest(bd_int, permutations = h)


bd_year <- betadisper(bray_sqrt_recruits, environment_data$Year_Collected)
permutest(bd_year, permutations = h)

```
Shelf
F = 4.96, p = 0.005

👉 Dispersion differs among shelf positions
Year
F = 7.13, p = 0.005

👉 Dispersion differs among years
Shelf × Year
F = 3.44, p = 0.005

👉 Dispersion differs among specific combinations

*Big takeaway*

All sources of variation (Shelf, Year, and their interaction) show differences in dispersion

This is important:

Not just one problematic factor
Heterogeneity is a fundamental feature of your recruits data.

The PERMANOVA results are not purely centroid-driven

Instead:

Differences among groups reflect a combination of:

Changes in community composition (centroids)
Changes in variability/heterogeneity (dispersion)

Shelf:
Differences in both composition and variability
Likely environmental gradients + patchiness
Year:
Strong dispersion signal →
Some years are highly variable recruitment years
Interaction:
Certain Shelf–Year combos are:
Highly variable
Others more stable.

Recruitment to tiles are highly variable across both space and time, with differences driven by both community composition and variability. As communities develop (juveniles), patterns become more temporally dynamic and interaction-driven. In adult assemblages (PIT), strong spatial structuring emerges, although variability among shelf positions remains high.
Tests for homogeneity of multivariate dispersion indicated significant differences among shelf positions (F = 4.96, p = 0.005), years (F = 7.13, p = 0.005), and Shelf × Year combinations (F = 3.44, p = 0.005). These results suggest that variability in recruit assemblages differs across both spatial and temporal scales. As PERMANOVA is sensitive to dispersion effects, the observed differences in community structure likely reflect a combination of shifts in composition and changes in within-group variability.

##1.3. NMDS

```{r}
set.seed(999)     # for reproducibility

nmds <- metaMDS(Recruits_sqrt, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress #0.15

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, environment_data[, c("Year_Collected", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2022" ~ "21/22"),
    Year = Deployment
  )
# Creating Vectors for driving the patterns
Recruits_envfit <- envfit(
  nmds,
  Recruits_sqrt,
  permutations = h #restrict permutations to year, change if using PERMANOVA from PRIMER
) 

#Saving the vectors as a data frame
Recruits_envfit <- as.data.frame(scores(Recruits_envfit, display = "vectors"))

Recruits_envfit <- Recruits_envfit |> 
  rownames_to_column(var = "label_family")
```

```{r}
grp <- interaction(scores$Shelf, scores$Year, drop = TRUE)
# Function to generate ellipse points for ggplot
ellipse_points <- function(ell_obj, npoints = 100) {
  do.call(rbind,
    lapply(names(ell_obj), function(g) {
      center <- ell_obj[[g]]$center      # centroid
      cov_mat <- ell_obj[[g]]$cov        # covariance matrix
      angles <- seq(0, 2*pi, length.out = npoints)
      circle <- cbind(cos(angles), sin(angles))
      # scale unit circle by sqrt of covariance (1 SD)
      ellipse <- t(center + t(circle %*% chol(cov_mat)))
      data.frame(NMDS1 = ellipse[,1],
                 NMDS2 = ellipse[,2],
                 Shelf = g)
    })
  )
}

ell <- ordiellipse(nmds, groups = grp, kind = "sd", draw = "none")
ell_df <- ellipse_points(ell, npoints = 100)

ell_df <- ell_df |>
  tidyr::separate(Shelf, into = c("Shelf", "Year"), sep = "\\.")
```

```{r}
p1 <- ggplot(data = scores, aes(x = NMDS1, y = NMDS2,colour = Shelf)) +
     geom_point(size = 1, alpha = 0.8) +
  geom_path(data = ell_df,
            aes(x = NMDS1, y = NMDS2, colour = Shelf, group = Shelf),
            linewidth = 0.6) +
    # stat_ellipse(aes(group = Shelf, colour = Shelf),
    #              level = 0.75, linewidth = 0.5) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.5, 0.05),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
)  +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 0.95,
         y = -0.95,
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1,
         vjust = 0,
         size = 3)+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))+
  facet_wrap(~Year, ncol = 1)



p2 <- ggplot()+
  geom_segment(data = Recruits_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Recruits_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_family),
                  inherit.aes = FALSE,
                  size = 2.5,
                  box.padding = 0.5, 
                  min.segment.length = Inf,
                  nudge_x = -0.1,
                  nudge_y = 0.1)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year Collected"
    )+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))

ST_NMDS <- ggarrange(p1, p2, ncol = 1, heights = c(4, 1), legend = "none")
ST_NMDS
```

#2. Juveniles

## 2.2. Composition
```{r}
Belt_10m <- Belt_10m |> 
  rename("Other Hard Coral" = Other_HC
  )
# Step 1: Create site-by-species matrix with Site_ID
site_by_belt <- Belt_10m %>%
  filter(Total_juveniles > 0) |> #filter to make sure that dissimilarities are non zero
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, Acroporidae, Pocilloporidae, Poritidae, `Other Hard Coral`
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))


# Step 2: Metadata (KEEP Site_ID)
metadata_belt <- site_by_belt %>%
  dplyr::select(Shelf, Year, Reef)

# Step 3: Species matrix
belt_matrix_filtered <- site_by_belt %>%
  dplyr::select(
    6:9
  )

belt_matrix_filtered <- sqrt(belt_matrix_filtered)

# Check whether any family has much higher abundances than other families
# summary(belt_matrix_filtered$Acroporidae)
# summary(belt_matrix_filtered$Pocilloporidae)
# summary(belt_matrix_filtered$Poritidae)
# summary(belt_matrix_filtered$`Other Hard Coral`)

#Think Ill just do raw abundances then. Ended up sqrt transforming to be consistent with recruits and benthic composition.
```

```{r}
# create a distance matrix
belt_dist <- vegdist(belt_matrix_filtered, method = "bray")

# To stratify by year (this asks "do shelf differences exist after accounting for year structure?") 
h <- how(blocks = metadata_belt$Year)                                                   # restrict permutations within years
# With respect to Shelf
bd <- betadisper(belt_dist, metadata_belt$Shelf) #calculates distance to centroids (dispersion)
#anova(bd) # Significant was <0.001
permutest(bd, 
          permutations = h) #PERMDISP #0.05 
```

```{r}
# To stratify by year (this asks "do shelf differences exist after accounting for year structure?") 
h <- how(blocks = metadata_belt$Year)                                                   # restrict permutations within years

#Bray curtis PERMANOVA
permanova_result <- adonis2(belt_matrix_filtered ~ Shelf, 
        data = metadata_belt, 
        permutations = h,
        method = "bray")   # PERMANOVA
```

```{r}
h <- how(blocks = metadata_belt$Reef)                                                   # restrict permutations within years

#Bray curtis PERMANOVA
adonis2(belt_matrix_filtered ~ Shelf*Year, 
        data = metadata_belt, 
        permutations = h,
        method = "bray", by = "margin")   # PERMANOVA, where main effects are not estimable independently of the interaction, 

adonis2(belt_matrix_filtered ~ Shelf*Year, 
        data = metadata_belt, 
        permutations = h,
        method = "bray", by = "terms")

mod_unrestricted <- adonis2(belt_matrix_filtered ~ Shelf*Year, 
        data = metadata_belt,
                           permutations = 999,
                           method = "bray",
                           by = "margin")
mod_restricted <- adonis2(belt_matrix_filtered ~ Shelf*Year, 
        data = metadata_belt, 
                         permutations = h,
                         method = "bray",
                         by = "margin")

mod_restricted #permutations restricted to reef, which there are 2 levels
mod_unrestricted #permutations were not restricted to reef

adonis2(belt_matrix_filtered ~ Shelf + Year, 
        data = metadata_belt, 
        permutations = h,
        method = "bray", by = "margin")
```

Key model is permutations = h, by "margin, with interaction term
Result:

Shelf × Year
R2=0.066 (~6.6%)
p = 0.005 -> significant

Same story as before (and now confirmed with transformation):

There is a significant interaction between Shelf and Year

Result:

Shelf: p = 0.090 (ns / marginal)
Year: p = 0.005

This is the key ecological signal:

Shelf differences are not consistent through time

If they were:

Shelf would stay strongly significant without the interaction

But instead:

It weakens -> becomes marginal

Sequential test (by = "terms")
Shows:

Shelf: R² = 0.20
Year: R² = 0.04
Interaction: R² = 0.066

But remember:
- Shelf is tested first, so this is inflated

Still, one useful takeaway:

Interaction explains more than Year
And a meaningful chunk overall (~6.6%)

Restricted (p = 0.005) vs unrestricted (p = 0.001) permutations:
Same conclusion:

Ignoring reef structure inflates significance
Your valid result = p = 0.005

Juvenile community structure varied significantly across the continental shelf, but these spatial patterns were not consistent through time, as indicated by a significant Shelf × Year interaction (R² = 0.066, p = 0.005). When the interaction was removed, Shelf was no longer significant (p = 0.090), suggesting that cross-shelf differences occurred only in certain years rather than representing a stable spatial pattern. The persistence of the interaction following square-root transformation indicates that this pattern is not driven solely by dominant taxa.

##2.2. Beta dispersion
Since the main result is shelf x year interaction
```{r}
# interaction(metadata_belt$Shelf, metadata_belt$Year)
belt_dist <- vegdist(belt_matrix_filtered, method = "bray")

bd <- betadisper(belt_dist, 
                 interaction(metadata_belt$Shelf, metadata_belt$Year))

h <- how(blocks = metadata_belt$Reef)

permutest(bd, permutations = h)
plot(bd)
```
Groups differ in variability → PERMANOVA may reflect dispersion differences

BUT (important nuance):

If interaction is significant AND dispersion is weak/moderate
→ you can still interpret, but cautiously
“Communities differ among Shelf × Year, but this may reflect differences in variability as well as composition”

```{r}
# TukeyHSD(bd)
```

## 2.3. NMDS juveniles

```{r}
nmds <- metaMDS(belt_matrix_filtered, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress #0.16

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, metadata_belt[, c("Year", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year == "2024-25" ~ "24/25",
      Year == "2023-24" ~ "23/24",
      Year == "2022-23" ~ "22/23",
      Year == "2021-22" ~ "21/22")
  )
```


```{r}
grp <- interaction(scores$Shelf, scores$Year, drop = TRUE)
# Function to generate ellipse points for ggplot
ellipse_points <- function(ell_obj, npoints = 100) {
  do.call(rbind,
    lapply(names(ell_obj), function(g) {
      center <- ell_obj[[g]]$center      # centroid
      cov_mat <- ell_obj[[g]]$cov        # covariance matrix
      angles <- seq(0, 2*pi, length.out = npoints)
      circle <- cbind(cos(angles), sin(angles))
      # scale unit circle by sqrt of covariance (1 SD)
      ellipse <- t(center + t(circle %*% chol(cov_mat)))
      data.frame(NMDS1 = ellipse[,1],
                 NMDS2 = ellipse[,2],
                 Shelf = g)
    })
  )
}

ell <- ordiellipse(nmds, groups = grp, kind = "sd", draw = "none")
ell_df <- ellipse_points(ell, npoints = 100)

ell_df <- ell_df |>
  tidyr::separate(Shelf, into = c("Shelf", "Year"), sep = "\\.")
```

```{r}
belt_A <- ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 1, alpha = 0.8) +
  geom_path(data = ell_df,
            aes(x = NMDS1, y = NMDS2, colour = Shelf, group = Shelf),
            linewidth = 0.6) +
  
    # stat_ellipse(aes(group = Shelf, colour = Shelf),
    #              level = 0.75, linewidth = 0.5) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.5, 0.05),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 0.95,     # right limit
         y = -0.75,       # bottom limit
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1.05, # push slightly inward
         vjust = -0.5, # lift slightly above bottom
         size = 3)+
  scale_x_continuous(limits = c(-0.85, 1))+
  scale_y_continuous(limits = c(-1.05, 0.95))+
  facet_wrap(~Year)

belt_A
```

# 3. Benthic

```{r}
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

PIT <- PIT %>%
  mutate(
    # Acroporidae = rowSums(select(., `Acropora - branching`:`Montipora`), na.rm = TRUE),
    # 
    # Pocilloporidae = rowSums(select(., Pocillopora:Stylophora), na.rm = TRUE),
    # 
    # Poritidae = rowSums(select(., `Porites - branching`:Goniopora), na.rm = TRUE),
    
    `Other Hard coral` = rowSums(select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE),
    
    `Soft coral` = rowSums(select(., Lobophytum:`Soft-other`), na.rm = TRUE),
    `Phaeophyceae` = rowSums(select(., Sargassum:Lobophora), na.rm = TRUE),
    `Rhodophyta` = rowSums(select(., Asparagopsis:Jania), na.rm = TRUE),
    `Chlorophyta`= rowSums(select(., Halimeda:Valonia), na.rm = TRUE),
    `Sand and Rubble` = rowSums(select(., Sand:Rubble), na.rm = TRUE),
    Turf = rowSums(select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE),
    Other = rowSums(select(., Sponge:Clam), na.rm = TRUE),
    Zoantharian = rowSums(select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  dplyr::select(Year, Shelf, Reef, Site, Transect, `Acropora - branching`:Goniopora, 93:101,CCA, Cyanobacteria, Halophila)

# PIT <- PIT |> 
#   mutate(
#     sum = rowSums(across(`Acropora - branching`:Halophila), na.rm = TRUE)
#   )
# 
# summary(PIT$sum)

PIT <- PIT |> 
  mutate(
    across(`Acropora - branching`:Halophila, ~ .x * 2)
  )
# Step 1: Create site-by-species matrix with Site_ID
site_by_PIT <- PIT %>%
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, 6:18, 20:29,
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  dplyr::select(Shelf, Year, Reef, Site)

# Step 3: Species matrix
PIT_matrix_filtered <- site_by_PIT %>%
  dplyr::select(
    `Acropora - branching`:Halophila
  )

PIT_matrix_filtered <- sqrt(PIT_matrix_filtered)
```

```{r}
h <- how(blocks = metadata_PIT$Reef)                                                   # restrict permutations within years

#Bray curtis PERMANOVA
mod_restricted <- adonis2(PIT_matrix_filtered ~ Shelf*Year, 
        data = metadata_PIT, 
        permutations = h,
        method = "bray", by = "margin")   # PERMANOVA, where main effects are not estimable independently of the interaction, 

adonis2(PIT_matrix_filtered ~ Shelf*Year, 
        data = metadata_PIT, 
        permutations = h,
        method = "bray", by = "terms")

mod_unrestricted <- adonis2(PIT_matrix_filtered ~ Shelf*Year, 
        data = metadata_PIT,
                           permutations = 999,
                           method = "bray",
                           by = "margin")


mod_restricted #permutations restricted to reef, which there are 2 levels
mod_unrestricted #permutations were not restricted to reef

adonis2(PIT_matrix_filtered ~ Shelf + Year, 
        data = metadata_PIT, 
        permutations = h,
        method = "bray", by = "margin")
```

Benthic community structure differed strongly across the continental shelf, with Shelf explaining approximately 50% of the variation. Temporal variation was also significant but comparatively small (~3%). Although a statistically significant Shelf × Year interaction was detected, its effect size was low (~2.8%), indicating that spatial patterns were largely consistent through time, with only minor temporal shifts.


##3.2. Beta dispersion
For PIT, your key PERMANOVA result was:

Strong Shelf effect (R² ~ 0.50)
Weak but significant Shelf × Year interaction (~2.8%)

Since the primary biological interpretation = Shelf differences
Since the main result is shelf x year interaction
```{r}
# interaction(metadata_belt$Shelf, metadata_belt$Year)
PIT_dist <- vegdist(PIT_matrix_filtered, method = "bray")

bd_shelf <- betadisper(PIT_dist, metadata_PIT$Shelf)

h <- how(blocks = metadata_PIT$Reef)

permutest(bd_shelf, permutations = h)

plot(bd_shelf)
```

Shelf (main effect)
F = 40.23
p = 0.005

 Very strong signal:
Dispersion differs substantially among shelf positions

```{r}
bd_int <- betadisper(PIT_dist, 
                     interaction(metadata_PIT$Shelf, metadata_PIT$Year))

permutest(bd_int, permutations = h)

plot(bd_int)
```

Shelf × Year
F = 6.31
p = 0.005

Also significant:

Variability differs among specific Shelf–Year combination

Because:

Shelf R² = 50% (very large)
Dispersion is also very strong

Interpretation becomes:

Cross-shelf differences are partly due to differences in community variability, not just shifts in mean composition

From plot(bd_shelf), you’ll probably see:

One shelf = tight clustering (low variance)
Another = very spread out (high variance)

Ecologically, this often means:

Outer shelf → more stable, consistent assemblages
Inner shelf → more variable / patchy / disturbed


Adult communities show strong cross-shelf structuring, but this is accompanied by substantial differences in variability among shelf positions. In contrast, juvenile assemblages are more temporally dynamic, with spatial patterns that shift among years and are also associated with changes in dispersion.
A test for homogeneity of multivariate dispersion revealed significant differences among shelf positions (F = 40.23, p = 0.005), indicating substantial variation in within-group dispersion. As PERMANOVA is sensitive to dispersion effects, the strong Shelf effect (R² = 0.50) likely reflects both differences in community composition and variability among shelf positions.
Dispersion also differed among Shelf × Year combinations (F = 6.31, p = 0.005), although this effect was weaker than the main shelf-level differences.

## 3.3. nMDS figure with facet by year
```{r}
nmds <- metaMDS(PIT_matrix_filtered, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, metadata_PIT[, c("Year", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year == "2024-25" ~ "24/25",
      Year == "2023-24" ~ "23/24",
      Year == "2022-23" ~ "22/23",
      Year == "2021-22" ~ "21/22")
  )
```

```{r}
grp <- interaction(scores$Shelf, scores$Year, drop = TRUE)
# Function to generate ellipse points for ggplot
ellipse_points <- function(ell_obj, npoints = 100) {
  do.call(rbind,
    lapply(names(ell_obj), function(g) {
      center <- ell_obj[[g]]$center      # centroid
      cov_mat <- ell_obj[[g]]$cov        # covariance matrix
      angles <- seq(0, 2*pi, length.out = npoints)
      circle <- cbind(cos(angles), sin(angles))
      # scale unit circle by sqrt of covariance (1 SD)
      ellipse <- t(center + t(circle %*% chol(cov_mat)))
      data.frame(NMDS1 = ellipse[,1],
                 NMDS2 = ellipse[,2],
                 Shelf = g)
    })
  )
}

ell <- ordiellipse(nmds, groups = grp, kind = "sd", draw = "none")
ell_df <- ellipse_points(ell, npoints = 100)

ell_df <- ell_df |>
  tidyr::separate(Shelf, into = c("Shelf", "Year"), sep = "\\.")
```


```{r}
# # Extract NMDS for plot
# site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
# scores <- cbind(site_scores, metadata_PIT[, c("Year", "Shelf", "Reef")])
# 
# scores_df <- as.data.frame(scores(nmds, display = "sites"))
# 
# scores_df$Shelf <- site_scores$Shelf
# scores_df$Year  <- site_scores$Year
# 
# site_scores$Shelf <- metadata_PIT$Shelf
# site_scores$Year  <- metadata_PIT$Year
```

```{r}
library(dplyr)

ell_list <- lapply(unique(site_scores$Year), function(yr) {
  
  idx <- which(site_scores$Year == yr)
  
  sub <- site_scores[idx, ]
  
  ell <- ordiellipse(
    nmds,
    groups = sub$Shelf,
    kind = "sd",
    draw = "none"
  )
  
  ell_df <- ellipse_points(ell)
  ell_df$Year <- yr
  
  ell_df
})

ell_df_all <- bind_rows(ell_list)
```

```{r}
Benthic_A <- ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 1, alpha = 0.8) +
  geom_path(data = ell_df,
            aes(x = NMDS1, y = NMDS2, colour = Shelf, group = Shelf),
            linewidth = 0.6) +
    # stat_ellipse(aes(group = Shelf, colour = Shelf),
    #              level = 0.75, linewidth = 0.5) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.5, 0.05),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 8),
  title = element_text(size = 8)
)+
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 1.05,     # right limit
         y = -1,       # bottom limit
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1.05, # push slightly inward
         vjust = -0.5, # lift slightly above bottom
         size = 3)+
  scale_x_continuous(limits = c(-1.2, 1.05))+
  scale_y_continuous(limits = c(-1, 0.9))+
  facet_wrap(~Year)
Benthic_A
```
