---
title: "PERMANOVA_WF"
author: "Monique White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
library("FactoMineR")
library("factoextra")
library(FactoMineR)
library(factoextra)
library(vegan) #to do NMDS
library(PCAmixdata)
library(ggfortify)
library(glmmTMB) # Generalized Linear Mixed Models using Template Model Builder
library(DHARMa) # Diagnostics for Hierarchical (multi-level/mixed) regression models
library(ggpubr)
library(ggeffects) #to be able to do ggpredict
library(ggh4x) # to specify x and y axis if desired
library(MuMIn) #model selection function
library(ggrepel)
library(ggordiplots) #nmds and ordination plots

# Import and clean data
## Recruitment data
Coral_settlement_nGBR_CS <- read.csv("~/Larval recruitment resources/Data Analysis/Coral-Settlement-Northern-GBR/Coral settlement_nGBR_CS.csv") %>%
  mutate(across(15:29, ~replace_na(.x, 0))) # Read in data

Coral_settlement_nGBR <- Coral_settlement_nGBR_CS %>% 
  filter(Shelf.posn != "Coral Sea", #remove coral sea data
        !Obs %in% c(1499, 1621, 1622)) #Remove loose tiles with heavy biofouling

Coral_settlement_nGBR <- Coral_settlement_nGBR %>% 
  rename("Shelf"= Shelf.posn,
         "Year_Collected"= Year..Jan.,
         "Site_#"= Site..,
         "Acroporidae" =Acropora,
         "Pocilloporidae" = Pocillopora,
         "Poritid" = Porites) |> 
  mutate(Year_Collected = factor(Year_Collected))
  

Coral_settlement_nGBR <- Coral_settlement_nGBR |>
  filter(Shelf == "Outer" | (Shelf %in% c("Inner", "Mid") & Tile %in% as.character(1:10)))

Coral_settlement_nGBR$Year_Collected <- factor(Coral_settlement_nGBR$Year_Collected)

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  mutate(
    Year = case_when(
      Year_Collected == "2022" ~ "2021-22",
      Year_Collected == "2023" ~ "2022-23",
      Year_Collected == "2024" ~ "2023-24",
      Year_Collected == "2025" ~ "2024-25",
      Year_Collected == "2026" ~ "2025-26"
  ),
  Deployment = case_when(
      Year_Collected == "2022" ~ "21/22",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2026" ~ "25/26"
  )) |> 
mutate(
  notes = case_when(
    notes %in% c("dislodged", "dislodged tile", "dettached tile") ~ "Dislodged",
    notes %in% c("lots of tube worms") ~ "Lots of tube worms",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
mutate(
  notes = case_when(
    !notes %in% c("Dislodged", "Lots of tube worms")~"No notes",
    TRUE ~ notes
  ))

Coral_settlement_nGBR <- Coral_settlement_nGBR |> 
  filter(notes != "Dislodged")


# Mean or mean no. tiles per site / year
Join_Coral <- Coral_settlement_nGBR |> 
  group_by(Shelf, Year, Site, Reef) |> 
  summarise(
    st_mean_acro = mean(Acroporidae),
    st_sd_acro = sd(Acroporidae),
    st_median_acro = median(Acroporidae),
    st_mean_poci = mean(Pocilloporidae),
    st_sd_poci = sd(Pocilloporidae),
    st_median_poci = median(Pocilloporidae),
    st_mean_por = mean(Poritid),
    st_sd_por = sd(Poritid),
    st_median_por = median(Poritid),
    st_mean_total = mean(total),
    st_sd_total = sd(total),
    Tiles = n()
  )

# Benthic Data

# Benthic Data
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    )
  )


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

PIT_detailed <- PIT |> 
  dplyr::select(1:93)

PIT <- PIT %>%
  mutate(
    Acroporidae = rowSums(select(., `Acropora - branching`:`Montipora`), na.rm = TRUE),
    
    Pocilloporidae = rowSums(select(., Pocillopora:Stylophora), na.rm = TRUE),
    
    Poritidae = rowSums(select(., `Porites - branching`:Goniopora), na.rm = TRUE),
    
    `Other Hard coral` = rowSums(select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE),
    
    `Soft coral` = rowSums(select(., Lobophytum:`Soft-other`), na.rm = TRUE),
    `Brown MA` = rowSums(select(., Sargassum:Lobophora), na.rm = TRUE),
    `Red MA` = rowSums(select(., Asparagopsis:Jania), na.rm = TRUE),
    `Green MA`= rowSums(select(., Halimeda:Valonia), na.rm = TRUE),
    `Sand and Rubble` = rowSums(select(., Sand:Rubble), na.rm = TRUE),
    Turf = rowSums(select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE),
    Other = rowSums(select(., Sponge:Clam), na.rm = TRUE),
    Zoantharian = rowSums(select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  select(Year, Shelf, Reef, Site, Transect, Acroporidae:Zoantharian,CCA, Cyanobacteria, Halophila)

PIT <- PIT |> 
  mutate(
    sum = rowSums(across(Acroporidae:Halophila), na.rm = TRUE)
  )

summary(PIT$sum)

PIT <- PIT |> 
  mutate(
    across(Acroporidae:Halophila, ~ .x * 2)
  )

join <- PIT |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_aco = mean(Acroporidae),
    sd_aco = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae)
  )

Combined <- left_join(Coral_settlement_nGBR, join)
# Juvenile data
Belt_10m <- read_csv("Belt_10m_final.csv") |> 
  mutate(across(8:28, ~replace_na(.x, 0)))

Belt_10m <- Belt_10m |> 
 mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    ))

Belt_10m <- Belt_10m |> filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25"))
Belt_10m <- Belt_10m |>
  mutate(
    Other_HC = Favia + Favites + Goniastrea + Astrea +
                 Echinopora + Galaxea + Turbinaria +
                 Psammacora + Pavona
  )

join_belt <- Belt_10m |> 
  group_by(Year, Shelf, Reef, Site) |> 
  summarise(
    mean_acro = mean(Acroporidae),
    sd_acro = sd(Acroporidae),
    mean_poci = mean(Pocilloporidae),
    sd_poci = sd(Pocilloporidae),
    mean_por = mean(Poritidae),
    sd_por = sd(Poritidae),
    mean_total = mean(Total_juveniles),
    sd_total = sd(Total_juveniles),
    Belt = n()
  )

Combined1 <- left_join(Coral_settlement_nGBR, join_belt)
```

# 1. Recruits
## 1.1 Composition of Vegetated tiles

```{r echo=FALSE}
# Coral_settlement_nGBR <- Coral_settlement_nGBR |>
#   rename("Other Hard Corals" = Other)
```

```{r echo=FALSE}
# Vegetated tiles 
Recruits <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  dplyr::select(Acroporidae, Pocilloporidae, Poritid, Other)

environment_data <- Coral_settlement_nGBR %>%
  filter(total >0) |> 
  dplyr::select(Year_Collected, Site, Reef, Shelf)

# environment_data |> count(Shelf, Year_Collected)
```

```{r}
# environment_data |> group_by(Shelf, Reef, Site, Year_Collected) |> 
# summarise(reps = n()) |> 
#   pivot_wider(id_cols = c("Shelf", "Site"), names_from = Year_Collected, values_from = reps) |> write.csv("output/vegetated_tiles.csv")
```

```{r echo=TRUE}
# Filtered recruit data
Recruits_sqrt <- sqrt(Recruits) #Squareroot transformed because Acroporidae is more abundant than the other taxa

recruits_sqrt_enviro <- bind_cols(environment_data, Recruits_sqrt)

h <- how(blocks = environment_data$Site)                                                   # restrict permutations within years

#Bray curtis PERMANOVA
adonis2(Recruits_sqrt ~ Shelf*Year_Collected, 
        data = environment_data, 
        permutations = h,
        method = "bray", by = "margin")   # PERMANOVA, where main effects are not estimable independently of the interaction, 

adonis2(Recruits_sqrt ~ Shelf*Year_Collected, 
        data = environment_data, 
        permutations = h,
        method = "bray", by = "terms") #
```

Shelf × Year:  
R² = 0.0487, F = 7.56, p = 0.005
There is a significant interaction between Shelf and Year, explaining ~4.9% of variation in recruit community structure.

Supporting structure (by = "terms")
Shelf   R² = 0.222  
Year    R² = 0.045  
Interaction R² = 0.049  

Shelf explains a large proportion of variation (~22%)
Year and interaction are similar in magnitude (~4–5%)
(when sequential order dependent), so shelf looks strong because it is fitted first

```{r}
mod_unrestricted <- adonis2(Recruits_sqrt ~ Shelf * Year_Collected, 
                           data = environment_data, 
                           permutations = 999,
                           method = "bray",
                           by = "margin")
mod_restricted <- adonis2(Recruits_sqrt ~ Shelf * Year_Collected, 
                         data = environment_data, 
                         permutations = h,
                         method = "bray",
                         by = "margin")

mod_restricted #permutations restricted to reef, which there are 2 levels
mod_unrestricted #permutations were not restricted to reef
```

```{r echo=TRUE}
adonis2(Recruits_sqrt ~ Shelf + Year_Collected, 
        data = environment_data, 
        permutations = h,
        method = "bray",
        by = "margin")
```

Unrestricted permutations inflate significance slightly, but do not change the conclusion
restricted p = 0.005  
unrestricted p = 0.001
Results were consistent between restricted and unrestricted permutations, although unrestricted permutations produced slightly lower p-values.

#### Additive model
Shelf: p = 0.005  
Year:  p = 0.005
BOTH Shelf and Year remain significant when the interaction is removed
Recruitment is structured by both space and time, but the strength or nature of spatial patterns varies among years

#### ROle of by = terms
Shelf   R² approx 0.22  
Year    R² approx equal 0.045  
Interaction approx equal 0.049  

PERMANOVA revealed a significant interaction between Shelf and Year (R² = 0.049, F = 7.56, p = 0.005), indicating that cross-shelf patterns in recruit assemblages varied among years.

Despite this interaction, both Shelf (R² = 0.224, p = 0.005) and Year (R² = 0.045, p = 0.005) remained significant in the additive model, demonstrating that spatial and temporal effects independently contributed to variation in recruitment.

Shelf explained a substantial proportion of variation when fitted sequentially (R² = 0.223), indicating a strong cross-shelf gradient in recruit assemblages. However, the significant interaction suggests that the magnitude and composition of this gradient varied through time.


##1.2. PERMDISP for Recruits
What that means for PERMDISP strategy

You’ve got:

Strong main effect (Shelf)
Moderate Year effect
Moderate interaction (~5%)

So you need to test both scales:
```{r}
#PERMDISP
# Distance matrix
bray_sqrt_recruits<- vegdist(Recruits_sqrt, method = "bray")

bd_shelf <- betadisper(bray_sqrt_recruits, environment_data$Shelf)
permutest(bd_shelf, permutations = h)

bd_int <- betadisper(bray_sqrt_recruits, 
                   interaction(environment_data$Shelf, environment_data$Year_Collected))
permutest(bd_int, permutations = h)


bd_year <- betadisper(bray_sqrt_recruits, environment_data$Year_Collected)
permutest(bd_year, permutations = h)
```
Shelf
F = 4.96, p = 0.005

Dispersion differs among shelf positions
Year
F = 7.13, p = 0.005

Dispersion differs among years
Shelf × Year
F = 3.44, p = 0.005

Dispersion differs among specific combinations

*Big takeaway*

All sources of variation (Shelf, Year, and their interaction) show differences in dispersion

This is important:

Not just one problematic factor
Heterogeneity is a fundamental feature of your recruits data.

The PERMANOVA results are not purely centroid-driven

Instead:
Differences among groups reflect a combination of:

Changes in community composition (centroids) and changes in variability/heterogeneity (dispersion)

Shelf:
Differences in both composition and variability
Likely environmental gradients + patchiness
Year:
Strong dispersion signal ->
Some years are highly variable recruitment years
Interaction:
Certain Shelf–Year combos are:
Highly variable
Others more stable.

Recruitment to tiles are highly variable across both space and time, with differences driven by both community composition and variability. As communities develop (juveniles), patterns become more temporally dynamic and interaction-driven. In adult assemblages (PIT), strong spatial structuring emerges, although variability among shelf positions remains high.
Tests for homogeneity of multivariate dispersion indicated significant differences among shelf positions (F = 4.96, p = 0.005), years (F = 7.13, p = 0.005), and Shelf × Year combinations (F = 3.44, p = 0.005). These results suggest that variability in recruit assemblages differs across both spatial and temporal scales. As PERMANOVA is sensitive to dispersion effects, the observed differences in community structure likely reflect a combination of shifts in composition and changes in within-group variability.

##1.3. NMDS

```{r}
set.seed(999)     # for reproducibility

nmds <- metaMDS(Recruits_sqrt, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress #0.15

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, environment_data[, c("Year_Collected", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year_Collected == "2025" ~ "24/25",
      Year_Collected == "2024" ~ "23/24",
      Year_Collected == "2023" ~ "22/23",
      Year_Collected == "2022" ~ "21/22"),
    Year = Deployment
  )
# Creating Vectors for driving the patterns
Recruits_envfit <- envfit(
  nmds,
  Recruits_sqrt,
  permutations = h #restrict permutations to year, change if using PERMANOVA from PRIMER
) 

#Saving the vectors as a data frame
Recruits_envfit <- as.data.frame(scores(Recruits_envfit, display = "vectors"))

Recruits_envfit <- Recruits_envfit |> 
  rownames_to_column(var = "label_family")
```

```{r}
grp <- interaction(scores$Shelf, scores$Year, drop = TRUE)
# Function to generate ellipse points for ggplot
ellipse_points <- function(ell_obj, npoints = 100) {
  do.call(rbind,
    lapply(names(ell_obj), function(g) {
      center <- ell_obj[[g]]$center      # centroid
      cov_mat <- ell_obj[[g]]$cov        # covariance matrix
      angles <- seq(0, 2*pi, length.out = npoints)
      circle <- cbind(cos(angles), sin(angles))
      # scale unit circle by sqrt of covariance (1 SD)
      ellipse <- t(center + t(circle %*% chol(cov_mat)))
      data.frame(NMDS1 = ellipse[,1],
                 NMDS2 = ellipse[,2],
                 Shelf = g)
    })
  )
}

ell <- ordiellipse(nmds, groups = grp, kind = "sd", draw = "none")
ell_df <- ellipse_points(ell, npoints = 100)

ell_df <- ell_df |>
  tidyr::separate(Shelf, into = c("Shelf", "Year"), sep = "\\.")
```

```{r}
p1 <- ggplot(data = scores, aes(x = NMDS1, y = NMDS2,colour = Shelf)) +
     geom_point(size = 1, alpha = 0.8) +
  geom_path(data = ell_df,
            aes(x = NMDS1, y = NMDS2, colour = Shelf, group = Shelf),
            linewidth = 0.6) +
    # stat_ellipse(aes(group = Shelf, colour = Shelf),
    #              level = 0.75, linewidth = 0.5) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.5, 0.05),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
)  +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 0.95,
         y = -0.95,
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1,
         vjust = 0,
         size = 3)+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))+
  facet_wrap(~Year, ncol = 2)



p2 <- ggplot()+
  geom_segment(data = Recruits_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Recruits_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_family),
                  inherit.aes = FALSE,
                  size = 2.5,
                  box.padding = 0.5, 
                  min.segment.length = Inf,
                  nudge_x = -0.1,
                  nudge_y = 0.1)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year Collected"
    )+
  scale_x_continuous(limits = c(-1, 1))+
  scale_y_continuous(limits = c(-1, 1))

ST_NMDS <- ggarrange(p1, p2, ncol = 1, heights = c(2, 1), legend = "none")
ST_NMDS
```

```{r eval=FALSE}
ggsave(ST_NMDS,
  filename = "figures/facet_Recruits_NMDS.PDF",
  height = 280, 
  units = "mm"
)
```


#2. Juveniles

## 2.2. Composition
```{r}
Belt_10m <- Belt_10m |> 
  rename("Other Hard Coral" = Other_HC
  )
```

```{r}
 #detailed juvenile
site_by_belt <- Belt_10m %>%
  filter(Total_juveniles > 0) |> #filter to make sure that dissimilarities are non zero
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, Acropora:Pavona
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

# Step 2: Metadata (KEEP Site_ID)
metadata_belt <- site_by_belt %>%
  dplyr::select(Shelf, Year, Reef, Site)

# metadata_belt|> group_by(Shelf, Reef, Site, Year) |> 
# summarise(reps = n()) |>
#   pivot_wider(id_cols = c("Shelf", "Site"), names_from = Year, values_from = reps) |> write.csv("output/belt_transects_with_juveniles.csv")

# Step 3: Species matrix
belt_matrix_filtered <- site_by_belt %>%
  dplyr::select(
    6:21
  )

#write.csv(belt_matrix_filtered, file = "output/detailed_juvenile_community_matrix.csv")
#write.csv(metadata_belt, file = "Matrices_for_PRIMER/belt_meta.csv")
```


```{r}
# Step 1: Create site-by-species matrix with Site_ID
site_by_belt <- Belt_10m %>%
  filter(Total_juveniles > 0) |> #filter to make sure that dissimilarities are non zero
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, Acroporidae, Pocilloporidae, Poritidae, `Other Hard Coral`
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

# Step 2: Metadata (KEEP Site_ID)
metadata_belt <- site_by_belt %>%
  dplyr::select(Shelf, Year, Reef, Site)

# metadata_belt|> group_by(Shelf, Reef, Site, Year) |> 
# summarise(reps = n()) |>
#   pivot_wider(id_cols = c("Shelf", "Site"), names_from = Year, values_from = reps) |> write.csv("output/belt_transects_with_juveniles.csv")

# Step 3: Species matrix
belt_matrix_filtered <- site_by_belt %>%
  dplyr::select(
    6:9
  )

belt_matrix_filtered <- sqrt(belt_matrix_filtered)

#Think Ill just do raw abundances then. Ended up sqrt transforming to be consistent with recruits and benthic composition.
```

```{r}
# create a distance matrix
belt_dist <- vegdist(belt_matrix_filtered, method = "bray")
# To stratify by year (this asks "do shelf differences exist after accounting for year structure?")
```

```{r echo=TRUE}
h <- how(blocks = metadata_belt$Site)                                                   # restrict permutations within years

#Bray curtis PERMANOVA
adonis2(belt_matrix_filtered ~ Shelf*Year, 
        data = metadata_belt, 
        permutations = h,
        method = "bray", by = "margin")   # PERMANOVA, where main effects are not estimable independently of the interaction, 
```

Shelf:Year   R² = 0.066, F = 3.61, p = 0.005
There is a significant interaction between Shelf and Year, explaining ~6.6% of variation in community structure.

```{r echo=TRUE}
adonis2(belt_matrix_filtered ~ Shelf*Year, 
        data = metadata_belt, 
        permutations = h,
        method = "bray", by = "terms")
```

Supporting structure, by = terms
Shelf        R² = 0.203
Year         R² = 0.040
Interaction  R² = 0.066

Shelf explains the most variation overall
But this is sequential (order-dependent)

“Shelf contributes strongly, but its effect varies through time”

```{r}
mod_unrestricted <- adonis2(belt_matrix_filtered ~ Shelf*Year, 
        data = metadata_belt,
                           permutations = 999,
                           method = "bray",
                           by = "margin")
mod_restricted <- adonis2(belt_matrix_filtered ~ Shelf*Year, 
        data = metadata_belt, 
                         permutations = h,
                         method = "bray",
                         by = "margin")

mod_restricted #permutations restricted to reef, which there are 2 levels
mod_unrestricted #permutations were not restricted to reef
```

```{r echo=TRUE}
adonis2(belt_matrix_filtered ~ Shelf + Year, 
        data = metadata_belt, 
        permutations = h,
        method = "bray", by = "margin")
```
Additive model
Shelf: p = 0.075 (ns)
Year:  p = 0.005 (**)
When the interaction is removed, Shelf is no longer significant
Cross-shelf differences are not consistent across years

Key result
Shelf:Year   R² = 0.066, F = 3.61, p = 0.005
There is a significant interaction between Shelf and Year, explaining ~6.6% of variation in community structure.
Main effects are not independently estimable due to the interaction

Permutational multivariate analysis of variance (PERMANOVA) revealed a significant interaction between Shelf and Year (R² = 0.066, F = 3.61, p = 0.005), indicating that cross-shelf patterns in community structure varied through time.

Although Shelf explained a substantial proportion of variation when fitted sequentially (R² = 0.203), this effect was not significant in the additive model (p = 0.075), suggesting that spatial differences among shelf positions were not consistent across years.

In contrast, Year remained significant in the additive model (R² = 0.040, p = 0.005), indicating temporal variation in community structure independent of spatial effects.

Your result is robust, not an artefact of permutation design, because there is the same pattern, just slightly stronger when unrestricted.
restricted:   p = 0.005  
unrestricted: p = 0.001


##2.2. Beta dispersion
Since the main result is shelf x year interaction
```{r}
# interaction(metadata_belt$Shelf, metadata_belt$Year)
belt_dist <- vegdist(belt_matrix_filtered, method = "bray")

bd <- betadisper(belt_dist, 
                 interaction(metadata_belt$Shelf, metadata_belt$Year))

h <- how(blocks = metadata_belt$Reef)

permutest(bd, permutations = h)
plot(bd)
```
Groups differ in variability -> PERMANOVA may reflect dispersion differences

BUT (important nuance):

If interaction is significant AND dispersion is weak/moderate
-> you can still interpret, but cautiously
“Communities differ among Shelf × Year, but this may reflect differences in variability as well as composition”


## 2.3. NMDS juveniles

```{r}
nmds <- metaMDS(belt_matrix_filtered, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress #0.16

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, metadata_belt[, c("Year", "Shelf", "Reef")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year == "2024-25" ~ "24/25",
      Year == "2023-24" ~ "23/24",
      Year == "2022-23" ~ "22/23",
      Year == "2021-22" ~ "21/22")
  )
```


```{r}
grp <- interaction(scores$Shelf, scores$Year, drop = TRUE)
# Function to generate ellipse points for ggplot
ellipse_points <- function(ell_obj, npoints = 100) {
  do.call(rbind,
    lapply(names(ell_obj), function(g) {
      center <- ell_obj[[g]]$center      # centroid
      cov_mat <- ell_obj[[g]]$cov        # covariance matrix
      angles <- seq(0, 2*pi, length.out = npoints)
      circle <- cbind(cos(angles), sin(angles))
      # scale unit circle by sqrt of covariance (1 SD)
      ellipse <- t(center + t(circle %*% chol(cov_mat)))
      data.frame(NMDS1 = ellipse[,1],
                 NMDS2 = ellipse[,2],
                 Shelf = g)
    })
  )
}

ell <- ordiellipse(nmds, groups = grp, kind = "sd", draw = "none")
ell_df <- ellipse_points(ell, npoints = 100)

ell_df <- ell_df |>
  tidyr::separate(Shelf, into = c("Shelf", "Year"), sep = "\\.")
```

```{r}
belt_A <- ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 1, alpha = 0.8) +
  geom_path(data = ell_df,
            aes(x = NMDS1, y = NMDS2, colour = Shelf, group = Shelf),
            linewidth = 0.6) +
  
    # stat_ellipse(aes(group = Shelf, colour = Shelf),
    #              level = 0.75, linewidth = 0.5) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.5, 0.05),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 9),
  title = element_text(size = 9)
) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 0.95,     # right limit
         y = -0.75,       # bottom limit
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1.05, # push slightly inward
         vjust = -0.5, # lift slightly above bottom
         size = 3)+
  scale_x_continuous(limits = c(-0.85, 1))+
  scale_y_continuous(limits = c(-1.05, 0.95))+
  facet_wrap(~Year, ncol =1)

belt_A
```

```{r}
h <- how(blocks = metadata_belt$Reef)

# Creating Vectors for driving the patterns
belt_envfit <- envfit(
  nmds,
  belt_matrix_filtered,
  permutations = h #restricts permuatations to a year
) 

# Extract p-values
pvals <- belt_envfit$vectors$pvals

# Extract r-squared values
r2 <- belt_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(belt_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- belt_envfit$vectors$r
vec_scores$pval <- belt_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

belt_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")
```

```{r}
belt_B <- ggplot()+
  geom_segment(data = belt_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = belt_envfit,
                  aes(x = NMDS1, y= NMDS2, label = label_category),
                  inherit.aes = FALSE,
                  size = 2.5,
                  box.padding = 0.1, 
                  min.segment.length = Inf,
                  nudge_x = -0.05,
                  nudge_y = -0.1)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year"
    )+
  scale_x_continuous(limits = c(-0.85, 1))+
  scale_y_continuous(limits = c(-1.05, 0.95))
belt_B
```

```{r}
belt_Figure <- ggarrange(
  belt_A, belt_B, labels = c("A", "B"), heights = c(4, 1), ncol = 1, legend = "none"
)

belt_Figure
```

```{r eval=FALSE}
ggsave(belt_Figure,
  filename = "figures/facet_belt_NMDS.PDF",
  height = 280, 
  units = "mm"
)
```

# 3. Benthic

```{r}
PIT <- PIT <- read_csv("PIT_final.csv") %>%
  mutate(across(8:94, ~replace_na(.x, 0))) %>%
  mutate(across(8:94, as.numeric))

PIT <- PIT %>% 
  filter(Year %in% c("2021-22", "2022-23", "2023-24", "2024-25", "2025-26"))

PIT <- PIT |> 
  mutate(
    Site = case_when(
      Site %in% c("Vicky-HS", "Vickys") ~ "Vicky-HS",
      Site %in% c("Turtle Beach", "TurtleB-CooksP") ~ "TurtleB-CooksP",
      Site %in% c("Lag2-Trimodal", "Lagoon") ~ "Lag2-Trimodal",
      Site == "Turtle South - Front (Sargassum)" ~ "TS_Crest",
      Site == "Turtle South - Coral" ~ "TS_Back",
      Site == "Turtle North - front" ~ "TN_Crest",
      TRUE ~ Site
    ),
    Reef = case_when(
      Site %in% c("Day 1", "Day 2", "Day 3") ~ "Day",
      Site %in% c("Hicks 1", "Hicks 2", "Hicks 3") ~ "Hicks",
      Site %in% c("Mac 1", "Mac 2", "Mac 3") ~ "MacGillivray",
      Site %in% c("TN_Back", "TN_coral", "TN_Crest") ~ "Turtle N",
      Site %in% c("TS_Back", "TS_Channel", "TS_Crest") ~ "Turtle S",
      Site %in% c("TurtleB-CooksP", "Lag2-Trimodal", "Vicky-HS") ~ "Lizard",
      TRUE ~ NA_character_
    ))


PIT<- PIT |> 
  mutate(
    Shelf = factor(Shelf),
    Reef = factor(Reef),
    Site = factor(Site))

PIT <- PIT %>%
  mutate(
    # Acroporidae = rowSums(select(., `Acropora - branching`:`Montipora`), na.rm = TRUE),
    # 
    # Pocilloporidae = rowSums(select(., Pocillopora:Stylophora), na.rm = TRUE),
    # 
    # Poritidae = rowSums(select(., `Porites - branching`:Goniopora), na.rm = TRUE),
    
    `Other Hard coral` = rowSums(select(., `Favia (plocoid)`:`other submassive`), na.rm = TRUE),
    
    `Soft coral` = rowSums(select(., Lobophytum:`Soft-other`), na.rm = TRUE),
    `Phaeophyceae` = rowSums(select(., Sargassum:Lobophora), na.rm = TRUE),
    `Rhodophyta` = rowSums(select(., Asparagopsis:Jania), na.rm = TRUE),
    `Chlorophyta`= rowSums(select(., Halimeda:Valonia), na.rm = TRUE),
    `Sand and Rubble` = rowSums(select(., Sand:Rubble), na.rm = TRUE),
    Turf = rowSums(select(., `Pavement/EAM`:`Dead massive`), na.rm = TRUE),
    Other = rowSums(select(., Sponge:Clam), na.rm = TRUE),
    Zoantharian = rowSums(select(., `Zoantharian - Red`:`Zoantharian - Green`), na.rm = TRUE)
  )

PIT <- PIT |> 
  dplyr::select(Year, Shelf, Reef, Site, Transect, `Acropora - branching`:Goniopora, 93:101,CCA, Cyanobacteria, Halophila)

# PIT <- PIT |> 
#   mutate(
#     sum = rowSums(across(`Acropora - branching`:Halophila), na.rm = TRUE)
#   )
# 
# summary(PIT$sum)

PIT <- PIT |> 
  mutate(
    across(`Acropora - branching`:Halophila, ~ .x * 2)
  )
# Step 1: Create site-by-species matrix with Site_ID
site_by_PIT <- PIT %>%
  dplyr::select(
    Year, Reef, Shelf, Site, Transect, 6:18, 20:29,
  ) %>%
  mutate(Site_ID = paste(Year, Reef, Site, Transect, sep = "_"))

# Step 2: Metadata (KEEP Site_ID)
metadata_PIT <- site_by_PIT %>%
  dplyr::select(Shelf, Year, Reef, Site)

# Step 3: Species matrix
PIT_matrix_filtered <- site_by_PIT %>%
  dplyr::select(
    `Acropora - branching`:Halophila
  )

PIT_matrix_filtered <- sqrt(PIT_matrix_filtered)
```

```{r}
h <- how(blocks = metadata_PIT$Site) # restrict permutations within years
#Bray curtis PERMANOVA
```

```{r echo=TRUE}
mod_restricted <- adonis2(PIT_matrix_filtered ~ Shelf*Year, 
        data = metadata_PIT, 
        permutations = h,
        method = "bray", by = "margin")   # PERMANOVA, where main effects are not estimable independently of the interaction, 



mod_unrestricted <- adonis2(PIT_matrix_filtered ~ Shelf*Year, 
        data = metadata_PIT,
        permutations = 999,
        method = "bray",
        by = "margin")

mod_restricted #permutations restricted to reef, which there are 2 levels
mod_unrestricted #permutations were not restricted to reef
```
#### Main result
Shelf × Year:  
R² = 0.0282, F = 2.48, p = 0.005
There is a significant interaction between Shelf and Year, but it explains a small proportion of variation (~2.8%).

Effect is robust, as indicated by significant p values when unrestricted
restricted:   p = 0.005  
unrestricted: p = 0.001


```{r echo=TRUE}
adonis2(PIT_matrix_filtered ~ Shelf*Year, 
        data = metadata_PIT, 
        permutations = h,
        method = "bray", by = "terms")

adonis2(PIT_matrix_filtered ~ Shelf + Year, 
        data = metadata_PIT, 
        permutations = h,
        method = "bray", by = "margin")
```
#### Additive model
Shelf: R² = 0.503, F = 127.65, p = 0.005  
Year:  R² = 0.029, F = 4.88,  p = 0.005  
Adult community structure is overwhelmingly structured by Shelf, with a smaller but significant temporal effect.
Shelf explains 50% of the variation - huge in ecological community

Shelf × Year: R² = 0.028, p = 0.005
Shelf = dominant driver
Year = secondary
Interaction = small but significant

#### Use of by terms, to understand structure
Shelf        R² = 0.509  
Year         R² = 0.029  
Interaction  R² = 0.028  
Shelf explains the vast majority of explained variation
Year and interaction contribute small but detectable effects
They align perfectly with the additive model
So interpretation is very robust.

Community structure is strongly and consistently differentiated across the shelf, with minor temporal fluctuations that do not fundamentally alter spatial patterns.

PERMANOVA revealed strong spatial structuring of adult assemblages, with Shelf explaining a substantial proportion of variation (R² = 0.503, F = 127.65, p = 0.005). Year also contributed significantly, but accounted for a much smaller proportion of variation (R² = 0.029, p = 0.005).

A significant Shelf × Year interaction was detected (R² = 0.028, p = 0.005), although its effect size was small relative to the main effects. This indicates that while temporal variability influences community structure, cross-shelf differences remain consistent through time.

## 3.2. Beta dispersion
For PIT, your key PERMANOVA result was:

Strong Shelf effect (R² ~ 0.50)
Weak but significant Shelf × Year interaction (~2.8%)

Since the primary biological interpretation = Shelf differences
Since the main result is shelf x year interaction
```{r}
# interaction(metadata_belt$Shelf, metadata_belt$Year)
PIT_dist <- vegdist(PIT_matrix_filtered, method = "bray")

bd_shelf <- betadisper(PIT_dist, metadata_PIT$Shelf)

h <- how(blocks = metadata_PIT$Site)

permutest(bd_shelf, permutations = h)

plot(bd_shelf)
```

Shelf (main effect)
F = 40.23
p = 0.005

Very strong signal:
Dispersion differs substantially among shelf positions

```{r}
bd_int <- betadisper(PIT_dist, 
                     interaction(metadata_PIT$Shelf, metadata_PIT$Year))

permutest(bd_int, permutations = h)

plot(bd_int)
```

Shelf × Year
F = 6.31
p = 0.005

Also significant:
Variability differs among specific Shelf–Year combination

Because:
Shelf R² = 50% (very large)
Dispersion is also very strong

Interpretation becomes:
Cross-shelf differences are partly due to differences in community variability, not just shifts in mean composition

From plot(bd_shelf), you’ll probably see:

One shelf = tight clustering (low variance)
Another = very spread out (high variance)

Ecologically, this often means:

Outer shelf -> more stable, consistent assemblages
Inner shelf -> more variable / patchy / disturbed

Adult communities show strong cross-shelf structuring, but this is accompanied by substantial differences in variability among shelf positions. In contrast, juvenile assemblages are more temporally dynamic, with spatial patterns that shift among years and are also associated with changes in dispersion.
A test for homogeneity of multivariate dispersion revealed significant differences among shelf positions (F = 40.23, p = 0.005), indicating substantial variation in within-group dispersion. As PERMANOVA is sensitive to dispersion effects, the strong Shelf effect (R² = 0.50) likely reflects both differences in community composition and variability among shelf positions.
Dispersion also differed among Shelf × Year combinations (F = 6.31, p = 0.005), although this effect was weaker than the main shelf-level differences.

## 3.3. nMDS figure with facet by year
```{r}
nmds <- metaMDS(PIT_matrix_filtered, distance = "bray", k = 2, trymax = 100)

#Check the Stress Value
nmds$stress

# Extract NMDS for plot
site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
scores <- cbind(site_scores, metadata_PIT[, c("Year", "Shelf", "Reef", "Site")])

scores <- scores |> 
  mutate(
    Deployment =
    case_when(
      Year == "2024-25" ~ "24/25",
      Year == "2023-24" ~ "23/24",
      Year == "2022-23" ~ "22/23",
      Year == "2021-22" ~ "21/22")
  )
```

```{r}
grp <- interaction(scores$Shelf, scores$Year, drop = TRUE)
# Function to generate ellipse points for ggplot
ellipse_points <- function(ell_obj, npoints = 100) {
  do.call(rbind,
    lapply(names(ell_obj), function(g) {
      center <- ell_obj[[g]]$center      # centroid
      cov_mat <- ell_obj[[g]]$cov        # covariance matrix
      angles <- seq(0, 2*pi, length.out = npoints)
      circle <- cbind(cos(angles), sin(angles))
      # scale unit circle by sqrt of covariance (1 SD)
      ellipse <- t(center + t(circle %*% chol(cov_mat)))
      data.frame(NMDS1 = ellipse[,1],
                 NMDS2 = ellipse[,2],
                 Shelf = g)
    })
  )
}

ell <- ordiellipse(nmds, groups = grp, kind = "sd", draw = "none")
ell_df <- ellipse_points(ell, npoints = 100)

ell_df <- ell_df |>
  tidyr::separate(Shelf, into = c("Shelf", "Year"), sep = "\\.")
```

```{r}
# # Extract NMDS for plot
# site_scores <- as.data.frame(vegan::scores(nmds, display = "sites"))
# scores <- cbind(site_scores, metadata_PIT[, c("Year", "Shelf", "Reef")])
# 
# scores_df <- as.data.frame(scores(nmds, display = "sites"))
# 
# scores_df$Shelf <- site_scores$Shelf
# scores_df$Year  <- site_scores$Year
# 
# site_scores$Shelf <- metadata_PIT$Shelf
# site_scores$Year  <- metadata_PIT$Year
```

```{r}
library(dplyr)

ell_list <- lapply(unique(site_scores$Year), function(yr) {
  
  idx <- which(site_scores$Year == yr)
  
  sub <- site_scores[idx, ]
  
  ell <- ordiellipse(
    nmds,
    groups = sub$Shelf,
    kind = "sd",
    draw = "none"
  )
  
  ell_df <- ellipse_points(ell)
  ell_df$Year <- yr
  
  ell_df
})

ell_df_all <- bind_rows(ell_list)
```

```{r}
Benthic_A <- ggplot(scores, aes(x = NMDS1, y = NMDS2,                          colour = Shelf)) +
     geom_point(size = 1, alpha = 0.8) +
  geom_path(data = ell_df,
            aes(x = NMDS1, y = NMDS2, colour = Shelf, group = Shelf),
            linewidth = 0.6) +
    # stat_ellipse(aes(group = Shelf, colour = Shelf),
    #              level = 0.75, linewidth = 0.5) +
     theme_bw(base_size = 14) +
     theme(
  panel.grid = element_blank(),
  legend.position = c(0.5, 0.05),   # x = centre, y = near bottom
  legend.justification = c(0.5, 0), # anchor legend by bottom-centre
  legend.background = element_rect(fill = "white", colour = NA),
  legend.direction = "horizontal",
  text = element_text(size = 8),
  title = element_text(size = 8)
)+
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf"
    ) +
  scale_colour_manual(values = c("#A8C85A", "#9e379f", "#C97B5F"))+
    annotate("text",
         x = 1.05,     # right limit
         y = -1,       # bottom limit
         label = paste0("Stress = ", round(nmds$stress, 3)),
         hjust = 1.05, # push slightly inward
         vjust = -0.5, # lift slightly above bottom
         size = 3)+
  scale_x_continuous(limits = c(-1.2, 1.05))+
  scale_y_continuous(limits = c(-1, 0.9))+
  facet_wrap(~Year, ncol =2)
Benthic_A
```

```{r}
h <- how(blocks = metadata_PIT$Reef)
# Creating Vectors for driving the patterns
Benthic_envfit <- envfit(
  nmds,
  PIT_matrix_filtered,
  permutations = h
) 

# Extract p-values
pvals <- Benthic_envfit$vectors$pvals

# Extract r-squared values
r2 <- Benthic_envfit$vectors$r
# Arrow coordinates
vec_scores <- as.data.frame(scores(Benthic_envfit, display = "vectors"))

# Add r2 and p-values
vec_scores$r2 <- Benthic_envfit$vectors$r
vec_scores$pval <- Benthic_envfit$vectors$pvals

vec_scores
#Saving the vectors as a data frame

vec_scores <- vec_scores |> 
  filter(pval < 0.05, r2 > 0.15) # filter to only show vectors that are significant drivers.

Benthic_envfit <- vec_scores |> 
  rownames_to_column(var = "label_category")

#Other, Zoantharian, Poritidae...
```


```{r}
Benthic_envfit$NMDS1_lab <- Benthic_envfit$NMDS1 * 1.1

Benthic_envfit$NMDS2_lab <- Benthic_envfit$NMDS2 * 1.1

Benthic_B <- ggplot()+
  geom_segment(data = Benthic_envfit, aes(x = 0, y = 0, xend=NMDS1, yend = NMDS2),
             arrow = arrow(length = unit(1/3, "picas")),
             linetype = "solid",
             linewidth = 0.5)+
  geom_text_repel(data = Benthic_envfit,
                  aes(x = NMDS1_lab, y= NMDS2_lab, label = label_category),
                  inherit.aes = FALSE,
                   size = 2.5,
    max.overlaps = Inf,
    force = 2,
    max.iter = 20000,
    box.padding = 0.3,
    point.padding = 0.2)+
  theme_bw() +
     theme(
        panel.grid = element_blank(),
         legend.position = "right",
         text = element_text(size = 9)
     ) +
     labs(
         x = "NMDS1",
        y = "NMDS2",
         colour = "Shelf",
        shape = "Year"
    )+
  scale_x_continuous(limits = c(-1.3, 1.05))+
  scale_y_continuous(limits = c(-1, 0.9))
```

```{r}
Benthic_Figure <- ggarrange(
  Benthic_A, Benthic_B, labels = c("A", "B"), heights = c(2, 1), ncol = 1
)

Benthic_Figure
```

```{r eval=FALSE}
ggsave(Benthic_Figure,
  filename = "figures/facet_benthic_NMDS.PDF",
  height = 280,
  width = 240,
  units = "mm"
)
```
